### 2025-02-26 ms.qmd
updates

---
docx: 
reference-doc: "./bib/tmpl.docx"
csl: "./bib/proceedings-of-the-royal-society-b.csl"
bibliography: "./bib/refs.bib"
execute:
  echo: false
  error: false
  cache: false
  warning: false
crossref:  
  fig-title: Figure    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Tab.   # (default is "Table")
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
```

```{r, cleandata}
#| label: cleandata
# Obtain the main df using "./R/1_data_process.R"
source(here("R", "data_process.R"))
```

```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the main database.
source(here("R", "func.R"))
#
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(k in 1:length(hormone)){
  for(l in 1:length(temperature)){
    list_name <- paste0(hormone[k], "_", temperature[l])
    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
  }
}
#
```

```{r, countclutches}
#| label: countclutches
# Count the number of clutches per species
#
clutches <- clean_df %>% 
  distinct(clutch) %>% 
  nrow()
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

## Introduction

Cognition encompasses how animals acquire, process, and store information, enabling perception, learning, memory, and decision-making [@shettleworth]. It is essential for survival and reproduction, allowing individuals to adapt to their environment effectively [@dukas_evolutionary_2004]. However, cognitive abilities can vary considerably between individuals, with differences arising from genetic factors, environmental conditions, or a combination of both [@sakata_neural_2000; @dukas_evolutionary_2004]. The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017]. For instance, plenty of studies have investigated how prenatal conditions can influence learning abilities and memory across different taxa [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020]. However, the impact of early-life environments on other cognitive domains has been less studied [but see @burger_antipredator_1998; @vila_pouca_quantity_2019]. A fundamental cognitive skill is acquiring information from the environment [@shettleworth]. Perception of different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific]. Variations in the ability to detect or recognize some relevant stimuli can have significant consequences for an individual’s fitness. For example, individuals better at detecting prey might be more competitive in foraging contexts, which will affect the energy they can obtain and allocate for other processes like growth or reproduction. Therefore, understanding how prenatal conditions affect the ability to detect prey can be essential to understanding the broader consequences of early-life environments on fitness. Yet, little is known about how prenatal conditions influence prey-related cognition [but see @burger_effects_1990].

The prenatal environment influences cognitive abilities because the brain is particularly sensitive to environmental inputs during the early stages of development [@zhu_prenatal_2004]. Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production [@amani2021perinatal], or brain structure [@amiel_effects_2017]. One key mechanism linking cognition and brain function is mitochondrial physiology and their effects on energy production and oxidative stress [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018]. Given the high energetic demands of cognitive processes [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], cognitive performance is expected to depend on efficient mitochondrial respiration. Additionally, high cognitive abilities are typically associated with increased neuron density and functionality [@lefebvre_taxonomic_2011; @amiel_effects_2017], both of which can be impaired by excessive ROS production and oxidative stress byproduct of mitochondrial function [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018]. For instance, @hara_presynaptic_2014 found that performance in visuospatial working-memory tasks correlated negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic bouton. These findings emphasize the role of mitochondrial action in shaping cognitive abilities. However, how the early environment influences mitochondrial physiology and how that translates into cognitive consequences remains largely unknown in non-mammalian animals [see @chaudhari2022early for a review]. In particular, whether and how these interactions affect perception abilities, such as prey detection, is still largely unexplored.

Maternal conditions can play a crucial role in shaping mitochondrial function in the offspring. Because mitochondria are maternally inherited, a mother’s physiological state can directly impact mitochondria through trans-generational effects [@zhu_prenatal_2004; @song2009prenatal; @picard_mitochondria_2014]. One key factor in this process is the action of glucocorticoids (GCs), hormones that regulate energy metabolism and stress responses [@sapolsky_how_2000]. Under stressful situations, animals instigate adaptive physiological and behavioural adjustments that involve increases in GCs. These GCs can be transmitted to the offspring and influence their mitochondrial functioning [@picard_mitochondria_2014]. For instance, prenatal stress or elevated yolk GCs increase oxidative stress in Sprague-Dawley rats (_Rattus norvegicus_) and domestic chickens (_Gallus domesticus_) [@zhu_prenatal_2004; @song2009prenatal; @haussmann_embryonic_2012].

In ectotherms, thermal conditions can be a significant source of maternal stress. When animals experience temperatures outside their optimal thermal range, they activate the physiological stress response, leading to increased GC production that can be passed to the offspring [see @Crino_2023]. However, the temperature can also directly affect offspring development beyond maternal effects, particularly during early life [@noble_developmental_2018; @crino2024eggs]. The prenatal thermal environment can play a critical role in shaping mitochondrial function, altering energy metabolism or oxidative stress [@stier2022experimental; @crino2024eggs]. For example, Red-eared slider turtle (_Trachemys scripta elegans_) hatchlings incubated at 26.5 ºC exhibited greater oxidative damage compared to those incubated at temperatures 0.7 to 1.2 ºC warmer [@treidel2016temperature]. Therefore, the combined effects of prenatal GCs and temperature can have profound consequences for mitochondrial function, with significant repercussions for brain development and cognition. Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities through mitochondrial physiology.

Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial functioning and prey detection in the delicate skink (_Lampropholis delicata_). We manipulated prenatal CORT levels and incubation temperature, then assessed hatchlings’ ability to detect chemical and visual prey stimuli through behavioral tests. After these tests, we examined mitochondrial function in different brain regions associated with processing chemical and visual cues. We hypothesized that prenatal CORT and temperature would interact to influence mitochondrial physiology and prey detection. GCs can alter mitochondrial function by increasing oxidative stress [reviewed in @costantini2011meta] and reducing mitochondrial efficiency [@gong_chronic_2011;  but see @crino2024eggs]. Similarly, high temperatures can decrease mitochondrial efficiency but also oxidative damage [@treidel2016temperature; @crino2024eggs]. Therefore, we predicted that prenatal CORT exposure would impair mitochondrial function by increasing ROS levels and reducing energy availability, disrupting cognitive performance. However, higher incubation temperatures would increase performance only when the cost of higher oxidative stress overcome the benefits of increased mitochondrial efficiency. In addition, we predicted complex interactions between CORT and incubation temperature, as both factors negatively impact mitochondrial efficiency but have opposite effects on oxidative stress. Finally, since early environmental effects on mitochondrial function can vary across brain regions [@coomber_independent_1997], we expected regional differences in mitochondrial physiology accross treatments, which in turn would lead to stimulus-dependent variations in detection abilities. Our study provides insights into how prenatal conditions shape mitochondrial physiology and cognitive abilities, emphasizing the importance of early-life environments in cognitive development.

## Methods

#### Husbandry

#### Animal husbandry  
*Breeding colony* -- The lizards tested came from a breeding colony established in the laboratory since 2019. This colony consisted of 270 adults of *L. delicata* housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure. Enclosures were provided with non-stick matting, shelter, and several small water dishes. Water was given daily, and the lizards were fed approx. 40 mid-size crickets (_Acheta domestica_) per enclosure three days a week. Crickets were dusted with calcium weekly and multivitamin and calcium biweekly. Room temperatures were set to 22-24 ºC, but to ensure a temperature gradient, we provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.

*Eggs collection and incubation* -- Between mid-October 2022 and the end of February 2023, we provide females with a place to lay the eggs by placing a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures. These boxes were checked three days a week for eggs. After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.

*Hatchlings* -- Incubators were checked three times a week for hatchlings. We measured and weighted lizards immediately after hatching. We employed a rule to obtain Snout-vent length (SVL) and Tail length (TL) to the nearest mm, and a digital scale (± 0.001g error) to measure mass. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with nonstick matting and a small water dish until the beginning of the experiment. During this period, lizards were given water dayly and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).

Two weeks before we started the tests (see below), lizards were moved to the experimental arena (see @fig-Methods) for acclimatization. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the middle of the arena. These arenas were placed in two rooms in six racks, each with its own CCTV system (device model DVR-HP210475) that allowed us to record lizards' behaviour during the experiment (see details below). The number of lizards per treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatization, lizards were fed with only one cricket per day dusted with calcium and multivitamin, and water was supplied *ad libitum*. We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The rooms temperature was set to between 22-24 Celsius. After the tests, animals were euthanized and mitochondrial physiology was analyzed in various brain regions (see Brain mitochondrial activity protocol below).

#### CORT and Temperature manipulation
To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A). Eggs were topically supplied with: a) 5 µL of crystalline corticosterone (Sigma, Cat. No. C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or b) an equal volume of 100% Ethanol (Control treatment). We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [@crino2024eggs]. Then, eggs were incubated in one of the two previously mentioned temperature regimes. These temperatures are within the natural limits in _L. delicata_ [@cheetham2011embryonic].

The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design. At least one egg per clutch was assigned to each treatment, while in cluthces bigger than four the remaining eggs were randomly assigned to one of the treatments. When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.

#### Prey discrimination tests
The experiment involved presenting lizards with various stimuli from different prey types, then recording and analyzing their behaviour towards each stimulus. We followed the methods employed by @desfilis2003stimulus and @recio_prey_2021, with some modifications relevant to the objectives of this study. Lizards were presented chemical or visual stimuli from a familiar and unfamiliar prey, with a total of four trials per individual (@fig-Methods C).

We used crickets (_A. domestica_) as the familiar prey and mealworm larvae (_Tenebrio molitor_) as the unfamiliar prey. None of the species are included in the natural diet of the lizards [@lunney1989diets], but both are reported to be eaten in lab conditions [@michelangeli2016sa]. We expected to see differences between known and unknown prey because previous experience may influence stimuli perception through habituation or sensitisation [@burger_effects_1990; @burger_effects_1991; @desfilis2003stimulus], and we included familiarity with the prey as a factor in our analyses (see below).

Each stimulus was presented inside a transparent plastic vessel containing a white, two-chambered device (see @fig-Methods B) made of polylactic acid. In chemical trials, the prey was placed in the closed chamber at the back of the device, making it invisible to the lizard, while in visual trials, the prey was placed in the front chamber. Holes in both the device and the front sides of the transparent vessel (see @fig-Methods B, C) allowed chemical cues to be released; however, these holes were sealed with silicone in the visual trials. To increase the availability of chemical cues, we glued a piece of filter paper (left for at least 8 hours in one of the prey's enclosures: _A. domestica_ or _T. molitor_) to the device during chemical trials. In visual trials, the filter paper was placed in an empty box for the same duration under identical conditions. In both chemical and visual trials, the prey remained inside the vessel to control for potential acoustic cues. The order of stimulus presentation was counterbalanced across treatments.

Each trial began by placing the experimental device in the side of the arena opposite to the shelter (see arena in @fig-Methods A), and then removing the shelter. The water cup had already been removed. We recorded the lizard's behaviour for approximately one hour. We assessed lizards' ability to detect each stimulus ('Detection latency') by recording the time from when the lizard resumed normal activity (i.e., walking for at least 5 consecutive seconds; T~0~ in @fig-Methods D) until the first interaction with the stimulus (T~D~ in @fig-Methods D). We consider an interatcion when the lizard touched the front of the vessel or the filter paper with its snout for more than five consecutive seconds.

To control for potential differences in hunger levels, all lizards fasted for two days before the experiment, a period considered harmless for this species [@young2022physiological]. Besides, after each trial, the lizards were given a cricket to assess their motivation to forage. The cricket was left in the enclosure for one hour, and we recorded whether the lizard ate it (recorded as 1) or not (recorded as 0). In `r sum(is.na(clean_df$motivation))` videos, the camera stopped recording before the end of the motivation test (T~f~ in @fig-Methods D), so motivation was recorded as NA. We used their performance in the motivation test as a covariate in the analyses (see below).

All trials were conducted between 1100 and 1300 h, when the lizards were most active. To control for potential effects of neophobia, we simulated test conditions for two days prior to the experiment by removing the shelter and water cup, and exposing the subjects to the vessel without a stimulus. This simulation lasted for 1 hour at the same time of day as the tests, but no behavior was recorded.

```{r, fig-Methods}
#| label: fig-Methods
#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
#
knitr::include_graphics("./Others/Methods.png")
```

#### Brain mitochondrial activity
Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation. We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard's response by turning it upside down, and by testing the pinching relfex in one of the front limbs. Lizards were decapitated with surgical scissors once they were not responsive. This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33). After decapitation, the head was opened and the brain was dissected. We extracted manually two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with lizards' chemical and visual perception, respectively [@wyneken2007reptilian]. Both regions were transferred immediately to 1.5mL centrifuge tubes containing 100µL of 1X PBS.

Both regions were transferred immediately to 1.5 mL centrifuge tubes containing 100 µL of cold 1X PBS and kept on ice until further processing. For the olfactory bulbs, the tissue was mechanically homogenized by placing the tissue in the well of a 100 µm mesh filter (pluriStrainer) affixed atop a 1.5 mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper. The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1X PBS to prepare a homogenate suspension. Optic tecta were mechanically homogenized the same way as olfactory bulbs, but were first enzymatically digested by incubating the tissue in 100µL of 125 U/mL collagenase (type II) for 30-min at 32°C. We enzymatically digested optic tecta but not olfactory bulbs because in a pilot study, enzymatic digestion prior to mechanical digestion improved the yield of viable cells in optic tecta homogenate, but did not noticeably affect the yield of viable cells in olfactory bulbs homogenate (unpublished data). Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500 µL 1x PBS. This step was to performed to remove cellular debris from homogenates. From each 500 µL suspension of homogenate collected on a given trial day, we first added 100 µL of homogenate to a pooled sample of each tissue type to use for single-color controls, and the remaining 40 0µL of homogenate was split among two 200 µL aliquots. One aliquot was used fresh to measure mitochondrial function (mitochondrial density, membrane potential, ROS), one aliquot was cryopreserved for later measurements of oxidative damage (8-OHdG, lipid peroxidation), and the third aliquot was cryopreserved for a different experiment. Aliquots were cryopreserved in 1 mL solution of 1% Neutral-Buffered Formalin (as a fixative agent), 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant), then stored at -20 °C until oxidative damage assays.

From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50 µL of homogenate in duplicates (2 wells per homogenate). To each replicate well, we added 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red. We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively. We then added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris. We then loaded 6 wells with 50 µL of homogenate taken from each pooled homogenate suspension (12 wells total), which were to be negative and single-color controls. One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5 µL of one of 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, 50 µM MitoSOX Red, or #µM Hoechst 33342 Nuclear Viability Dye. Any remaining pooled homogenate was fixed and frozen as previously described. We incubated the loaded plate at 32 °C for 30 minutes to stain and then diluted the samples with 50 µL cold 1x PBS to halt the staining process. Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours. Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).

Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples. On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in near-boiling water. We washed each thawed sample twice, the first time resuspending the pelleted cells in 1 mL warm 1X Tris-EDTA, and the second time in 200 µL warm 1X PBS. Pooled homogenate was divided into 100 µL aliquots to serve as single-color controls. Thawed, resuspended samples were first stained with 20 µL (10 µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32 °C for 20 minutes. Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin. We incubated the samples at 32 °C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200 µL 1X PBS. We added 20 µL of 70 µM 8-OHdG Polyclonal Antibody to each sample, after which we left the homogenate overnight (~12 hours) for the antibody to bind to 8-OHdG, a marker of oxidative damage on DNA. The following day we counterstained the cells with 20 µL of #µM (100 µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32 °C for 20 minutes. After the cells had been tagged with 8-OHdG antibodies and counterstained, we washed the cells once more and resuspended the pellet in 400 µL of 1X PBS. Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate. Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated. We then loaded a 96-well plate with 100 µL of each single-color control and 100 µL in duplicate of each sample. We performed all flow cytometry assays on samples within 48-hours of thawing the samples.

All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors. Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v. 8.0.1) and BD CS&T fluorescent beads (Lot No. 30664) diluted at 2 drops to 150 µL 1X PBS. During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector. These thresholds were chosen to filter small debris or inviable or non-intact cells from our observations. The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment. Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events. For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels. For the oxidative damage assay, we recorded data from the following channels: FSC (425; forward scatter), SSC (300; side scatter), Alexa-Fluor 488 (275; 8-OHdG Antibody + Alexa-Fluor 488 conjugate), BUV-496 (525; Hoechst 33342), and PE-Cy5 (850; BODIPY 665/676). Fluorescent intensity data was collected via the BD FACSDiva (v. 8.0.1) software, with no compensation applied during data collection, and all on a linear scale (detectable range of 0-252166). We recorded data for both the area and height of the fluorescent signal, but only used the area in downstream analyses, with height being recorded for the sake of quality control. Data was exported from BD FACSDiva as individual *.fcs (“flow cytometry standard”) files for each sample, then imported into FlowJo (v. 10.1) for processing.

In FlowJo v. 10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events. We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains. We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103). For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103). For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103). When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population. We used the backgated population for compensation of fluorescent spillover between different fluorescent probes. To account for fluorescent spillover, we used a traditional compensation matrix using the compensation function of FlowJo v.10.1. We identified the “positive” population for each channel as the brightest ~2.5% of the distribution of observations in the respective single-color control for that channel and used unstained controls as a universal negative. We visually inspected the compensation matrix and its effects on population distributions for under- and over-compensation, whereupon we changed the compensation matrix manually until data was properly compensated. We applied the compensation matrix to all samples for downstream processing. Following compensation, we again gated the data following the same process as for backgating but using the compensated parameters for each channel. Following gating, we exported the geometric mean (mean fluorescent intensity; MFI) and robust confidence-values for each channel for each sample. For analysis, we exported summary statistics of only the area of the fluorescent signal. Although we exported robust confidence values for checking repeatability between replicate samples, we used the mean and standard deviation of geometric means among replicates for each individual as our main response variables in analyses.

#### Statistical analyses

We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, mitochondrial potential, ROS, DNA damage, lipid peroxidation, and detection latency) separately. We fitted first a set of preliminary models where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction. We also added to these models the sex and age of the lizards at the time where the trials started (for detection latency), or when the lizards were euthanized (for all mitochondrial-related variables). For detection latency we also included as fixed factors lizards' previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation, as CORT can impact apetite [@conde2018stress]. We also included for all the models the clutch identity as a random factor. Since detection latency was fitted for both types of prey in a repeated measures design, these models also included individual identity as a random factor. Structure and results of these models are provided on Tables S1-S12 in Supplementary Material. After the preliminary models, we fitted a set of definitive models where those factors that were not significant in the preliminary models were excluded except the main interest factors: CORT, temperature and their interaction. Random factors remain the same in the definitive models. All the response variables were mean centered and standardized by dividing by two times the standard deviation. Before standardization, mitochondrial density, DNA damage, lipid peroxidation, and detection latency were log-transformed.

All the models were run using the brm function from the _brms_ package [@burkner2017brms] in R (version 4.4.0) [@R]. We ran four parallel MCMC chains of 8000 iterations for each model, with a warmup period of 2000 iterations. The error structure was modelled assuming a gaussian distribution for all variables. To test for differences between treatments, we made contrasts between treatments using the posterior distributions of the parameters. We employed pMCMC values to tests if the contrasts were different from zero [@endo2019introduction]. We considered an effect statistically significant if pMCMC < 0.05.

To further explore the relationships between mitochondrial physiology and detection latency, we followed Structural Equations Modelling approach. We fitted a multivariate model using the _brms_ package [@burkner2017brms] in R (version 4.4.0) [@R] for each brain region/stimulus separately. We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem_results_OB or @fig-sem_results_OT). We also included for each variable those factors that were significant in our definitive models but we excluded the non-significant ones. Because experience with prey did not affect lizards' behaviour (see Tables S11, S12), and to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors. Clutch identity was included as a random factor for all variables. We ran four parallel MCMC chains of 8000 iterations for each model, with a warmup period of 2000 iterations. The error structure was modelled assuming a gaussian distribution. Rescor was set as true to control for potential correlations between the residuals of the variables.

We obtained direct, indirect, and total effects of one variable on another from the posterior parameters. We estimated direct effects using posterior estimates from the SEM models. Specifically, direct effects correspond to the posterior estimates of a predictor's effect on the response variable. For example, the direct effect of mitochondrial density on detection latency is given by the posterior estimate of detection latency regressed on mitochondrial density. Indirect effects were computed by multiplying the direct effect of the predictor on the mediator by the direct effect of the mediator on the response variable. For instance, the indirect effect of ROS production on detection latency was obtained by summing the products of (i) the direct effect of ROS on DNA damage multiplied by the direct effect of DNA damage on detection latency, and (ii) the direct effect of ROS on lipid peroxidation multiplied by the direct effect of lipid peroxidation on detection latency. Total effects were the sum of direct and indirect effects. Refer to @fig-sem_results_OB or @fig-sem_results_OT for a visual representation of the SEM models.

## Results

Final sample size for mitochondrial assays was 80 lizards, 20 per treatment. These lizards were subjected each to 4 tests getting a total of 320 behavioural observations, out of which `r sum(is.na(clean_df$t_D))` were recorded as NA. These animals came from a total of `r clutches` clutches.

```{r, models_mitochondrial}
#| label: models_mitochondrial
#
# Run models mitochondrial physiology (each region separately)
#
var_m <- c("mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
regions <- c("OB", "OT")
for (p in var_m){
  for (h in regions){
    if (h == "OB"){
      df <- clean_df %>% filter(region == "OB")
      l <- "OB"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    } else {
      df <- clean_df %>% filter(region == "OT")
      l <- "OT"
      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
        formula <- paste0(p, "~ cort*temp + (1|clutch)")
      } else if (p == "mean_dnadamage"){
        formula <- paste0(p, "~ cort*temp + sex + (1|clutch)")
      } else {
        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
      }
    }

    pmodel_name <- paste0("m_def_", p, "_", h)
    assign(pmodel_name, fit_m(df = df,
                              cat = "def",
                              var = p,
                              formula = formula,
                              fam = gaussian(),
                              label = l,
                              refit = FALSE),
          envir = .GlobalEnv)  # Assign to the global environment
  }
} 
```

```{r, models_behaviour}
#| label: models_behaviour
# Fitting the model and extraction of posteriors for Detection Latency (log-normal).
source(here("R", "func.R"))
#
#
## Run model behaviour for both stimuli separately
#
beh_df <- clean_df
stimuli <- c("Chemical", "Visual")
for (s in stimuli){
  df_b <- beh_df %>% filter(stimulus == s)
  formula_t_D <- t_D ~ cort*temp + (1|clutch) + (1|lizard_id)

  pmodel_name <- paste0("m_def_t_D_", s)
  assign(pmodel_name, fit_m(df = df_b,
                             cat = "def",
                             var = "t_D",
                             formula = formula_t_D,
                             fam = gaussian(),
                             label = s,
                             refit = FALSE),
          envir = .GlobalEnv)  # Assign to the global environment
}
```

```{r, organise_posteriors}
#| label: organise_posteriors
#
# Organising the posteriors of the previous models to fit the tables below
#
source(here("R", "func.R"))
#
# Databases for each region/stimulus
post_OB <- data.frame()
post_OT <- data.frame()
# Names posteriors:
region <- c("OB", "OT")
names_OB <- c("m_def_mean_mitodensity_OB", "m_def_mean_potential_OB", "m_def_mean_ros_OB", "m_def_mean_dnadamage_OB", "m_def_mean_peroxidation_OB", "m_def_t_D_Chemical")
names_OT <- c("m_def_mean_mitodensity_OT", "m_def_mean_potential_OT", "m_def_mean_ros_OT", "m_def_mean_dnadamage_OT", "m_def_mean_peroxidation_OT", "m_def_t_D_Visual")
#
# Organising the results
for (r in region) {
  model_select <- get(paste0("names_", r))   
  for (pos in model_select) {
    model <- get(pos)      # Get the model from the global environment
    post_result <- tidy_post(model)        # Apply tidy_post to each model
    # Add a new column to identify the region and model
    post_result$Region <- r
    post_result$Model <- pos
    # Append to the appropriate data frame
    if (r == "OB") {
      post_OB <- bind_rows(post_OB, post_result)
    } else {
      post_OT <- bind_rows(post_OT, post_result)
    }
  }
}
```

```{r, values_posteriors}
#| label: values_posteriors
#
# Extracting the posteriors for the models and the values of interest. Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
#
source(here("R", "func.R"))
#
# A) Olfactory bulbs/Chemical stimulus
#
MD_OB <- post_values(m_def_mean_mitodensity_OB, "none")
MP_OB <- post_values(m_def_mean_potential_OB, "none")
ROS_OB <- post_values(m_def_mean_ros_OB, "none")
DNA_OB <- post_values(m_def_mean_dnadamage_OB, "none")
LP_OB <- post_values(m_def_mean_peroxidation_OB, "none")
DET_OB <- post_values(m_def_t_D_Chemical, "none")
#
# B) Optic tecta/Visual stimulus
MD_OT <- post_values(m_def_mean_mitodensity_OT, "none")
MP_OT <- post_values(m_def_mean_potential_OT, "none")
ROS_OT <- post_values(m_def_mean_ros_OT, "none")
DNA_OT <- post_values(m_def_mean_dnadamage_OT, "sex")
LP_OT <- post_values(m_def_mean_peroxidation_OT, "none")
DET_OT <- post_values(m_def_t_D_Visual, "none")
```

#### Mitochondrial physiology

Results for mitochondrial analyses are presented on Tables S13-S15 (Supplementary Material) in means of R^2 scores and the estimates of each predictor in the difinitive models. Preliminary models can be found in Tables S1-S12 (Supplementary Material). @fig-results_energy and @fig-results_oxidative summarize the effects of early conditions on mitochondrial fuction.

_Mitochondrial density_: neither CORT (Contrast β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(MD_OB$Control_Cold, MD_OB$Control_Hot)) - mean(c(MD_OB$CORT_Cold, MD_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OB$Control_Cold, MD_OB$Control_Hot) - c(MD_OB$CORT_Cold, MD_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(MD_OT$Control_Cold, MD_OT$Control_Hot)) - mean(c(MD_OT$CORT_Cold, MD_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OT$Control_Cold, MD_OT$Control_Hot) - c(MD_OT$CORT_Cold, MD_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(MD_OB$CORT_Hot, MD_OB$Control_Hot)) - mean(c(MD_OB$CORT_Cold, MD_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OB$CORT_Hot, MD_OB$Control_Hot) - c(MD_OB$CORT_Cold, MD_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(MD_OT$CORT_Hot, MD_OT$Control_Hot)) - mean(c(MD_OT$CORT_Cold, MD_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OT$CORT_Hot, MD_OT$Control_Hot) - c(MD_OT$CORT_Cold, MD_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(MD_OB$Control_Hot) - mean(MD_OB$CORT_Hot)) - (mean(MD_OB$Control_Cold) - mean(MD_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MD_OB$Control_Hot - MD_OB$CORT_Hot) - (MD_OB$Control_Cold - MD_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(MD_OT$Control_Hot) - mean(MD_OT$CORT_Hot)) - (mean(MD_OT$Control_Cold) - mean(MD_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MD_OT$Control_Hot - MD_OT$CORT_Hot) - (MD_OT$Control_Cold - MD_OT$CORT_Cold)), 3, equal = TRUE)`) had significant effects on mitochondrial density in neither of the two brain regions extracted (@fig-results_energy A, C).

_Mitochondrial potential_: the effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(MP_OB$Control_Cold, MP_OB$Control_Hot)) - mean(c(MP_OB$CORT_Cold, MP_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OB$Control_Cold, MP_OB$Control_Hot) - c(MP_OB$CORT_Cold, MP_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(MP_OT$Control_Cold, MP_OT$Control_Hot)) - mean(c(MP_OT$CORT_Cold, MP_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OT$Control_Cold, MP_OT$Control_Hot) - c(MP_OT$CORT_Cold, MP_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(MP_OB$CORT_Hot, MP_OB$Control_Hot)) - mean(c(MP_OB$CORT_Cold, MP_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OB$CORT_Hot, MP_OB$Control_Hot) - c(MP_OB$CORT_Cold, MP_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(MP_OT$CORT_Hot, MP_OT$Control_Hot)) - mean(c(MP_OT$CORT_Cold, MP_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OT$CORT_Hot, MP_OT$Control_Hot) - c(MP_OT$CORT_Cold, MP_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(MP_OB$Control_Hot) - mean(MP_OB$CORT_Hot)) - (mean(MP_OB$Control_Cold) - mean(MP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MP_OB$Control_Hot - MP_OB$CORT_Hot) - (MP_OB$Control_Cold - MP_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(MP_OT$Control_Hot) - mean(MP_OT$CORT_Hot)) - (mean(MP_OT$Control_Cold) - mean(MP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MP_OT$Control_Hot - MP_OT$CORT_Hot) - (MP_OT$Control_Cold - MP_OT$CORT_Cold)), 3, equal = TRUE)`) were not significant on membrane potential neither for the Olfactory bulbs or the Optic tecta (@fig-results_energy B, D).

```{r, fig-results_energy}
#| label: fig-results_energy
#| fig-cap: "Estimates of mitochondrial density (A, C) and mitochondrial potential (B, D) in the olfactory bulbs (A, B) and optic tecta (C, D) of L. delicata hatchlings as a function of the different prenatal conditions. Black dots indicate the posterior mean, and the bars represent the SD of the estimates. The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions. Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
#| fig-name: "fig-results_energy"
#
source(here("R", "func.R"))
#
# A) Plotting the results for OB/Chemical stimulus using the df from before
plot_density_OB <- plotting(MD_OB, "Mit density")
plot_potential_OB <- plotting(MP_OB, "Mit potential")
#
legend_plot <- get_legend(plot_density_OB + theme(legend.position = "right"))
#
fig_OB_energy <- plot_grid(plot_density_OB, NULL, plot_potential_OB, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
#
# B) Plotting the results for OT/Visual stimulus
plot_density_OT <- plotting(MD_OT, "Mit density")
plot_potential_OT <- plotting(MP_OT, "Mit potential")
#
fig_OT_energy <- plot_grid(plot_density_OT, NULL, plot_potential_OT, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
#
# C) Merging plots
#
# Create the figure grid with extra space for images
fig_results_energy <- plot_grid(
  fig_OB_energy, fig_OT_energy,
  nrow = 2)
# Final composition: Merge everything, adding images and legend
final_plot_energy <- ggdraw(fig_results_energy) +
  # Insert OB image in the top right
  draw_image("./Others/OB.png", x = 0.7, y = 0.8, width = 0.12, height = 0.2) +
  # Insert OT image in the middle right
  draw_image("./Others/OT.png", x = 0.7, y = 0.3, width = 0.11, height = 0.18) +
  # Insert legend at the bottom-right
  draw_grob(legend_plot, x = 0.85, y = 0.15, width = 0.0001, height = 0.0001) +
  # Insert title for each plot
  annotate("text", x = 0.032, y = 0.965, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.39, y = 0.965, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.032, y = 0.47, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.39, y = 0.47, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold")
#
# Print final plot
ggsave(here("./output/figures/text/results_energy.png"), plot = final_plot_energy, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_energy.png")
```

_ROS_: we did not find significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(ROS_OB$Control_Cold, ROS_OB$Control_Hot)) - mean(c(ROS_OB$CORT_Cold, ROS_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OB$Control_Cold, ROS_OB$Control_Hot) - c(ROS_OB$CORT_Cold, ROS_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(ROS_OT$Control_Cold, ROS_OT$Control_Hot)) - mean(c(ROS_OT$CORT_Cold, ROS_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OT$Control_Cold, ROS_OT$Control_Hot) - c(ROS_OT$CORT_Cold, ROS_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(ROS_OB$CORT_Hot, ROS_OB$Control_Hot)) - mean(c(ROS_OB$CORT_Cold, ROS_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OB$CORT_Hot, ROS_OB$Control_Hot) - c(ROS_OB$CORT_Cold, ROS_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(ROS_OT$CORT_Hot, ROS_OT$Control_Hot)) - mean(c(ROS_OT$CORT_Cold, ROS_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OT$CORT_Hot, ROS_OT$Control_Hot) - c(ROS_OT$CORT_Cold, ROS_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(ROS_OB$Control_Hot) - mean(ROS_OB$CORT_Hot)) - (mean(ROS_OB$Control_Cold) - mean(ROS_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((ROS_OB$Control_Hot - ROS_OB$CORT_Hot) - (ROS_OB$Control_Cold - ROS_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(ROS_OT$Control_Hot) - mean(ROS_OT$CORT_Hot)) - (mean(ROS_OT$Control_Cold) - mean(ROS_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((ROS_OT$Control_Hot - ROS_OT$CORT_Hot) - (ROS_OT$Control_Cold - ROS_OT$CORT_Cold)), 3, equal = TRUE)`) on ROS levels in Olfactory bulbs nor Optic tecta (@fig-results_oxidative A, D).

_DNA damage_: there were not significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(DNA_OB$Control_Cold, DNA_OB$Control_Hot)) - mean(c(DNA_OB$CORT_Cold, DNA_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OB$Control_Cold, DNA_OB$Control_Hot) - c(DNA_OB$CORT_Cold, DNA_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(DNA_OT$Control_Cold, DNA_OT$Control_Hot)) - mean(c(DNA_OT$CORT_Cold, DNA_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OT$Control_Cold, DNA_OT$Control_Hot) - c(DNA_OT$CORT_Cold, DNA_OT$CORT_Hot)), 3, equal = TRUE)`) or temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(DNA_OB$CORT_Hot, DNA_OB$Control_Hot)) - mean(c(DNA_OB$CORT_Cold, DNA_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OB$CORT_Hot, DNA_OB$Control_Hot) - c(DNA_OB$CORT_Cold, DNA_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(DNA_OT$CORT_Hot, DNA_OT$Control_Hot)) - mean(c(DNA_OT$CORT_Cold, DNA_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OT$CORT_Hot, DNA_OT$Control_Hot) - c(DNA_OT$CORT_Cold, DNA_OT$Control_Cold)), 3, equal = TRUE)`) in none of the regions. The CORT-temperature interaction was not significant for the Olfactory bulbs ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OB$Control_Hot) - mean(DNA_OB$CORT_Hot)) - (mean(DNA_OB$Control_Cold) - mean(DNA_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OB$Control_Hot - DNA_OB$CORT_Hot) - (DNA_OB$Control_Cold - DNA_OB$CORT_Cold)), 3, equal = TRUE)`). However, there was a significant interaction between CORT and temperature in the Optic tecta ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot)) - (mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OT$Control_Hot - DNA_OT$CORT_Hot) - (DNA_OT$Control_Cold - DNA_OT$CORT_Cold)), 3, equal = TRUE)`), where CORT had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`) and decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E). Additionally, we found a significant effect of sex on DNA damage in the Optic tecta, and an effect of age on DNA damage in the Olfactory bulbs (Tables S13-S15 Supplementary Material).

_Lipid peroxidation_: we did not see significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(LP_OB$Control_Cold, LP_OB$Control_Hot)) - mean(c(LP_OB$CORT_Cold, LP_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OB$Control_Cold, LP_OB$Control_Hot) - c(LP_OB$CORT_Cold, LP_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(LP_OT$Control_Cold, LP_OT$Control_Hot)) - mean(c(LP_OT$CORT_Cold, LP_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OT$Control_Cold, LP_OT$Control_Hot) - c(LP_OT$CORT_Cold, LP_OT$CORT_Hot)), 3, equal = TRUE)`) ot temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(LP_OB$CORT_Hot, LP_OB$Control_Hot)) - mean(c(LP_OB$CORT_Cold, LP_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OB$CORT_Hot, LP_OB$Control_Hot) - c(LP_OB$CORT_Cold, LP_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(LP_OT$CORT_Hot, LP_OT$Control_Hot)) - mean(c(LP_OT$CORT_Cold, LP_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OT$CORT_Hot, LP_OT$Control_Hot) - c(LP_OT$CORT_Cold, LP_OT$Control_Cold)), 3, equal = TRUE)`) on lipid peroxidation in either the Olfactory bulbs nor the Optic tecta. However, the interaction between CORT and temperature was significant in the Optic tecta ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot)) - (mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OT$Control_Hot - LP_OT$CORT_Hot) - (LP_OT$Control_Cold - LP_OT$CORT_Cold)), 3, equal = TRUE)`), but not in the Olfactory bulbs ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OB$Control_Hot) - mean(LP_OB$CORT_Hot)) - (mean(LP_OB$Control_Cold) - mean(LP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OB$Control_Hot - LP_OB$CORT_Hot) - (LP_OB$Control_Cold - LP_OB$CORT_Cold)), 3, equal = TRUE)`) (@fig-results_oxidative C, F). We found that, in the Optic tecta, CORT has opposite effects on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), but differences between groups were not significant (see @fig-results_oxidative C).  In addition, age had significant effects on lipid peroxidation for both regions (Tables S13-S15 Supplementary Material).

```{r, fig-results_oxidative}
#| label: fig-results_oxidative
#| fig-cap: "Estimates of ROS (A, D), DNA damage (B, E), and lipid peroxidation (C, F) in the olfactory bulbs (A - C) and optic tecta (D - F) of L. delicata hatchlings as a function of the different prenatal conditions. Black dots indicate the posterior mean, and the bars represent the SD of the estimates. The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions. Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
#| fig-name: "fig-results_oxidative"
#
source(here("R", "func.R"))
#
# A) Plotting the results for OB/Chemical stimulus
plot_ros_OB <- plotting(ROS_OB, "ROS")
plot_dnadamage_OB <- plotting(DNA_OB, "DNA damage")
plot_peroxidation_OB <- plotting(LP_OB, "Lipid peroxidation")
#
legend_plot <- get_legend(plot_ros_OB + theme(legend.position = "right"))
#
fig_OB_oxidative_top <- plot_grid(plot_ros_OB, NULL, NULL, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, NULL, plot_peroxidation_OB, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
fig_OB_oxidative <- plot_grid(fig_OB_oxidative_top, fig_OB_oxidative_bottom,
                    nrow = 2)
#
# B) Plotting the results for OT/Visual stimulus
plot_ros_OT <- plotting(ROS_OT, "ROS")
plot_dnadamage_OT <- plotting(DNA_OT, "DNA damage")
plot_peroxidation_OT <- plotting(LP_OT, "Lipid peroxidation")
#
fig_OT_oxidative_top <- plot_grid(plot_ros_OT, NULL, NULL, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, NULL, plot_peroxidation_OT, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
fig_OT_oxidative <- plot_grid(fig_OT_oxidative_top, fig_OT_oxidative_bottom,
                    nrow = 2)
#
# C) Merging plots
#
# Create the figure grid with extra space for images
fig_results_oxidative <- plot_grid(
  fig_OB_oxidative, NULL, fig_OT_oxidative,
  nrow = 3, rel_heights = c(1, 0.05, 1))
# Final composition: Merge everything, adding images and legend
final_plot_oxidative <- ggdraw(fig_results_oxidative) +
  # Insert OB image in the top right
  draw_image("./Others/OB.png", x = 0.7, y = 0.795, width = 0.15, height = 0.25) +
  # Insert OT image in the middle right
  draw_image("./Others/OT.png", x = 0.7, y = 0.305, width = 0.14, height = 0.23) +
  # Insert legend at the bottom-right
  draw_grob(legend_plot, x = 0.85, y = 0.1, width = 0.0001, height = 0.0001) +
  # Insert title for each plot
  annotate("text", x = 0.0325, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0325, y = 0.735, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.3925, y = 0.735, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0325, y = 0.485, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.0325, y = 0.235, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.3925, y = 0.235, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("segment", x = 0.32, xend= 0.32, y = 0.17, yend = 0.215, size = 0.5, colour = "black") +
  annotate("segment", x = 0.31, xend= 0.32, y = 0.17, yend = 0.17, size = 0.5, colour = "black") +
  annotate("segment", x = 0.31, xend= 0.32, y = 0.215, yend = 0.215, size = 0.5, colour = "black") +
  annotate("text", x = 0.33, y = 0.187, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
#
# Print final plot
ggsave(here("./output/figures/text/results_oxidative.png"), plot = final_plot_oxidative, width = 21, height = 21, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_oxidative.png")
```

#### Prey detection

Results for behaviour analyses are presented on Tables S13-S15 (Supplementary Material) in means of R^2 scores and the estimates of each predictor in the difinitive models. Preliminary models can be found in Tables S1-S12 (Supplementary Material). @fig-results_behaviour summarizes the effects of early conditions on detection latency for both types of stimulus.

Lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature (β~Hot~ - β~Cold~: mean = `r format_dec(mean(c(DET_OB$CORT_Hot, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$CORT_Hot, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$Control_Cold)), 3, equal = TRUE)`), nor the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DET_OB$Control_Hot) - mean(DET_OB$CORT_Hot)) - (mean(DET_OB$Control_Cold) - mean(DET_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DET_OB$Control_Hot - DET_OB$CORT_Hot) - (DET_OB$Control_Cold - DET_OB$CORT_Cold)), 3, equal = TRUE)`) in chemo-detection (@fig-results_behaviour A). In contrast, there were no significant effects of CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OT$Control_Cold, DET_OT$Control_Hot)) - mean(c(DET_OT$CORT_Cold, DET_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OT$Control_Cold, DET_OT$Control_Hot) - c(DET_OT$CORT_Cold, DET_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~: mean = `r format_dec(mean(c(DET_OT$CORT_Hot, DET_OT$Control_Hot)) - mean(c(DET_OT$CORT_Cold, DET_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OT$CORT_Hot, DET_OT$Control_Hot) - c(DET_OT$CORT_Cold, DET_OT$Control_Cold)), 3, equal = TRUE)`), or their interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DET_OT$Control_Hot) - mean(DET_OT$CORT_Hot)) - (mean(DET_OT$Control_Cold) - mean(DET_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DET_OT$Control_Hot - DET_OT$CORT_Hot) - (DET_OT$Control_Cold - DET_OT$CORT_Cold)), 3, equal = TRUE)`) on the detection latency of visual stimulus (@fig-results_behaviour B).

```{r, fig-results_behaviour}
#| label: fig-results_behaviour
#| fig-cap: "Estimates of detection latency of chemical (A) and visual (B) stimulus by L. delicata hatchlings as a function of the different prenatal conditions. Black dots indicate the posterior mean, and the bars represent the SD of the estimates. The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions. Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
#| fig-name: "fig-results"
#
source(here("R", "func.R"))
# A) Plotting the results for OB/Chemical stimulus
plot_t_D_Chemical <- plotting(DET_OB, "Detection latency")
#
# B) Plotting the results for OT/Visual stimulus
plot_t_D_Visual <- plotting(DET_OT, "Detection latency")
#
legend_plot <- get_legend(plot_t_D_Chemical + theme(legend.position = "right"))
#
# C) Merging plots
#
# Create the figure grid with extra space for images
fig_results_behaviour <- plot_grid(plot_t_D_Chemical, NULL, plot_t_D_Visual, NULL,
                    nrow = 1, rel_widths = c(0.8, 0.5, 0.8, 0.7))
# Final composition: Merge everything, adding images and legend
final_plot_behaviour <- ggdraw(fig_results_behaviour) +
  # Insert CS image in the top middle
  draw_image("./Others/CS.png", x = 0.165, y = 0.47, width = 0.4, height = 0.5) +
  # Insert VS image in the top right
  draw_image("./Others/VS.png", x = 0.6, y = 0.47, width = 0.4, height = 0.5) +
  # Insert legend at the bottom-right
  draw_grob(legend_plot, x = 0.885, y = 0.23, width = 0.0001, height = 0.0001) +
  # Insert title for each plot
  annotate("text", x = 0.035, y = 0.96, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
  annotate("text", x = 0.5, y = 0.96, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +  
  annotate("segment", x = 0.27, xend= 0.27, y = 0.69, yend = 0.88, size = 0.5, colour = "black") +
  annotate("segment", x = 0.266, xend= 0.27, y = 0.69, yend = 0.69, size = 0.5, colour = "black") +
  annotate("segment", x = 0.266, xend= 0.27, y = 0.88, yend = 0.88, size = 0.5, colour = "black") +
  annotate("text", x = 0.28, y = 0.76, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
  annotate("segment", x = 0.27, xend= 0.27, y = 0.32, yend = 0.51, size = 0.5, colour = "black") +
  annotate("segment", x = 0.266, xend= 0.27, y = 0.32, yend = 0.32, size = 0.5, colour = "black") +
  annotate("segment", x = 0.266, xend= 0.27, y = 0.51, yend = 0.51, size = 0.5, colour = "black") +
  annotate("text", x = 0.28, y = 0.39, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
#
# Print final plot
ggsave(here("./output/figures/text/results_behaviour.png"), plot = final_plot_behaviour, width = 21, height = 7, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/results_behaviour.png")
```

#### SEM results

```{r, model_sem_OB}
#| label: model_sem_OB
# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
# To simplify the random factors, and since there is no effect of experience with the prey in prey detection, latency was averaged across the two preys. Every other predictor added to the model was based on previous brms for each variable separated.
# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
#
source(here("R", "func.R"))
#
SEM_df <- clean_df %>%
  group_by(lizard_id, prey) %>%
  mutate(t_D = mean(t_D, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(prey == "Unknown")
#
#
data_sem_OB <- SEM_df %>%
  filter(region == "OB")
# Create the models
#
refit <- FALSE
#
#
if(refit){
  m_OB <- brm(
    bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
    bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch)) +
    bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)) +
    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
  set_rescor(TRUE),
  family = gaussian(),
  data = data_sem_OB, 
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OB, file = here("output/m_SEM/m_OB.rds"))
} else {
  m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
}
```

```{r, sem_tidy_OB}
#| label: sem_tidy_OB
source(here("R", "func.R"))
#
# I am extracting here all the values for getting the total effects of each of the variables in the model. I am using the posterior values for each of the variables to get the total effects assuming that:
## total effect = direct effect + indirect effect + residual correlation
# In other words:
## total effect = 
# Extract the posteriors for the SEM model
post_sem_OB <- as_draws_df(m_OB) 
#
#### A) Get the direct paths per each variable
# 
# A.1) Detection
OB_coeff_cort_det <- post_sem_OB$b_tD_cortCORT
det_sem_OB_control <- post_sem_OB %>%
  dplyr::select(-b_tD_cortCORT) %>%
  dplyr::sample_n(size = 12000, replace = FALSE) # We take here 12000 random values for all the variables to get estimates for Control treatment, which is the reference level
det_sem_OB_cort <- post_sem_OB %>%
  mutate(across(everything(), ~.x + b_tD_cortCORT)) %>%
  dplyr::sample_n(size = 12000, replace = FALSE) %>% # We take here 12000 random values for all the variables AFTER adding to b_tD_cortCORT to all columns to get estimates for CORT lizards, which is NOT the reference level
  dplyr::select(-b_tD_cortCORT)
det_sem_OB_controlled <- bind_rows(det_sem_OB_control, det_sem_OB_cort)
#
OB_coeff_mitodensity_det <- det_sem_OB_controlled$b_tD_mean_mitodensity
OB_coeff_potential_det <- det_sem_OB_controlled$b_tD_mean_potential
OB_coeff_dna_det <- det_sem_OB_controlled$b_tD_mean_dnadamage
OB_coeff_perox_det <- det_sem_OB_controlled$b_tD_mean_peroxidation
# 
# A.2) DNA damage
OB_coeff_age_dna <- post_sem_OB$b_meandnadamage_age_euthanasia
OB_coeff_ros_dna <- post_sem_OB$b_meandnadamage_mean_ros
#
# A.3) Lipid peroxidation
OB_coeff_age_perox <- post_sem_OB$b_meanperoxidation_age_euthanasia
OB_coeff_ros_perox <- post_sem_OB$b_meanperoxidation_mean_ros
#
# A.4) ROS
OB_coeff_mitodensity_ros <- post_sem_OB$b_meanros_mean_mitodensity
OB_coeff_potential_ros <- post_sem_OB$b_meanros_mean_potential
#
#
#### B) Get the indirect paths for each variable
#
# B.1) Detection
OB_undir_ros_det <- OB_coeff_ros_dna * OB_coeff_dna_det + OB_coeff_ros_perox * OB_coeff_perox_det
OB_undir_mitodensity_det <- OB_coeff_mitodensity_ros * OB_coeff_ros_dna * OB_coeff_dna_det + OB_coeff_mitodensity_ros * OB_coeff_ros_perox * OB_coeff_perox_det
OB_undir_potential_det <- OB_coeff_potential_ros * OB_coeff_ros_dna * OB_coeff_dna_det + OB_coeff_potential_ros * OB_coeff_ros_perox * OB_coeff_perox_det
OB_undir_age_det <- OB_coeff_age_dna * OB_coeff_dna_det + OB_coeff_age_perox * OB_coeff_perox_det
#
# B.2) DNA damage
OB_undir_mitodensity_dna <- OB_coeff_mitodensity_ros * OB_coeff_ros_dna
OB_undir_potential_dna <- OB_coeff_potential_ros * OB_coeff_ros_dna
#
# B.3) Lipid peroxidation
OB_undir_mitodensity_perox <- OB_coeff_mitodensity_ros * OB_coeff_ros_perox
OB_undir_potential_perox <- OB_coeff_potential_ros * OB_coeff_ros_perox
#
#
#### C) Get the total effects for each variable
#
# C.1) Detection
OB_total_cort_det <- OB_coeff_cort_det
OB_total_mitodensity_det <- OB_undir_mitodensity_det
OB_total_potential_det <- OB_undir_potential_det
OB_total_ros_det <- OB_undir_ros_det
OB_total_dna_det <- OB_coeff_dna_det
OB_total_perox_det <- OB_coeff_perox_det
OB_total_age_det <- OB_undir_age_det
#
# C.2) DNA damage
OB_total_age_dna <- OB_coeff_age_dna
OB_total_ros_dna <- OB_coeff_ros_dna
OB_total_mitodensity_dna <- OB_undir_mitodensity_dna
OB_total_potential_dna <- OB_undir_potential_dna
#
# C.3) Lipid peroxidation
OB_total_age_perox <- OB_coeff_age_perox
OB_total_ros_perox <- OB_coeff_ros_perox
OB_total_mitodensity_perox <- OB_undir_mitodensity_perox
OB_total_potential_perox <- OB_undir_potential_perox
#
# C.4) ROS
OB_total_mitodensity_ros <- OB_coeff_mitodensity_ros
OB_total_potential_ros <- OB_coeff_potential_ros
#
#
# D) Create a df with the values for each variable
#
# D.1) Detection (for example)
det_results_semOB <- data.frame(
  reference_variable = rep("t_D", 7),
  predictor_modulator = c("cort",
                          "age",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS",
                          "mean_dnadamage",
                          "mean_peroxidation"),
  direct_effects = I(list(OB_coeff_cort_det,
                          NA,
                          OB_coeff_mitodensity_det,
                          OB_coeff_potential_det,
                          NA,
                          OB_coeff_dna_det,
                          OB_coeff_perox_det)),
  indirect_effects = I(list(NA,
                          OB_undir_age_det,
                          OB_undir_mitodensity_det,
                          OB_undir_potential_det,
                          OB_undir_ros_det,
                          NA,
                          NA)),
  total_effects = I(list(OB_total_cort_det,
                        OB_total_age_det,
                        OB_total_mitodensity_det,
                        OB_total_potential_det,
                        OB_total_ros_det,
                        OB_total_dna_det,
                        OB_total_perox_det))
  ) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
  mutate(source = "det_results_semOB")
#
# D.2) DNA damage
dna_results_semOB <- data.frame(
  reference_variable = rep("DNA damage", 4),
  predictor_modulator = c("age",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS"),
  direct_effects = I(list(OB_coeff_age_dna,
                          NA,
                          NA,
                          OB_coeff_ros_dna)),
  indirect_effects = I(list(NA,
                          OB_undir_mitodensity_dna,
                          OB_undir_potential_dna,
                          NA)),
  total_effects = I(list(OB_total_age_dna,
                        OB_total_mitodensity_dna,
                        OB_total_potential_dna,
                        OB_total_ros_dna))) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
  mutate(source = "dna_results_semOB")
#
# D.3) Lipid peroxidation
perox_results_semOB <- data.frame(
  reference_variable = rep("lipid peroxidation", 4),
  predictor_modulator = c("age",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS"),
  direct_effects = I(list(OB_coeff_age_perox,
                          NA,
                          NA,
                          OB_coeff_ros_perox)),
  indirect_effects = I(list(NA,
                          OB_undir_mitodensity_perox,
                          OB_undir_potential_perox,
                          NA)),
  total_effects = I(list(OB_total_age_perox,
                        OB_total_mitodensity_perox,
                        OB_total_potential_perox,
                        OB_total_ros_perox))) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
  mutate(source = "perox_results_semOB")
#
# D.4) ROS
ros_results_semOB <- data.frame(
  reference_variable = rep("ROS", 2),
  predictor_modulator = c("mean_mitodensity",
                          "mean_potential"),
  direct_effects = I(list(OB_coeff_mitodensity_ros,
                          OB_coeff_potential_ros)),
  indirect_effects = I(list(NA,
                          NA)),
  total_effects = I(list(OB_total_mitodensity_ros,
                        OB_total_potential_ros))) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
  mutate(source = "ros_results_semOB")
#
#
# E) Merge everything into a single df
#
sem_results_OB <- bind_rows(det_results_semOB, dna_results_semOB, perox_results_semOB, ros_results_semOB)
```

Results for SEM models are presented on Tables S16-S18 (Supplementary Material) in means of R^2 scores and the direct, indirect, and total effects for all variables. @fig-sem_results_OB and @fig-sem_results_OT summarize direct effects between the variables included in the model for bot brain regions and stimuli.

_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found significant increases of ROS with increases in mitochondrial density (Direct effect estimation: β~D~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_det), 3, equal = TRUE)`) and mitochondrial potential (β~D~ = `r format_dec(mean(OB_coeff_potential_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_potential_det), 3, equal = TRUE)`) (see @fig-sem_results_OB). However, there were no significant direct effects of ROS on DNA damage or lipid peroxidation; neither of mitochondrial density, mitochondrial potential, DNA damage, or lipid peroxidation on the detection abilities (see @fig-sem_results_OB and Table S17 in Supplementary Material). Age increased DNA damage (β~D~ = `r format_dec(mean(OB_coeff_age_dna), 3)`m pMCMC `r format_p(pmcmc(OB_coeff_age_dna), 3, equal = TRUE)`) and lipid peroxidation (β~D~ = `r format_dec(mean(OB_coeff_age_perox), 3)`m pMCMC `r format_p(pmcmc(OB_coeff_age_perox), 3, equal = TRUE)`) in the Olfactory bulbs, but there was no indirect effect on detecion latency (Indirect effect estimation β~I~ = `format_dec(mean(OB_undir_age_det), 3)`m pMCMC `r format_p(pmcmc(OB_undir_age_det), 3, equal = TRUE)`) Additionally, CORT shows a week effect on detection latency (β~D~ = `r format_dec(mean(OB_coeff_cort_det), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_cort_det), 3, equal = TRUE)`), suggesting that CORT treated animals detect chemical stimulus faster; but the effect here was not significant.

```{r, fig-sem_results_OB}
#| label: fig-sem_results_OB
#| fig-cap: "Structural Equation Models for OB/Chemical stimulus"
#| fig-name: "fig-sem_results_OB"
#
# A) Getting all the direct coefficients
OB_cort_det <- paste0(format_dec(mean(OB_coeff_cort_det), 3),
                  " [", format_dec(quantile(OB_coeff_cort_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_cort_det, 0.95), 3), "]")
OB_density_det <- paste0(format_dec(mean(OB_coeff_mitodensity_det), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_det, 0.95), 3), "]")
OB_potential_det <- paste0(format_dec(mean(OB_coeff_potential_det), 3),
                  " [", format_dec(quantile(OB_coeff_potential_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_det, 0.95), 3), "]")
OB_dna_det <- paste0(format_dec(mean(OB_coeff_dna_det), 3),
                  " [", format_dec(quantile(OB_coeff_dna_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_dna_det, 0.95), 3), "]")
OB_perox_det <- paste0(format_dec(mean(OB_coeff_perox_det), 3),
                  " [", format_dec(quantile(OB_coeff_perox_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_perox_det, 0.95), 3), "]")
#
OB_age_dna <- paste0(format_dec(mean(OB_coeff_age_dna), 3),
                  " [", format_dec(quantile(OB_coeff_age_dna, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_age_dna, 0.95), 3), "]")
OB_ros_dna <- paste0(format_dec(mean(OB_coeff_ros_dna), 3),
                  " [", format_dec(quantile(OB_coeff_ros_dna, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_dna, 0.95), 3), "]")
#
OB_age_perox <- paste0(format_dec(mean(OB_coeff_age_perox), 3),
                  " [", format_dec(quantile(OB_coeff_age_perox, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_age_perox, 0.95), 3), "]")
OB_ros_perox <- paste0(format_dec(mean(OB_coeff_ros_perox), 3),
                  " [", format_dec(quantile(OB_coeff_ros_perox, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_perox, 0.95), 3), "]")
#
OB_density_ros <- paste0(format_dec(mean(OB_coeff_mitodensity_ros), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_ros, 0.95), 3), "]")
OB_potential_ros <- paste0(format_dec(mean(OB_coeff_potential_ros), 3),
                  " [", format_dec(quantile(OB_coeff_potential_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_ros, 0.95), 3), "]")
#
imgOB <- readPNG(here("Others", "SEM_OB.png"))
plot_SEM_OB <- rasterGrob(imgOB, interpolate = TRUE)
#
fig_SEM_OB <- ggdraw(plot_SEM_OB) +
  annotate("text", x = 0.815, y = 0.59, label = OB_cort_det, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.35, y = 0.26, label = OB_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.35, y = 0.918, label = OB_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.745, label = OB_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.455, label = OB_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.55, y = 0.538, label = OB_age_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.813, label = OB_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.57, y = 0.115, label = OB_age_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.365, label = OB_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.285, y = 0.355, label = OB_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.285, y = 0.825, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/SEM_OB.png")
```

```{r, models_sem_OT}
#| label: models_sem_OT
# Making the same models above but for OT/Visual stimulus
#
source(here("R", "func.R"))
#
#
data_sem_OT <- SEM_df %>%
  filter(region == "OT")
#
# Create the models
#
refit <- FALSE
#
if(refit){
  m_OT <- brm(
    bf(t_D ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
    bf(mean_dnadamage ~ cort*temp + sex + mean_ros + (1|clutch)) +
    bf(mean_peroxidation ~ cort*temp + age_euthanasia + mean_ros + (1|clutch)) +
    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
  set_rescor(TRUE),
  family = gaussian(),
  data = data_sem_OT, 
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OT, file = here("output/m_SEM/m_OT.rds"))
} else {
  m_OT <- readRDS(here("output/m_SEM/m_OT.rds"))
}
#
```

```{r, sem_tidy_OT}
#| label: sem_tidy_OT
source(here("R", "func.R"))
#
# I am extracting here all the values for getting the total effects of each of the variables in the model. I am using the posterior values for each of the variables to get the total effects assuming that:
## total effect = direct effect + indirect effect + residual correlation
# In other words:
## total effect = 
# Extract the posteriors for the SEM model
post_sem_OT <- as_draws_df(m_OT) 
#
#### A) Get the direct paths per each variable
# 
# A.1) Detection
OT_coeff_mitodensity_det <- post_sem_OT$b_tD_mean_mitodensity
OT_coeff_potential_det <- post_sem_OT$b_tD_mean_potential
OT_coeff_dna_det <- post_sem_OT$b_tD_mean_dnadamage
OT_coeff_perox_det <- post_sem_OT$b_tD_mean_peroxidation
#
# A.2) DNA damage
#
# Getting CORT coefficient for both temperatures controlling for each sex
dna_sem_OT_male <- post_sem_OT %>%
  dplyr::select(-b_meandnadamage_sexFemale)
dna_sem_OT_fem <- post_sem_OT %>%
  mutate(across(everything(), ~.x + b_meandnadamage_sexFemale)) %>%
  dplyr::select(-b_meandnadamage_sexFemale) # Add effect of sex to everything to see estimates on females (males the reference level for sex)

dna_sem_OT_male_trt <- dna_sem_OT_male %>%
  dplyr::sample_n(size = 12000, replace = FALSE)
dna_sem_OT_fem_trt <- dna_sem_OT_fem %>%
  dplyr::sample_n(size = 12000, replace = FALSE)

dna_sem_OT_trt <- bind_rows(dna_sem_OT_male_trt, dna_sem_OT_fem_trt)

OT_coeff_cort_dna_cold <- dna_sem_OT_trt$b_meandnadamage_cortCORT
OT_coeff_cort_dna_hot <- dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_cortCORT + dna_sem_OT_trt$b_meandnadamage_tempHot + dna_sem_OT_trt$`b_meandnadamage_cortCORT:tempHot` - (dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_tempHot)
# Getting ros coefficient controlling for each level of the cort-temp interaction and controlling for each sex
dna_sem_OT_male_controlcold <- dna_sem_OT_male %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_controlhot <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_cortcold <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_corthot <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)

dna_sem_OT_fem_controlcold <- dna_sem_OT_fem %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_controlhot <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_cortcold <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_corthot <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)

dna_sem_OT_controlled <- bind_rows(dna_sem_OT_male_cortcold,
                                  dna_sem_OT_male_corthot,
                                  dna_sem_OT_male_controlcold,
                                  dna_sem_OT_male_controlhot,
                                  dna_sem_OT_fem_cortcold,
                                  dna_sem_OT_fem_corthot,
                                  dna_sem_OT_fem_controlcold,
                                  dna_sem_OT_fem_controlhot)

OT_coeff_ros_dna <- dna_sem_OT_controlled$b_meandnadamage_mean_ros
#
# Getting sex coefficient controlling for each level of the cort-temp interaction
dna_sem_OT_controlcold <- post_sem_OT %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_cortcold <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_controlhot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_corthot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)

dna_sem_OT_sex <- bind_rows(dna_sem_OT_cortcold,
                            dna_sem_OT_controlcold,
                            dna_sem_OT_corthot,
                            dna_sem_OT_controlhot)

OT_coeff_sex_dna <- dna_sem_OT_sex$b_meandnadamage_sexFemale
#
# A.3) Lipid peroxidation
OT_coeff_cort_perox_cold <- post_sem_OT$b_meanperoxidation_cortCORT
OT_coeff_cort_perox_hot <- post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_cortCORT + post_sem_OT$b_meanperoxidation_tempHot + post_sem_OT$`b_meanperoxidation_cortCORT:tempHot` - (post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_tempHot)

# Getting the coefficient of age and ros controlling for each level of the cort-temp interaction
perox_sem_OT_controlcold <- post_sem_OT %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_cortcold <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_controlhot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_tempHot)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_corthot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT + b_meanperoxidation_tempHot +
                `b_meanperoxidation_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)

perox_sem_OT_controlled <- bind_rows(perox_sem_OT_cortcold,
                            perox_sem_OT_controlcold,
                            perox_sem_OT_corthot,
                            perox_sem_OT_controlhot)

OT_coeff_ros_perox <- perox_sem_OT_controlled$b_meanperoxidation_mean_ros
OT_coeff_age_perox <- perox_sem_OT_controlled$b_meanperoxidation_age_euthanasia
#
# A.4) ROS
OT_coeff_mitodensity_ros <- post_sem_OT$b_meanros_mean_mitodensity
OT_coeff_potential_ros <- post_sem_OT$b_meanros_mean_potential
#
#
#### B) Get the indirect paths for each variable
#
# B.1) Detection
OT_undir_cort_det_COLD <- OT_coeff_cort_dna_cold * OT_coeff_dna_det + OT_coeff_cort_perox_cold * OT_coeff_perox_det
OT_undir_cort_det_HOT <- OT_coeff_cort_dna_hot * OT_coeff_dna_det + OT_coeff_cort_perox_hot * OT_coeff_perox_det
OT_undir_age_det <- OT_coeff_age_perox * OT_coeff_perox_det
OT_undir_sex_det <- OT_coeff_sex_dna * OT_coeff_dna_det
OT_undir_ros_det <- OT_coeff_ros_dna * OT_coeff_dna_det + OT_coeff_ros_perox * OT_coeff_perox_det
OT_undir_mitodensity_det <- OT_coeff_mitodensity_ros * OT_coeff_ros_dna * OT_coeff_dna_det + OT_coeff_mitodensity_ros * OT_coeff_ros_perox * OT_coeff_perox_det
OT_undir_potential_det <- OT_coeff_potential_ros * OT_coeff_ros_dna * OT_coeff_dna_det + OT_coeff_potential_ros * OT_coeff_ros_perox * OT_coeff_perox_det
#
# B.2) DNA damage
OT_undir_mitodensity_dna <- OT_coeff_mitodensity_ros * OT_coeff_ros_dna
OT_undir_potential_dna <- OT_coeff_potential_ros * OT_coeff_ros_dna
#
# B.3) Lipid peroxidation
OT_undir_mitodensity_perox <- OT_coeff_mitodensity_ros * OT_coeff_ros_perox
OT_undir_potential_perox <- OT_coeff_potential_ros * OT_coeff_ros_perox
#
#
#### C) Get the total effects for each variable
#
# C.1) Detection
OT_total_cort_det_COLD <- OT_undir_cort_det_COLD
OT_total_cort_det_HOT <- OT_undir_cort_det_HOT
OT_total_age_det <- OT_undir_age_det
OT_total_sex_det <- OT_undir_sex_det
OT_total_mitodensity_det <- OT_undir_mitodensity_det
OT_total_potential_det <- OT_undir_potential_det
OT_total_ros_det <- OT_undir_ros_det 
OT_total_dna_det <- OT_coeff_dna_det 
OT_total_perox_det <- OT_coeff_perox_det 
#
# C.2) DNA damage
OT_total_cort_dna_COLD <- OT_coeff_cort_dna_cold
OT_total_cort_dna_HOT <- OT_coeff_cort_dna_hot
OT_total_sex_dna <- OT_coeff_sex_dna
OT_total_ros_dna <- OT_coeff_ros_dna 
OT_total_mitodensity_dna <- OT_undir_mitodensity_dna
OT_total_potential_dna <- OT_undir_potential_dna
#
# C.3) Lipid peroxidation
OT_total_cort_perox_COLD <- OT_coeff_cort_perox_cold
OT_total_cort_perox_HOT <- OT_coeff_cort_perox_hot
OT_total_age_perox <- OT_coeff_age_perox
OT_total_ros_perox <- OT_coeff_ros_perox 
OT_total_mitodensity_perox <- OT_undir_mitodensity_perox
OT_total_potential_perox <- OT_undir_potential_perox
#
# C.4) ROS
OT_total_mitodensity_ros <- OT_coeff_mitodensity_ros
OT_total_potential_ros <- OT_coeff_potential_ros
#
#
# D) Create a df with the values for each variable
#
# D.1) Detection
det_results_semOT <- data.frame(
  reference_variable = rep("t_D", 9),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "age",
                          "sex",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS",
                          "mean_dnadamage",
                          "mean_peroxidation"),
  direct_effects = I(list(NA,
                          NA,
                          NA,
                          NA,
                          OT_coeff_mitodensity_det,
                          OT_coeff_potential_det,
                          NA,
                          OT_coeff_dna_det,
                          OT_coeff_perox_det)),
  indirect_effects = I(list(OT_undir_cort_det_COLD,
                          OT_undir_cort_det_HOT,
                          OT_undir_age_det,
                          OT_undir_sex_det,
                          OT_undir_mitodensity_det,
                          OT_undir_potential_det,
                          OT_undir_ros_det,
                          NA,
                          NA)),
  total_effects = I(list(OT_total_cort_det_COLD,
                        OT_total_cort_det_HOT,
                        OT_total_age_det,
                        OT_total_sex_det,
                        OT_total_mitodensity_det,
                        OT_total_potential_det,
                        OT_total_ros_det,
                        OT_total_dna_det,
                        OT_total_perox_det))
  ) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)
  )  %>%
  mutate(source = "det_results_semOT")
#
# D.2) DNA damage
dna_results_semOT <- data.frame(
  reference_variable = rep("DNA damage", 6),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "sex",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS"),
  direct_effects = I(list(OT_coeff_cort_dna_cold,
                          OT_coeff_cort_dna_hot,
                          OT_coeff_sex_dna,
                          NA,
                          NA,
                          OT_coeff_ros_dna)),
  indirect_effects = I(list(NA,
                          NA,
                          NA,
                          OT_undir_mitodensity_dna,
                          OT_undir_potential_dna,
                          NA)),
  total_effects = I(list(OT_total_cort_dna_COLD,
                        OT_total_cort_dna_HOT,
                        OT_total_sex_dna,
                        OT_total_mitodensity_dna,
                        OT_total_potential_dna,
                        OT_total_ros_dna))
  ) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)
  )  %>%
  mutate(source = "dna_results_semOT")
#
# D.3) Lipid peroxidation
perox_results_semOT <- data.frame(
  reference_variable = rep("Lipid peroxidation", 6),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "age",
                          "mean_mitodensity",
                          "mean_potential",
                          "ROS"),
  direct_effects = I(list(OT_coeff_cort_perox_cold,
                          OT_coeff_cort_perox_hot,
                          OT_coeff_age_perox,
                          NA,
                          NA,
                          OT_coeff_ros_perox)),
  indirect_effects = I(list(NA,
                          NA,
                          NA,
                          OT_undir_mitodensity_perox,
                          OT_undir_potential_perox,
                          NA)),
  total_effects = I(list(OT_total_cort_perox_COLD,
                        OT_total_cort_perox_HOT,
                        OT_total_age_perox,
                        OT_total_mitodensity_perox,
                        OT_total_potential_perox,
                        OT_total_ros_perox))
  ) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)
  ) %>%
  mutate(source = "perox_results_semOT")
#
# D.4) ROS
ros_results_semOT <- data.frame(
  reference_variable = rep("ROS", 2),
  predictor_modulator = c("mean_mitodensity",
                          "mean_potential"),
  direct_effects = I(list(OT_coeff_mitodensity_ros,
                          OT_coeff_potential_ros)),
  indirect_effects = I(list(NA,
                          NA)),
  total_effects = I(list(OT_total_mitodensity_ros,
                        OT_total_potential_ros))
  ) %>% 
  group_by(predictor_modulator) %>%
  summarize(
    # Summarizing each list column (direct_effects, indirect_effects, etc.)
    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)
  )  %>%
  mutate(source = "ros_results_semOT")
#
# E) Merge everything into a single df
#
sem_results_OT <- bind_rows(det_results_semOT, dna_results_semOT, perox_results_semOT, ros_results_semOT)
```

_Detection of visual stimulus and its relation with mitochondrial physiology in Optic tecta_: mitochondrial density had big significant effects on ROS production (β~D~ = `r format_dec(mean(OT_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OT_coeff_mitodensity_det), 3, equal = TRUE)`). However, we did not find any significant direct or indirect effects of ROS on detection latency, DNA damage, or lipid peroxidation (see @fig-sem_results_OT and Table S18 in Supplementary Material). Mitochondrial potential did not affect ROS or any other variable in the model (see @fig-sem_results_OT and Table S18 in Supplementary Material). We also did not see any effects of ROS production on DNA damage or lipid peroxidation, neither of DNA damage or lipid peroxidation on detection latency. In addition, we found weak non significant effects of sex and CORT on DNA damage, neither age and CORT on lipid peroxidation (see @fig-sem_results_OT and Table S18 in Supplementary Material).

```{r, fig-sem_results_OT}
#| label: fig-sem_results_OT
#| fig-cap: "Structural Equation Models for OT/Visual stimulus"
#| fig-name: "fig-sem_results_OT"
#
# A) Getting all the direct coefficients
OT_density_det <- paste0(format_dec(mean(OT_coeff_mitodensity_det), 3),
                  " [", format_dec(quantile(OT_coeff_mitodensity_det, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_mitodensity_det, 0.95), 3), "]")
OT_potential_det <- paste0(format_dec(mean(OT_coeff_potential_det), 3),
                  " [", format_dec(quantile(OT_coeff_potential_det, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_potential_det, 0.95), 3), "]")
OT_dna_det <- paste0(format_dec(mean(OT_coeff_dna_det), 3),
                  " [", format_dec(quantile(OT_coeff_dna_det, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_dna_det, 0.95), 3), "]")
OT_perox_det <- paste0(format_dec(mean(OT_coeff_perox_det), 3),
                  " [", format_dec(quantile(OT_coeff_perox_det, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_perox_det, 0.95), 3), "]")
#
OT_cort_dna_cold <- paste0(format_dec(mean(OT_coeff_cort_dna_cold), 3),
                      " [", format_dec(quantile(OT_coeff_cort_dna_cold, 0.05), 3),
                      ", ", format_dec(quantile(OT_coeff_cort_dna_cold, 0.95), 3), "]")
OT_cort_dna_hot <- paste0(format_dec(mean(OT_coeff_cort_dna_hot), 3),
                      " [", format_dec(quantile(OT_coeff_cort_dna_hot, 0.05), 3),
                     ", ", format_dec(quantile(OT_coeff_cort_dna_hot, 0.95), 3), "]")
OT_sex_dna <- paste0(format_dec(mean(OT_coeff_sex_dna), 3),
                  " [", format_dec(quantile(OT_coeff_sex_dna, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_sex_dna, 0.95), 3), "]")
OT_ros_dna <- paste0(format_dec(mean(OT_coeff_ros_dna), 3),
                  " [", format_dec(quantile(OT_coeff_ros_dna, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_ros_dna, 0.95), 3), "]")
#
OT_cort_perox_cold <- paste0(format_dec(mean(OT_coeff_cort_perox_cold), 3),
                          " [", format_dec(quantile(OT_coeff_cort_perox_cold, 0.05), 3),
                          ", ", format_dec(quantile(OT_coeff_cort_perox_cold, 0.95), 3), "]")
OT_cort_perox_hot <- paste0(format_dec(mean(OT_coeff_cort_perox_hot), 3),
                      " [", format_dec(quantile(OT_coeff_cort_perox_hot, 0.05), 3),
                     ", ", format_dec(quantile(OT_coeff_cort_perox_hot, 0.95), 3), "]")
OT_age_perox <- paste0(format_dec(mean(OT_coeff_age_perox), 3),
                  " [", format_dec(quantile(OT_coeff_age_perox, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_age_perox, 0.95), 3), "]")
OT_ros_perox <- paste0(format_dec(mean(OT_coeff_ros_perox), 3),
                  " [", format_dec(quantile(OT_coeff_ros_perox, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_ros_perox, 0.95), 3), "]")
#
OT_density_ros <- paste0(format_dec(mean(OT_coeff_mitodensity_ros), 3),
                  " [", format_dec(quantile(OT_coeff_mitodensity_ros, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_mitodensity_ros, 0.95), 3), "]")
OT_potential_ros <- paste0(format_dec(mean(OT_coeff_potential_ros), 3),
                  " [", format_dec(quantile(OT_coeff_potential_ros, 0.05), 3),
                  ", ", format_dec(quantile(OT_coeff_potential_ros, 0.95), 3), "]")
#
#
imgOT <- readPNG(here("Others", "SEM_OT.png"))
plot_SEM_OT <- rasterGrob(imgOT, interpolate = TRUE)
#
fig_SEM_OT <- ggdraw(plot_SEM_OT) +
  annotate("text", x = 0.35, y = 0.26, label = OT_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.35, y = 0.918, label = OT_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.745, label = OT_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.455, label = OT_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.58, y = 0.533, label = OT_sex_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.813, label = OT_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.79, y = 0.533, label = OT_cort_dna_cold, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen", fontface = "italic") +
  annotate("text", x = 0.79, y = 0.567, label = OT_cort_dna_hot, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.57, y = 0.115, label = OT_age_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.365, label = OT_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.78, y = 0.27, label = OT_cort_perox_cold, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen", fontface = "italic") +
  annotate("text", x = 0.78, y = 0.305, label = OT_cort_perox_hot, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.285, y = 0.355, label = OT_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.285, y = 0.825, label = OT_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
ggsave(here("./output/figures/text/SEM_OT.png"), plot = fig_SEM_OT, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/SEM_OT.png")
```

## Discussion

- Early environmental conditions did not have significant effects on mitochondrial density, potential, nor ROS production in neither of the brain regions studied.

- Mitocondrial density was a good predictor of ROS production in both regions, but mitochondrial potential was only a good predictor in Olfactory bulbs. We might be missing interactions with other variables that could explain the differences between regions.

- Early conditions did not affect DNA damage on Olfactory bulbs, but had a significant effect on DNA damage in Optic tecta. CORT reduced DNA damage in cold incubated lizards but had no effects at high incubation temperatures. This effect was not mediated by ROS production or any other mitochondrial-related measurement. It could be because of the way CORT can redistribute resources to cope with stressors, and in this case, it could be reducing DNA damage in the brain.

- Early conditions had a weak effect on lipid peroxidation in both regions. The effect of CORT seems to increase lipid peroxidation at low temperatures, and increase it at higher. This patern is similar in both Olfactory bulbs and Optic tecta. It could be related to the way lipids are organised at different temperatures, and how CORT can affect the way they are metabolised.

- Age has significant effects on DNA damage and lipid peroxidation in Olfactory bulbs (as expected), but a weak effect on lipid peroxidation in Optic tecta. Females had lower DNA damage in Olfactory bulbs (marginally significant) and slightly higher in the Optic tecta. Aging increases accumulative damage. Males could be investing more in repairing or avoiding damage in chemosensory areas because they could be more important for sexual-related behaviours.

- Oxidative damage was not influence by ROS production or any other mitochondrial-related measurement in either of the brain regions studied. It could be that measuring only H2O2 is too reductionist, or the action of antioxidants, which, at the same time, could potentially explained the patterns mentioned before.

- Detection latency decreased with CORT exposure when animals were presented chemical stimulus, but not with visual. It could be related to the fact that motivation was not totally controlled. Motivation tests presented both types of stimulus, but it could be that stressed animals were more motivated to search for food and, therefore, interacted faster with the vessel in chemical tests.

- Performance in the behavioral tests was not affected by mitochondrial physiology. It could be that the brain regions studied are not directly related to the behaviors tested, that other parameters are more relevant, or that the oxidatie damage is not high enought for detecting an effect.

# References
<div id="refs"></div>

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

# Supplementary Material

#### Results of preliminary models

```{r, models_preliminary}
#| label: models_preliminary
# Fitting intial models to see if sex and age are relevant for our models
source(here("R", "func.R"))
#
#
# Run models mitochondrial physiology (each region separately)
#
var_m <- c("mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
regions <- c("OB", "OT")
formula_list_ <- list()
for (p in var_m){
  formula_list_[[p]] <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
  for (h in regions){
    if (h == "OB"){
      df <- clean_df %>% filter(region == "OB")
      l <- "OB"
    } else {
      df <- clean_df %>% filter(region == "OT")
      l <- "OT"
    }

  pmodel_name <- paste0("m_prel_", p, "_", h)
  assign(pmodel_name, fit_m(df = df,
                             cat = "prel",
                             var = p,
                             formula = formula_list_[[p]],
                             fam = gaussian(),
                             label = l,
                             refit = FALSE),
          envir = .GlobalEnv)  # Assign to the global environment
  }
}
#
#
# Run model behaviour (stimuli separated)
beh_df <- clean_df
stimulus <- c("Chemical", "Visual")
formula_t_D <- t_D ~ motivation + motivation:cort + cort*temp + age_trial + sex + prey +(1|clutch) + (1|lizard_id) 
for (k in stimulus){
  df <- beh_df %>% filter(stimulus == k)
  pmodel_name <- paste0("m_prel_t_D_", k)
  assign(pmodel_name, fit_m(df = df,
                            cat = "prel",
                            var = "t_D",
                            formula = formula_t_D,
                            fam = gaussian(),
                            label = k,
                            refit = FALSE),
        envir = .GlobalEnv)  # Assign to the global environment
}
```

Table S1. Preliminary results of the models testing for Mitochondrial Density in Olfactory Bulbs.

```{r, results_preliminary_mitdensity_OB}
#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OB"
#| label: results_preliminary_mitdensity
#
sum_m_mitdensity_OB_prel <- m_prel_mean_mitodensity_OB %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws()  %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_mitdensity_OB_prel)
```

Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S2. Preliminary results of the models testing for Mitochondrial Density in Optic Tecta.
```{r, results_preliminary_mitdensity_OT}
#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OT"
#| label: results_preliminary_mitdensity_OT
#
sum_m_mitdensity_OT_prel <- m_prel_mean_mitodensity_OT %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_mitdensity_OT_prel)
```

Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S3. Preliminary results of the models testing for Mitochondrial Potential in Olfactory Bulbs.
```{r, results_preliminary_potential_OB}
#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OB"
#| label: results_preliminary_potential_OB
#
sum_m_potential_OB_prel <- m_prel_mean_potential_OB %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_potential_OB_prel)
```

Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S4. Preliminary results of the models testing for Mitochondrial Potential in Optic Tecta.
```{r, results_preliminary_potential_OT}
#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OT"
#| label: results_preliminary_potential_OT
#
sum_m_potential_OT_prel <- m_prel_mean_potential_OT %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_potential_OT_prel)
```

Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S5. Preliminary results of the models testing for ROS Production in Olfactory Bulbs.
```{r, results_preliminary_ros_OB}
#| tbl-cap: "Preliminary results of the models testing for ROS Production in OB"
#| label: results_preliminary_ros_OB
#
sum_m_ros_OB_prel <- m_prel_mean_ros_OB %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_ros_OB_prel)
```

Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S6. Preliminary results of the models testing for ROS Production in Optic Tecta.
```{r, results_preliminary_ros_OT}
#| tbl-cap: "Preliminary results of the models testing for ROS Production in OT"
#| label: results_preliminary_ros_OT
#
sum_m_ros_OT_prel <- m_prel_mean_ros_OT %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_ros_OT_prel)
```

Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S7. Preliminary results of the models testing for DNA Damage in Olfactory Bulbs.
```{r, results_preliminary_dnadamage_OB}
#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OB"
#| label: results_preliminary_dnadamage_OB
#
sum_m_dnadamage_OB_prel <- m_prel_mean_dnadamage_OB %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_dnadamage_OB_prel)
```

Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates an effect of sex and age, so they were included in the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S8. Preliminary results of the models testing for DNA Damage in Optic Tecta.
```{r, results_preliminary_dnadamage_OT}
#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OT"
#| label: results_preliminary_dnadamage_OT
#
sum_m_dnadamage_OT_prel <- m_prel_mean_dnadamage_OT %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_dnadamage_OT_prel)
```

Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of age, so it was discarded from the final models. However, we saw an effect of sex and we included it in our final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S9. Preliminary results of the models testing for Lipid Peroxidation in Olfactory Bulbs.
```{r, results_preliminary_peroxidation_OB}
#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OB"
#| label: results_preliminary_peroxidation_OB
#
sum_m_peroxidation_OB_prel <- m_prel_mean_peroxidation_OB %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_peroxidation_OB_prel)
```

Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex, so it was discarded from the final models. However, we saw an effect of age and we included it in our final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S10. Preliminary results of the models testing for Lipid Peroxidation in Optic Tecta.
```{r, results_preliminary_peroxidation_OT}
#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OT"
#| label: results_preliminary_peroxidation_OT
#
sum_m_peroxidation_OT_prel <- m_prel_mean_peroxidation_OT %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_peroxidation_OT_prel)
```

Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex, so it was discarded from the final models. However, we saw an effect of age and we included it in our final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S11. Preliminary results of the models testing for Detection Latency (t_D) of Chemical stimuli. 
```{r, results_preliminary_t_D_Chemical}
#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Chemical stimuli"
#| label: results_preliminary_t_D_Chemical
# 
sum_m_t_D_prel_Chem <- m_prel_t_D_Chemical %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_t_D_prel_Chem)
```

Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S12. Preliminary results of the models testing for Detection Latency (t_D) of Visual stimuli.
```{r, results_preliminary_t_D_Visual}
#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Visual stimuli"
#| label: results_preliminary_t_D_Visual
#
sum_m_t_D_prel_Vis <- m_prel_t_D_Visual %>%
  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
  summarise_draws() %>%
  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
#
flextable(sum_m_t_D_prel_Vis)
```

Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Results of the main models

Table S13. R^2 values of the definite models for both regions and stimuli.
```{r, table_bayesR2}
#| label: table_bayesR2
#| tbl-cap: "BayesR2 values of the definite models for both regions and stimuli."
#
data_bayes <- data.frame(
  Model = character(0),
  Mean = numeric(0),
  Error = numeric(0),
  Q2_5 = numeric(0),
  Q97_5 = numeric(0)
)
models <- c("mean_mitodensity_def_OB",
            "mean_potential_def_OB",
            "mean_ros_def_OB",
            "mean_dnadamage_def_OB",
            "mean_peroxidation_def_OB",
            "t_D_def_Chemical",
            "mean_mitodensity_def_OT",
            "mean_potential_def_OT",
            "mean_ros_def_OT",
            "mean_dnadamage_def_OT",
            "mean_peroxidation_def_OT",
            "t_D_def_Visual")
#
for (m in models){
  mod <- readRDS(here("output/models/", paste0(m, ".rds")))
  bayes <- bayes_R2(mod)
  data_bayes <- rbind(data_bayes, data.frame(
    Model = m,
    Mean = format_dec(bayes[1], 3),
    Error = format_dec(bayes[2], 3),
    Q2_5 = format_dec(bayes[3], 3),
    Q97_5 = format_dec(bayes[4], 3)
  ))
}
#
bayes_table_df <- data_bayes %>%
  mutate(Region = gsub(".*_", "", Model)    # Extract everything after the last "_"
  ) %>%
  mutate(Model = factor(Model,
                        levels = c("mean_mitodensity_def_OB",
                                  "mean_potential_def_OB",
                                  "mean_ros_def_OB",
                                  "mean_dnadamage_def_OB",
                                  "mean_peroxidation_def_OB",
                                  "t_D_def_Chemical",
                                  "mean_mitodensity_def_OT",
                                  "mean_potential_def_OT",
                                  "mean_ros_def_OT",
                                  "mean_dnadamage_def_OT",
                                  "mean_peroxidation_def_OT",
                                  "t_D_def_Visual"),
                        labels = c("m_def_mean_mitodensity_OB" = "Mit density",
                                  "m_def_mean_potential_OB" = "Mit potential",
                                  "m_def_mean_ros_OB" = "ROS",
                                  "m_def_mean_dnadamage_OB" = "DNA damage",
                                  "m_def_mean_peroxidation_OB" = "Peroxidation",
                                  "m_def_t_D_Chemical" = "Detection lat",
                                  "m_def_mean_mitodensity_OT" = "Mit density",
                                  "m_def_mean_potential_OT" = "Mit potential",
                                  "m_def_mean_ros_OT" = "ROS",
                                  "m_def_mean_dnadamage_OT" = "DNA damage",
                                  "m_def_mean_peroxidation_OT" = "Peroxidation",
                                  "m_def_t_D_Visual" = "Detection lat")),
      Region = factor(Region, levels = c("OB", "Chemical", "OT", "Visual"),
                      labels = c("OB" = "Olfactory bulbs",
                                "Chemical" = "Chemical",
                                "OT" = "Optic tecta",
                                "Visual" = "Visual"))) %>%
  dplyr::select(Region, Model, Mean, Error, Q2_5, Q97_5) %>%
  arrange(Region, Model)
#
#
# Create the table
#
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
bayes_table <- flextable(bayes_table_df) %>%
  set_header_labels(
    Region = "Region/Stimulus",
    Mean = "Mean",
    Error = "Error",
    Q2_5 = "2.5%",
    Q97_5 = "97.5%") %>%
  align(align = "center", j = c(3:5), part = "body") %>%
  align(align = "center", j = c(1:5), part = "header") %>%
  flextable::compose(i = c(2:5,8:11), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = 6, part = "body") %>% 
  autofit()
#
bayes_table
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S14. Results of the definitive models for Olfactory Bulbs/Chemical stimulus.
```{r, tbl-results_OB_table}
#| label: results_OB_table
#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
#| tbl-name: "results_OB"
#| tbl-label: "results_OB"
source(here("R", "func.R"))
# 
# A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB)
post_OB_refined <- refine_post(post_OB) %>%
  arrange(Variable, Predictors)
#
# B) Create table
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
OB_table <- flextable(post_OB_refined) %>%
  align(align = "center", j = c(3:5), part = "body") %>%
  align(align = "center", j = c(1:5), part = "header") %>%
  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
  flextable::compose(i = c(2:4,6:8,10:12,14:18,20:23,25:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  autofit()
#
OB_table
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S15. Results of the definitive models for Optic Tecta/Visual stimulus.
```{r, tbl-results_OT_table}
#| label: results_OT_table
#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
#| tbl-name: "results_OT"
#| tbl-label: "results_OT"
source(here("R", "func.R"))
# 
# A) Refining the df summarizing the posteriors for OT/Visual stimulus (post_OT)
post_OT_refined <- refine_post(post_OT) %>%
  arrange(Variable, Predictors)
#
# B) Create table
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
OT_table <- flextable(post_OT_refined) %>%
  align(align = "center", j = c(3:5), part = "body") %>%
  align(align = "center", j = c(1:5), part = "header") %>%
  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
  flextable::compose(i = c(2:4,6:8,10:12,14:17,19:22,24:26), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  autofit()
#
OT_table
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Main models diagnostics (plots)

```{r , plotmod_mitdensity_OB, out.width="70%", fig.align="center"}
#| label: plotmod_mitdensity_OB
#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
#
mod <- readRDS(here("output/models/mean_mitodensity_def_OB.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
```

Fig S1. Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs.
Formula: mean_mitodensity ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_mitdensity_OT, out.width="70%", fig.align="center"}
#| label: plotmod_mitdensity_OT
#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta."
#
mod <- readRDS(here("output/models/mean_mitodensity_def_OT.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
```

Fig S2. Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta.
Formula: mean_mitodensity ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_potential_OB, out.width="70%", fig.align="center"}
#| label: plotmod_potential_OB
#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
#
mod <- readRDS(here("output/models/mean_potential_def_OB.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
```

Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
Formula: mean_potential ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_potential_OT, out.width="70%", fig.align="center"}
#| label: plotmod_potential_OT
#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
#
mod <- readRDS(here("output/models/mean_potential_def_OT.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
```

Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
Formula: mean_potential ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_ros_OB, out.width="70%", fig.align="center"}
#| label: plotmod_ros_OB
#| caption: "Posterior predictive checks for the model of ROS Production in Olfactory Bulbs."
#
mod <- readRDS(here("output/models/mean_ros_def_OB.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
```

Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
Formula: mean_ros ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_ros_OT, out.width="70%", fig.align="center"}
#| label: plotmod_ros_OT
#| caption: "Posterior predictive checks for the model of ROS Production in Optic Tecta."
#
mod <- readRDS(here("output/models/mean_ros_def_OT.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
```

Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
Formula: mean_ros ~ cort * temp + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_dnadamage_OB, out.width="70%", fig.align="center"}
#| label: plotmod_dnadamage_OB
#| caption: "Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs."
#
mod <- readRDS(here("output/models/mean_dnadamage_def_OB.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
```

Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_dnadamage_OT, out.width="70%", fig.align="center"}
#| label: plotmod_dnadamage_OT
#| caption: "Posterior predictive checks for the model of DNA Damage in Optic Tecta."
#
mod <- readRDS(here("output/models/mean_dnadamage_def_OT.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
```

Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_peroxidation_OB, out.width="70%", fig.align="center"}
#| label: plotmod_peroxidation_OB
#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs."
#
mod <- readRDS(here("output/models/mean_peroxidation_def_OB.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
```

Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_peroxidation_OT, out.width="70%", fig.align="center"}
#| label: plotmod_peroxidation_OT
#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta."
#
mod <- readRDS(here("output/models/mean_peroxidation_def_OT.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
```

Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_tD_Chemical, out.width="70%", fig.align="center"}
#| label: plotmod_tD_Chemical
#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials."
#
mod <- readRDS(here("output/models/t_D_def_Chemical.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
```

Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r , plotmod_tD_Visual, out.width="70%", fig.align="center"}
#| label: plotmod_tD_Visual
#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials."
#
mod <- readRDS(here("output/models/t_D_def_Visual.rds"))
plot_mod <- plot(mod, plot = FALSE)
#
plot_mod_1 <- plot_mod[[1]]
plot_mod_2 <- plot_mod[[2]]
#
fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
```

Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Results of the multivariate models

Table S16. R^2 values of the multivariate models.
```{r, table_bayesR2_SEM}
#| label: table_bayesR2_SEM
#| tbl-cap: "BayesR2 values of the multivariate models"
#
data_bayes_SEM <- data.frame(
  Model = character(0),
  Response = character(0),
  Mean = numeric(0),
  Error = numeric(0),
  Q2_5 = numeric(0),
  Q97_5 = numeric(0)
)
#
models <- c("m_OB", "m_OT")  
for (m in models) {
  mod <- readRDS(here("output/m_SEM/", paste0(m, ".rds")))

  # Extract bayes_R2() for each response variable in the multivariate model
  bayes_df_SEM <- as.data.frame(bayes_R2(mod))
  bayes_df_SEM$Response <- rownames(bayes_df_SEM)  # Keep track of response variable names

  # Add model identifier and format values
  bayes_df_SEM <- bayes_df_SEM %>%
    mutate(
      Model = m,
      Mean = format_dec(Estimate, 3),
      Error = format_dec(`Est.Error`, 3),
      Q2_5 = format_dec(`Q2.5`, 3),
      Q97_5 = format_dec(`Q97.5`, 3)
    ) %>%
    dplyr::select(Model, Response, Mean, Error, Q2_5, Q97_5)

  # Append to the main data frame
  data_bayes_SEM <- rbind(data_bayes_SEM, bayes_df_SEM)
}
#
bayes_table_SEM_df <- data_bayes_SEM %>%
  mutate(Response = gsub("^R2", "", Response)) %>%
  mutate(Response = factor(Response,
                        levels = c("tD",
                                  "meandnadamage",
                                  "meanperoxidation",
                                  "meanros"),
                        labels = c("tD" = "Detection lat",
                                  "meandnadamage" = "DNA damage",
                                  "meanperoxidation" = "Peroxidation",
                                  "meanros" = "ROS")),
      Model = factor(Model, levels = c("m_OB", "m_OT"),
                      labels = c("OB" = "Olfactory bulbs / Chemical stimulus",
                                "OT" = "Optic tecta / Visual stimulus"))) %>%
  dplyr::select(Model, Response, Mean, Error, Q2_5, Q97_5) %>%
  arrange(Model, Response)
#
#
# 
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
bayes_table_SEM <- flextable(bayes_table_SEM_df) %>%
  set_header_labels(
    Region = "Region/Stimulus",
    Mean = "Mean",
    Error = "Error",
    Q2_5 = "2.5%",
    Q97_5 = "97.5%") %>%
  align(align = "center", j = c(3:6), part = "body") %>%
  align(align = "center", part = "header") %>%
  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = 4, part = "body") %>% 
  autofit()
#
bayes_table_SEM
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S17. Results of multivariate models for OB/Chemical stimulus
```{r, table_sem_results_OB}
#| label: table_sem_results_OB
#| fig-cap: "Structural Equation Models for OB/Chemical stimulus"
#
# Table created from the df sem_results_OB (see above)
source(here("R", "func.R"))
#
# Modify database
table_semOB_df <- sem_results_OB %>%
  mutate(`Direct effects` = paste0(mean_direct_effects, " [", q5_direct, ", ", q95_direct, "]"),
         `Indirect effects` = paste0(mean_indirect_effects, " [", q5_indirect, ", ", q95_indirect, "]"),
         `Total effects` = paste0(mean_total_effects, " [", q5_total, ", ", q95_total, "]")) %>%
  dplyr::select(source, predictor_modulator, `Direct effects`, `Indirect effects`, `Total effects`) %>%
  mutate(
      predictor_modulator = case_when(
        predictor_modulator == "mean_mitodensity" ~ "Mitochondrial density",
        predictor_modulator == "mean_potential" ~ "Mitochondrial potential",
        predictor_modulator == "mean_dnadamage" ~ "DNA damage",
        predictor_modulator == "mean_peroxidation" ~ "Lipid peroxidation",
        predictor_modulator == "ROS" ~ "ROS",
        predictor_modulator == "cort" ~ "CORT",
        predictor_modulator == "age" ~ "Age",
        TRUE ~ predictor_modulator
      ),
      source = case_when(
        source == "det_results_semOB" ~ "Detection latency",
        source == "dna_results_semOB" ~ "DNA damage",
        source == "perox_results_semOB" ~ "Lipid peroxidation",
        source == "ros_results_semOB" ~ "ROS",
        TRUE ~ source
      ),
      `Direct effects` = case_when(`Direct effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Direct effects`),
      `Indirect effects` = case_when(`Indirect effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Indirect effects`),
      `Total effects` = case_when(`Total effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Total effects`)) %>%
  rename(Predictor = predictor_modulator, Response = source) %>%
  mutate(Response = factor(Response, levels = c("Detection latency", "DNA damage", "Lipid peroxidation", "ROS")),
         Predictor = factor(Predictor, levels = c("CORT",
                                                "Age",
                                                "Mitochondrial density",
                                                "Mitochondrial potential",
                                                "ROS",
                                                "DNA damage",
                                                "Lipid peroxidation"))) %>%
  arrange(Response, Predictor)
#
# Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
table_semOB <- flextable(table_semOB_df) %>%
  align(align = "center", part = "header") %>%
  flextable::compose(i = c(2:7, 9:11, 13:15, 17), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = c(7, 11, 15), part = "body") %>% 
  autofit()
#
table_semOB
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table S18. Results of multivariate models for OT/Visual stimulus
```{r, table_sem_results_OT}
#| label: table_sem_results_OT
#| fig-cap: "Structural Equation Models for OT/Visual stimulus"
#
# Table created from the df sem_results_OT (see above)
source(here("R", "func.R"))
#
# Modify database
table_semOT_df <- sem_results_OT %>%
  mutate(`Direct effects` = paste0(mean_direct_effects, " [", q5_direct, ", ", q95_direct, "]"),
         `Indirect effects` = paste0(mean_indirect_effects, " [", q5_indirect, ", ", q95_indirect, "]"),
         `Total effects` = paste0(mean_total_effects, " [", q5_total, ", ", q95_total, "]")) %>%
  dplyr::select(source, predictor_modulator, `Direct effects`, `Indirect effects`, `Total effects`) %>%
  mutate(
      predictor_modulator = case_when(
        predictor_modulator == "mean_mitodensity" ~ "Mitochondrial density",
        predictor_modulator == "mean_potential" ~ "Mitochondrial potential",
        predictor_modulator == "mean_dnadamage" ~ "DNA damage",
        predictor_modulator == "mean_peroxidation" ~ "Lipid peroxidation",
        predictor_modulator == "ROS" ~ "ROS",
        predictor_modulator == "cort" ~ "CORT",
        predictor_modulator == "age" ~ "Age",
        predictor_modulator == "sex" ~ "Sex",
        predictor_modulator == "cort_cold" ~ "CORT when Cold",
        predictor_modulator == "cort_hot" ~ "CORT when Hot",
        TRUE ~ predictor_modulator
      ),
      source = case_when(
        source == "det_results_semOT" ~ "Detection latency",
        source == "dna_results_semOT" ~ "DNA damage",
        source == "perox_results_semOT" ~ "Lipid peroxidation",
        source == "ros_results_semOT" ~ "ROS",
        TRUE ~ source
      ),
      `Direct effects` = case_when(`Direct effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Direct effects`),
      `Indirect effects` = case_when(`Indirect effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Indirect effects`),
      `Total effects` = case_when(`Total effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Total effects`)) %>%
  rename(Predictor = predictor_modulator, Response = source) %>%
  mutate(Response = factor(Response, levels = c("Detection latency", "DNA damage", "Lipid peroxidation", "ROS")),
         Predictor = factor(Predictor, levels = c("CORT when Cold",
                                                "CORT when Hot",
                                                "Age",
                                                "Sex",
                                                "Mitochondrial density",
                                                "Mitochondrial potential",
                                                "ROS",
                                                "DNA damage",
                                                "Lipid peroxidation"))) %>%
  arrange(Response, Predictor)
#
# Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
table_semOT <- flextable(table_semOT_df) %>%
  align(align = "center", part = "header") %>%
  flextable::compose(i = c(2:9, 11:15, 17:21, 23), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = c(9, 15, 21), part = "body") %>% 
  autofit()
#
table_semOT
```

### 2025-02-26 R/data_process.R
updates

@@ -1,0 +1,87 @@
+####################################
+# Data_process
+####################################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2,
+               lme4, zoo, lmerTest, broom, tidybayes, forcats)
+#
+source(here("R", "func.R"))
+# A) Processing behaviour_df
+data_discr <- read.csv(here("./data/data_main.csv")) %>%
+  mutate(temp = gsub("[AB]_", "", trt),
+         cort = gsub("_[2][38]", "", trt))  %>%
+  mutate(temp = factor(temp,
+                       levels = c("23", "28"),
+                       labels = c("23" = "Cold", "28" = "Hot"))) %>%
+  mutate(cort = factor(cort,
+                       levels = c("A", "B"),
+                       labels = c("A" = "Control", "B" = "CORT"))) %>%
+  mutate(test = factor(test,
+                       levels = c("CC", "CV", "TC", "TV"),
+                       labels = c("CC" = "Chemical_Familiar",
+                                  "CV" = "Visual_Familiar",
+                                  "TC" = "Chemical_Unknown",
+                                  "TV" = "Visual_Unknown"))) %>%
+  mutate(stimulus = gsub("_Familiar|_Unknown", "", test),
+         prey = gsub("Chemical_|Visual_", "", test)) %>%
+  mutate(sex = factor(sex,
+                      levels = c("m", "f"),
+                      labels = c("m" = "Male", "f" = "Female"))) %>%
+  mutate(t_D = as.numeric(t_D)) %>%
+  dplyr::select(lizard_id, clutch, sex, temp, cort, age_trial, age_euthanasia,
+                trial, stimulus, prey, t_D, motivation, Id_flow) %>%
+  data.frame()
+#
+#
+# B) Processing and merging 'mitochondrial activity' (data/mit_act) and
+# 'mitochondrial damage' (data/mit_damage) dfs
+#
+mit_act <- read.csv(here("./data/mit_act.csv")) %>%
+  group_by(Id_flow, plate, region) %>%
+  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)),
+            .groups = "drop") %>%
+  ungroup() %>%
+  dplyr::select(region,
+                Id_flow,
+                mean_mitodensity,
+                mean_potential,
+                mean_ros,
+                mean_size) %>%
[…]

### 2025-02-26 R/exploratory.R

@@ -1,0 +1,149 @@
+####################################
+# Exploratory analyses
+####################################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes,
+ggh4x, cowplot, fitdistrplus, MASS, goftest)
+#
+source(here("R", "func.R"))
+#
+# 
+data_expl <- df_merged %>%
+       mutate(treatment = paste(cort, temp, sep = "-"))
+#
+# A) Make histograms for the main variables and figures factorizing main predictors
+var <- c("t_D", "mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
+
+hist_list <- list()
+for (i in var) {
+  bin_width <- (max(data_expl[[i]], na.rm = TRUE) - min(data_expl[[i]], na.rm = TRUE)) / sqrt(length(data_expl[[i]]))
+  # Labels
+  if (i == "t_D"){
+     lab <- "Detection latency (s)"
+  } else if (i == "mean_mitodensity") {
+     lab <- "Mitochondrial density"
+  } else if (i == "mean_potential") {
+     lab <- "Mitochondrial potential"
+  } else if (i == "mean_ros") {
+     lab <- "ROS production"
+  } else if (i == "mean_dnadamage") {
+     lab <- "DNA damage"
+  } else if (i == "mean_peroxidation") {
+     lab <- "Lipid peroxidation"
+  }
+  # Histograms
+  hist_plot <- ggplot(data_expl, aes(x = .data[[i]])) +
+    geom_histogram(binwidth = bin_width, fill = "#062d00", color = "black", alpha = 0.5) +
+    theme_classic() +
+    labs(x = lab, y = "Counts") +
+    theme(axis.title = element_text(size = 12, family = "Times"),
+          axis.text = element_text(size = 10, family = "Times"))
+  hist_list[[i]] <- hist_plot
+}
+fig_hist <- plot_grid(plotlist = hist_list, ncol = 3)
+#
+for (i in var) {
+  if(i == "t_D"){
+    plot_age <- ggplot(data_expl, aes(x = age_trial, y = t_D, color = treatment)) +
+      geom_point(alpha = 0.8) +
+      scale_color_manual(values = c("CORT-Cold"="#00008B", "Control-Cold"="#68bde1", 
+                                   "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +
[…]

### 2025-02-26 R/func.r

@@ -1,0 +1,317 @@
+#################### 
+####################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom,
+tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest)
+#
+####################
+####################
+# Extract sample size per group
+#' @title sample
+#' @description Estract sample size per group
+#' @param df Database
+#' @param corti To select cort treatment ("CORT"/"Control")
+#' @param therm To select temp treatment ("Cold"/"Hot")
+sample <- function(df, corti, therm){
+  sample_size <- df %>%
+    filter(cort == corti, temp == therm) %>%
+    distinct(lizard_id) %>%
+    summarise(total_count = n()) %>%
+    pull(total_count)
+  return(sample_size)
+}
+####################
+####################
+# Function to generate four Q-Q plots for one variable
+#' @title qq_plots_single 
+#' @description Generate four Q-Q plots for one variable
+#' @param df Database
+#' @param vle Variable chosen
+#' @param label Label assigned to each variable
+#' @return fig_qq
+qq_plots_single <- function(df, vle, label) {
+  # Remove NAs
+  plot_df <- data.frame(value = na.omit(df[[vle]])) 
+
+    # Normal Q-Q Plot
+    p1 <- ggplot(plot_df, aes(sample = value)) +
+      stat_qq() + 
+      stat_qq_line(color = "red") +
+      ggtitle(paste(label, "- Normal Q-Q Plot")) +
+      theme_classic()
+
+    # Log-Normal Q-Q Plot
+    p2 <- ggplot(plot_df, aes(sample = log(value +1))) +
+      stat_qq() + 
+      stat_qq_line(color = "red") +
+      ggtitle(paste(label, "- Log-Normal Q-Q Plot")) +
+      theme_classic()
+
+    # Exponential Q-Q Plot
+    p3 <- ggplot(plot_df, aes(sample = value)) +
+      stat_qq(distribution = qexp, dparams = list(rate = 1/mean(plot_df$value))) +
+      stat_qq_line(distribution = qexp, dparams = list(rate = 1/mean(plot_df$value)), color = "red") +
+      ggtitle(paste(label, "- Exponential Q-Q Plot")) +
+      theme_classic()
+
+    # Arrange plots in a grid
[…]

### 2025-03-09 ms.qmd
updates

@@
 data_sem_OB <- SEM_df %>%
-  filter(region == "OB")
+  filter(region == "OB") %>%
+  mutate(obs = as.integer(c(1:80))) %>%
+  mutate(vec = rep(0, length(obs)))
 # Create the models
@@
   m_OB <- brm(
-    bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(t_D | se(vec, sigma = TRUE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+    bf(mean_ros | se(vec, sigma = TRUE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
@@
 }
+#
+#
+#
+#
+#Chcking residual correlations
+m_OB <- brm(
+  bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+  bf(mean_dnadamage | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
+  bf(mean_peroxidation | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+  bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+set_rescor(FALSE),
+family = gaussian(),
+data = data_sem_OB,
+chains = 2, cores = 4, iter = 2000, warmup = 1000,
+control = list(adapt_delta = 0.99, max_treedepth = 11))
+#
+m_OB2 <- brm(
+  bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+  bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
+  bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+  bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+set_rescor(FALSE),
+family = gaussian(),
+data = data_sem_OB,
+chains = 2, cores = 4, iter = 2000, warmup = 1000,
+prior = c(
+    prior(normal(0, 1e-10), class = "sigma")),
+control = list(adapt_delta = 0.99, max_treedepth = 11))
+#
+#
+m_OB_rescorT <- readRDS(here("output/m_SEM/m_OB_rescorT.rds"))
 ```

### 2025-03-09 .vdoc.r

@@ -466,7 +466,9 @@ SEM_df <- clean_df %>%
 #
 #
 data_sem_OB <- SEM_df %>%
-  filter(region == "OB")
+  filter(region == "OB") %>%
+  mutate(obs = as.integer(c(1:80))) %>%
+  mutate(vec = rep(0, length(obs)))
 # Create the models
 #
 refit <- FALSE
@@ -474,13 +476,13 @@ refit <- FALSE
 #
 if(refit){
   m_OB <- brm(
-    bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(t_D | se(vec, sigma = TRUE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+    bf(mean_ros | se(vec, sigma = TRUE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
-  data = data_sem_OB, 
+  data = data_sem_OB,
   chains = 4, cores = 4, iter = 8000, warmup = 2000,
   control = list(adapt_delta = 0.99, max_treedepth = 11))
   # Save the model
@@ -488,6 +490,37 @@ if(refit){
 } else {
   m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
 }
+#
+#
+#
+#
+#Chcking residual correlations
+m_OB <- brm(
+  bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
[…]

### 2025-03-16 ms.qmd
updates

--- 2025-02-26 ms.qmd
@@
 ---
+title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
 docx: 
@@
 
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Daniel W.A. Noble^1^
+
+^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
+^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
+$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
+
+ORCID:
+
+Pablo Recio ORCID: 0000-0002-5890-0218  
+Dalton C. Leibold ORCID: 0000-0001-9645-2033  
+Ondi L. Crino ORCID: 0000-0001-5700-1387  
+Christopher R. Friesen ORCID: 0000-0001-5338-7454  
+Daniel W.A. Noble ORCID: 0000-0001-9460-8743
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+## Abstract
+Cognitive processes such as the ability to perceive prey, are crucial for survival and reproduction.
+However, early-life conditions like stress-related hormones or thermal environments, can shape cognitive abilities by influencing brain structure and function.
+Mitochondrial physiology is a central mechanism underlying these effects.
+Here, we investigated the combined influence of prenatal corticosterone (CORT) — a key stress hormone in reptiles — and incubation temperature on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
+We manipulated egg CORT levels and incubation temperature, then assessed hatchlings’ ability to detect chemical and visual prey stimuli.
+Using flow cytometry, we measured mitochondrial function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta — two brain regions involved in sensory processing.
+We predicted that CORT exposure and high temperatures would interact to shape perception of prey via changes in mitochondrial respiration and oxidative damage.
+While mitochondrial function remained robust to early conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta but not the olfactory bulbs.
+Additionally, CORT-treated lizards responded faster when presented with chemical cues but not visual stimuli.
+However, mitochondrial function and oxidative stress were not directly related to prey detection performance.
+These findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
+Future research should explore alternative mechanisms linking early-life stress, metabolic function, and cognitive performance.
 
@@
 The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017].
-For instance, plenty of studies have investigated how prenatal conditions can influence learning abilities and memory across different taxa [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020].
-However, the impact of early-life environments on other cognitive domains has been less studied [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
-A fundamental cognitive skill is acquiring information from the environment [@shettleworth].
-Perception of different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
-Variations in the ability to detect or recognize some relevant stimuli can have significant consequences for an individual’s fitness.
-For example, individuals better at detecting prey might be more competitive in foraging contexts, which will affect the energy they can obtain and allocate for other processes like growth or reproduction.
+For instance,  studies have identified how prenatal conditions can influence learning abilities and memory across different taxa [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020].
+However, the impact of early-life environments on other cognitive domains has been poorly studied [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
+Perception is an often overlooked, but fundamental skill [@shettleworth].
+Perceiving different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
+Variation in the ability to detect or recognize relevant stimuli can have significant consequences for an individual’s fitness [@pyke1984optimal].
+For example, visual acuity in the bluegill sunfish (_Lepomis macrochirus_) allow animals to detect prey at larger distances, which can increase their foraging efficiency and energy intake [@hairston1982fish].
 Therefore, understanding how prenatal conditions affect the ability to detect prey can be essential to understanding the broader consequences of early-life environments on fitness.
@@
 
-The prenatal environment influences cognitive abilities because the brain is particularly sensitive to environmental inputs during the early stages of development [@zhu_prenatal_2004].
-Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production [@amani2021perinatal], or brain structure [@amiel_effects_2017].
-One key mechanism linking cognition and brain function is mitochondrial physiology and their effects on energy production and oxidative stress [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental inputs during early stages of development [@zhu_prenatal_2004].
+Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production [@amani2021perinatal], or brain structure and function [@amiel_effects_2017].
+Long-lasting effects on energy production and oxidative stress in the brain may be a key mechanism by which cognitive function is impacted by developmental environments  [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
 Given the high energetic demands of cognitive processes [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], cognitive performance is expected to depend on efficient mitochondrial respiration.
-Additionally, high cognitive abilities are typically associated with increased neuron density and functionality [@lefebvre_taxonomic_2011; @amiel_effects_2017], both of which can be impaired by excessive ROS production and oxidative stress byproduct of mitochondrial function [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
-For instance, @hara_presynaptic_2014 found that performance in visuospatial working-memory tasks correlated negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic bouton.
+Additionally, high cognitive abilities are typically associated with increased neuron density and functionality [@lefebvre_taxonomic_2011; @amiel_effects_2017], both of which can be impaired by excessive reactive oxygen species (ROS) production and oxidative stress [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+For instance, Hara et al.
+(2014) found that performance in visuospatial working-memory tasks correlated negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic terminal.
 These findings emphasize the role of mitochondrial action in shaping cognitive abilities.
-However, how the early environment influences mitochondrial physiology and how that translates into cognitive consequences remains largely unknown in non-mammalian animals [see @chaudhari2022early for a review].
-In particular, whether and how these interactions affect perception abilities, such as prey detection, is still largely unexplored.
+However, how the early environment influences mitochondrial physiology and how it may translate into cognitive consequences remains largely unknown in non-mammalian animals [see @chaudhari2022early for a review].
+In particular, whether and how these interactions affect perceptual abilities, such as prey detection, is still largely unexplored.
 
@@
 
-In ectotherms, thermal conditions can be a significant source of maternal stress.
-When animals experience temperatures outside their optimal thermal range, they activate the physiological stress response, leading to increased GC production that can be passed to the offspring [see @Crino_2023].
-However, the temperature can also directly affect offspring development beyond maternal effects, particularly during early life [@noble_developmental_2018; @crino2024eggs].
+In ectotherms, thermal stress experienced by mothers can increase maternal circulating glucocorticoids.
+Activation of the physiological stress response and higher GC concentrations can subsequently impact offspring [see @Crino_2023].
+However, temperature is also known to directly affect offspring development beyond maternal effects, particularly during early life [@noble_developmental_2018; @crino2024eggs].
 The prenatal thermal environment can play a critical role in shaping mitochondrial function, altering energy metabolism or oxidative stress [@stier2022experimental; @crino2024eggs].
 For example, Red-eared slider turtle (_Trachemys scripta elegans_) hatchlings incubated at 26.5 ºC exhibited greater oxidative damage compared to those incubated at temperatures 0.7 to 1.2 ºC warmer [@treidel2016temperature].
-Therefore, the combined effects of prenatal GCs and temperature can have profound consequences for mitochondrial function, with significant repercussions for brain development and cognition.
-Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities through mitochondrial physiology.
-
-Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial functioning and prey detection in the delicate skink (_Lampropholis delicata_).
-We manipulated prenatal CORT levels and incubation temperature, then assessed hatchlings’ ability to detect chemical and visual prey stimuli through behavioral tests.
-After these tests, we examined mitochondrial function in different brain regions associated with processing chemical and visual cues.
-We hypothesized that prenatal CORT and temperature would interact to influence mitochondrial physiology and prey detection.
-GCs can alter mitochondrial function by increasing oxidative stress [reviewed in @costantini2011meta] and reducing mitochondrial efficiency [@gong_chronic_2011;  but see @crino2024eggs].
-Similarly, high temperatures can decrease mitochondrial efficiency but also oxidative damage [@treidel2016temperature; @crino2024eggs].
-Therefore, we predicted that prenatal CORT exposure would impair mitochondrial function by increasing ROS levels and reducing energy availability, disrupting cognitive performance.
-However, higher incubation temperatures would increase performance only when the cost of higher oxidative stress overcome the benefits of increased mitochondrial efficiency.
-In addition, we predicted complex interactions between CORT and incubation temperature, as both factors negatively impact mitochondrial efficiency but have opposite effects on oxidative stress.
-Finally, since early environmental effects on mitochondrial function can vary across brain regions [@coomber_independent_1997], we expected regional differences in mitochondrial physiology accross treatments, which in turn would lead to stimulus-dependent variations in detection abilities.
-Our study provides insights into how prenatal conditions shape mitochondrial physiology and cognitive abilities, emphasizing the importance of early-life environments in cognitive development.
+Therefore, the combined effects of prenatal GCs and temperature can have profound consequences for metabolic function, with significant repercussions for brain development and cognition.
+Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities by developmental programming of metabolic pathways.
+
+Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
+We manipulated eggs' CORT levels and incubation temperature, then assessed hatchlings’ prey detection ability and mitochondrial function in brain regions related to processing chemical and visual cues.
+We hypothesized that CORT would impair prey detection through mitochondrial pysiology by decreasing respiration efficiency [@gong_chronic_2011;  but see @crino2024eggs] and increasing oxidative stress [reviewed in @costantini2011meta].
+In contrast, we predicted that low incubation temperatures would only enhance beahvioural performance if the benefits of mitochondrial efficiency outweigh the costs of higher oxidative stress [@treidel2016temperature; @crino2024eggs].
+We also expected complex interactions between CORT and temperature, as they have opposing effects on mitochondrial efficiency.
+Finally, we predicted region-dependet effects of early-life conditions on mitochondrial function [@coomber_independent_1997] that would lead to stimulus-specific differences in prey detection.
 
 ## Methods
-
-#### Husbandry
 
@@
 We measured and weighted lizards immediately after hatching.
-We employed a rule to obtain Snout-vent length (SVL) and Tail length (TL) to the nearest mm, and a digital scale (± 0.001g error) to measure mass.
+We used a ruler to measure snout-vent length (SVL) and tail length (TL) to the nearest mm, and a digital scale (± 0.001g error) to measure mass.
 We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with nonstick matting and a small water dish until the beginning of the experiment.
-During this period, lizards were given water dayly and received 3-6 small *A.
+During this period, lizards were given water daily and received 3-6 small *A.
 domestica* crickets three times a week.
@@
 
-Two weeks before we started the tests (see below), lizards were moved to the experimental arena (see @fig-Methods) for acclimatization.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the middle of the arena.
-These arenas were placed in two rooms in six racks, each with its own CCTV system (device model DVR-HP210475) that allowed us to record lizards' behaviour during the experiment (see details below).
-The number of lizards per treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatization, lizards were fed with only one cricket per day dusted with calcium and multivitamin, and water was supplied *ad libitum*.
-We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
-The rooms temperature was set to between 22-24 Celsius.
-After the tests, animals were euthanized and mitochondrial physiology was analyzed in various brain regions (see Brain mitochondrial activity protocol below).
-
 #### CORT and Temperature manipulation
-To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods A).
-Eggs were topically supplied with: a) 5 µL of crystalline corticosterone (Sigma, Cat.
+
+To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them under one of two temperature regimes (see @fig-Methods A).
+We used a partial split clutch design where eggs from a given clutch were distributed equally across the four treatments when clutch sizes were larger than four and randomly across treatments when less than four.
+Eggs were topically supplied with either: a) 5 µL of crystalline corticosterone (Sigma, Cat.
 No.
 C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or b) an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [@crino2024eggs].
-Then, eggs were incubated in one of the two previously mentioned temperature regimes.
+We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [@crino2024eggs].
+Eggs were then incubated at either cold  (23 ± 3 ºC) or hot (28 ± 3 ºC) incubation temperatures in a 2x2 factorial design (Fig.
+1 A).
 These temperatures are within the natural limits in _L.
@@
 
-The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design.
-At least one egg per clutch was assigned to each treatment, while in cluthces bigger than four the remaining eggs were randomly assigned to one of the treatments.
-When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
-
 #### Prey discrimination tests
+Two weeks before we started the tests (see below), lizards were moved to the experimental arena (see Fig.
+1) for acclimatization.
+The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the middle of the arena.
+These arenas were placed in two rooms in six racks, each with its own CCTV system (device model DVR-HP210475) that allowed us to record lizards’ behaviour during the experiment (see details below).
+The number of lizards per treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
+During acclimatization, lizards were fed with only one cricket per day dusted with calcium and multivitamin, and water was supplied ad libitum.
+We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
+The room temperature was set to between 22-24 Celsius.
+After the tests, animals were euthanized and metabolic function was analyzed in various brain regions (see Brain mitochondrial activity protocol below).
+
 The experiment involved presenting lizards with various stimuli from different prey types, then recording and analyzing their behaviour towards each stimulus.
@@
 domestica_) as the familiar prey and mealworm larvae (_Tenebrio molitor_) as the unfamiliar prey.
-None of the species are included in the natural diet of the lizards [@lunney1989diets], but both are reported to be eaten in lab conditions [@michelangeli2016sa].
 We expected to see differences between known and unknown prey because previous experience may influence stimuli perception through habituation or sensitisation [@burger_effects_1990; @burger_effects_1991; @desfilis2003stimulus], and we included familiarity with the prey as a factor in our analyses (see below).
 
-Each stimulus was presented inside a transparent plastic vessel containing a white, two-chambered device (see @fig-Methods B) made of polylactic acid.
+Each stimulus was presented inside a transparent plastic vessel containing a white, two-chambered device (see @fig-Methods B) made of polylactic acid (PLA).
 In chemical trials, the prey was placed in the closed chamber at the back of the device, making it invisible to the lizard, while in visual trials, the prey was placed in the front chamber.
@@
 
-Each trial began by placing the experimental device in the side of the arena opposite to the shelter (see arena in @fig-Methods A), and then removing the shelter.
+Each trial began by placing the experimental device at the side of the arena opposite to the shelter (see arena in @fig-Methods A), and then removing the shelter.
 The water cup had already been removed.
@@
 We assessed lizards' ability to detect each stimulus ('Detection latency') by recording the time from when the lizard resumed normal activity (i.e., walking for at least 5 consecutive seconds; T~0~ in @fig-Methods D) until the first interaction with the stimulus (T~D~ in @fig-Methods D).
-We consider an interatcion when the lizard touched the front of the vessel or the filter paper with its snout for more than five consecutive seconds.
-
-To control for potential differences in hunger levels, all lizards fasted for two days before the experiment, a period considered harmless for this species [@young2022physiological].
-Besides, after each trial, the lizards were given a cricket to assess their motivation to forage.
+Lizards were considered to have interacted with the object when the lizard touched the front of the vessel or the filter paper with its snout for more than five consecutive seconds.
+
+To control for potential differences in hunger levels, all lizards were fasted for two days before the experiment, a period considered harmless for this species [@young2022physiological].
+After each trial, the lizards were also given a cricket to assess their motivation to forage.
 The cricket was left in the enclosure for one hour, and we recorded whether the lizard ate it (recorded as 1) or not (recorded as 0).
@@
 
-#### Brain mitochondrial activity
+#### Brain metabolic function
+
+_Brain dissection and homogenization_
 Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
-We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard's response by turning it upside down, and by testing the pinching relfex in one of the front limbs.
-Lizards were decapitated with surgical scissors once they were not responsive.
+We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
+Lizards without responses were decapitated with surgical scissors.
 This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
 After decapitation, the head was opened and the brain was dissected.
-We extracted manually two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with lizards' chemical and visual perception, respectively [@wyneken2007reptilian].
-Both regions were transferred immediately to 1.5mL centrifuge tubes containing 100µL of 1X PBS.
-
-Both regions were transferred immediately to 1.5 mL centrifuge tubes containing 100 µL of cold 1X PBS and kept on ice until further processing.
+We extracted manually two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with chemical and visual perception in lizards [@wyneken2007reptilian].
+Both regions were transferred immediately to 1.5mL centrifuge tubes containing 100µL of 1X phosphate buffered saline (PBS).
+
+Tissue suspected in PBS was then homogenized.
 For the olfactory bulbs, the tissue was mechanically homogenized by placing the tissue in the well of a 100 µm mesh filter (pluriStrainer) affixed atop a 1.5 mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
-The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1X PBS to prepare a homogenate suspension.
-Optic tecta were mechanically homogenized the same way as olfactory bulbs, but were first enzymatically digested by incubating the tissue in 100µL of 125 U/mL collagenase (type II) for 30-min at 32°C.
-We enzymatically digested optic tecta but not olfactory bulbs because in a pilot study, enzymatic digestion prior to mechanical digestion improved the yield of viable cells in optic tecta homogenate, but did not noticeably affect the yield of viable cells in olfactory bulbs homogenate (unpublished data).
-Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500 µL 1x PBS.
-This step was to performed to remove cellular debris from homogenates.
-From each 500 µL suspension of homogenate collected on a given trial day, we first added 100 µL of homogenate to a pooled sample of each tissue type to use for single-color controls, and the remaining 40 0µL of homogenate was split among two 200 µL aliquots.
-One aliquot was used fresh to measure mitochondrial function (mitochondrial density, membrane potential, ROS), one aliquot was cryopreserved for later measurements of oxidative damage (8-OHdG, lipid peroxidation), and the third aliquot was cryopreserved for a different experiment.
-Aliquots were cryopreserved in 1 mL solution of 1% Neutral-Buffered Formalin (as a fixative agent), 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant), then stored at -20 °C until oxidative damage assays.
-
-From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50 µL of homogenate in duplicates (2 wells per homogenate).
-To each replicate well, we added 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
+The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1X.
+Optic tecta were mechanically homogenized the same way as olfactory bulbs, but were first enzymatically digested by incubating the tissue in 100µL of 125 U/mL collagenase (type II).
+
+Homogenates were split among two aliquots: one was used fresh to measure mitochondrial density, membrane potential and ROS, while the other one was cryopreserved for later measurements of DNA damage and lipid peroxidation.
+Cryopreservation was made by suspending the homogenates in 1 mL solution of 1% Neutral-Buffered Formalin, 1X Tris-EDTA, and 10% DMSO, then stored at -20 °C until oxidative damage assays.
+
+_Metabolic function_
+Fresh homogenate suspensions were stained with 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
 We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
-We then added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
-We then loaded 6 wells with 50 µL of homogenate taken from each pooled homogenate suspension (12 wells total), which were to be negative and single-color controls.
-One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5 µL of one of 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, 50 µM MitoSOX Red, or #µM Hoechst 33342 Nuclear Viability Dye.
-Any remaining pooled homogenate was fixed and frozen as previously described.
-We incubated the loaded plate at 32 °C for 30 minutes to stain and then diluted the samples with 50 µL cold 1x PBS to halt the staining process.
-Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours.
-Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).
-
+We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
+Within two hours after dissection, these samples were analyzed by flow cytometry.
+
+_Oxidative Damage_
 Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
-On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in near-boiling water.
-We washed each thawed sample twice, the first time resuspending the pelleted cells in 1 mL warm 1X Tris-EDTA, and the second time in 200 µL warm 1X PBS.
-Pooled homogenate was divided into 100 µL aliquots to serve as single-color controls.
-Thawed, resuspended samples were first stained with 20 µL (10 µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32 °C for 20 minutes.
-Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin.
-We incubated the samples at 32 °C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200 µL 1X PBS.
-We added 20 µL of 70 µM 8-OHdG Polyclonal Antibody to each sample, after which we left the homogenate overnight (~12 hours) for the antibody to bind to 8-OHdG, a marker of oxidative damage on DNA.
-The following day we counterstained the cells with 20 µL of #µM (100 µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32 °C for 20 minutes.
-After the cells had been tagged with 8-OHdG antibodies and counterstained, we washed the cells once more and resuspended the pellet in 400 µL of 1X PBS.
-Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate.
-Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated.
-We then loaded a 96-well plate with 100 µL of each single-color control and 100 µL in duplicate of each sample.
-We performed all flow cytometry assays on samples within 48-hours of thawing the samples.
-
-All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
-Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v.
-8.0.1) and BD CS&T fluorescent beads (Lot No.
-30664) diluted at 2 drops to 150 µL 1X PBS.
-During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector.
-These thresholds were chosen to filter small debris or inviable or non-intact cells from our observations.
-The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment.
-Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events.
-For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels.
-For the oxidative damage assay, we recorded data from the following channels: FSC (425; forward scatter), SSC (300; side scatter), Alexa-Fluor 488 (275; 8-OHdG Antibody + Alexa-Fluor 488 conjugate), BUV-496 (525; Hoechst 33342), and PE-Cy5 (850; BODIPY 665/676).
-Fluorescent intensity data was collected via the BD FACSDiva (v.
-8.0.1) software, with no compensation applied during data collection, and all on a linear scale (detectable range of 0-252166).
-We recorded data for both the area and height of the fluorescent signal, but only used the area in downstream analyses, with height being recorded for the sake of quality control.
-Data was exported from BD FACSDiva as individual *.fcs (“flow cytometry standard”) files for each sample, then imported into FlowJo (v.
+On the day of oxidative damage assays, we rapidly thawed frozen samples, removed the cryopreservation solution (see above), and resuspended samples in 1 mL warm 1X Tris-EDTA.
+Afterward, the Tris-EDTA was removed, and samples were resuspended in 200 µL of warm 1X PBS containing 20 µL of #µM Hoechst 33342 Nuclear Viability Dye (for cell viability) and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor (to measure lipid peroxidation).
+Following staining, we permeabilized the cell membranes incubating the samples in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin.
+After permeabilization, we stained the samples with 20 µL of 70 µM 8-OHdG Polyclonal Antibody to measure oxidative damage on DNA.
+We left the homogenate overnight (~12 hours).
+The following day we counterstained the cells with 20 µL of #µM (100 µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyze the samples in the flow cytometer.
+
+_Flow Cytometry_
+Flow cytometry assays were performed using a Becton Dickson LSRFortessa X-20 flow cytometer with the default wavelength filters on detectors.
+The detectors and voltage settings used in data acquisition for each were determined during pilot trials and kept consistent throughout the experiment.
+Data was imported into FlowJo (v.
 10.1) for processing.
-
-In FlowJo v.
-10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events.
-We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains.
-We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103).
-For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103).
-For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103).
-When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population.
-We used the backgated population for compensation of fluorescent spillover between different fluorescent probes.
-To account for fluorescent spillover, we used a traditional compensation matrix using the compensation function of FlowJo v.10.1.
-We identified the “positive” population for each channel as the brightest ~2.5% of the distribution of observations in the respective single-color control for that channel and used unstained controls as a universal negative.
-We visually inspected the compensation matrix and its effects on population distributions for under- and over-compensation, whereupon we changed the compensation matrix manually until data was properly compensated.
-We applied the compensation matrix to all samples for downstream processing.
-Following compensation, we again gated the data following the same process as for backgating but using the compensated parameters for each channel.
-Following gating, we exported the geometric mean (mean fluorescent intensity; MFI) and robust confidence-values for each channel for each sample.
-For analysis, we exported summary statistics of only the area of the fluorescent signal.
-Although we exported robust confidence values for checking repeatability between replicate samples, we used the mean and standard deviation of geometric means among replicates for each individual as our main response variables in analyses.
+We obtained the mean fluorescent intensity for mitochondrial density, mitochondrial potential, ROS production, DNA damage, and lipid peroxidation.
+For further details on the homogenization, staining, or flow cytometry assays, see Methods: flow cytometry in Supplementary Material
 
 #### Statistical analyses
-
 We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, mitochondrial potential, ROS, DNA damage, lipid peroxidation, and detection latency) separately.
-We fitted first a set of preliminary models where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction.
-We also added to these models the sex and age of the lizards at the time where the trials started (for detection latency), or when the lizards were euthanized (for all mitochondrial-related variables).
-For detection latency we also included as fixed factors lizards' previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation, as CORT can impact apetite [@conde2018stress].
-We also included for all the models the clutch identity as a random factor.
-Since detection latency was fitted for both types of prey in a repeated measures design, these models also included individual identity as a random factor.
-Structure and results of these models are provided on Tables S1-S12 in Supplementary Material.
-After the preliminary models, we fitted a set of definitive models where those factors that were not significant in the preliminary models were excluded except the main interest factors: CORT, temperature and their interaction.
-Random factors remain the same in the definitive models.
-All the response variables were mean centered and standardized by dividing by two times the standard deviation.
+We first fitted a set of preliminary models for all the interest varaibles (detection latency, mitochondrial density, mitochondrial potential, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction .
+Models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables).
+For detection latency, we also included previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation as fixed effects.
+We included the interaction between CORT and motivation because CORT can impact appetite [@conde2018stress].
+All models included clutch identity as a random factor to account for clutch effects.
+For models of detection latency we also included a random effect of lizard identity because of our repeated measures design.
+The structure and results of these models are provided on Tables S1-S12 of the Supplementary Material.
+After the preliminary models, we simplified models by re-fitting to exclude factors that were not significant.
+However, our models always included the main interest factors: CORT, temperature and their interaction.
+Random factors remain the same for the final models.
+All the response variables were mean centered and standardized by dividing by two times the standard deviation [@gelman2008scaling].
 Before standardization, mitochondrial density, DNA damage, lipid peroxidation, and detection latency were log-transformed.
 
-All the models were run using the brm function from the _brms_ package [@burkner2017brms] in R (version 4.4.0) [@R].
+All models were fit using using brm models from stan [@stan] in R (version 4.4.
+0) [@R].
+The error structure was modelled assuming a gaussian distribution for all variables.
 We ran four parallel MCMC chains of 8000 iterations for each model, with a warmup period of 2000 iterations.
-The error structure was modelled assuming a gaussian distribution for all variables.
-To test for differences between treatments, we made contrasts between treatments using the posterior distributions of the parameters.
-We employed pMCMC values to tests if the contrasts were different from zero [@endo2019introduction].
+To test for differences between treatments, we made contrasts between treatments using the posterior distribution of relevant parameters.
+Significance was assessed by calculating the probability distribution parameter estimates and contrasts to test whether were they different from zero [@endo2019introduction].
 We considered an effect statistically significant if pMCMC < 0.05.
 
-To further explore the relationships between mitochondrial physiology and detection latency, we followed Structural Equations Modelling approach.
-We fitted a multivariate model using the _brms_ package [@burkner2017brms] in R (version 4.4.0) [@R] for each brain region/stimulus separately.
-We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem_results_OB or @fig-sem_results_OT).
-We also included for each variable those factors that were significant in our definitive models but we excluded the non-significant ones.
-Because experience with prey did not affect lizards' behaviour (see Tables S11, S12), and to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
+To further explore the relationships between mitochondrial physiology and detection latency, we used a Structural Equations Modelling (SEM) approach.
+We fitted a multivariate brm from stan [@stan] in R (version 4.4.0) [@R] for each brain region/stimulus separately.
+We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem_results_OB and @fig-sem_results_OT).
+Because experience with prey did not affect lizards’ behaviour (see Tables S11, S12), and to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
 Clutch identity was included as a random factor for all variables.
-We ran four parallel MCMC chains of 8000 iterations for each model, with a warmup period of 2000 iterations.
-The error structure was modelled assuming a gaussian distribution.
-Rescor was set as true to control for potential correlations between the residuals of the variables.
-
-We obtained direct, indirect, and total effects of one variable on another from the posterior parameters.
-We estimated direct effects using posterior estimates from the SEM models.
-Specifically, direct effects correspond to the posterior estimates of a predictor's effect on the response variable.
-For example, the direct effect of mitochondrial density on detection latency is given by the posterior estimate of detection latency regressed on mitochondrial density.
-Indirect effects were computed by multiplying the direct effect of the predictor on the mediator by the direct effect of the mediator on the response variable.
-For instance, the indirect effect of ROS production on detection latency was obtained by summing the products of (i) the direct effect of ROS on DNA damage multiplied by the direct effect of DNA damage on detection latency, and (ii) the direct effect of ROS on lipid peroxidation multiplied by the direct effect of lipid peroxidation on detection latency.
-Total effects were the sum of direct and indirect effects.
-Refer to @fig-sem_results_OB or @fig-sem_results_OT for a visual representation of the SEM models.
+For our SEM model, the error structure was modelled assuming a gaussian distribution and we estimated residual correlations among all variables.
+
+We obtained direct, indirect, and total effects from posterior estimates in the multivariate models.
+Direct effects represent the posterior estimates of a predictor’s effect on the response variable, while indirect effects were calculated by multiplying these effects on a mediator by the mediator’s effect on the response [@kline2005principles].
+Total effects are the sum of direct and indirect effects.
 
 ## Results
-
-Final sample size for mitochondrial assays was 80 lizards, 20 per treatment.
-These lizards were subjected each to 4 tests getting a total of 320 behavioural observations, out of which `r sum(is.na(clean_df$t_D))` were recorded as NA.
-These animals came from a total of `r clutches` clutches.
+Final sample size for mitochondrial assays was 80 lizards from a total of `r clutches` clutches (n = 20 per treatment).
+Each lizard was subjected to 4 tests for a total of n = 320 behavioural observations (n = `r sum(is.na(clean_df$t_D))` with missing data).
 
@@
 
-#### Mitochondrial physiology
-
-Results for mitochondrial analyses are presented on Tables S13-S15 (Supplementary Material) in means of R^2 scores and the estimates of each predictor in the difinitive models.
-Preliminary models can be found in Tables S1-S12 (Supplementary Material).
-@fig-results_energy and @fig-results_oxidative summarize the effects of early conditions on mitochondrial fuction.
-
-_Mitochondrial density_: neither CORT (Contrast β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(MD_OB$Control_Cold, MD_OB$Control_Hot)) - mean(c(MD_OB$CORT_Cold, MD_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OB$Control_Cold, MD_OB$Control_Hot) - c(MD_OB$CORT_Cold, MD_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(MD_OT$Control_Cold, MD_OT$Control_Hot)) - mean(c(MD_OT$CORT_Cold, MD_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OT$Control_Cold, MD_OT$Control_Hot) - c(MD_OT$CORT_Cold, MD_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(MD_OB$CORT_Hot, MD_OB$Control_Hot)) - mean(c(MD_OB$CORT_Cold, MD_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OB$CORT_Hot, MD_OB$Control_Hot) - c(MD_OB$CORT_Cold, MD_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(MD_OT$CORT_Hot, MD_OT$Control_Hot)) - mean(c(MD_OT$CORT_Cold, MD_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MD_OT$CORT_Hot, MD_OT$Control_Hot) - c(MD_OT$CORT_Cold, MD_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(MD_OB$Control_Hot) - mean(MD_OB$CORT_Hot)) - (mean(MD_OB$Control_Cold) - mean(MD_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MD_OB$Control_Hot - MD_OB$CORT_Hot) - (MD_OB$Control_Cold - MD_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(MD_OT$Control_Hot) - mean(MD_OT$CORT_Hot)) - (mean(MD_OT$Control_Cold) - mean(MD_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MD_OT$Control_Hot - MD_OT$CORT_Hot) - (MD_OT$Control_Cold - MD_OT$CORT_Cold)), 3, equal = TRUE)`) had significant effects on mitochondrial density in neither of the two brain regions extracted (@fig-results_energy A, C).
-
-_Mitochondrial potential_: the effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(MP_OB$Control_Cold, MP_OB$Control_Hot)) - mean(c(MP_OB$CORT_Cold, MP_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OB$Control_Cold, MP_OB$Control_Hot) - c(MP_OB$CORT_Cold, MP_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(MP_OT$Control_Cold, MP_OT$Control_Hot)) - mean(c(MP_OT$CORT_Cold, MP_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OT$Control_Cold, MP_OT$Control_Hot) - c(MP_OT$CORT_Cold, MP_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(MP_OB$CORT_Hot, MP_OB$Control_Hot)) - mean(c(MP_OB$CORT_Cold, MP_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OB$CORT_Hot, MP_OB$Control_Hot) - c(MP_OB$CORT_Cold, MP_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(MP_OT$CORT_Hot, MP_OT$Control_Hot)) - mean(c(MP_OT$CORT_Cold, MP_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(MP_OT$CORT_Hot, MP_OT$Control_Hot) - c(MP_OT$CORT_Cold, MP_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(MP_OB$Control_Hot) - mean(MP_OB$CORT_Hot)) - (mean(MP_OB$Control_Cold) - mean(MP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MP_OB$Control_Hot - MP_OB$CORT_Hot) - (MP_OB$Control_Cold - MP_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(MP_OT$Control_Hot) - mean(MP_OT$CORT_Hot)) - (mean(MP_OT$Control_Cold) - mean(MP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((MP_OT$Control_Hot - MP_OT$CORT_Hot) - (MP_OT$Control_Cold - MP_OT$CORT_Cold)), 3, equal = TRUE)`) were not significant on membrane potential neither for the Olfactory bulbs or the Optic tecta (@fig-results_energy B, D).
+#### Does prenatal CORT and temperature affect mitochondrial function and oxidative stress in the brain?
+Models explained between 37.8 to 50.4 % of the variation in mitochondrial function and 39.9 to 59 % for oxidative stress (see Table S1).
+
+_Olfactory bulbs_: age significantly increased DNA damage and lipid peroxidation (Table S3).
+However, we did not find any significant effects of CORT, temperature, or their interaction on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2).
 
@@
 #
-legend_plot <- get_legend(plot_density_OB + theme(legend.position = "right"))
-#
 fig_OB_energy <- plot_grid(plot_density_OB, NULL, plot_potential_OB, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
+                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
 #
@@
 fig_OT_energy <- plot_grid(plot_density_OT, NULL, plot_potential_OT, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
-#
+                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
+#
+# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
+plot_legend_bottom <- plotting(MP_OB, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
+legend_plot <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
 # C) Merging plots
@@
 fig_results_energy <- plot_grid(
-  fig_OB_energy, fig_OT_energy,
-  nrow = 2)
+  fig_OB_energy, fig_OT_energy, NULL,
+  nrow = 3, rel_heights = c(0.9, 0.9, 0.15))
 # Final composition: Merge everything, adding images and legend
@@
   # Insert OB image in the top right
-  draw_image("./Others/OB.png", x = 0.7, y = 0.8, width = 0.12, height = 0.2) +
+  draw_image("./Others/OB.png", x = 0.87, y = 0.79, width = 0.12, height = 0.2) +
   # Insert OT image in the middle right
-  draw_image("./Others/OT.png", x = 0.7, y = 0.3, width = 0.11, height = 0.18) +
+  draw_image("./Others/OT.png", x = 0.87, y = 0.34, width = 0.11, height = 0.18) +
   # Insert legend at the bottom-right
-  draw_grob(legend_plot, x = 0.85, y = 0.15, width = 0.0001, height = 0.0001) +
+  draw_grob(legend_plot, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001) +
   # Insert title for each plot
-  annotate("text", x = 0.032, y = 0.965, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.39, y = 0.965, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.032, y = 0.47, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.39, y = 0.47, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold")
+  annotate("text", x = 0.032, y = 0.97, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.487, y = 0.97, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.032, y = 0.515, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.49, y = 0.515, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold")
 #
@@
 
-_ROS_: we did not find significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(ROS_OB$Control_Cold, ROS_OB$Control_Hot)) - mean(c(ROS_OB$CORT_Cold, ROS_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OB$Control_Cold, ROS_OB$Control_Hot) - c(ROS_OB$CORT_Cold, ROS_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(ROS_OT$Control_Cold, ROS_OT$Control_Hot)) - mean(c(ROS_OT$CORT_Cold, ROS_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OT$Control_Cold, ROS_OT$Control_Hot) - c(ROS_OT$CORT_Cold, ROS_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(ROS_OB$CORT_Hot, ROS_OB$Control_Hot)) - mean(c(ROS_OB$CORT_Cold, ROS_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OB$CORT_Hot, ROS_OB$Control_Hot) - c(ROS_OB$CORT_Cold, ROS_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(ROS_OT$CORT_Hot, ROS_OT$Control_Hot)) - mean(c(ROS_OT$CORT_Cold, ROS_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(ROS_OT$CORT_Hot, ROS_OT$Control_Hot) - c(ROS_OT$CORT_Cold, ROS_OT$Control_Cold)), 3, equal = TRUE)`), or the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Olfactory bulbs: mean = `r format_dec((mean(ROS_OB$Control_Hot) - mean(ROS_OB$CORT_Hot)) - (mean(ROS_OB$Control_Cold) - mean(ROS_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((ROS_OB$Control_Hot - ROS_OB$CORT_Hot) - (ROS_OB$Control_Cold - ROS_OB$CORT_Cold)), 3, equal = TRUE)`; [(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)] for Optic tecta: mean = `r format_dec((mean(ROS_OT$Control_Hot) - mean(ROS_OT$CORT_Hot)) - (mean(ROS_OT$Control_Cold) - mean(ROS_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((ROS_OT$Control_Hot - ROS_OT$CORT_Hot) - (ROS_OT$Control_Cold - ROS_OT$CORT_Cold)), 3, equal = TRUE)`) on ROS levels in Olfactory bulbs nor Optic tecta (@fig-results_oxidative A, D).
-
-_DNA damage_: there were not significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(DNA_OB$Control_Cold, DNA_OB$Control_Hot)) - mean(c(DNA_OB$CORT_Cold, DNA_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OB$Control_Cold, DNA_OB$Control_Hot) - c(DNA_OB$CORT_Cold, DNA_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(DNA_OT$Control_Cold, DNA_OT$Control_Hot)) - mean(c(DNA_OT$CORT_Cold, DNA_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OT$Control_Cold, DNA_OT$Control_Hot) - c(DNA_OT$CORT_Cold, DNA_OT$CORT_Hot)), 3, equal = TRUE)`) or temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(DNA_OB$CORT_Hot, DNA_OB$Control_Hot)) - mean(c(DNA_OB$CORT_Cold, DNA_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OB$CORT_Hot, DNA_OB$Control_Hot) - c(DNA_OB$CORT_Cold, DNA_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(DNA_OT$CORT_Hot, DNA_OT$Control_Hot)) - mean(c(DNA_OT$CORT_Cold, DNA_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DNA_OT$CORT_Hot, DNA_OT$Control_Hot) - c(DNA_OT$CORT_Cold, DNA_OT$Control_Cold)), 3, equal = TRUE)`) in none of the regions.
-The CORT-temperature interaction was not significant for the Olfactory bulbs ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OB$Control_Hot) - mean(DNA_OB$CORT_Hot)) - (mean(DNA_OB$Control_Cold) - mean(DNA_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OB$Control_Hot - DNA_OB$CORT_Hot) - (DNA_OB$Control_Cold - DNA_OB$CORT_Cold)), 3, equal = TRUE)`).
-However, there was a significant interaction between CORT and temperature in the Optic tecta ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot)) - (mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OT$Control_Hot - DNA_OT$CORT_Hot) - (DNA_OT$Control_Cold - DNA_OT$CORT_Cold)), 3, equal = TRUE)`), where CORT had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`) and decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E).
-Additionally, we found a significant effect of sex on DNA damage in the Optic tecta, and an effect of age on DNA damage in the Olfactory bulbs (Tables S13-S15 Supplementary Material).
-
-_Lipid peroxidation_: we did not see significant effects of CORT (β~Control~ - β~CORT~ for Olfactory bulbs: mean = `r format_dec(mean(c(LP_OB$Control_Cold, LP_OB$Control_Hot)) - mean(c(LP_OB$CORT_Cold, LP_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OB$Control_Cold, LP_OB$Control_Hot) - c(LP_OB$CORT_Cold, LP_OB$CORT_Hot)), 3, equal = TRUE)`; β~Control~ - β~CORT~ for Optic tecta: mean = `r format_dec(mean(c(LP_OT$Control_Cold, LP_OT$Control_Hot)) - mean(c(LP_OT$CORT_Cold, LP_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OT$Control_Cold, LP_OT$Control_Hot) - c(LP_OT$CORT_Cold, LP_OT$CORT_Hot)), 3, equal = TRUE)`) ot temperature (β~Hot~ - β~Cold~ for Olfactory bulbs: mean = `r format_dec(mean(c(LP_OB$CORT_Hot, LP_OB$Control_Hot)) - mean(c(LP_OB$CORT_Cold, LP_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OB$CORT_Hot, LP_OB$Control_Hot) - c(LP_OB$CORT_Cold, LP_OB$Control_Cold)), 3, equal = TRUE)`; β~Hot~ - β~Cold~ for Optic tecta: mean = `r format_dec(mean(c(LP_OT$CORT_Hot, LP_OT$Control_Hot)) - mean(c(LP_OT$CORT_Cold, LP_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(LP_OT$CORT_Hot, LP_OT$Control_Hot) - c(LP_OT$CORT_Cold, LP_OT$Control_Cold)), 3, equal = TRUE)`) on lipid peroxidation in either the Olfactory bulbs nor the Optic tecta.
-However, the interaction between CORT and temperature was significant in the Optic tecta ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot)) - (mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OT$Control_Hot - LP_OT$CORT_Hot) - (LP_OT$Control_Cold - LP_OT$CORT_Cold)), 3, equal = TRUE)`), but not in the Olfactory bulbs ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OB$Control_Hot) - mean(LP_OB$CORT_Hot)) - (mean(LP_OB$Control_Cold) - mean(LP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OB$Control_Hot - LP_OB$CORT_Hot) - (LP_OB$Control_Cold - LP_OB$CORT_Cold)), 3, equal = TRUE)`) (@fig-results_oxidative C, F).
-We found that, in the Optic tecta, CORT has opposite effects on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), but differences between groups were not significant (see @fig-results_oxidative C).
-In addition, age had significant effects on lipid peroxidation for both regions (Tables S13-S15 Supplementary Material).
+_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, mitochondrial potential, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
+However, the interaction of CORT and temperature have significant effects on DNA damage in the optic tecta (see @fig-results_oxidative).
+For DNA damage, CORT elevations had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`), but it decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E).
+Prenatal CORT had opposite effects on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), but differences between groups were not significant (see @fig-results_oxidative C).
+In addition, we found that females had higher DNA damage levels, and lipid peroxidation increased with age (see Table S4).
 
@@
 #
-legend_plot <- get_legend(plot_ros_OB + theme(legend.position = "right"))
-#
-fig_OB_oxidative_top <- plot_grid(plot_ros_OB, NULL, NULL, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
-fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, NULL, plot_peroxidation_OB, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
+fig_OB_oxidative_top <- plot_grid(plot_ros_OB, NULL, NULL,
+                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
+fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, NULL, plot_peroxidation_OB,
+                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
 fig_OB_oxidative <- plot_grid(fig_OB_oxidative_top, fig_OB_oxidative_bottom,
@@
 #
-fig_OT_oxidative_top <- plot_grid(plot_ros_OT, NULL, NULL, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
-fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, NULL, plot_peroxidation_OT, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8, 0.8))
+fig_OT_oxidative_top <- plot_grid(plot_ros_OT, NULL, NULL,
+                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
+fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, NULL, plot_peroxidation_OT,
+                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
 fig_OT_oxidative <- plot_grid(fig_OT_oxidative_top, fig_OT_oxidative_bottom,
@@
 #
+# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
+plot_legend_bottom <- plotting(ROS_OB, "ROS") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
+legend_plot <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
 # C) Merging plots
@@
 fig_results_oxidative <- plot_grid(
-  fig_OB_oxidative, NULL, fig_OT_oxidative,
-  nrow = 3, rel_heights = c(1, 0.05, 1))
+  fig_OB_oxidative, NULL, fig_OT_oxidative, NULL,
+  nrow = 4, rel_heights = c(1, 0.05, 1, 0.1))
 # Final composition: Merge everything, adding images and legend
@@
   # Insert OB image in the top right
-  draw_image("./Others/OB.png", x = 0.7, y = 0.795, width = 0.15, height = 0.25) +
+  draw_image("./Others/OB.png", x = 0.56, y = 0.755, width = 0.2, height = 0.3) +
   # Insert OT image in the middle right
-  draw_image("./Others/OT.png", x = 0.7, y = 0.305, width = 0.14, height = 0.23) +
+  draw_image("./Others/OT.png", x = 0.56, y = 0.28, width = 0.19, height = 0.28) +
   # Insert legend at the bottom-right
-  draw_grob(legend_plot, x = 0.85, y = 0.1, width = 0.0001, height = 0.0001) +
+  draw_grob(legend_plot, x = 0.5, y = 0.025, width = 0.0001, height = 0.0001) +
   # Insert title for each plot
-  annotate("text", x = 0.0325, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0325, y = 0.735, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.3925, y = 0.735, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0325, y = 0.485, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0325, y = 0.235, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.3925, y = 0.235, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("segment", x = 0.32, xend= 0.32, y = 0.17, yend = 0.215, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.31, xend= 0.32, y = 0.17, yend = 0.17, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.31, xend= 0.32, y = 0.215, yend = 0.215, size = 0.5, colour = "black") +
-  annotate("text", x = 0.33, y = 0.187, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
+  annotate("text", x = 0.0365, y = 0.986, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0365, y = 0.755, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.567, y = 0.755, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0365, y = 0.5, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0365, y = 0.268, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.567, y = 0.268, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("segment", x = 0.46, xend= 0.46, y = 0.21, yend = 0.26, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.45, xend= 0.46, y = 0.21, yend = 0.21, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.45, xend= 0.46, y = 0.26, yend = 0.26, size = 0.5, colour = "black") +
+  annotate("text", x = 0.47, y = 0.228, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
 #
 # Print final plot
-ggsave(here("./output/figures/text/results_oxidative.png"), plot = final_plot_oxidative, width = 21, height = 21, units = "cm", dpi = 600, bg = "white")
+ggsave(here("./output/figures/text/results_oxidative.png"), plot = final_plot_oxidative, width = 18, height = 24, units = "cm", dpi = 600, bg = "white")
 knitr::include_graphics("./output/figures/text/results_oxidative.png")
@@
 
-#### Prey detection
-
-Results for behaviour analyses are presented on Tables S13-S15 (Supplementary Material) in means of R^2 scores and the estimates of each predictor in the difinitive models.
-Preliminary models can be found in Tables S1-S12 (Supplementary Material).
-@fig-results_behaviour summarizes the effects of early conditions on detection latency for both types of stimulus.
-
-Lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature (β~Hot~ - β~Cold~: mean = `r format_dec(mean(c(DET_OB$CORT_Hot, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$CORT_Hot, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$Control_Cold)), 3, equal = TRUE)`), nor the interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DET_OB$Control_Hot) - mean(DET_OB$CORT_Hot)) - (mean(DET_OB$Control_Cold) - mean(DET_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DET_OB$Control_Hot - DET_OB$CORT_Hot) - (DET_OB$Control_Cold - DET_OB$CORT_Cold)), 3, equal = TRUE)`) in chemo-detection (@fig-results_behaviour A).
-In contrast, there were no significant effects of CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OT$Control_Cold, DET_OT$Control_Hot)) - mean(c(DET_OT$CORT_Cold, DET_OT$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OT$Control_Cold, DET_OT$Control_Hot) - c(DET_OT$CORT_Cold, DET_OT$CORT_Hot)), 3, equal = TRUE)`), temperature (β~Hot~ - β~Cold~: mean = `r format_dec(mean(c(DET_OT$CORT_Hot, DET_OT$Control_Hot)) - mean(c(DET_OT$CORT_Cold, DET_OT$Control_Cold)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OT$CORT_Hot, DET_OT$Control_Hot) - c(DET_OT$CORT_Cold, DET_OT$Control_Cold)), 3, equal = TRUE)`), or their interaction ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DET_OT$Control_Hot) - mean(DET_OT$CORT_Hot)) - (mean(DET_OT$Control_Cold) - mean(DET_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DET_OT$Control_Hot - DET_OT$CORT_Hot) - (DET_OT$Control_Cold - DET_OT$CORT_Cold)), 3, equal = TRUE)`) on the detection latency of visual stimulus (@fig-results_behaviour B).
+#### Does prenatal CORT and temperature affect prey detection?
+Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).
+
+_Chemical stimulus_: lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature, nor the interaction in chemo-detection (@fig-results_behaviour A and Table S2).
 
@@
 #
-legend_plot <- get_legend(plot_t_D_Chemical + theme(legend.position = "right"))
-#
+# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
+plot_legend_bottom <- plotting(DET_OB, "Detection latency") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
+legend_plot <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
 # C) Merging plots
@@
 # Create the figure grid with extra space for images
-fig_results_behaviour <- plot_grid(plot_t_D_Chemical, NULL, plot_t_D_Visual, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.5, 0.8, 0.7))
+fig_results_det <- plot_grid(plot_t_D_Chemical, NULL, plot_t_D_Visual, NULL,
+                    nrow = 1, rel_widths = c(1, 0.4, 1, 0.35))
+fig_results_behaviour <- plot_grid(fig_results_det, NULL,
+  nrow = 2, rel_heights = c(1, 0.15))
 # Final composition: Merge everything, adding images and legend
@@
   # Insert CS image in the top middle
-  draw_image("./Others/CS.png", x = 0.165, y = 0.47, width = 0.4, height = 0.5) +
+  draw_image("./Others/CS.png", x = 0.19, y = 0.36, width = 0.48, height = 0.6) +
   # Insert VS image in the top right
-  draw_image("./Others/VS.png", x = 0.6, y = 0.47, width = 0.4, height = 0.5) +
+  draw_image("./Others/VS.png", x = 0.6, y = 0.35, width = 0.65, height = 0.6) +
   # Insert legend at the bottom-right
-  draw_grob(legend_plot, x = 0.885, y = 0.23, width = 0.0001, height = 0.0001) +
+  draw_grob(legend_plot, x = 0.45, y = 0.07, width = 0.0001, height = 0.0001) +
   # Insert title for each plot
   annotate("text", x = 0.035, y = 0.96, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.5, y = 0.96, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +  
-  annotate("segment", x = 0.27, xend= 0.27, y = 0.69, yend = 0.88, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.266, xend= 0.27, y = 0.69, yend = 0.69, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.266, xend= 0.27, y = 0.88, yend = 0.88, size = 0.5, colour = "black") +
-  annotate("text", x = 0.28, y = 0.76, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
-  annotate("segment", x = 0.27, xend= 0.27, y = 0.32, yend = 0.51, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.266, xend= 0.27, y = 0.32, yend = 0.32, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.266, xend= 0.27, y = 0.51, yend = 0.51, size = 0.5, colour = "black") +
-  annotate("text", x = 0.28, y = 0.39, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
+  annotate("text", x = 0.542, y = 0.96, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +  
+  annotate("segment", x = 0.34, xend= 0.34, y = 0.73, yend = 0.91, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.335, xend= 0.34, y = 0.73, yend = 0.73, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.335, xend= 0.34, y = 0.91, yend = 0.91, size = 0.5, colour = "black") +
+  annotate("text", x = 0.35, y = 0.795, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
+  annotate("segment", x = 0.292, xend= 0.292, y = 0.418, yend = 0.598, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.287, xend= 0.292, y = 0.418, yend = 0.418, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.287, xend= 0.292, y = 0.598, yend = 0.598, size = 0.5, colour = "black") +
+  annotate("text", x = 0.302, y = 0.471, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
 #
@@
 
-#### SEM results
-
+_Visual stimulus_: there were no significant effects of CORT, temperature, or their interaction on the detection latency of visual stimulus (@fig-results_behaviour B and Table S2).
+
+#### How are mitochondrial physiology and oxidative stress related to prey detection?
 ```{r, model_sem_OB}
@@
 data_sem_OB <- SEM_df %>%
-  filter(region == "OB")
+  filter(region == "OB") %>%
+  mutate(obs = as.integer(c(1:80))) %>%
+  mutate(vec = rep(1, length(obs)))
 # Create the models
@@
   m_OB <- brm(
-    bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
@@
 
-Results for SEM models are presented on Tables S16-S18 (Supplementary Material) in means of R^2 scores and the direct, indirect, and total effects for all variables.
-@fig-sem_results_OB and @fig-sem_results_OT summarize direct effects between the variables included in the model for bot brain regions and stimuli.
-
-_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found significant increases of ROS with increases in mitochondrial density (Direct effect estimation: β~D~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_det), 3, equal = TRUE)`) and mitochondrial potential (β~D~ = `r format_dec(mean(OB_coeff_potential_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_potential_det), 3, equal = TRUE)`) (see @fig-sem_results_OB).
-However, there were no significant direct effects of ROS on DNA damage or lipid peroxidation; neither of mitochondrial density, mitochondrial potential, DNA damage, or lipid peroxidation on the detection abilities (see @fig-sem_results_OB and Table S17 in Supplementary Material).
-Age increased DNA damage (β~D~ = `r format_dec(mean(OB_coeff_age_dna), 3)`m pMCMC `r format_p(pmcmc(OB_coeff_age_dna), 3, equal = TRUE)`) and lipid peroxidation (β~D~ = `r format_dec(mean(OB_coeff_age_perox), 3)`m pMCMC `r format_p(pmcmc(OB_coeff_age_perox), 3, equal = TRUE)`) in the Olfactory bulbs, but there was no indirect effect on detecion latency (Indirect effect estimation β~I~ = `format_dec(mean(OB_undir_age_det), 3)`m pMCMC `r format_p(pmcmc(OB_undir_age_det), 3, equal = TRUE)`) Additionally, CORT shows a week effect on detection latency (β~D~ = `r format_dec(mean(OB_coeff_cort_det), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_cort_det), 3, equal = TRUE)`), suggesting that CORT treated animals detect chemical stimulus faster; but the effect here was not significant.
+_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found no significant effects of mitochondrial function or oxidative stress on prey detection.
+There were no significant effects between any other of the variables either (see @fig-sem_results_OB and Table S5).
 
@@
 data_sem_OT <- SEM_df %>%
-  filter(region == "OT")
+  filter(region == "OT") %>%
+  mutate(obs = as.integer(c(1:80))) %>%
+  mutate(vec = rep(1, length(obs)))
 #
@@
   m_OT <- brm(
-    bf(t_D ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ cort*temp + sex + mean_ros + (1|clutch)) +
-    bf(mean_peroxidation ~ cort*temp + age_euthanasia + mean_ros + (1|clutch)) +
-    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|p|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ cort*temp + sex + mean_ros + (1|clutch) + (1|q|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ cort*temp + age_euthanasia + mean_ros + (1|clutch) + (1|q|obs)) +
+    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
@@
 }
-#
 ```
@@
 
-_Detection of visual stimulus and its relation with mitochondrial physiology in Optic tecta_: mitochondrial density had big significant effects on ROS production (β~D~ = `r format_dec(mean(OT_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OT_coeff_mitodensity_det), 3, equal = TRUE)`).
-However, we did not find any significant direct or indirect effects of ROS on detection latency, DNA damage, or lipid peroxidation (see @fig-sem_results_OT and Table S18 in Supplementary Material).
-Mitochondrial potential did not affect ROS or any other variable in the model (see @fig-sem_results_OT and Table S18 in Supplementary Material).
-We also did not see any effects of ROS production on DNA damage or lipid peroxidation, neither of DNA damage or lipid peroxidation on detection latency.
-In addition, we found weak non significant effects of sex and CORT on DNA damage, neither age and CORT on lipid peroxidation (see @fig-sem_results_OT and Table S18 in Supplementary Material).
+_Detection of visual stimulus and its relation with mitochondrial physiology in Optic tecta_: we found no significant relationships within mitochondrial variables or between mitochondrial physiology and cognitive performance (see @fig-sem_results_OT and Table S6).
 
@@
 ## Discussion
-
-- Early environmental conditions did not have significant effects on mitochondrial density, potential, nor ROS production in neither of the brain regions studied.
-
-- Mitocondrial density was a good predictor of ROS production in both regions, but mitochondrial potential was only a good predictor in Olfactory bulbs.
-We might be missing interactions with other variables that could explain the differences between regions.
-
-- Early conditions did not affect DNA damage on Olfactory bulbs, but had a significant effect on DNA damage in Optic tecta.
-CORT reduced DNA damage in cold incubated lizards but had no effects at high incubation temperatures.
-This effect was not mediated by ROS production or any other mitochondrial-related measurement.
-It could be because of the way CORT can redistribute resources to cope with stressors, and in this case, it could be reducing DNA damage in the brain.
-
-- Early conditions had a weak effect on lipid peroxidation in both regions.
-The effect of CORT seems to increase lipid peroxidation at low temperatures, and increase it at higher.
-This patern is similar in both Olfactory bulbs and Optic tecta.
-It could be related to the way lipids are organised at different temperatures, and how CORT can affect the way they are metabolised.
-
-- Age has significant effects on DNA damage and lipid peroxidation in Olfactory bulbs (as expected), but a weak effect on lipid peroxidation in Optic tecta.
-Females had lower DNA damage in Olfactory bulbs (marginally significant) and slightly higher in the Optic tecta.
-Aging increases accumulative damage.
-Males could be investing more in repairing or avoiding damage in chemosensory areas because they could be more important for sexual-related behaviours.
-
-- Oxidative damage was not influence by ROS production or any other mitochondrial-related measurement in either of the brain regions studied.
-It could be that measuring only H2O2 is too reductionist, or the action of antioxidants, which, at the same time, could potentially explained the patterns mentioned before.
-
-- Detection latency decreased with CORT exposure when animals were presented chemical stimulus, but not with visual.
-It could be related to the fact that motivation was not totally controlled.
-Motivation tests presented both types of stimulus, but it could be that stressed animals were more motivated to search for food and, therefore, interacted faster with the vessel in chemical tests.
-
-- Performance in the behavioral tests was not affected by mitochondrial physiology.
-It could be that the brain regions studied are not directly related to the behaviors tested, that other parameters are more relevant, or that the oxidatie damage is not high enought for detecting an effect.
+In this study, we examined the impact of prenatal temperature and CORT elevations on lizards' ability to detect different stimuli from prey together with the impacts of early conditions on mitochondrial metabolic function and oxidative stress in two brain regions associated to perception of visual and chemical stimuli.
+Our findings suggest that mitochondrial metabolic function remained largely resilient to early environmental conditions, yet oxidative damage and behavioral responses exhibited region-specific and context-dependent variations.
+Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioral performance.
+
+#### Mitochondrial function and ROS production is robust to early life conditions
+We found no significant effects of early environmental conditions on mitochondrial density, membrane potential, or ROS production in either of the brain regions studied.
+This suggests that, contrary to our predictions, mitochondrial function is robust to prenatal temperature and CORT.
+Previous studies have shown that high incubation temperatures - but not CORT - decreased mitochondrial efficiency in the liver of _L.
+delicata_ [@crino2024eggs].
+Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing efficiency [@rice2002brain].
+Given the brain's high energy demands [@mcnay_decreases_2000; @mann_coupling_2021], it is plausible that mitochondria are particularly well-protected, which may help explain their resilience towards environmental inputs.
+Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.
+
+#### The interactive effects of prenatal CORT and temperature on oxidative damage are region-dependent
+We found that DNA damage and lipid peroxidation were affected in region-specific ways.
+Early conditions did not significantly influence oxidative stress in the olfactory bulbs, but had a pronounced effect on DNA damage and lipid peroxidation in the optic tecta.
+
+CORT exposure reduced DNA damage in optic tecta from cold-incubated lizards, but had no effect at higher incubation temperatures.
+This pattern suggests that CORT might play a protective role in specific thermal conditions.
+Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — these resources can be directed towards antioxidant activity or DNA repairing mechanisms.
+In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms.
+
+In comparison, prenatal CORT elevations increased lipid peroxidation in cold incubated lizards but reduced it in those incubated at higher temperatures.
+Although the two hormonal treatments did not significantly differ within each incubation temperature, the direction of the CORT effect was the opposite between temperatures.
+One possible reason for this pattern could be temperature-dependent changes in membrane lipid composition or fluidity, which may affect how lipids respond to oxidative stress [@zeis2019temperature].
+For instance, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids [@zeis2019temperature].
+Future studies should explore how temperature-dependent effects on membrane composition influence oxidative damage susceptibility.
+
+#### Detection of chemical, but not visual, stimulus improved with prenatal CORT exposure
+Detection latency decreased with CORT exposure when animals were presented with a chemical stimulus, but not with a visual stimulus.
+This effect was not related to alterations in mitochondrial physiology in the olfactory bulbs.
+Instead, we hypothesize that it may be driven by differences in motivation.
+Although lizards were fasted for two days before testing and we specifically tested for motivation, we may have overlooked factors related to the the ecological relevance of the cues, the animals' internal state, or the interaction between both factors.
+Prenatal CORT is expected to increase hunger through resource mobilization [@spencer2008post; @cossin-sevrin_effect_2022; but see @crino_corticosterone_2014-learn].
+This could have lead to increase motivation in CORT-treated lizards, which could have switched to a more active foraging mode, affecting the way they respond to chemical stimuli.
+For example, species of European lizards (family Lacertidae) that spend more time actively looking for food tend to rely more on chemoreception [@verwaijen2007relationships].
+Further studies need to be conducted to understand how CORT can change foraging strategies in lizards and how this can affect their perception abilities.
+
+#### Mitochondrial function and oxidative stress do not affect detection abilities
+Despite the relevance of mitochondrial function and oxidative damage on cognitive abilities in other taxa [@zhu_prenatal_2004; @cao2014ampk; @hara_presynaptic_2014] we found no significant relationships between mitochondrial physiology and prey detection in _L.
+delicata_.
+This suggest that the brain regions studied are not directly related to the behaviors tested, or that other mechanisms are more relevant for preceiving prey stimuli.
+Alternatively, mitochondrial physiology may influence detection abilities only beyond a certain threshold, which may not have been reached in our study.
+
+In this regard, mitochondria are predicted to impact cognitive performance by providing energy for neural activity or by regulating oxidative stress [@siegel1994basic; @mcnay_decreases_2000; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018; @mann_coupling_2021; @alexandrov_neuronal_2022].
+However, the lack of effect of mitochondrial density and mitochondrial potential on detection latency abilities in _L.
+delicata_ suggests that energy production is not a limiting factor for prey perception, or that _L.
+delicata_ can sustain this cognitive function despite fluctuations in energy availability.
+Similarly, the levels of damage in the brain regions studied may not have been high enough to impair prey perception in this species or compensatory mechanisms may have counteracted any negative effects of oxidative stress.
+However, oxidative damage can accumulate over time, potentially pronouncing the effects of oxidative damage on cognition over time [@terman2006oxidative; @hara_presynaptic_2014].
+In fact, we saw a significant effect of age on DNA damage in olfactory bulbs, and on lipid peroxidation in both regions.
+Future research should investigate the long-term consequences of mitochondrial activity on brain health and the potential mechanisms that sustain cognitive resilience in _L.
+delicata_.
+
+#### Conclusions
+Prenatal incubation and CORT exposure had region-specific effects in the brain of _L.
+delicata_.
+While mitochondrial function remained resilient to early environmental conditions, oxidative damage and behavioral responses exhibited context-dependent variations.
+Our findings suggest that the brain regions studied are not directly related to the behaviors tested, or that other mechanisms are more relevant for preceiving prey stimuli.
+Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.
+
+### Ethics
+Both the breeding animals and the experimental lizards were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate levels of humidity.
+Euthanasia was performed by peritoneal injection of a 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
+We monitored the animals to ensure there was no irritation from the agent as indicated by distressed animals.
+Before disposing of the lizard, we confirmed the absence of righting response and pinching reflex in one of the front limbs.
+All the protocols complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
+
+### Data accessibility  
+All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT-Temp_PreyD](https://github.com/Pablo-Recio/CORT_Temp_PreyD).
+
+### Declaration of AI use
+We declare Chat GPT was used mainly for questions related to coding and data analyses.
+All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+
+### Authors' contributions
+P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: conceptualization, methodology, data collection, data curation, writing—review and editing; O.C.: conceptualization, methodology, writing—review and editing; C.F.: conceptualization, methodology, funding acquisition, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.
+All authors gave final approval for publication and agreed to be held accountable for the work performed therein.
+
+### Conflict of interest declaration  
+We declare we have no competing interests.
+
+### Funding
+This work was supported by a National Australian University fellowship (to P.R. and D.C.L.), and the Australian Research Council (grant no. DP210101152) to D.N. and C.R.F.
+
+### Acknowledgements
+We thank Dr.
+Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
+We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards.
+We also thank ANU MakerSpace, where we designed and built the prototypes of the 3D printed feeders.
+Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.
 
@@
 
-#### Results of preliminary models
-
-```{r, models_preliminary}
-#| label: models_preliminary
-# Fitting intial models to see if sex and age are relevant for our models
-source(here("R", "func.R"))
-#
-#
-# Run models mitochondrial physiology (each region separately)
-#
-var_m <- c("mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
-regions <- c("OB", "OT")
-formula_list_ <- list()
-for (p in var_m){
-  formula_list_[[p]] <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
-  for (h in regions){
-    if (h == "OB"){
-      df <- clean_df %>% filter(region == "OB")
-      l <- "OB"
-    } else {
-      df <- clean_df %>% filter(region == "OT")
-      l <- "OT"
-    }
-
-  pmodel_name <- paste0("m_prel_", p, "_", h)
-  assign(pmodel_name, fit_m(df = df,
-                             cat = "prel",
-                             var = p,
-                             formula = formula_list_[[p]],
-                             fam = gaussian(),
-                             label = l,
-                             refit = FALSE),
-          envir = .GlobalEnv)  # Assign to the global environment
-  }
-}
-#
-#
-# Run model behaviour (stimuli separated)
-beh_df <- clean_df
-stimulus <- c("Chemical", "Visual")
-formula_t_D <- t_D ~ motivation + motivation:cort + cort*temp + age_trial + sex + prey +(1|clutch) + (1|lizard_id) 
-for (k in stimulus){
-  df <- beh_df %>% filter(stimulus == k)
-  pmodel_name <- paste0("m_prel_t_D_", k)
-  assign(pmodel_name, fit_m(df = df,
-                            cat = "prel",
-                            var = "t_D",
-                            formula = formula_t_D,
-                            fam = gaussian(),
-                            label = k,
-                            refit = FALSE),
-        envir = .GlobalEnv)  # Assign to the global environment
-}
-```
-
-Table S1. Preliminary results of the models testing for Mitochondrial Density in Olfactory Bulbs.
-
-```{r, results_preliminary_mitdensity_OB}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OB"
-#| label: results_preliminary_mitdensity
-#
-sum_m_mitdensity_OB_prel <- m_prel_mean_mitodensity_OB %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws()  %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_mitdensity_OB_prel)
-```
-
-Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S2. Preliminary results of the models testing for Mitochondrial Density in Optic Tecta.
-```{r, results_preliminary_mitdensity_OT}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OT"
-#| label: results_preliminary_mitdensity_OT
-#
-sum_m_mitdensity_OT_prel <- m_prel_mean_mitodensity_OT %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_mitdensity_OT_prel)
-```
-
-Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S3. Preliminary results of the models testing for Mitochondrial Potential in Olfactory Bulbs.
-```{r, results_preliminary_potential_OB}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OB"
-#| label: results_preliminary_potential_OB
-#
-sum_m_potential_OB_prel <- m_prel_mean_potential_OB %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_potential_OB_prel)
-```
-
-Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S4. Preliminary results of the models testing for Mitochondrial Potential in Optic Tecta.
-```{r, results_preliminary_potential_OT}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OT"
-#| label: results_preliminary_potential_OT
-#
-sum_m_potential_OT_prel <- m_prel_mean_potential_OT %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_potential_OT_prel)
-```
-
-Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S5. Preliminary results of the models testing for ROS Production in Olfactory Bulbs.
-```{r, results_preliminary_ros_OB}
-#| tbl-cap: "Preliminary results of the models testing for ROS Production in OB"
-#| label: results_preliminary_ros_OB
-#
-sum_m_ros_OB_prel <- m_prel_mean_ros_OB %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_ros_OB_prel)
-```
-
-Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S6. Preliminary results of the models testing for ROS Production in Optic Tecta.
-```{r, results_preliminary_ros_OT}
-#| tbl-cap: "Preliminary results of the models testing for ROS Production in OT"
-#| label: results_preliminary_ros_OT
-#
-sum_m_ros_OT_prel <- m_prel_mean_ros_OT %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_ros_OT_prel)
-```
-
-Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S7. Preliminary results of the models testing for DNA Damage in Olfactory Bulbs.
-```{r, results_preliminary_dnadamage_OB}
-#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OB"
-#| label: results_preliminary_dnadamage_OB
-#
-sum_m_dnadamage_OB_prel <- m_prel_mean_dnadamage_OB %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_dnadamage_OB_prel)
-```
-
-Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates an effect of sex and age, so they were included in the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S8. Preliminary results of the models testing for DNA Damage in Optic Tecta.
-```{r, results_preliminary_dnadamage_OT}
-#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OT"
-#| label: results_preliminary_dnadamage_OT
-#
-sum_m_dnadamage_OT_prel <- m_prel_mean_dnadamage_OT %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_dnadamage_OT_prel)
-```
-
-Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of age, so it was discarded from the final models.
-However, we saw an effect of sex and we included it in our final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S9. Preliminary results of the models testing for Lipid Peroxidation in Olfactory Bulbs.
-```{r, results_preliminary_peroxidation_OB}
-#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OB"
-#| label: results_preliminary_peroxidation_OB
-#
-sum_m_peroxidation_OB_prel <- m_prel_mean_peroxidation_OB %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_peroxidation_OB_prel)
-```
-
-Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of sex, so it was discarded from the final models.
-However, we saw an effect of age and we included it in our final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S10. Preliminary results of the models testing for Lipid Peroxidation in Optic Tecta.
-```{r, results_preliminary_peroxidation_OT}
-#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OT"
-#| label: results_preliminary_peroxidation_OT
-#
-sum_m_peroxidation_OT_prel <- m_prel_mean_peroxidation_OT %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_peroxidation_OT_prel)
-```
-
-Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of sex, so it was discarded from the final models.
-However, we saw an effect of age and we included it in our final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S11. Preliminary results of the models testing for Detection Latency (t_D) of Chemical stimuli. 
-```{r, results_preliminary_t_D_Chemical}
-#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Chemical stimuli"
-#| label: results_preliminary_t_D_Chemical
-# 
-sum_m_t_D_prel_Chem <- m_prel_t_D_Chemical %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_t_D_prel_Chem)
-```
-
-Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S12. Preliminary results of the models testing for Detection Latency (t_D) of Visual stimuli.
-```{r, results_preliminary_t_D_Visual}
-#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Visual stimuli"
-#| label: results_preliminary_t_D_Visual
-#
-sum_m_t_D_prel_Vis <- m_prel_t_D_Visual %>%
-  dplyr::select(-starts_with("r_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_t_D_prel_Vis)
-```
-
-Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-#### Results of the main models
-
-Table S13. R^2 values of the definite models for both regions and stimuli.
+#### Methods: Flow cytometry
+Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
+We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
+Lizards without responses were decapitated with surgical scissors.
+This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
+After decapitation, the head was opened and the brain was dissected.
+We extracted manually two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with chemical and visual perception in lizards [@wyneken2007reptilian].
+Both regions were transferred immediately to 1.5mL centrifuge tubes containing 100µL of 1X phosphate buffered saline (PBS).
+
+Tissue suspended in PBS was then homogenized.
+For the olfactory bulbs, the tissue was mechanically homogenized by placing the tissue in the well of a 100 µm mesh filter (pluriStrainer) affixed atop a 1.5 mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
+The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1X PBS to prepare a homogenate suspension.
+Optic tecta were mechanically homogenized the same way as olfactory bulbs, but were first enzymatically digested by incubating the tissue in 100µL of 125 U/mL collagenase (type II) for 30-min at 32°C.
+We enzymatically digested optic tecta but not olfactory bulbs because in a pilot study, enzymatic digestion prior to mechanical digestion improved the yield of viable cells in optic tecta homogenate, but did not noticeably affect the yield of viable cells in olfactory bulbs homogenate (unpublished data).
+Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500 µL 1x PBS.
+This step was to performed to remove cellular debris from homogenates.
+From each 500 µL suspension of homogenate collected on a given trial day, we first added 100 µL of homogenate to a pooled sample of each tissue type to use for single-color controls, and the remaining 40 0µL of homogenate was split among two 200 µL aliquots.
+One aliquot was used fresh to measure mitochondrial function (mitochondrial density, membrane potential, ROS), one aliquot was cryopreserved for later measurements of oxidative damage (8-OHdG, lipid peroxidation), and the third aliquot was cryopreserved for a different experiment.
+Aliquots were cryopreserved in 1 mL solution of 1% Neutral-Buffered Formalin (as a fixative agent), 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant), then stored at -20 °C until oxidative damage assays.
+
+From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50 µL of homogenate in duplicates (2 wells per homogenate).
+To each replicate well, we added 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
+We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
+We then added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
+We then loaded 6 wells with 50 µL of homogenate taken from each pooled homogenate suspension (12 wells total), which were to be negative and single-color controls.
+One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5 µL of one of 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, 50 µM MitoSOX Red, or #µM Hoechst 33342 Nuclear Viability Dye.
+Any remaining pooled homogenate was fixed and frozen as previously described.
+We incubated the loaded plate at 32 °C for 30 minutes to stain and then diluted the samples with 50 µL cold 1x PBS to halt the staining process.
+Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours.
+Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).
+
+Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in near-boiling water.
+We washed each thawed sample twice, the first time resuspending the pelleted cells in 1 mL warm 1X Tris-EDTA, and the second time in 200 µL warm 1X PBS.
+Pooled homogenate was divided into 100 µL aliquots to serve as single-color controls.
+Thawed, resuspended samples were first stained with 20 µL (10 µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32 °C for 20 minutes.
+Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin.
+We incubated the samples at 32 °C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200 µL 1X PBS.
+We added 20 µL of 70 µM 8-OHdG Polyclonal Antibody to each sample, after which we left the homogenate overnight (~12 hours) for the antibody to bind to 8-OHdG, a marker of oxidative damage on DNA.
+The following day we counterstained the cells with 20 µL of #µM (100 µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32 °C for 20 minutes.
+After the cells had been tagged with 8-OHdG antibodies and counterstained, we washed the cells once more and resuspended the pellet in 400 µL of 1X PBS.
+Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate.
+Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated.
+We then loaded a 96-well plate with 100 µL of each single-color control and 100 µL in duplicate of each sample.
+We performed all flow cytometry assays on samples within 48-hours of thawing the samples.
+
+All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
+Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v.
+8.0.1) and BD CS&T fluorescent beads (Lot No.
+30664) diluted at 2 drops to 150 µL 1X PBS.
+During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector.
+These thresholds were chosen to filter small debris or inviable or non-intact cells from our observations.
+The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment.
+Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events.
+For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels.
+For the oxidative damage assay, we recorded data from the following channels: FSC (425; forward scatter), SSC (300; side scatter), Alexa-Fluor 488 (275; 8-OHdG Antibody + Alexa-Fluor 488 conjugate), BUV-496 (525; Hoechst 33342), and PE-Cy5 (850; BODIPY 665/676).
+Fluorescent intensity data was collected via the BD FACSDiva (v.
+8.0.1) software, with no compensation applied during data collection, and all on a linear scale (detectable range of 0-252166).
+We recorded data for both the area and height of the fluorescent signal, but only used the area in downstream analyses, with height being recorded for the sake of quality control.
+Data was exported from BD FACSDiva as individual *.fcs (“flow cytometry standard”) files for each sample, then imported into FlowJo (v.
+10.1) for processing.
+
+In FlowJo v.
+10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events.
+We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains.
+We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103).
+For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103).
+For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103).
+When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population.
+We used the backgated population for compensation of fluorescent spillover between different fluorescent probes.
+To account for fluorescent spillover, we used a traditional compensation matrix using the compensation function of FlowJo v.10.1.
+We identified the “positive” population for each channel as the brightest ~2.5% of the distribution of observations in the respective single-color control for that channel and used unstained controls as a universal negative.
+We visually inspected the compensation matrix and its effects on population distributions for under- and over-compensation, whereupon we changed the compensation matrix manually until data was properly compensated.
+We applied the compensation matrix to all samples for downstream processing.
+Following compensation, we again gated the data following the same process as for backgating but using the compensated parameters for each channel.
+Following gating, we exported the geometric mean (mean fluorescent intensity; MFI) and robust confidence-values for each channel for each sample.
+For analysis, we exported summary statistics of only the area of the fluorescent signal.
+Although we exported robust confidence values for checking repeatability between replicate samples, we used the mean and standard deviation of geometric means among replicates for each individual as our main response variables in analyses.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Results of the final models
+
+Table S1. R^2^ values of the definite models for both regions and stimuli.
 ```{r, table_bayesR2}
@@
 
-Table S14. Results of the definitive models for Olfactory Bulbs/Chemical stimulus.
+Table S2.
+Estimate of the contrasts between treatments for all the variables analysed, and pMCMC values testing the hypothesis that contrast is different from 0.
+In bold, the values that are significant at pMCMC < 0.05.
+```{r tbl-contrasts}
+#| label: tbl-contrasts
+#| tbl-cap: "Contrasts between prenatal conditions for mitochondrial physiology and learning."
+#
+source(here("R", "func.R"))
+#
+# A) Organise df 
+var <- c("MD", "MP", "ROS", "DNA", "LP", "DET")
+region <- c("OB", "OT")
+#
+data_table <- data.frame()
+for(v in var){
+  for(r in region){
+    x <- paste0(v, "_", r)
+    df <- get(x)
+    Temperature <- format_dec(mean(c(df$CORT_Hot, df$Control_Hot)) - mean(c(df$CORT_Cold, df$Control_Cold)), 3)
+    pMCMC_temp <- format_p(pmcmc(c(df$CORT_Hot, df$Control_Hot) - c(df$CORT_Cold, df$Control_Cold)), 3, equal = FALSE)
+    CORT <- format_dec(mean(c(df$Control_Hot, df$Control_Cold)) - mean(c(df$CORT_Hot, df$CORT_Cold)), 3)
+    pMCMC_cort <- format_p(pmcmc(c(df$Control_Hot, df$Control_Cold) - c(df$CORT_Hot, df$CORT_Cold)), 3, equal = FALSE)
+    Interaction <- format_dec((mean(df$Control_Hot) - mean(df$CORT_Hot)) - (mean(df$Control_Cold) - mean(df$CORT_Cold)), 3)
+    pMCMC_int <- format_p(pmcmc((df$Control_Hot - df$CORT_Hot) - (df$Control_Cold - df$CORT_Cold)), 3, equal = FALSE)
+    data_temp <- data.frame(Variable = x,
+                          Temperature = as.numeric(Temperature),
+                          pMCMC_temp = pMCMC_temp,
+                          CORT = as.numeric(CORT),
+                          pMCMC_cort = pMCMC_cort,
+                          Interaction = as.numeric(Interaction),
+                          pMCMC_int = pMCMC_int)
+    data_table <- dplyr::bind_rows(data_table, data_temp)
+  }
+}
+# Modify the df
+data_table_final <- data_table %>%
+  pivot_longer(cols = c(Temperature, CORT, Interaction), 
+               names_to = "Predictor", 
+               values_to = "Contrast") %>%
+  mutate(
+    # Extract the pMCMC values from the corresponding columns
+    `pMCMC contrast` = case_when(
+      Predictor == "Temperature" ~ pMCMC_temp,
+      Predictor == "CORT" ~ pMCMC_cort,
+      Predictor == "Interaction" ~ pMCMC_int
+    )
+  ) %>%
+  separate(col = Variable, into = c("Variable", "Region"), sep = "_") %>%
+  mutate(
+    Variable = case_when(
+      Variable == "MD" ~ "Mit density",
+      Variable == "MP" ~ "Mit potential",
+      Variable == "ROS" ~ "ROS",
+      Variable == "DNA" ~ "DNA damage",
+      Variable == "LP" ~ "Lipid peroxidation",
+      Variable == "DET" ~ "Detection latency",
+      TRUE ~ Variable
+      )
+    ) %>%
+  mutate(Variable = factor(Variable, levels = c("Mit density", "Mit potential", "ROS", "DNA damage", "Lipid peroxidation", "Detection latency"))) %>%
+  mutate(
+    Region = case_when(
+      Region == "OB" ~ "Olfactory bulbs / Chemical",
+      Region == "OT" ~ "Optic tecta / Visual",
+      TRUE ~ Region
+    )
+  ) %>%
+  dplyr::select(Region, Variable, Predictor, Contrast, `pMCMC contrast`) %>%
+  arrange(Region, Variable)
+#
+# C) Make the contrasts table:
+#
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+contrast_table <- flextable(data_table_final) %>%
+  align(align = "center", j = c(4,5), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>%
+  bold(~`pMCMC contrast` < 0.05, j = c("pMCMC contrast", "Contrast", "Predictor"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC contrast` <0.001, j = c("pMCMC contrast", "Contrast", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:18,20:36), j = 1, value = as_paragraph(""), part = "body") %>%
+  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24,26,27,29,30,32,33,35,36),
+                    j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
+  flextable::hline(i = c(3,6,9,12,15,21,24,27,30,33,36), j = c(2:5), part = "body") %>% 
+  flextable::hline(i = 18, part = "body") %>%
+  autofit()
+#
+contrast_table
+```
+
+Contrasts were done by:  
+_Temperature_: β~Hot~ - β~Cold~  
+_CORT_: β~CORT~ - β~Control~    
+_Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S3. Summary of the final models for Olfactory Bulbs/Chemical stimulus.
 ```{r, tbl-results_OB_table}
@@
 
-Table S15. Results of the definitive models for Optic Tecta/Visual stimulus.
+Table S4. Summary of the final models for Optic Tecta/Visual stimulus.
 ```{r, tbl-results_OT_table}
@@
 
-#### Main models diagnostics (plots)
-
-```{r , plotmod_mitdensity_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_mitdensity_OB
-#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_mitodensity_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
-```
-
-Fig S1. Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs.
-Formula: mean_mitodensity ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_mitdensity_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_mitdensity_OT
-#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_mitodensity_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
-```
-
-Fig S2. Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta.
-Formula: mean_mitodensity ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_potential_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_potential_OB
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_potential_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
-```
-
-Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
-Formula: mean_potential ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_potential_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_potential_OT
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_potential_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
-```
-
-Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
-Formula: mean_potential ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_ros_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_ros_OB
-#| caption: "Posterior predictive checks for the model of ROS Production in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_ros_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
-```
-
-Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
-Formula: mean_ros ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_ros_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_ros_OT
-#| caption: "Posterior predictive checks for the model of ROS Production in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_ros_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
-```
-
-Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
-Formula: mean_ros ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_dnadamage_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_dnadamage_OB
-#| caption: "Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_dnadamage_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
-```
-
-Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
-Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_dnadamage_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_dnadamage_OT
-#| caption: "Posterior predictive checks for the model of DNA Damage in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_dnadamage_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
-```
-
-Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
-Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_peroxidation_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_peroxidation_OB
-#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_peroxidation_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
-```
-
-Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
-Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_peroxidation_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_peroxidation_OT
-#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_peroxidation_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
-```
-
-Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
-Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_tD_Chemical, out.width="70%", fig.align="center"}
-#| label: plotmod_tD_Chemical
-#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials."
-#
-mod <- readRDS(here("output/models/t_D_def_Chemical.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
-```
-
-Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
-Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_tD_Visual, out.width="70%", fig.align="center"}
-#| label: plotmod_tD_Visual
-#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials."
-#
-mod <- readRDS(here("output/models/t_D_def_Visual.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
-```
-
-Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
-Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
 #### Results of the multivariate models
 
-Table S16. R^2 values of the multivariate models.
-```{r, table_bayesR2_SEM}
-#| label: table_bayesR2_SEM
-#| tbl-cap: "BayesR2 values of the multivariate models"
-#
-data_bayes_SEM <- data.frame(
-  Model = character(0),
-  Response = character(0),
-  Mean = numeric(0),
-  Error = numeric(0),
-  Q2_5 = numeric(0),
-  Q97_5 = numeric(0)
-)
-#
-models <- c("m_OB", "m_OT")  
-for (m in models) {
-  mod <- readRDS(here("output/m_SEM/", paste0(m, ".rds")))
-
-  # Extract bayes_R2() for each response variable in the multivariate model
-  bayes_df_SEM <- as.data.frame(bayes_R2(mod))
-  bayes_df_SEM$Response <- rownames(bayes_df_SEM)  # Keep track of response variable names
-
-  # Add model identifier and format values
-  bayes_df_SEM <- bayes_df_SEM %>%
-    mutate(
-      Model = m,
-      Mean = format_dec(Estimate, 3),
-      Error = format_dec(`Est.Error`, 3),
-      Q2_5 = format_dec(`Q2.5`, 3),
-      Q97_5 = format_dec(`Q97.5`, 3)
-    ) %>%
-    dplyr::select(Model, Response, Mean, Error, Q2_5, Q97_5)
-
-  # Append to the main data frame
-  data_bayes_SEM <- rbind(data_bayes_SEM, bayes_df_SEM)
-}
-#
-bayes_table_SEM_df <- data_bayes_SEM %>%
-  mutate(Response = gsub("^R2", "", Response)) %>%
-  mutate(Response = factor(Response,
-                        levels = c("tD",
-                                  "meandnadamage",
-                                  "meanperoxidation",
-                                  "meanros"),
-                        labels = c("tD" = "Detection lat",
-                                  "meandnadamage" = "DNA damage",
-                                  "meanperoxidation" = "Peroxidation",
-                                  "meanros" = "ROS")),
-      Model = factor(Model, levels = c("m_OB", "m_OT"),
-                      labels = c("OB" = "Olfactory bulbs / Chemical stimulus",
-                                "OT" = "Optic tecta / Visual stimulus"))) %>%
-  dplyr::select(Model, Response, Mean, Error, Q2_5, Q97_5) %>%
-  arrange(Model, Response)
-#
-#
-# 
-## Table format
-set_flextable_defaults(
- font.family = "Times New Roman",
- font.size = 10)
-#
-bayes_table_SEM <- flextable(bayes_table_SEM_df) %>%
-  set_header_labels(
-    Region = "Region/Stimulus",
-    Mean = "Mean",
-    Error = "Error",
-    Q2_5 = "2.5%",
-    Q97_5 = "97.5%") %>%
-  align(align = "center", j = c(3:6), part = "body") %>%
-  align(align = "center", part = "header") %>%
-  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = 4, part = "body") %>% 
-  autofit()
-#
-bayes_table_SEM
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S17. Results of multivariate models for OB/Chemical stimulus
+Table S5. Compiled direct, indirect, and total effects of multivariate models for OB/Chemical stimulus
 ```{r, table_sem_results_OB}
 #| label: table_sem_results_OB
-#| fig-cap: "Structural Equation Models for OB/Chemical stimulus"
+#| fig-cap: ""
 #
@@
 
-Table S18. Results of multivariate models for OT/Visual stimulus
+Table S6. Compiled direct, indirect, and total effects  of multivariate models for OT/Visual stimulus
 ```{r, table_sem_results_OT}
 #| label: table_sem_results_OT
-#| fig-cap: "Structural Equation Models for OT/Visual stimulus"
+#| fig-cap: ""
 #
@@
 ```
+
+#### Results of preliminary models
+
+```{r, models_preliminary}
+#| label: models_preliminary
+# Fitting intial models to see if sex and age are relevant for our models
+source(here("R", "func.R"))
+#
+#
+# Run models mitochondrial physiology (each region separately)
+#
+var_m <- c("mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
+regions <- c("OB", "OT")
+formula_list_ <- list()
+for (p in var_m){
+  formula_list_[[p]] <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
+  for (h in regions){
+    if (h == "OB"){
+      df <- clean_df %>% filter(region == "OB")
+      l <- "OB"
+    } else {
+      df <- clean_df %>% filter(region == "OT")
+      l <- "OT"
+    }
+
+  pmodel_name <- paste0("m_prel_", p, "_", h)
+  assign(pmodel_name, fit_m(df = df,
+                             cat = "prel",
+                             var = p,
+                             formula = formula_list_[[p]],
+                             fam = gaussian(),
+                             label = l,
+                             refit = FALSE),
+          envir = .GlobalEnv)  # Assign to the global environment
+  }
+}
+#
+#
+# Run model behaviour (stimuli separated)
+beh_df <- clean_df
+stimulus <- c("Chemical", "Visual")
+formula_t_D <- t_D ~ motivation + motivation:cort + cort*temp + age_trial + sex + prey +(1|clutch) + (1|lizard_id) 
+for (k in stimulus){
+  df <- beh_df %>% filter(stimulus == k)
+  pmodel_name <- paste0("m_prel_t_D_", k)
+  assign(pmodel_name, fit_m(df = df,
+                            cat = "prel",
+                            var = "t_D",
+                            formula = formula_t_D,
+                            fam = gaussian(),
+                            label = k,
+                            refit = FALSE),
+        envir = .GlobalEnv)  # Assign to the global environment
+}
+```
+
+Table S7. Summary of the preliminary model for Mitochondrial Density in Olfactory Bulbs.
+```{r, results_preliminary_mitdensity_OB}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OB"
+#| label: results_preliminary_mitdensity
+#
+sum_m_mitdensity_OB_prel <- m_prel_mean_mitodensity_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws()  %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_mitdensity_OB_prel)
+```
+
+Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S8. Summary of the preliminary model for Mitochondrial Density in Optic Tecta.
+```{r, results_preliminary_mitdensity_OT}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density in OT"
+#| label: results_preliminary_mitdensity_OT
+#
+sum_m_mitdensity_OT_prel <- m_prel_mean_mitodensity_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_mitdensity_OT_prel)
+```
+
+Model formula: mean_mitodensity ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S9. Summary of the preliminary model for Mitochondrial Potential in Olfactory Bulbs.
+```{r, results_preliminary_potential_OB}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OB"
+#| label: results_preliminary_potential_OB
+#
+sum_m_potential_OB_prel <- m_prel_mean_potential_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_potential_OB_prel)
+```
+
+Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S10. Summary of the preliminary model for Mitochondrial Potential in Optic Tecta.
+```{r, results_preliminary_potential_OT}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OT"
+#| label: results_preliminary_potential_OT
+#
+sum_m_potential_OT_prel <- m_prel_mean_potential_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_potential_OT_prel)
+```
+
+Model formula: mean_potential ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S11. Summary of the preliminary model for ROS Production in Olfactory Bulbs.
+```{r, results_preliminary_ros_OB}
+#| tbl-cap: "Preliminary results of the models testing for ROS Production in OB"
+#| label: results_preliminary_ros_OB
+#
+sum_m_ros_OB_prel <- m_prel_mean_ros_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_ros_OB_prel)
+```
+
+Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S12. Summary of the preliminary model for ROS Production in Optic Tecta.
+```{r, results_preliminary_ros_OT}
+#| tbl-cap: "Preliminary results of the models testing for ROS Production in OT"
+#| label: results_preliminary_ros_OT
+#
+sum_m_ros_OT_prel <- m_prel_mean_ros_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_ros_OT_prel)
+```
+
+Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S13. Summary of the preliminary model for DNA Damage in Olfactory Bulbs.
+```{r, results_preliminary_dnadamage_OB}
+#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OB"
+#| label: results_preliminary_dnadamage_OB
+#
+sum_m_dnadamage_OB_prel <- m_prel_mean_dnadamage_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_dnadamage_OB_prel)
+```
+
+Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates an effect of sex and age, so they were included in the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S14. Summary of the preliminary model for DNA Damage in Optic Tecta.
+```{r, results_preliminary_dnadamage_OT}
+#| tbl-cap: "Preliminary results of the models testing for DNA Damage in OT"
+#| label: results_preliminary_dnadamage_OT
+#
+sum_m_dnadamage_OT_prel <- m_prel_mean_dnadamage_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_dnadamage_OT_prel)
+```
+
+Model formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of age, so it was discarded from the final models.
+However, we saw an effect of sex and we included it in our final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S15. Summary of the preliminary model for Lipid Peroxidation in Olfactory Bulbs.
+```{r, results_preliminary_peroxidation_OB}
+#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OB"
+#| label: results_preliminary_peroxidation_OB
+#
+sum_m_peroxidation_OB_prel <- m_prel_mean_peroxidation_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_peroxidation_OB_prel)
+```
+
+Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of sex, so it was discarded from the final models.
+However, we saw an effect of age and we included it in our final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S16. Summary of the preliminary model for Lipid Peroxidation in Optic Tecta.
+```{r, results_preliminary_peroxidation_OT}
+#| tbl-cap: "Preliminary results of the models testing for Lipid Peroxidation in OT"
+#| label: results_preliminary_peroxidation_OT
+#
+sum_m_peroxidation_OT_prel <- m_prel_mean_peroxidation_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_peroxidation_OT_prel)
+```
+
+Model formula: mean_peroxidation ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of sex, so it was discarded from the final models.
+However, we saw an effect of age and we included it in our final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S17. Summary of the preliminary model for Detection Latency (t_D) of Chemical stimuli. 
+```{r, results_preliminary_t_D_Chemical}
+#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Chemical stimuli"
+#| label: results_preliminary_t_D_Chemical
+# 
+sum_m_t_D_prel_Chem <- m_prel_t_D_Chemical %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_t_D_prel_Chem)
+```
+
+Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S18. Summary of the preliminary model for Detection Latency (t_D) of Visual stimuli.
+```{r, results_preliminary_t_D_Visual}
+#| tbl-cap: "Preliminary results of the models testing for Detection Latency (t_D) of Visual stimuli"
+#| label: results_preliminary_t_D_Visual
+#
+sum_m_t_D_prel_Vis <- m_prel_t_D_Visual %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_t_D_prel_Vis)
+```
+
+Model formula: t_D~ motivation + cort:motivation + cort * temp + age_trial + sex + prey + (1|clutch) + (1|lizard_id).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of age, sex, familiarity with prey (prey), motivation, or the cort-motivation interaction, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Final models diagnostics (plots)
+
+```{r , plotmod_mitdensity_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_mitdensity_OB
+#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_mitodensity_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
+```
+
+Fig S1. Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs.
+Formula: mean_mitodensity ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_mitdensity_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_mitdensity_OT
+#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_mitodensity_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
+```
+
+Fig S2. Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta.
+Formula: mean_mitodensity ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_potential_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_potential_OB
+#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_potential_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
+```
+
+Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_potential_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_potential_OT
+#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_potential_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
+```
+
+Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_ros_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_ros_OB
+#| caption: "Posterior predictive checks for the model of ROS Production in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_ros_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
+```
+
+Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_ros_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_ros_OT
+#| caption: "Posterior predictive checks for the model of ROS Production in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_ros_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
+```
+
+Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_dnadamage_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_dnadamage_OB
+#| caption: "Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_dnadamage_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
+```
+
+Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
+Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_dnadamage_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_dnadamage_OT
+#| caption: "Posterior predictive checks for the model of DNA Damage in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_dnadamage_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
+```
+
+Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
+Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_peroxidation_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_peroxidation_OB
+#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_peroxidation_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
+```
+
+Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
+Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_peroxidation_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_peroxidation_OT
+#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_peroxidation_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
+```
+
+Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
+Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_tD_Chemical, out.width="70%", fig.align="center"}
+#| label: plotmod_tD_Chemical
+#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials."
+#
+mod <- readRDS(here("output/models/t_D_def_Chemical.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
+```
+
+Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
+Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_tD_Visual, out.width="70%", fig.align="center"}
+#| label: plotmod_tD_Visual
+#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials."
+#
+mod <- readRDS(here("output/models/t_D_def_Visual.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
+```
+
+Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
+Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-03-16 R/func.r

@@ -303,6 +303,7 @@ plotting <- function(df, lab){
   geom_segment(data = db_bars, aes(y = Mean - SD, yend = Mean + SD, x = Treatment, xend = Treatment), size = 1.5, color = "black") +
   ylim(min(db_violin$Estimate), max(db_violin$Estimate)) +
   coord_flip() +
+  guides(fill = guide_legend(reverse = TRUE)) +
   theme_classic() +
   labs(y = lab, x = "Treatments") +
   theme(plot.margin = margin(3, 3, 3, 3, "mm"),

### 2025-04-01 ms.qmd
updates

@@
   m_OB <- brm(
-    bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
-    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
-    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
+    bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|p|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch)+ (1|p|obs)) +
     bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
@@
 # A.1) Detection
-OB_coeff_cort_det <- post_sem_OB$b_tD_cortCORT
-det_sem_OB_control <- post_sem_OB %>%
-  dplyr::select(-b_tD_cortCORT) %>%
-  dplyr::sample_n(size = 12000, replace = FALSE) # We take here 12000 random values for all the variables to get estimates for Control treatment, which is the reference level
-det_sem_OB_cort <- post_sem_OB %>%
-  mutate(across(everything(), ~.x + b_tD_cortCORT)) %>%
-  dplyr::sample_n(size = 12000, replace = FALSE) %>% # We take here 12000 random values for all the variables AFTER adding to b_tD_cortCORT to all columns to get estimates for CORT lizards, which is NOT the reference level
-  dplyr::select(-b_tD_cortCORT)
-det_sem_OB_controlled <- bind_rows(det_sem_OB_control, det_sem_OB_cort)
-#
-OB_coeff_mitodensity_det <- det_sem_OB_controlled$b_tD_mean_mitodensity
-OB_coeff_potential_det <- det_sem_OB_controlled$b_tD_mean_potential
-OB_coeff_dna_det <- det_sem_OB_controlled$b_tD_mean_dnadamage
-OB_coeff_perox_det <- det_sem_OB_controlled$b_tD_mean_peroxidation
+OB_coeff_mitodensity_det <- post_sem_OB$b_tD_mean_mitodensity
+OB_coeff_potential_det <- post_sem_OB$b_tD_mean_potential
+OB_coeff_dna_det <- post_sem_OB$b_tD_mean_dnadamage
+OB_coeff_perox_det <- post_sem_OB$b_tD_mean_peroxidation
 # 
 # A.2) DNA damage
-OB_coeff_age_dna <- post_sem_OB$b_meandnadamage_age_euthanasia
 OB_coeff_ros_dna <- post_sem_OB$b_meandnadamage_mean_ros
@@
 # A.3) Lipid peroxidation
-OB_coeff_age_perox <- post_sem_OB$b_meanperoxidation_age_euthanasia
 OB_coeff_ros_perox <- post_sem_OB$b_meanperoxidation_mean_ros
@@
 OB_undir_potential_det <- OB_coeff_potential_ros * OB_coeff_ros_dna * OB_coeff_dna_det + OB_coeff_potential_ros * OB_coeff_ros_perox * OB_coeff_perox_det
-OB_undir_age_det <- OB_coeff_age_dna * OB_coeff_dna_det + OB_coeff_age_perox * OB_coeff_perox_det
 #
@@
 # C.1) Detection
-OB_total_cort_det <- OB_coeff_cort_det
 OB_total_mitodensity_det <- OB_undir_mitodensity_det
@@
 OB_total_perox_det <- OB_coeff_perox_det
-OB_total_age_det <- OB_undir_age_det
 #
 # C.2) DNA damage
-OB_total_age_dna <- OB_coeff_age_dna
 OB_total_ros_dna <- OB_coeff_ros_dna
@@
 # C.3) Lipid peroxidation
-OB_total_age_perox <- OB_coeff_age_perox
 OB_total_ros_perox <- OB_coeff_ros_perox
@@
 det_results_semOB <- data.frame(
-  reference_variable = rep("t_D", 7),
-  predictor_modulator = c("cort",
-                          "age",
-                          "mean_mitodensity",
+  reference_variable = rep("t_D", 5),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
@@
                           "mean_peroxidation"),
-  direct_effects = I(list(OB_coeff_cort_det,
-                          NA,
-                          OB_coeff_mitodensity_det,
+  direct_effects = I(list(OB_coeff_mitodensity_det,
                           OB_coeff_potential_det,
@@
                           OB_coeff_perox_det)),
-  indirect_effects = I(list(NA,
-                          OB_undir_age_det,
-                          OB_undir_mitodensity_det,
+  indirect_effects = I(list(OB_undir_mitodensity_det,
                           OB_undir_potential_det,
@@
                           NA)),
-  total_effects = I(list(OB_total_cort_det,
-                        OB_total_age_det,
-                        OB_total_mitodensity_det,
+  total_effects = I(list(OB_total_mitodensity_det,
                         OB_total_potential_det,
@@
 dna_results_semOB <- data.frame(
-  reference_variable = rep("DNA damage", 4),
-  predictor_modulator = c("age",
-                          "mean_mitodensity",
+  reference_variable = rep("DNA damage", 3),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
                           "ROS"),
-  direct_effects = I(list(OB_coeff_age_dna,
-                          NA,
+  direct_effects = I(list(NA,
                           NA,
                           OB_coeff_ros_dna)),
-  indirect_effects = I(list(NA,
-                          OB_undir_mitodensity_dna,
+  indirect_effects = I(list(OB_undir_mitodensity_dna,
                           OB_undir_potential_dna,
                           NA)),
-  total_effects = I(list(OB_total_age_dna,
-                        OB_total_mitodensity_dna,
+  total_effects = I(list(OB_total_mitodensity_dna,
                         OB_total_potential_dna,
@@
 perox_results_semOB <- data.frame(
-  reference_variable = rep("lipid peroxidation", 4),
-  predictor_modulator = c("age",
-                          "mean_mitodensity",
+  reference_variable = rep("lipid peroxidation", 3),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
                           "ROS"),
-  direct_effects = I(list(OB_coeff_age_perox,
-                          NA,
+  direct_effects = I(list(NA,
                           NA,
                           OB_coeff_ros_perox)),
-  indirect_effects = I(list(NA,
-                          OB_undir_mitodensity_perox,
+  indirect_effects = I(list(OB_undir_mitodensity_perox,
                           OB_undir_potential_perox,
                           NA)),
-  total_effects = I(list(OB_total_age_perox,
-                        OB_total_mitodensity_perox,
+  total_effects = I(list(OB_total_mitodensity_perox,
                         OB_total_potential_perox,
@@
 # A) Getting all the direct coefficients
-OB_cort_det <- paste0(format_dec(mean(OB_coeff_cort_det), 3),
-                  " [", format_dec(quantile(OB_coeff_cort_det, 0.05), 3),
-                  ", ", format_dec(quantile(OB_coeff_cort_det, 0.95), 3), "]")
 OB_density_det <- paste0(format_dec(mean(OB_coeff_mitodensity_det), 3),
@@
 #
-OB_age_dna <- paste0(format_dec(mean(OB_coeff_age_dna), 3),
-                  " [", format_dec(quantile(OB_coeff_age_dna, 0.05), 3),
-                  ", ", format_dec(quantile(OB_coeff_age_dna, 0.95), 3), "]")
 OB_ros_dna <- paste0(format_dec(mean(OB_coeff_ros_dna), 3),
@@
 #
-OB_age_perox <- paste0(format_dec(mean(OB_coeff_age_perox), 3),
-                  " [", format_dec(quantile(OB_coeff_age_perox, 0.05), 3),
-                  ", ", format_dec(quantile(OB_coeff_age_perox, 0.95), 3), "]")
 OB_ros_perox <- paste0(format_dec(mean(OB_coeff_ros_perox), 3),
@@
 #
-imgOB <- readPNG(here("Others", "SEM_OB.png"))
+imgOB <- readPNG(here("Others", "SEM.png"))
 plot_SEM_OB <- rasterGrob(imgOB, interpolate = TRUE)
@@
 fig_SEM_OB <- ggdraw(plot_SEM_OB) +
-  annotate("text", x = 0.815, y = 0.59, label = OB_cort_det, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.35, y = 0.26, label = OB_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.35, y = 0.918, label = OB_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.83, y = 0.745, label = OB_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.83, y = 0.455, label = OB_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.55, y = 0.538, label = OB_age_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.525, y = 0.813, label = OB_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.57, y = 0.115, label = OB_age_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.525, y = 0.365, label = OB_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.285, y = 0.355, label = OB_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.285, y = 0.825, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
-ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
+  draw_image("./Others/OB.png", x = 0.855, y = 0.67, width = 0.18, height = 0.3) +
+  annotate("text", x = 0.52, y = 0.075, label = OB_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.52, y = 0.94, label = OB_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.83, y = 0.715, label = OB_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.83, y = 0.335, label = OB_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.525, y = 0.805, label = OB_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.515, y = 0.215, label = OB_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.282, y = 0.2, label = OB_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.282, y = 0.82, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
+ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
 knitr::include_graphics("./output/figures/text/SEM_OB.png")
@@
     bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|p|obs)) +
-    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ cort*temp + sex + mean_ros + (1|clutch) + (1|q|obs)) +
-    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ cort*temp + age_euthanasia + mean_ros + (1|clutch) + (1|q|obs)) +
+    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|q|obs)) +
+    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|q|obs)) +
     bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
@@
 # A.2) DNA damage
-#
-# Getting CORT coefficient for both temperatures controlling for each sex
-dna_sem_OT_male <- post_sem_OT %>%
-  dplyr::select(-b_meandnadamage_sexFemale)
-dna_sem_OT_fem <- post_sem_OT %>%
-  mutate(across(everything(), ~.x + b_meandnadamage_sexFemale)) %>%
-  dplyr::select(-b_meandnadamage_sexFemale) # Add effect of sex to everything to see estimates on females (males the reference level for sex)
-
-dna_sem_OT_male_trt <- dna_sem_OT_male %>%
-  dplyr::sample_n(size = 12000, replace = FALSE)
-dna_sem_OT_fem_trt <- dna_sem_OT_fem %>%
-  dplyr::sample_n(size = 12000, replace = FALSE)
-
-dna_sem_OT_trt <- bind_rows(dna_sem_OT_male_trt, dna_sem_OT_fem_trt)
-
-OT_coeff_cort_dna_cold <- dna_sem_OT_trt$b_meandnadamage_cortCORT
-OT_coeff_cort_dna_hot <- dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_cortCORT + dna_sem_OT_trt$b_meandnadamage_tempHot + dna_sem_OT_trt$`b_meandnadamage_cortCORT:tempHot` - (dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_tempHot)
-# Getting ros coefficient controlling for each level of the cort-temp interaction and controlling for each sex
-dna_sem_OT_male_controlcold <- dna_sem_OT_male %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_male_controlhot <- dna_sem_OT_male %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_male_cortcold <- dna_sem_OT_male %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_male_corthot <- dna_sem_OT_male %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
-                `b_meandnadamage_cortCORT:tempHot`)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-
-dna_sem_OT_fem_controlcold <- dna_sem_OT_fem %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_fem_controlhot <- dna_sem_OT_fem %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_fem_cortcold <- dna_sem_OT_fem %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-dna_sem_OT_fem_corthot <- dna_sem_OT_fem %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
-                `b_meandnadamage_cortCORT:tempHot`)) %>%
-  dplyr::sample_n(size = 3000, replace = FALSE)
-
-dna_sem_OT_controlled <- bind_rows(dna_sem_OT_male_cortcold,
-                                  dna_sem_OT_male_corthot,
-                                  dna_sem_OT_male_controlcold,
-                                  dna_sem_OT_male_controlhot,
-                                  dna_sem_OT_fem_cortcold,
-                                  dna_sem_OT_fem_corthot,
-                                  dna_sem_OT_fem_controlcold,
-                                  dna_sem_OT_fem_controlhot)
-
-OT_coeff_ros_dna <- dna_sem_OT_controlled$b_meandnadamage_mean_ros
-#
-# Getting sex coefficient controlling for each level of the cort-temp interaction
-dna_sem_OT_controlcold <- post_sem_OT %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-dna_sem_OT_cortcold <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-dna_sem_OT_controlhot <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-dna_sem_OT_corthot <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
-                `b_meandnadamage_cortCORT:tempHot`)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-
-dna_sem_OT_sex <- bind_rows(dna_sem_OT_cortcold,
-                            dna_sem_OT_controlcold,
-                            dna_sem_OT_corthot,
-                            dna_sem_OT_controlhot)
-
-OT_coeff_sex_dna <- dna_sem_OT_sex$b_meandnadamage_sexFemale
+OT_coeff_ros_dna <- post_sem_OT$b_meandnadamage_mean_ros
 #
 # A.3) Lipid peroxidation
-OT_coeff_cort_perox_cold <- post_sem_OT$b_meanperoxidation_cortCORT
-OT_coeff_cort_perox_hot <- post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_cortCORT + post_sem_OT$b_meanperoxidation_tempHot + post_sem_OT$`b_meanperoxidation_cortCORT:tempHot` - (post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_tempHot)
-
-# Getting the coefficient of age and ros controlling for each level of the cort-temp interaction
-perox_sem_OT_controlcold <- post_sem_OT %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-perox_sem_OT_cortcold <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-perox_sem_OT_controlhot <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_tempHot)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-perox_sem_OT_corthot <- post_sem_OT %>%
-  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT + b_meanperoxidation_tempHot +
-                `b_meanperoxidation_cortCORT:tempHot`)) %>%
-  dplyr::sample_n(size = 6000, replace = FALSE)
-
-perox_sem_OT_controlled <- bind_rows(perox_sem_OT_cortcold,
-                            perox_sem_OT_controlcold,
-                            perox_sem_OT_corthot,
-                            perox_sem_OT_controlhot)
-
-OT_coeff_ros_perox <- perox_sem_OT_controlled$b_meanperoxidation_mean_ros
-OT_coeff_age_perox <- perox_sem_OT_controlled$b_meanperoxidation_age_euthanasia
+OT_coeff_ros_perox <- post_sem_OT$b_meanperoxidation_mean_ros
 #
@@
 # B.1) Detection
-OT_undir_cort_det_COLD <- OT_coeff_cort_dna_cold * OT_coeff_dna_det + OT_coeff_cort_perox_cold * OT_coeff_perox_det
-OT_undir_cort_det_HOT <- OT_coeff_cort_dna_hot * OT_coeff_dna_det + OT_coeff_cort_perox_hot * OT_coeff_perox_det
-OT_undir_age_det <- OT_coeff_age_perox * OT_coeff_perox_det
-OT_undir_sex_det <- OT_coeff_sex_dna * OT_coeff_dna_det
 OT_undir_ros_det <- OT_coeff_ros_dna * OT_coeff_dna_det + OT_coeff_ros_perox * OT_coeff_perox_det
@@
 # C.1) Detection
-OT_total_cort_det_COLD <- OT_undir_cort_det_COLD
-OT_total_cort_det_HOT <- OT_undir_cort_det_HOT
-OT_total_age_det <- OT_undir_age_det
-OT_total_sex_det <- OT_undir_sex_det
 OT_total_mitodensity_det <- OT_undir_mitodensity_det
@@
 # C.2) DNA damage
-OT_total_cort_dna_COLD <- OT_coeff_cort_dna_cold
-OT_total_cort_dna_HOT <- OT_coeff_cort_dna_hot
-OT_total_sex_dna <- OT_coeff_sex_dna
 OT_total_ros_dna <- OT_coeff_ros_dna 
@@
 # C.3) Lipid peroxidation
-OT_total_cort_perox_COLD <- OT_coeff_cort_perox_cold
-OT_total_cort_perox_HOT <- OT_coeff_cort_perox_hot
-OT_total_age_perox <- OT_coeff_age_perox
 OT_total_ros_perox <- OT_coeff_ros_perox 
@@
 det_results_semOT <- data.frame(
-  reference_variable = rep("t_D", 9),
-  predictor_modulator = c("cort_cold",
-                          "cort_hot",
-                          "age",
-                          "sex",
-                          "mean_mitodensity",
+  reference_variable = rep("t_D", 5),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
@@
                           "mean_peroxidation"),
-  direct_effects = I(list(NA,
-                          NA,
-                          NA,
-                          NA,
-                          OT_coeff_mitodensity_det,
+  direct_effects = I(list(OT_coeff_mitodensity_det,
                           OT_coeff_potential_det,
@@
                           OT_coeff_perox_det)),
-  indirect_effects = I(list(OT_undir_cort_det_COLD,
-                          OT_undir_cort_det_HOT,
-                          OT_undir_age_det,
-                          OT_undir_sex_det,
-                          OT_undir_mitodensity_det,
+  indirect_effects = I(list(OT_undir_mitodensity_det,
                           OT_undir_potential_det,
@@
                           NA)),
-  total_effects = I(list(OT_total_cort_det_COLD,
-                        OT_total_cort_det_HOT,
-                        OT_total_age_det,
-                        OT_total_sex_det,
-                        OT_total_mitodensity_det,
+  total_effects = I(list(OT_total_mitodensity_det,
                         OT_total_potential_det,
@@
 dna_results_semOT <- data.frame(
-  reference_variable = rep("DNA damage", 6),
-  predictor_modulator = c("cort_cold",
-                          "cort_hot",
-                          "sex",
-                          "mean_mitodensity",
+  reference_variable = rep("DNA damage", 3),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
                           "ROS"),
-  direct_effects = I(list(OT_coeff_cort_dna_cold,
-                          OT_coeff_cort_dna_hot,
-                          OT_coeff_sex_dna,
-                          NA,
+  direct_effects = I(list(NA,
                           NA,
                           OT_coeff_ros_dna)),
-  indirect_effects = I(list(NA,
-                          NA,
-                          NA,
-                          OT_undir_mitodensity_dna,
+  indirect_effects = I(list(OT_undir_mitodensity_dna,
                           OT_undir_potential_dna,
                           NA)),
-  total_effects = I(list(OT_total_cort_dna_COLD,
-                        OT_total_cort_dna_HOT,
-                        OT_total_sex_dna,
-                        OT_total_mitodensity_dna,
+  total_effects = I(list(OT_total_mitodensity_dna,
                         OT_total_potential_dna,
@@
 perox_results_semOT <- data.frame(
-  reference_variable = rep("Lipid peroxidation", 6),
-  predictor_modulator = c("cort_cold",
-                          "cort_hot",
-                          "age",
-                          "mean_mitodensity",
+  reference_variable = rep("Lipid peroxidation", 3),
+  predictor_modulator = c("mean_mitodensity",
                           "mean_potential",
                           "ROS"),
-  direct_effects = I(list(OT_coeff_cort_perox_cold,
-                          OT_coeff_cort_perox_hot,
-                          OT_coeff_age_perox,
-                          NA,
+  direct_effects = I(list(NA,
                           NA,
                           OT_coeff_ros_perox)),
-  indirect_effects = I(list(NA,
-                          NA,
-                          NA,
-                          OT_undir_mitodensity_perox,
+  indirect_effects = I(list(OT_undir_mitodensity_perox,
                           OT_undir_potential_perox,
                           NA)),
-  total_effects = I(list(OT_total_cort_perox_COLD,
-                        OT_total_cort_perox_HOT,
-                        OT_total_age_perox,
-                        OT_total_mitodensity_perox,
+  total_effects = I(list(OT_total_mitodensity_perox,
                         OT_total_potential_perox,
@@
 #
-OT_cort_dna_cold <- paste0(format_dec(mean(OT_coeff_cort_dna_cold), 3),
-                      " [", format_dec(quantile(OT_coeff_cort_dna_cold, 0.05), 3),
-                      ", ", format_dec(quantile(OT_coeff_cort_dna_cold, 0.95), 3), "]")
-OT_cort_dna_hot <- paste0(format_dec(mean(OT_coeff_cort_dna_hot), 3),
-                      " [", format_dec(quantile(OT_coeff_cort_dna_hot, 0.05), 3),
-                     ", ", format_dec(quantile(OT_coeff_cort_dna_hot, 0.95), 3), "]")
-OT_sex_dna <- paste0(format_dec(mean(OT_coeff_sex_dna), 3),
-                  " [", format_dec(quantile(OT_coeff_sex_dna, 0.05), 3),
-                  ", ", format_dec(quantile(OT_coeff_sex_dna, 0.95), 3), "]")
 OT_ros_dna <- paste0(format_dec(mean(OT_coeff_ros_dna), 3),
@@
 #
-OT_cort_perox_cold <- paste0(format_dec(mean(OT_coeff_cort_perox_cold), 3),
-                          " [", format_dec(quantile(OT_coeff_cort_perox_cold, 0.05), 3),
-                          ", ", format_dec(quantile(OT_coeff_cort_perox_cold, 0.95), 3), "]")
-OT_cort_perox_hot <- paste0(format_dec(mean(OT_coeff_cort_perox_hot), 3),
-                      " [", format_dec(quantile(OT_coeff_cort_perox_hot, 0.05), 3),
-                     ", ", format_dec(quantile(OT_coeff_cort_perox_hot, 0.95), 3), "]")
-OT_age_perox <- paste0(format_dec(mean(OT_coeff_age_perox), 3),
-                  " [", format_dec(quantile(OT_coeff_age_perox, 0.05), 3),
-                  ", ", format_dec(quantile(OT_coeff_age_perox, 0.95), 3), "]")
 OT_ros_perox <- paste0(format_dec(mean(OT_coeff_ros_perox), 3),
@@
 #
-imgOT <- readPNG(here("Others", "SEM_OT.png"))
+imgOT <- readPNG(here("Others", "SEM.png"))
 plot_SEM_OT <- rasterGrob(imgOT, interpolate = TRUE)
@@
 fig_SEM_OT <- ggdraw(plot_SEM_OT) +
-  annotate("text", x = 0.35, y = 0.26, label = OT_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.35, y = 0.918, label = OT_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.83, y = 0.745, label = OT_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.83, y = 0.455, label = OT_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.58, y = 0.533, label = OT_sex_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.525, y = 0.813, label = OT_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.79, y = 0.533, label = OT_cort_dna_cold, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen", fontface = "italic") +
-  annotate("text", x = 0.79, y = 0.567, label = OT_cort_dna_hot, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.57, y = 0.115, label = OT_age_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.525, y = 0.365, label = OT_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.78, y = 0.27, label = OT_cort_perox_cold, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen", fontface = "italic") +
-  annotate("text", x = 0.78, y = 0.305, label = OT_cort_perox_hot, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
-  annotate("text", x = 0.285, y = 0.355, label = OT_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
-  annotate("text", x = 0.285, y = 0.825, label = OT_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
-ggsave(here("./output/figures/text/SEM_OT.png"), plot = fig_SEM_OT, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
+  draw_image("./Others/OT.png", x = 0.86, y = 0.67, width = 0.16, height = 0.3) +
+  annotate("text", x = 0.52, y = 0.075, label = OT_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.52, y = 0.94, label = OT_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.83, y = 0.715, label = OT_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.83, y = 0.335, label = OT_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.525, y = 0.805, label = OT_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.515, y = 0.215, label = OT_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.285, y = 0.2, label = OT_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
+  annotate("text", x = 0.285, y = 0.82, label = OT_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
+ggsave(here("./output/figures/text/SEM_OT.png"), plot = fig_SEM_OT, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
 knitr::include_graphics("./output/figures/text/SEM_OT.png")
@@
 
-#### Results of the multivariate models
-
-Table S5. Compiled direct, indirect, and total effects of multivariate models for OB/Chemical stimulus
+#### Results of the SEM
+
+Table S5. Compiled direct, indirect, and total effects of SEM for OB/Chemical stimulus
 ```{r, table_sem_results_OB}
@@
         predictor_modulator == "ROS" ~ "ROS",
-        predictor_modulator == "cort" ~ "CORT",
-        predictor_modulator == "age" ~ "Age",
         TRUE ~ predictor_modulator
@@
   mutate(Response = factor(Response, levels = c("Detection latency", "DNA damage", "Lipid peroxidation", "ROS")),
-         Predictor = factor(Predictor, levels = c("CORT",
-                                                "Age",
-                                                "Mitochondrial density",
+         Predictor = factor(Predictor, levels = c("Mitochondrial density",
                                                 "Mitochondrial potential",
@@
   align(align = "center", part = "header") %>%
-  flextable::compose(i = c(2:7, 9:11, 13:15, 17), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(7, 11, 15), part = "body") %>% 
+  flextable::compose(i = c(2:5,7,8,10,11,13), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(5,8,11), part = "body") %>% 
   autofit()
@@
 
-Table S6. Compiled direct, indirect, and total effects  of multivariate models for OT/Visual stimulus
+Table S6. Compiled direct, indirect, and total effects  of SEM for OT/Visual stimulus
 ```{r, table_sem_results_OT}
@@
         predictor_modulator == "ROS" ~ "ROS",
-        predictor_modulator == "cort" ~ "CORT",
-        predictor_modulator == "age" ~ "Age",
-        predictor_modulator == "sex" ~ "Sex",
-        predictor_modulator == "cort_cold" ~ "CORT when Cold",
-        predictor_modulator == "cort_hot" ~ "CORT when Hot",
         TRUE ~ predictor_modulator
@@
   mutate(Response = factor(Response, levels = c("Detection latency", "DNA damage", "Lipid peroxidation", "ROS")),
-         Predictor = factor(Predictor, levels = c("CORT when Cold",
-                                                "CORT when Hot",
-                                                "Age",
-                                                "Sex",
-                                                "Mitochondrial density",
+         Predictor = factor(Predictor, levels = c("Mitochondrial density",
                                                 "Mitochondrial potential",
@@
   align(align = "center", part = "header") %>%
-  flextable::compose(i = c(2:9, 11:15, 17:21, 23), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = c(9, 15, 21), part = "body") %>% 
+  flextable::compose(i = c(2:5,7,8,10,11,13), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(5,8,11), part = "body") %>% 
   autofit()

### 2025-04-01 .vdoc.r

@@ -1,2330 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-04-03 .vdoc.r
updates

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,118 +46,118 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-04 ms.qmd
updates

@@
 
-_Metabolic function_
+_Metabolic function and ROS_
 Fresh homogenate suspensions were stained with 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
-We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
+We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential - a metric of mitochondrial metabolic capacity [@martinez2016tca] - and superoxide (ROS) production, respectively.
 We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
@@
 10.1) for processing.
-We obtained the mean fluorescent intensity for mitochondrial density, mitochondrial potential, ROS production, DNA damage, and lipid peroxidation.
+We obtained the mean fluorescent intensity for mitochondrial density, metabolic capacity, ROS production, DNA damage, and lipid peroxidation.
 For further details on the homogenization, staining, or flow cytometry assays, see Methods: flow cytometry in Supplementary Material
@@
 #### Statistical analyses
-We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, mitochondrial potential, ROS, DNA damage, lipid peroxidation, and detection latency) separately.
-We first fitted a set of preliminary models for all the interest varaibles (detection latency, mitochondrial density, mitochondrial potential, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction .
+We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and detection latency) separately.
+We first fitted a set of preliminary models for all the interest varaibles (detection latency, mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction .
 Models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables).
@@
 #| label: fig-results_energy
-#| fig-cap: "Estimates of mitochondrial density (A, C) and mitochondrial potential (B, D) in the olfactory bulbs (A, B) and optic tecta (C, D) of L.
+#| fig-cap: "Estimates of mitochondrial density (A, C) and metabolic capacity (B, D) in the olfactory bulbs (A, B) and optic tecta (C, D) of L.
 delicata hatchlings as a function of the different prenatal conditions.
@@
 
-_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, mitochondrial potential, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
+_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
 However, the interaction of CORT and temperature have significant effects on DNA damage in the optic tecta (see @fig-results_oxidative).
@@
 
-_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found no significant effects of mitochondrial function or oxidative stress on prey detection.
-There were no significant effects between any other of the variables either (see @fig-sem_results_OB and Table S5).
+_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
+However, there were no significant relationships between mitochondrial function or oxidative stress on prey detection, nor between other any other variables (see @fig-sem_results_OB and Table S5).
 
@@
 delicata_ [@crino2024eggs].
-Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing efficiency [@rice2002brain].
+Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing metabolic efficiency [@rice2002brain; @martinez2016tca].
 Given the brain's high energy demands [@mcnay_decreases_2000; @mann_coupling_2021], it is plausible that mitochondria are particularly well-protected, which may help explain their resilience towards environmental inputs.
@@
 In this regard, mitochondria are predicted to impact cognitive performance by providing energy for neural activity or by regulating oxidative stress [@siegel1994basic; @mcnay_decreases_2000; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018; @mann_coupling_2021; @alexandrov_neuronal_2022].
-However, the lack of effect of mitochondrial density and mitochondrial potential on detection latency abilities in _L.
+However, the lack of effect of mitochondrial density and metabolic capacity on detection latency abilities in _L.
 delicata_ suggests that energy production is not a limiting factor for prey perception, or that _L.
@@
 Similarly, the levels of damage in the brain regions studied may not have been high enough to impair prey perception in this species or compensatory mechanisms may have counteracted any negative effects of oxidative stress.
-However, oxidative damage can accumulate over time, potentially pronouncing the effects of oxidative damage on cognition over time [@terman2006oxidative; @hara_presynaptic_2014].
+**FOR INSTANCE, ANIMALS MAY UPREGULATE ANTIOXIDANTS FUNCTION WHEN HIGHER RESPIRATION (ONDI'S COMMENT)**
+**Talk about mitochondrial density and its effect on ROS**
+Despite not seeing any effect of oxidative damage, since DNA damage or lipid peroxidation can accumulate over time, their effects on cognition can pronounce over time [@terman2006oxidative; @hara_presynaptic_2014].
 In fact, we saw a significant effect of age on DNA damage in olfactory bulbs, and on lipid peroxidation in both regions.
@@
 While mitochondrial function remained resilient to early environmental conditions, oxidative damage and behavioral responses exhibited context-dependent variations.
-Our findings suggest that the brain regions studied are not directly related to the behaviors tested, or that other mechanisms are more relevant for preceiving prey stimuli.
+Our findings suggest that the brain regions studied are not directly related to the behaviors tested, or that other factors, like motivation, are more relevant for preceiving prey stimuli.
 Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.
@@
 Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
-On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in near-boiling water.
+On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in warm water.
 We washed each thawed sample twice, the first time resuspending the pelleted cells in 1 mL warm 1X Tris-EDTA, and the second time in 200 µL warm 1X PBS.
@@
         predictor_modulator == "mean_mitodensity" ~ "Mitochondrial density",
-        predictor_modulator == "mean_potential" ~ "Mitochondrial potential",
+        predictor_modulator == "mean_potential" ~ "Metabolic capacity",
         predictor_modulator == "mean_dnadamage" ~ "DNA damage",
@@
          Predictor = factor(Predictor, levels = c("Mitochondrial density",
-                                                "Mitochondrial potential",
+                                                "Metabolic capacity",
                                                 "ROS",
@@
 
-Table S9. Summary of the preliminary model for Mitochondrial Potential in Olfactory Bulbs.
+Table S9. Summary of the preliminary model for Metabolic capacity in Olfactory Bulbs.
 ```{r, results_preliminary_potential_OB}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OB"
+#| tbl-cap: "Preliminary results of the models testing for metabolic capacity in OB"
 #| label: results_preliminary_potential_OB
@@
 
-Table S10. Summary of the preliminary model for Mitochondrial Potential in Optic Tecta.
+Table S10. Summary of the preliminary model for Metabolic capacity in Optic Tecta.
 ```{r, results_preliminary_potential_OT}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential in OT"
+#| tbl-cap: "Preliminary results of the models testing for Metabolic capacity in OT"
 #| label: results_preliminary_potential_OT
@@
 #| label: plotmod_potential_OB
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
+#| caption: "Posterior predictive checks for the model of Metabolic capacity in Olfactory Bulbs."
 #
@@
 
-Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
+Fig S3. Posterior predictive checks for the model of Metabolic capacity in Olfactory Bulbs.
 Formula: mean_potential ~ cort * temp + (1|clutch)
@@
 #| label: plotmod_potential_OT
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
+#| caption: "Posterior predictive checks for the model of Metabolic capacity in Optic Tecta."
 #
@@
 
-Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
+Fig S4. Posterior predictive checks for the model of Metabolic capacity in Optic Tecta.
 Formula: mean_potential ~ cort * temp + (1|clutch)
@@
 ```
+
+#### Brain validation
+
+To ensure that neurons were not unintentionally lost during homogenization, we performed a small pilot study with four different lizards.
+We euthanized four random lizards and prpeared the homogenates from the same brain regions as in the experiment.
+The homogenates were then dyed with a specific neuronal marker as described in Farrow et al.
+(2021) and Storks et al.
+(2023).
+We employed flow cytometry and **fluorescent optical microscopy** to check for the presence of neurons in the homogenates.
+
+Euthanasia and homogenization for each of brain region followed the procedures outlined above (see Methods: Flow cytometry).
+Before staining the samples, cells were also fixed and permeabilized as described previously (see Methods: Flow cytometry).
+Following permeabilization, the samples were centrifuged at 1000 RCF for 10 minutes and the resulting pellet was resuspended in 100 µL of a 1:100 NeuN + Alexa488 fluorescent conjugate solution to dye the neuronal nuclei (Farrow et al.
+2021; Storks et al.
+2023).
+The samples were incubated at +4 ºC overnight.
+
+The following day we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of PBS.
+Samples were split in two: 50 µL from each was reserved for examination under a **MICROSCOPE**, while the remaining 50 µL from each sample was pooled into a single tube for flow cytometry analysis.
+
+Samples examined under the microscope showed a clear presence of neuronal nuclei in the homogenates (Fig.
+S13).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r format_dec(mean(mit_df$size), 3)`, 95% CI = [`r format_dec(quantile(mit_df$size, 0.025), 3)`, `r format_dec(quantile(mit_df$size, 0.975), 3)`], n = 80; size neurons = ).
+
+```{r, fig-S13}
+#| label: fig-S13
+#| fig-cap: "Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488.
+The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells.
+In panel A, NeuN-positive cells for the homogenate of the Olfactory Bulb (OB).
+In panel B, NeuN-positive cells for the homogenate of the Optic Tecta (OT)."
+#
+#
+imgOB <- image_read(here("output/figures/microscope/OB_neurons_black.jpg"))  %>%
+  image_resize("600x400!")
+plot_OB <- rasterGrob(imgOB, interpolate = TRUE)
+#
+imgOT <- image_read(here("output/figures/microscope/OT_neurons_black.jpg")) %>%
+  image_resize("600x400!")
+plot_OT <- rasterGrob(imgOT, interpolate = TRUE)
+#
+plot_neun <- plot_grid(plot_OB, plot_OT, nrow = 1)
+#
+fig_neun <- ggdraw(plot_neun) +
+  annotate("text", x = 0.05, y = 0.92, label = "A", hjust = 1, vjust = 1, size = 4, family = "Times", color = "white") +
+  annotate("text", x = 0.55, y = 0.92, label = "B", hjust = 1, vjust = 1, size = 4, family = "Times", color = "white") 
+ggsave(here("./output/figures/suppl/Figure_SX13.png"), plot = fig_neun, width = 15, height = 5, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX13.png")
+```
+
+Figure S13. Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells.
+
+**REFERENCES**
+
+Farrow LF, Andronicos NM, McDonald PG, Hamlin AS. 2021. Quantitative determination of neuronal size and density using flow cytometry. _Journal of Neuroscience Methods_, **352**, 109081.
+
+Storks L, Powell BJ, Leal M.
+2023.
+Peeking inside the lizard brain: neuron numbers in Anolis and its implications for cognitive performance and vertebrate brain evolution.
+_Integrative and Comparative Biology_, **63(1)**, 223-237.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-04-04 .vdoc.r

@@ -1104,6 +1104,8 @@ knitr::include_graphics("./output/figures/text/SEM_OT.png")
 #
 #
 #
+#
+#
 cat("\\newpage")
 #
 #
@@ -2092,6 +2094,53 @@ knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
 #
 #
 cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-S13
+#| fig-cap: "Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells. In panel A, NeuN-positive cells for the homogenate of the Olfactory Bulb (OB). In panel B, NeuN-positive cells for the homogenate of the Optic Tecta (OT)."
+#
+#
+imgOB <- image_read(here("output/figures/microscope/OB_neurons_black.jpg"))  %>%
+  image_resize("600x400!")
+plot_OB <- rasterGrob(imgOB, interpolate = TRUE)
+#
+imgOT <- image_read(here("output/figures/microscope/OT_neurons_black.jpg")) %>%
+  image_resize("600x400!")
+plot_OT <- rasterGrob(imgOT, interpolate = TRUE)
+#
+plot_neun <- plot_grid(plot_OB, plot_OT, nrow = 1)
+#
+fig_neun <- ggdraw(plot_neun) +
+  annotate("text", x = 0.05, y = 0.92, label = "A", hjust = 1, vjust = 1, size = 4, family = "Times", color = "white") +
+  annotate("text", x = 0.55, y = 0.92, label = "B", hjust = 1, vjust = 1, size = 4, family = "Times", color = "white") 
+ggsave(here("./output/figures/suppl/Figure_SX13.png"), plot = fig_neun, width = 15, height = 5, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX13.png")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
 #
 #
 #

### 2025-04-07 ms.qmd
updates

@@
 
+```{r, neurons_size}
+#| label: neurons_size
+# Getting the size of the neurons from the validated data
+# Olfactory Bulbs
+# A) Experiment cells
+cells_exp_OB <- mit_df %>% filter(region == "OB")
+mean_cells_size_OB <- format_dec(mean(cells_exp_OB$mean_size), 3)
+min_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.025), 3)
+max_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OB <- read.csv(here("./data/size_neurons_OB.csv")) %>%
+  filter(X == "Olfactory Bulb_Tube_001_001.fcs")
+neurons_size_OB <- log(neurons_OB$neurons...Geometric.Mean..FSC.A.)
+##
+##
+# Optic Tecta
+# A) Experiment cells
+cells_exp_OT <- mit_df %>% filter(region == "OT")
+mean_cells_size_OT <- format_dec(mean(cells_exp_OT$mean_size), 3)
+min_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.025), 3)
+max_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OT <- read.csv(here("./data/size_neurons_OT.csv")) %>%
+  filter(X == "Optical Tecta_Tube_002_002.fcs")
+neurons_size_OT <- log(neurons_OT$cells.neurons...Geometric.Mean..FSC.A.)
+```
+
 Samples examined under the microscope showed a clear presence of neuronal nuclei in the homogenates (Fig.
 S13).
-Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r format_dec(mean(mit_df$size), 3)`, 95% CI = [`r format_dec(quantile(mit_df$size, 0.025), 3)`, `r format_dec(quantile(mit_df$size, 0.975), 3)`], n = 80; size neurons = ).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment for OB = `r format_dec(mean(mit_df$mean_size), 3)`, 95% CI = [`r format_dec(quantile(mit_df$mean_size, 0.025), 3)`, `r format_dec(quantile(mit_df$mean_size, 0.975), 3)`], n = 80; size neurons = ).
 

### 2025-04-07 .vdoc.r

@@ -2105,6 +2105,33 @@ cat("\\newpage")
 #
 #
 #
+#| label: neurons_size
+# Getting the size of the neurons from the validated data
+# Olfactory Bulbs
+# A) Experiment cells
+cells_exp_OB <- mit_df %>% filter(region == "OB")
+mean_cells_size_OB <- format_dec(mean(cells_exp_OB$mean_size), 3)
+min_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.025), 3)
+max_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OB <- read.csv(here("./data/size_neurons_OB.csv")) %>%
+  filter(X == "Olfactory Bulb_Tube_001_001.fcs")
+neurons_size_OB <- log(neurons_OB$neurons...Geometric.Mean..FSC.A.)
+##
+##
+# Optic Tecta
+# A) Experiment cells
+cells_exp_OT <- mit_df %>% filter(region == "OT")
+mean_cells_size_OT <- format_dec(mean(cells_exp_OT$mean_size), 3)
+min_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.025), 3)
+max_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OT <- read.csv(here("./data/size_neurons_OT.csv")) %>%
+  filter(X == "Optical Tecta_Tube_002_002.fcs")
+neurons_size_OT <- log(neurons_OT$neurons...Geometric.Mean..FSC.A.)
+#
+#
+#
 #
 #
 #| label: fig-S13

### 2025-04-11 ms.qmd
updates

--- 2025-04-04 ms.qmd
@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Daniel W.A. Noble^1^
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
 
 ^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
-^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^2^ Flinders University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
 ^3^ University of Wollongong, Wollongong, NSW 2500, Australia   
+^4^ Environmental Futures University of Wollongong, Wollongong, NSW 2500, Australia
 $\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
@@
 Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
-We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
+We injected intraperitoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
 Lizards without responses were decapitated with surgical scissors.
@@
 
-Homogenates were split among two aliquots: one was used fresh to measure mitochondrial density, membrane potential and ROS, while the other one was cryopreserved for later measurements of DNA damage and lipid peroxidation.
+Homogenates were split among two aliquots: one was used fresh to measure mitochondrial density, membrane potential - a metric of metabolic capacity [@martinez2016tca] - and ROS, while the other one was cryopreserved for later measurements of DNA damage and lipid peroxidation.
 Cryopreservation was made by suspending the homogenates in 1 mL solution of 1% Neutral-Buffered Formalin, 1X Tris-EDTA, and 10% DMSO, then stored at -20 °C until oxidative damage assays.
@@
 Fresh homogenate suspensions were stained with 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
-We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential - a metric of mitochondrial metabolic capacity [@martinez2016tca] - and superoxide (ROS) production, respectively.
-We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
+We used these fluorescent probes as indicators of mitochondrial density, metabolic capacity, and superoxide (ROS) production, respectively.
+We also added 10 µg/mL Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
 Within two hours after dissection, these samples were analyzed by flow cytometry.
@@
 On the day of oxidative damage assays, we rapidly thawed frozen samples, removed the cryopreservation solution (see above), and resuspended samples in 1 mL warm 1X Tris-EDTA.
-Afterward, the Tris-EDTA was removed, and samples were resuspended in 200 µL of warm 1X PBS containing 20 µL of #µM Hoechst 33342 Nuclear Viability Dye (for cell viability) and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor (to measure lipid peroxidation).
-Following staining, we permeabilized the cell membranes incubating the samples in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin.
-After permeabilization, we stained the samples with 20 µL of 70 µM 8-OHdG Polyclonal Antibody to measure oxidative damage on DNA.
-We left the homogenate overnight (~12 hours).
-The following day we counterstained the cells with 20 µL of #µM (100 µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyze the samples in the flow cytometer.
+Afterward, the Tris-EDTA was removed, and samples were resuspended in 200 µL of warm 1X PBS containing 10 µg/mL Hoechst 33342 Nuclear Viability Dye (for cell viability) and 100 µM BODIPY 665/676 Lipid Peroxidation Sensor (to measure lipid peroxidation).
+Following staining, we permeabilized the cell membranes incubating the samples in 200 µL warm 1X PBS containing 20 µM digitonin.
+Following permeabilization, we stained the samples with 20 µL of 70 µM 8-OHdG Polyclonal Antibody to measure oxidative damage on DNA.
+We left the homogenate overnight (~12 hours) at 4 ºC.
+The following day we counterstained the cells with 20 µL of 100 µg/mL H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyze the samples in the flow cytometer.
 
@@
 We obtained the mean fluorescent intensity for mitochondrial density, metabolic capacity, ROS production, DNA damage, and lipid peroxidation.
-For further details on the homogenization, staining, or flow cytometry assays, see Methods: flow cytometry in Supplementary Material
+For further details on the homogenization, staining, or flow cytometry assays, see Methods: flow cytometry in Supplementary Material.
+We validated that our homogenates contained neurons in a pilot study using dyes specifically targeting neuronal nuclei (See _Brain validation in Supplementary Material_).
+This pilot study also ensured that our gating strategy identified these neurons using Flow cytometry.
 
@@
 0) [@R].
-The error structure was modelled assuming a gaussian distribution for all variables.
+The error structure was modelled assuming a Gaussian distribution for all variables.
 We ran four parallel MCMC chains of 8000 iterations for each model, with a warmup period of 2000 iterations.
@@
 Clutch identity was included as a random factor for all variables.
-For our SEM model, the error structure was modelled assuming a gaussian distribution and we estimated residual correlations among all variables.
+For our SEM model, the error structure was modelled assuming a Gaussian distribution and we estimated residual correlations among all variables.
 
@@
 #### Mitochondrial function and ROS production is robust to early life conditions
-We found no significant effects of early environmental conditions on mitochondrial density, membrane potential, or ROS production in either of the brain regions studied.
+We found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
 This suggests that, contrary to our predictions, mitochondrial function is robust to prenatal temperature and CORT.
@@
 Similarly, the levels of damage in the brain regions studied may not have been high enough to impair prey perception in this species or compensatory mechanisms may have counteracted any negative effects of oxidative stress.
-**FOR INSTANCE, ANIMALS MAY UPREGULATE ANTIOXIDANTS FUNCTION WHEN HIGHER RESPIRATION (ONDI'S COMMENT)**
-**Talk about mitochondrial density and its effect on ROS**
+
 Despite not seeing any effect of oxidative damage, since DNA damage or lipid peroxidation can accumulate over time, their effects on cognition can pronounce over time [@terman2006oxidative; @hara_presynaptic_2014].
@@
 Both the breeding animals and the experimental lizards were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate levels of humidity.
-Euthanasia was performed by peritoneal injection of a 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
+Euthanasia was performed by intraperitoneal injection of a 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
 We monitored the animals to ensure there was no irritation from the agent as indicated by distressed animals.
@@
 ### Acknowledgements
-We thank Dr.
-Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
+We thank Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
 We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards.
 We also thank ANU MakerSpace, where we designed and built the prototypes of the 3D printed feeders.
+The authors acknowledge Microscopy Australia (ROR: 042mm0k03) at the Centre for Advanced Microscopy, The Australian National University, a facility enabled by NCRIS and university support.
+We specifically thank Dr Angus Rae and Dr Daryl Webb for their help with microscopy protocols.
 Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.
@@
 #### Methods: Flow cytometry
-Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
-We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
+**Brain extraction**: Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
+We injected intraperitoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
 Lizards without responses were decapitated with surgical scissors.
@@
 
-Tissue suspended in PBS was then homogenized.
+**Homogenates**: Tissue suspended in PBS was then homogenized.
 For the olfactory bulbs, the tissue was mechanically homogenized by placing the tissue in the well of a 100 µm mesh filter (pluriStrainer) affixed atop a 1.5 mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
@@
 Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500 µL 1x PBS.
-This step was to performed to remove cellular debris from homogenates.
+This step was performed to remove cellular debris from homogenates.
 From each 500 µL suspension of homogenate collected on a given trial day, we first added 100 µL of homogenate to a pooled sample of each tissue type to use for single-color controls, and the remaining 40 0µL of homogenate was split among two 200 µL aliquots.
@@
 
-From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50 µL of homogenate in duplicates (2 wells per homogenate).
+**Staining fresh samples**: From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50 µL of homogenate in duplicates (2 wells per homogenate).
 To each replicate well, we added 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
@@
 Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours.
-Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).
-
-Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (DCL personal observation).
+
+**Staining cryopreserved samples**: Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in warm water.
@@
 Thawed, resuspended samples were first stained with 20 µL (10 µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32 °C for 20 minutes.
-Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet in 200 µL warm 1X PBS containing 5 mM (6.14 mg/mL) digitonin.
+Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet in 200 µL warm 1X PBS containing 20 µM digitonin.
 We incubated the samples at 32 °C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200 µL 1X PBS.
@@
 Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate.
-Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated.
+Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugate.
 We then loaded a 96-well plate with 100 µL of each single-color control and 100 µL in duplicate of each sample.
@@
 
-All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
+**Flow cytometry**: All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
 Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v.
 8.0.1) and BD CS&T fluorescent beads (Lot No.
-30664) diluted at 2 drops to 150 µL 1X PBS.
+30664) diluted at 1 drop of 150 µL 1X PBS.
 During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector.
@@
 The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment.
-Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events.
+Voltages were chosen to center the distribution of observations in each channel at 10^3^ fluorescent intensity and reduce observations of off-scale (<10^1^ or >10^5^) events.
 For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels.
@@
 
-In FlowJo v.
-10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events.
+**Data processing**: In FlowJo v.
+10.1 we first transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (10^1^-10^5^) to remove any off-scale events.
 We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains.
-We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103).
-For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103).
-For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103).
+We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 10^3^).
+For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 10^3^) and actively respiring (PE and PerCP-Cy5-5 > 10^3^).
+For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 10^3^) and lipid peroxidation (PerCP > 10^3^).
 When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population.
@@
 
-To ensure that neurons were not unintentionally lost during homogenization, we performed a small pilot study with four different lizards.
-We euthanized four random lizards and prpeared the homogenates from the same brain regions as in the experiment.
+To ensure that neurons were not unintentionally lost during homogenization, we performed a pilot study where we euthanized four lizards and prepared the homogenates from the same brain regions as in the experiment.
 The homogenates were then dyed with a specific neuronal marker as described in Farrow et al.
@@
 (2023).
-We employed flow cytometry and **fluorescent optical microscopy** to check for the presence of neurons in the homogenates.
-
-Euthanasia and homogenization for each of brain region followed the procedures outlined above (see Methods: Flow cytometry).
-Before staining the samples, cells were also fixed and permeabilized as described previously (see Methods: Flow cytometry).
-Following permeabilization, the samples were centrifuged at 1000 RCF for 10 minutes and the resulting pellet was resuspended in 100 µL of a 1:100 NeuN + Alexa488 fluorescent conjugate solution to dye the neuronal nuclei (Farrow et al.
+We employed flow cytometry and fluorescence microscopy to check for the presence of neurons in the homogenates.
+
+We followed the procedures outlined above (see _Methods: Flow cytometry_) for euthanasia and homogenization.
+Cells were also fixed and permeabilized as described previously before staining them (see _Methods: Flow cytometry_).
+Following permeabilization, the samples were centrifuged at 1000 RCF for 10 minutes and the resulting pellet was resuspended in 100 µL of a 1:100 dilution of NeuN + Alexa488 fluorescent conjugate to dye the neuronal nuclei (Farrow et al.
 2021; Storks et al.
@@
 The following day we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of PBS.
-Samples were split in two: 50 µL from each was reserved for examination under a **MICROSCOPE**, while the remaining 50 µL from each sample was pooled into a single tube for flow cytometry analysis.
+Samples were split in two: 50 µL from each was reserved for examination under a Zeiss AxioObserver Z1 microscope, while the remaining 50 µL from each sample was pooled into a single tube for flow cytometry analysis, which allowed us to validate that our gating strategy was correct.
+
+```{r, neurons_size}
+#| label: neurons_size
+# Getting the size of the neurons from the validated data
+# Olfactory Bulbs
+# A) Experiment cells
+cells_exp_OB <- mit_df %>% filter(region == "OB")
+mean_cells_size_OB <- format_dec(mean(cells_exp_OB$mean_size), 3)
+min_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.025), 3)
+max_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OB <- read.csv(here("./data/size_neurons_OB.csv")) %>%
+  filter(X == "Olfactory Bulb_Tube_001_001.fcs")
+neurons_size_OB <- log10(neurons_OB$neurons...Geometric.Mean..FSC.A.)
+##
+##
+# Optic Tecta
+# A) Experiment cells
+cells_exp_OT <- mit_df %>% filter(region == "OT")
+mean_cells_size_OT <- format_dec(mean(cells_exp_OT$mean_size), 3)
+min_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.025), 3)
+max_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.975), 3)
+# B) Neurons validation
+neurons_OT <- read.csv(here("./data/size_neurons_OT.csv")) %>%
+  filter(X == "Optical Tecta_Tube_002_002.fcs")
+neurons_size_OT <- log10(neurons_OT$cells.neurons...Geometric.Mean..FSC.A.)
+```
 
@@
 S13).
-Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r format_dec(mean(mit_df$size), 3)`, 95% CI = [`r format_dec(quantile(mit_df$size, 0.025), 3)`, `r format_dec(quantile(mit_df$size, 0.975), 3)`], n = 80; size neurons = ).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment in OB = `r mean_cells_size_OB`, 95% CI = [`r min_cells_OB`, `r max_cells_OB`], n = 80; size neurons in OB = `r neurons_size_OB`; mean size cells experiment in OT = `r mean_cells_size_OT`, 95% CI = [`r min_cells_OT`, `r max_cells_OT`], n = 80; size neurons in OB = `r neurons_size_OT`).
 
@@
 
-Figure S13. Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells.
+Figure S13.
+Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488.
+Images were taken on a Zeiss AxioObserver Z1, equipped with Zeiss Axiocam 506 monochrome camera.
+A Zeiss 38HE fluorescent  filter set (450-490nm Ex, 500-550nm Em.) was used  in conjunction with a 63x 1.4 NA.
 

### 2025-04-11 .vdoc.r

@@ -2116,7 +2116,7 @@ max_cells_OB <- format_dec(quantile(cells_exp_OB$mean_size, 0.975), 3)
 # B) Neurons validation
 neurons_OB <- read.csv(here("./data/size_neurons_OB.csv")) %>%
   filter(X == "Olfactory Bulb_Tube_001_001.fcs")
-neurons_size_OB <- log(neurons_OB$neurons...Geometric.Mean..FSC.A.)
+neurons_size_OB <- log10(neurons_OB$neurons...Geometric.Mean..FSC.A.)
 ##
 ##
 # Optic Tecta
@@ -2128,7 +2128,7 @@ max_cells_OT <- format_dec(quantile(cells_exp_OT$mean_size, 0.975), 3)
 # B) Neurons validation
 neurons_OT <- read.csv(here("./data/size_neurons_OT.csv")) %>%
   filter(X == "Optical Tecta_Tube_002_002.fcs")
-neurons_size_OT <- log(neurons_OT$neurons...Geometric.Mean..FSC.A.)
+neurons_size_OT <- log10(neurons_OT$cells.neurons...Geometric.Mean..FSC.A.)
 #
 #
 #

### 2025-04-26 .vdoc.09edb856-0da8-4995-9b5a-653fd3a8cbde.r
updates

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.10bc6297-a8ad-409f-aa88-1cb13c4cf8da.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.12984a79-ff87-4881-b97f-a4925130febf.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.18c70ae5-acc9-4694-8b06-2fcc16bc3143.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.1b52420e-782e-4b97-98dd-416c1744e1d7.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.25903e27-93b7-488a-8d2f-63720eaa83a6.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.35d0bc13-aeb7-45a0-b5ff-b6cd899105a5.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.3f01a3f4-576e-456a-af43-98932cf9ab46.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.3f40807f-c645-43fb-a4eb-6ce8ca04abef.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.4ba2b539-11d8-40fa-99f2-65b85652f0d5.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.5429e2bf-6352-45bd-ac37-187136747f29.r

@@ -1,0 +1,2488 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: cleandata
+# Obtain the main df using "./R/1_data_process.R"
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+#
+#
+#
+#| label: countclutches
+# Count the number of clutches per species
+#
+clutches <- clean_df %>% 
+  distinct(clutch) %>% 
+  nrow()
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
+#
[…]

### 2025-04-26 .vdoc.5c4ede2f-975f-4451-8748-95acc1cc2418.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.5d90369b-a2cf-4460-a460-8780e6cc1be4.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.6b146559-d45b-4ee8-bc4b-04ba62816d7a.r

@@ -1,0 +1,2488 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: cleandata
+# Obtain the main df using "./R/1_data_process.R"
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+#
+#
+#
+#| label: countclutches
+# Count the number of clutches per species
+#
+clutches <- clean_df %>% 
+  distinct(clutch) %>% 
+  nrow()
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
+#
[…]

### 2025-04-26 .vdoc.6cfe5f65-cce8-4b70-8f4a-0e1921d04887.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.7271b7d2-cd48-4559-aaaa-5bcbaff2cf2a.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.741daf43-e78d-4706-bdab-4509965804bd.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.8663a4d6-464a-4c8d-9a54-a1645db5f633.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.87f5184e-c0f8-4127-aa48-3d9280b6a6f7.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.8b65f7b7-a8b4-4222-be00-5d56ab3a71dc.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.a244228e-02bc-44f6-b0f8-f1fc1fdceb69.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.a5851b27-6dfd-4a12-a684-cb937acdf86a.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.ce9d0a43-d463-4712-ab9f-128fd9b3ec29.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.d62a989d-5f9b-4d2d-ab85-4c8a559af425.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.d6bf1e5b-1064-4533-ac83-cad6043d6f7b.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 .vdoc.d85663b3-8e54-442f-8b29-95f2bea2c336.r

@@ -1,35 +1,35 @@
----
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
-docx: 
-reference-doc: "./bib/tmpl.docx"
-csl: "./bib/proceedings-of-the-royal-society-b.csl"
-bibliography: "./bib/refs.bib"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-crossref:  
-  fig-title: Figure    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Tab.   # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, cleandata}
+#
+#
+#
 #| label: cleandata
 # Obtain the main df using "./R/1_data_process.R"
 source(here("R", "data_process.R"))
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -46,119 +46,119 @@ for(k in 1:length(hormone)){
   }
 }
 #
-```
-
-```{r, countclutches}
+#
+#
+#
 #| label: countclutches
 # Count the number of clutches per species
 #
 clutches <- clean_df %>% 
   distinct(clutch) %>% 
   nrow()
-```
-
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
-
-^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
[…]

### 2025-04-26 R/data_process.R

@@ -76,7 +76,7 @@ df_merged <- merge(data_discr, mit_df, by = c("stimulus", "Id_flow"))
 #
 log_var <- c("t_D", "mean_mitodensity", "mean_dnadamage", "mean_peroxidation")
 df_merged_log <- df_merged %>%
-  mutate(across(all_of(log_var), ~ log(. + 1)))
+  mutate(across(all_of(log_var), ~ log10(. + 1)))
 #
 # Convert age variables to numeric and standardize all numerical variables
 clean_df <- df_merged_log %>%

### 2025-04-28 ms.qmd
updates

@@
 
+#### What did average mitochondria do?
+
+When obtaining data using FlowJo, we got the geometric mean value for each event recorded in the flow cytometer after gatting for positive values in the dyes employed (see _Methods: Flow cytometry_).
+However, this is not informative about the effects per mitochondria.
+In other words, we might be loosing some resolution about the effects at a mitochondrial level rather than a cellular level.
+To further investigate the effects per mitochondrial unit, we repeated the models above but dividing each variable by mean mitochondrial density obtained for each individual.
+
+```{r, data_mitochondrial_controlled}
+#| label: data_mitochondrial_controlled
+# 
+mit_controlled <- mit_df %>%
+    mutate(Id_flow = as.character(Id_flow)) %>%
+    mutate(mean_dnadamage = log10(mean_dnadamage + 1)) %>%
+    mutate(mean_peroxidation = log10(mean_peroxidation + 1)) %>%
+    mutate(across(where(~ is.numeric(.)), ~.x / mean_mitodensity))
+#
+df_controlled <- merge(data_discr, mit_controlled, by = c("stimulus", "Id_flow")) %>%
+  mutate(across(c(age_trial, age_euthanasia), as.numeric)) %>%
+  mutate(across(where(~ is.double(.) & !is.integer(.)),
+                ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE))))
+```
+
+```{r, models_mitochondrial_controlled}
+#| label: models_mitochondrial_controlled
+#
+# Run models mitochondrial physiology from df_controlled (each region separately)
+#
+var_m_controlled <- c("mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
+regions <- c("OB", "OT")
+for (p in var_m_controlled){
+  for (h in regions){
+    if (h == "OB"){
+      df <- df_controlled %>% filter(region == "OB")
+      l <- "OB"
+      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
+        formula <- paste0(p, "~ cort*temp + (1|clutch)")
+      } else if (p == "mean_dnadamage"){
+        formula <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
+      } else {
+        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
+      }
+    } else {
+      df <- df_controlled %>% filter(region == "OT")
+      l <- "OT"
+      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
+        formula <- paste0(p, "~ cort*temp + (1|clutch)")
+      } else if (p == "mean_dnadamage"){
+        formula <- paste0(p, "~ cort*temp + sex + (1|clutch)")
+      } else {
+        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
+      }
+    }
+
+    pmodel_name <- paste0("m_controlled_", p, "_", h)
+    assign(pmodel_name, fit_m(df = df,
+                              cat = "controlled",
+                              var = p,
+                              formula = formula,
+                              fam = gaussian(),
+                              label = l,
+                              refit = FALSE),
+          envir = .GlobalEnv)  # Assign to the global environment
+  }
+} 
+```
+
+```{r, organise_posteriors_controlled}
+#| label: organise_posteriors_controlled
+#
+# Organising the posteriors of the previous models to fit the tables below
+#
+source(here("R", "func.R"))
+#
+# Databases for each region/stimulus
+post_OB_cont <- data.frame()
+post_OT_cont <- data.frame()
+# Names posteriors:
+region <- c("OB", "OT")
+names_OB_cont <- c("m_controlled_mean_potential_OB", "m_controlled_mean_ros_OB", "m_controlled_mean_dnadamage_OB", "m_controlled_mean_peroxidation_OB")
+names_OT_cont <- c("m_controlled_mean_potential_OT", "m_controlled_mean_ros_OT", "m_controlled_mean_dnadamage_OT", "m_controlled_mean_peroxidation_OT")
+#
+# Organising the results
+for (r in region) {
+  model_select <- get(paste0("names_", r, "_cont"))   
+  for (pos in model_select) {
+    model <- get(pos)      # Get the model from the global environment
+    post_result <- tidy_post(model)        # Apply tidy_post to each model
+    # Add a new column to identify the region and model
+    post_result$Region <- r
+    post_result$Model <- pos
+    # Append to the appropriate data frame
+    if (r == "OB") {
+      post_OB_cont <- bind_rows(post_OB_cont, post_result)
+    } else {
+      post_OT_cont <- bind_rows(post_OT_cont, post_result)
+    }
+  }
+}
+```
+
+```{r, values_posteriors_controlled}
+#| label: values_posteriors_controlled
+#
+# Extracting the posteriors for the models and the values of interest.
+Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
+#
+source(here("R", "func.R"))
+#
+# A) Olfactory bulbs/Chemical stimulus
+#
+MP_OB_con <- post_values(m_controlled_mean_potential_OB, "none")
+ROS_OB_con <- post_values(m_controlled_mean_ros_OB, "none")
+DNA_OB_con <- post_values(m_controlled_mean_dnadamage_OB, "none")
+LP_OB_con <- post_values(m_controlled_mean_peroxidation_OB, "none")
+#
+# B) Optic tecta/Visual stimulus
+MP_OT_con <- post_values(m_controlled_mean_potential_OT, "none")
+ROS_OT_con <- post_values(m_controlled_mean_ros_OT, "none")
+DNA_OT_con <- post_values(m_controlled_mean_dnadamage_OT, "sex")
+LP_OT_con <- post_values(m_controlled_mean_peroxidation_OT, "none")
+```
+
+Table SXX.
+Estimate of the contrasts between treatments for all the variables analysed, and pMCMC values testing the hypothesis that contrast is different from 0.
+In bold, the values that are significant at pMCMC < 0.05.
+```{r tbl-contrasts_controlled}
+#| label: tbl-contrasts_controlled
+#| tbl-cap: "Contrasts between prenatal conditions for mitochondrial physiology and learning."
+#
+source(here("R", "func.R"))
+#
+# A) Organise df 
+var <- c("MP", "ROS", "DNA", "LP")
+region <- c("OB", "OT")
+#
+data_table_controlled <- data.frame()
+for(v in var){
+  for(r in region){
+    x <- paste0(v, "_", r, "_con")
+    df <- get(x)
+    Temperature <- format_dec(mean(c(df$CORT_Hot, df$Control_Hot)) - mean(c(df$CORT_Cold, df$Control_Cold)), 3)
+    pMCMC_temp <- format_p(pmcmc(c(df$CORT_Hot, df$Control_Hot) - c(df$CORT_Cold, df$Control_Cold)), 3, equal = FALSE)
+    CORT <- format_dec(mean(c(df$Control_Hot, df$Control_Cold)) - mean(c(df$CORT_Hot, df$CORT_Cold)), 3)
+    pMCMC_cort <- format_p(pmcmc(c(df$Control_Hot, df$Control_Cold) - c(df$CORT_Hot, df$CORT_Cold)), 3, equal = FALSE)
+    Interaction <- format_dec((mean(df$Control_Hot) - mean(df$CORT_Hot)) - (mean(df$Control_Cold) - mean(df$CORT_Cold)), 3)
+    pMCMC_int <- format_p(pmcmc((df$Control_Hot - df$CORT_Hot) - (df$Control_Cold - df$CORT_Cold)), 3, equal = FALSE)
+    data_temp <- data.frame(Variable = x,
+                          Temperature = as.numeric(Temperature),
+                          pMCMC_temp = pMCMC_temp,
+                          CORT = as.numeric(CORT),
+                          pMCMC_cort = pMCMC_cort,
+                          Interaction = as.numeric(Interaction),
+                          pMCMC_int = pMCMC_int)
+    data_table_controlled <- dplyr::bind_rows(data_table_controlled, data_temp)
+  }
+}
+# Modify the df
+data_table_controlled_final <- data_table_controlled %>%
+  pivot_longer(cols = c(Temperature, CORT, Interaction), 
+               names_to = "Predictor", 
+               values_to = "Contrast") %>%
+  mutate(
+    # Extract the pMCMC values from the corresponding columns
+    `pMCMC contrast` = case_when(
+      Predictor == "Temperature" ~ pMCMC_temp,
+      Predictor == "CORT" ~ pMCMC_cort,
+      Predictor == "Interaction" ~ pMCMC_int
+    )
+  ) %>%
+  separate(col = Variable, into = c("Variable", "Region", "control"), sep = "_") %>%
+  mutate(
+    Variable = case_when(
+      Variable == "MP" ~ "Mit potential",
+      Variable == "ROS" ~ "ROS",
+      Variable == "DNA" ~ "DNA damage",
+      Variable == "LP" ~ "Lipid peroxidation",
+      Variable == "DET" ~ "Detection latency",
+      TRUE ~ Variable
+      )
+    ) %>%
+  mutate(Variable = factor(Variable, levels = c("Mit potential", "ROS", "DNA damage", "Lipid peroxidation"))) %>%
+  mutate(
+    Region = case_when(
+      Region == "OB" ~ "Olfactory bulbs",
+      Region == "OT" ~ "Optic tecta",
+      TRUE ~ Region
+    )
+  ) %>%
+  dplyr::select(Region, Variable, Predictor, Contrast, `pMCMC contrast`) %>%
+  arrange(Region, Variable)
+#
+# C) Make the contrasts table:
+#
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+contrast_table_controlled <- flextable(data_table_controlled_final) %>%
+  align(align = "center", j = c(4,5), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>%
+  bold(~`pMCMC contrast` < 0.05, j = c("pMCMC contrast", "Contrast", "Predictor"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC contrast` <0.001, j = c("pMCMC contrast", "Contrast", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:12,14:24), j = 1, value = as_paragraph(""), part = "body") %>%
+  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24),
+                    j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
+  flextable::hline(i = c(3,6,9,12,15,18,21), j = c(2:5), part = "body") %>% 
+  flextable::hline(i = 12, part = "body") %>%
+  autofit()
+#
+contrast_table_controlled
+```
+
+Contrasts were done by:  
+_Temperature_: β~Hot~ - β~Cold~  
+_CORT_: β~CORT~ - β~Control~    
+_Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table SXX+1. Summary of the final models for Olfactory Bulbs/Chemical stimulus.
+```{r, tbl-results_OB_table_cont}
+#| label: results_OB_table_cont
+#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
+#| tbl-name: "results_OB_cont"
+#| tbl-label: "results_OB_cont"
+source(here("R", "func.R"))
+# 
+# A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB_cont)
+variable_order <- c("Mit potential",
+                     "ROS",
+                     "DNA damage",
+                     "Peroxidation")
+predictor_order <- c("b_Intercept",
+                     "b_cortCORT",
+                     "b_tempHot",
+                     "b_cortCORT:tempHot",
+                     "b_age_euthanasia",
+                     "b_sexFemale")
+post_OB_cont_refined <- post_OB_cont %>%
+    mutate(
+      Model = case_when(
+        Model == "m_controlled_mean_potential_OB" ~ "Mit potential",
+        Model == "m_controlled_mean_ros_OB" ~ "ROS",
+        Model == "m_controlled_mean_dnadamage_OB" ~ "DNA damage",
+        Model == "m_controlled_mean_peroxidation_OB" ~ "Peroxidation",
+        Model == "m_controlled_mean_potential_OT" ~ "Mit potential",
+        Model == "m_controlled_mean_ros_OT" ~ "ROS",
+        Model == "m_controlled_mean_dnadamage_OT" ~ "DNA damage",
+        Model == "m_controlled_mean_peroxidation_OT" ~ "Peroxidation",
+        TRUE ~ Model
+      )
+    ) %>%
+    mutate(CI_95 = paste0("[", CI_lower, " , ", CI_upper, "]")) %>%
+    rename(
+      Variable = Model,
+      Predictors = Predictor,
+      'Estimate Mean' = Mean, 
+      '95% CI' = CI_95, 
+      PMCMC = PMCMC
+    ) %>%
+    dplyr::select(
+      Variable,
+      Predictors,
+      'Estimate Mean',
+      '95% CI',
+      PMCMC) %>%
+    mutate(Variable = factor(Variable, levels = variable_order)) %>%
+    mutate(Predictors = factor(Predictors, levels = predictor_order)) %>%
+    arrange(Variable, Predictors)
+#
+# B) Create table
+## Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+OB_table_controlled <- flextable(post_OB_cont_refined) %>%
+  align(align = "center", j = c(3:5), part = "body") %>%
+  align(align = "center", j = c(1:5), part = "header") %>%
+  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:4,6:8,10:14,16:19), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  autofit()
+#
+OB_table_controlled
+```
+
+```{r, fig-results_OBcont}
+#| label: fig-results_OBcont
+#| fig-cap: "NEW_OB"
+#
+source(here("R", "func.R"))
+#
+# A) Plotting the results for OB/Chemical stimulus using the df from before
+plot_potential_OB_con <- plotting(MP_OB_con, "Mit potential")
+plot_ROS_OB_con <- plotting(ROS_OB_con, "ROS")
+plot_DNA_OB_con <- plotting(DNA_OB_con, "DNA damage")
+plot_LP_OB_con <- plotting(LP_OB_con, "Lipid peroxidation")
+#
+fig_OB_con <- plot_grid(plot_potential_OB_con, plot_ROS_OB_con,
+                    plot_DNA_OB_con, plot_LP_OB_con, NULL, NULL,
+                    ncol = 2, nrow = 3, rel_heights = c(1, 1, 0.2))
+#
+# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
+plot_legend_bottom_con <- plotting(MP_OB_con, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable_con <- ggplot_gtable(ggplot_build(plot_legend_bottom_con))
+legend_plot_con <- gtable_con$grobs[[which(sapply(gtable_con$grobs, function(x) x$name) == "guide-box")]]
+# 
+# Final composition: Merge everything, adding images and legend
+plot_OB_con <- ggdraw(fig_OB_con +
+  # Insert legend at the bottom-centered
+  draw_grob(legend_plot_con, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001))
+# Print final plot
+ggsave(here("./output/figures/suppl/new_models/plot_OB_con.png"), plot = plot_OB_con, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/new_models/plot_OB_con.png")
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table SXX+2. Summary of the final models for Optic Tecta/Visual stimulus.
+```{r, tbl-results_OT_table_cont}
+#| label: results_OT_table_cont
+#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
+#| tbl-name: "results_OT_cont"
+#| tbl-label: "results_OT_cont"
+source(here("R", "func.R"))
+# 
+# A) Refining the df summarizing the posteriors for OT/Visual stimulus (post_OT_cont)
+variable_order <- c("Mit potential",
+                     "ROS",
+                     "DNA damage",
+                     "Peroxidation")
+predictor_order <- c("b_Intercept",
+                     "b_cortCORT",
+                     "b_tempHot",
+                     "b_cortCORT:tempHot",
+                     "b_age_euthanasia",
+                     "b_sexFemale")
+post_OT_cont_refined <- post_OT_cont %>%
+    mutate(
+      Model = case_when(
+        Model == "m_controlled_mean_potential_OB" ~ "Mit potential",
+        Model == "m_controlled_mean_ros_OB" ~ "ROS",
+        Model == "m_controlled_mean_dnadamage_OB" ~ "DNA damage",
+        Model == "m_controlled_mean_peroxidation_OB" ~ "Peroxidation",
+        Model == "m_controlled_mean_potential_OT" ~ "Mit potential",
+        Model == "m_controlled_mean_ros_OT" ~ "ROS",
+        Model == "m_controlled_mean_dnadamage_OT" ~ "DNA damage",
+        Model == "m_controlled_mean_peroxidation_OT" ~ "Peroxidation",
+        TRUE ~ Model
+      )
+    ) %>%
+    mutate(CI_95 = paste0("[", CI_lower, " , ", CI_upper, "]")) %>%
+    rename(
+      Variable = Model,
+      Predictors = Predictor,
+      'Estimate Mean' = Mean, 
+      '95% CI' = CI_95, 
+      PMCMC = PMCMC
+    ) %>%
+    dplyr::select(
+      Variable,
+      Predictors,
+      'Estimate Mean',
+      '95% CI',
+      PMCMC) %>%
+    mutate(Variable = factor(Variable, levels = variable_order)) %>%
+    mutate(Predictors = factor(Predictors, levels = predictor_order)) %>%
+    arrange(Variable, Predictors)
+#
+# B) Create table
+## Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+OT_table_controlled <- flextable(post_OT_cont_refined) %>%
+  align(align = "center", j = c(3:5), part = "body") %>%
+  align(align = "center", j = c(1:5), part = "header") %>%
+  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:4,6:8,10:13,15:18), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  autofit()
+#
+OT_table_controlled
+```
+
+```{r, fig-results_OTcont}
+#| label: fig-results_OTcont
+#| fig-cap: "NEW_OT"
+#
+source(here("R", "func.R"))
+#
+# A) Plotting the results for OT/Visual stimulus using the df from before
+plot_potential_OT_con <- plotting(MP_OT_con, "Mit potential")
+plot_ROS_OT_con <- plotting(ROS_OT_con, "ROS")
+plot_DNA_OT_con <- plotting(DNA_OT_con, "DNA damage")
+plot_LP_OT_con <- plotting(LP_OT_con, "Lipid peroxidation")
+#
+fig_OT_con <- plot_grid(plot_potential_OT_con, plot_ROS_OT_con,
+                    plot_DNA_OT_con, plot_LP_OT_con, NULL, NULL,
+                    ncol = 2, nrow = 3, rel_heights = c(1, 1, 0.2))
+#
+# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
+plot_legend_bottom_con <- plotting(MP_OT_con, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable_con <- ggplot_gtable(ggplot_build(plot_legend_bottom_con))
+legend_plot_con <- gtable_con$grobs[[which(sapply(gtable_con$grobs, function(x) x$name) == "guide-box")]]
+# 
+# Final composition: Merge everything, adding images and legend
+plot_OT_con <- ggdraw(fig_OT_con +
+  # Insert legend at the bottom-centered
+  draw_grob(legend_plot_con, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001))
+# Print final plot
+ggsave(here("./output/figures/suppl/new_models/plot_OT_con.png"), plot = plot_OT_con, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/new_models/plot_OT_con.png")
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Brain validation

### 2025-04-28 .vdoc.575d29ab-ba89-49fd-a5b2-a330671121e4.r

@@ -1,0 +1,2609 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: cleandata
+# Obtain the main df using "./R/1_data_process.R"
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+#
+#
+#
+#| label: countclutches
+# Count the number of clutches per species
+#
+clutches <- clean_df %>% 
+  distinct(clutch) %>% 
+  nrow()
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
+#
[…]

### 2025-04-28 .vdoc.d55671d3-f564-4e5a-b6fe-a86373a95df6.r

@@ -1,0 +1,2609 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: cleandata
+# Obtain the main df using "./R/1_data_process.R"
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+#
+#
+#
+#| label: countclutches
+# Count the number of clutches per species
+#
+clutches <- clean_df %>% 
+  distinct(clutch) %>% 
+  nrow()
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
+#
[…]

### 2025-05-25 ms.qmd
Update ms.qmd

@@
 ---
-title: "The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard"
+title: "The effect of prenatal environment on brain metabolic function and perception in a lizard"
 docx: 
@@
 However, early-life conditions like stress-related hormones or thermal environments, can shape cognitive abilities by influencing brain structure and function.
-Mitochondrial physiology is a central mechanism underlying these effects.
+Mitochondrial physiology is thought to be a central mechanism underlying these effects.
 Here, we investigated the combined influence of prenatal corticosterone (CORT) — a key stress hormone in reptiles — and incubation temperature on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
-We manipulated egg CORT levels and incubation temperature, then assessed hatchlings’ ability to detect chemical and visual prey stimuli.
-Using flow cytometry, we measured mitochondrial function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta — two brain regions involved in sensory processing.
+We manipulated egg CORT levels and incubation temperature, then assessed the ability of lizards to detect chemical and visual prey stimuli.
+Using flow cytometry, we measured metabolic function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta — two brain regions involved in sensory processing.
 We predicted that CORT exposure and high temperatures would interact to shape perception of prey via changes in mitochondrial respiration and oxidative damage.
-While mitochondrial function remained robust to early conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta but not the olfactory bulbs.
+While metabolic function remained robust to early life conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta but not the olfactory bulbs.
 Additionally, CORT-treated lizards responded faster when presented with chemical cues but not visual stimuli.
-However, mitochondrial function and oxidative stress were not directly related to prey detection performance.
-These findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
+Our findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
 Future research should explore alternative mechanisms linking early-life stress, metabolic function, and cognitive performance.
@@
 
-Cognition encompasses how animals acquire, process, and store information, enabling perception, learning, memory, and decision-making [@shettleworth].
-It is essential for survival and reproduction, allowing individuals to adapt to their environment effectively [@dukas_evolutionary_2004].
+Cognition encompasses the ways in which animals acquire, process, and store information, enabling perception, learning, memory, and decision-making [@shettleworth].
+It is essential for survival and reproduction, allowing individuals to adapt to changing environments [@dukas_evolutionary_2004].
 However, cognitive abilities can vary considerably between individuals, with differences arising from genetic factors, environmental conditions, or a combination of both [@sakata_neural_2000; @dukas_evolutionary_2004].
 The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017].
-For instance,  studies have identified how prenatal conditions can influence learning abilities and memory across different taxa [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020].
-However, the impact of early-life environments on other cognitive domains has been poorly studied [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
-Perception is an often overlooked, but fundamental skill [@shettleworth].
+However, the impact of early-life environments have primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other equally important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
+In particular, perception is an often overlooked, but fundamental skill [@shettleworth].
 Perceiving different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
-Variation in the ability to detect or recognize relevant stimuli can have significant consequences for an individual’s fitness [@pyke1984optimal].
-For example, visual acuity in the bluegill sunfish (_Lepomis macrochirus_) allow animals to detect prey at larger distances, which can increase their foraging efficiency and energy intake [@hairston1982fish].
-Therefore, understanding how prenatal conditions affect the ability to detect prey can be essential to understanding the broader consequences of early-life environments on fitness.
-Yet, little is known about how prenatal conditions influence prey-related cognition [but see @burger_effects_1990].
+Variations in the ability to detect or recognize relevant stimuli can have significant consequences for an individual’s fitness [@pyke1984optimal].
+For example, visual acuity in the bluegill sunfish (_Lepomis macrochirus_) allow animals to detect prey at large distances, which can increase their foraging efficiency and energy intake [@hairston1982fish].
+Therefore, understanding how prenatal conditions affect the ability to detect prey can be important to understanding the broader consequences of early-life environments on fitness.
 
@@
 For instance, Hara et al.
-(2014) found that performance in visuospatial working-memory tasks correlated negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic terminal.
+(2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic terminal.
 These findings emphasize the role of mitochondrial action in shaping cognitive abilities.
-However, how the early environment influences mitochondrial physiology and how it may translate into cognitive consequences remains largely unknown in non-mammalian animals [see @chaudhari2022early for a review].
+However, how the early environment influences mitochondrial physiology and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
 In particular, whether and how these interactions affect perceptual abilities, such as prey detection, is still largely unexplored.
 
-Maternal conditions can play a crucial role in shaping mitochondrial function in the offspring.
+**Maternal conditions can play a crucial role in shaping mitochondrial function in the offspring.
 Because mitochondria are maternally inherited, a mother’s physiological state can directly impact mitochondria through trans-generational effects [@zhu_prenatal_2004; @song2009prenatal; @picard_mitochondria_2014].
@@
 These GCs can be transmitted to the offspring and influence their mitochondrial functioning [@picard_mitochondria_2014].
-For instance, prenatal stress or elevated yolk GCs increase oxidative stress in Sprague-Dawley rats (_Rattus norvegicus_) and domestic chickens (_Gallus domesticus_) [@zhu_prenatal_2004; @song2009prenatal; @haussmann_embryonic_2012].
-
-In ectotherms, thermal stress experienced by mothers can increase maternal circulating glucocorticoids.
+For instance, prenatal stress or elevated yolk GCs increase oxidative stress in Sprague-Dawley rats (_Rattus norvegicus_) and domestic chickens (_Gallus domesticus_) [@zhu_prenatal_2004; @song2009prenatal; @haussmann_embryonic_2012].**
+
+**In ectotherms, thermal stress experienced by mothers can increase maternal circulating glucocorticoids.
 Activation of the physiological stress response and higher GC concentrations can subsequently impact offspring [see @Crino_2023].
@@
 Therefore, the combined effects of prenatal GCs and temperature can have profound consequences for metabolic function, with significant repercussions for brain development and cognition.
-Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities by developmental programming of metabolic pathways.
-
-Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
-We manipulated eggs' CORT levels and incubation temperature, then assessed hatchlings’ prey detection ability and mitochondrial function in brain regions related to processing chemical and visual cues.
-We hypothesized that CORT would impair prey detection through mitochondrial pysiology by decreasing respiration efficiency [@gong_chronic_2011;  but see @crino2024eggs] and increasing oxidative stress [reviewed in @costantini2011meta].
-In contrast, we predicted that low incubation temperatures would only enhance beahvioural performance if the benefits of mitochondrial efficiency outweigh the costs of higher oxidative stress [@treidel2016temperature; @crino2024eggs].
+Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities by developmental programming of metabolic pathways.**
+
+Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial function and prey detection in a lizard (_Lampropholis delicata_).
+We manipulated egg CORT levels and incubation in a fully factorial design.
+We then assessed hatchling prey detection ability and quantified metabolic function in brain regions related to processing chemical and visual cues.
+We hypothesized that CORT would impair prey detection through changes in metabolic function by decreasing respiration efficiency [@gong_chronic_2011;  but see @crino2024eggs] and increasing oxidative stress [reviewed in @costantini2011meta].
+**In contrast, we predicted that low incubation temperatures would only enhance behavioural performance if the benefits of mitochondrial efficiency outweigh the costs of higher oxidative stress [@treidel2016temperature; @crino2024eggs].
 We also expected complex interactions between CORT and temperature, as they have opposing effects on mitochondrial efficiency.
-Finally, we predicted region-dependet effects of early-life conditions on mitochondrial function [@coomber_independent_1997] that would lead to stimulus-specific differences in prey detection.
+Finally, we predicted region-dependent effects of early-life conditions on mitochondrial function [@coomber_independent_1997] that would lead to stimulus-specific differences in prey detection.**
 
@@
 #### Animal husbandry  
-*Breeding colony* -- The lizards tested came from a breeding colony established in the laboratory since 2019.
-This colony consisted of 270 adults of *L.
+*Breeding colony* -- The lizards tested came from a breeding colony established in the laboratory in 2019.
+The colony consisted of 270 adults of *L.
 delicata* housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure.
 Enclosures were provided with non-stick matting, shelter, and several small water dishes.
-Water was given daily, and the lizards were fed approx.
+Water was given daily, and lizards were fed approx.
 40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
@@
 
-*Eggs collection and incubation* -- Between mid-October 2022 and the end of February 2023, we provide females with a place to lay the eggs by placing a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures.
+*Egg collection and incubation* -- Between mid-October 2022 and the end of February 2023, we provide females with a place to lay the eggs by placing a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures.
 These boxes were checked three days a week for eggs.
-After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a digital scale ± 0.001g error.
-Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.
+After collection, we measured the length and width of eggs with a digital caliper to the nearest 0.1 mm and weighed them with a digital scale (± 0.001 g).
+Eggs were treated with either CORT or a vehicle control (see CORT and temperature manipulation below) and placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+Cups were covered with cling wrap to retain moisture and left at one of two different incubation temperatures (see CORT and temperature manipulation below) until hatching.
 
@@
 We measured and weighted lizards immediately after hatching.
-We used a ruler to measure snout-vent length (SVL) and tail length (TL) to the nearest mm, and a digital scale (± 0.001g error) to measure mass.
+We used a ruler to measure snout-vent length (SVL) and tail length (TL) to the nearest mm, and a digital scale to measure mass (± 0.001 g).
 We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with nonstick matting and a small water dish until the beginning of the experiment.
@@
 C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or b) an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [@crino2024eggs].
+We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels **~3.7x higher than control eggs** [@crino2024eggs].
 Eggs were then incubated at either cold  (23 ± 3 ºC) or hot (28 ± 3 ºC) incubation temperatures in a 2x2 factorial design (Fig.
@@
 1) for acclimatization.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the middle of the arena.
-These arenas were placed in two rooms in six racks, each with its own CCTV system (device model DVR-HP210475) that allowed us to record lizards’ behaviour during the experiment (see details below).
-The number of lizards per treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatization, lizards were fed with only one cricket per day dusted with calcium and multivitamin, and water was supplied ad libitum.
+The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish in the middle of the arena.
+Arenas were placed in two rooms on six racks, each with its own CCTV system (device model DVR-HP210475) that allowed us to record lizard behaviour during the experiment (see details below).
+The number of lizards per treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard on the rack.
+During acclimatization, lizards were fed with only one cricket per day dusted with calcium and multivitamin, and water was supplied _ad libitum_.
 We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
@@
 
-The experiment involved presenting lizards with various stimuli from different prey types, then recording and analyzing their behaviour towards each stimulus.
-We followed the methods employed by @desfilis2003stimulus and @recio_prey_2021, with some modifications relevant to the objectives of this study.
-Lizards were presented chemical or visual stimuli from a familiar and unfamiliar prey, with a total of four trials per individual (@fig-Methods C).
-
+The experiment involved presenting lizards with chemical and visual stimuli from different familiar and unfamiliar prey, then recording and analyzing their behaviour towards each stimulus (@fig-Methods C).
 We used crickets (_A.
@@
 
-_Brain dissection and homogenization_
-Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
+_Brain dissection and homogenization_: Two months after the completion of the tests, we euthanized lizards using an injectable anaesthetic followed by decapitation.
 We injected intraperitoneally 10 mg/kg of a 10 mg/mL alfaxan solution and then, after several minutes, we evaluated the lizard’s righting response and pinching reflex in one of the front limbs.
 Lizards without responses were decapitated with surgical scissors.
-This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
+This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022/33).
 After decapitation, the head was opened and the brain was dissected.
-We extracted manually two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with chemical and visual perception in lizards [@wyneken2007reptilian].
+We extracted two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with chemical and visual perception in lizards [@wyneken2007reptilian].
 Both regions were transferred immediately to 1.5mL centrifuge tubes containing 100µL of 1X phosphate buffered saline (PBS).
 
-Tissue suspected in PBS was then homogenized.
+Tissue suspented in PBS was then homogenized.
 For the olfactory bulbs, the tissue was mechanically homogenized by placing the tissue in the well of a 100 µm mesh filter (pluriStrainer) affixed atop a 1.5 mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
-The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1X.
+The resulting olfactory bulb homogenate was then rinsed through the filter with 1 mL of cold 1XPBS.
 Optic tecta were mechanically homogenized the same way as olfactory bulbs, but were first enzymatically digested by incubating the tissue in 100µL of 125 U/mL collagenase (type II).
@@
 
-_Metabolic function and ROS_
-Fresh homogenate suspensions were stained with 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
+_Metabolism and ROS_: Fresh homogenate suspensions were stained with 5 µL of a fluorescent probe mix containing equal parts 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
 We used these fluorescent probes as indicators of mitochondrial density, metabolic capacity, and superoxide (ROS) production, respectively.
 We also added 10 µg/mL Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
-Within two hours after dissection, these samples were analyzed by flow cytometry.
-
-_Oxidative Damage_
-Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+Within two hours after dissection, these samples were analyzed by flow cytometry (Becton Dickson LSRFortessa X-20).
+
+_Oxidative Damage_: Assays of oxidative damage from cryopreserved samples were performed **within 6 months** of the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples, removed the cryopreservation solution (see above), and resuspended samples in 1 mL warm 1X Tris-EDTA.
@@
 
-_Flow Cytometry_
-Flow cytometry assays were performed using a Becton Dickson LSRFortessa X-20 flow cytometer with the default wavelength filters on detectors.
+_Flow Cytometry_: Flow cytometry assays were performed using a Becton Dickson LSRFortessa X-20 flow cytometer with the default wavelength filters on detectors.
 The detectors and voltage settings used in data acquisition for each were determined during pilot trials and kept consistent throughout the experiment.
@@
 We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and detection latency) separately.
-We first fitted a set of preliminary models for all the interest varaibles (detection latency, mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction .
+We first fit a set of preliminary models for all the interest variables (detection latency, mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction.
 Models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables).
@@
 
-All models were fit using using brm models from stan [@stan] in R (version 4.4.
+All models were fit using using the package _brm_ which insterfaces with Stan [@stan] in R (version 4.4.
 0) [@R].
@@
 To test for differences between treatments, we made contrasts between treatments using the posterior distribution of relevant parameters.
-Significance was assessed by calculating the probability distribution parameter estimates and contrasts to test whether were they different from zero [@endo2019introduction].
+Significance was assessed by using the posterior distributions of parameter estimates and contrasts to test whether were they different from zero [@endo2019introduction].
 We considered an effect statistically significant if pMCMC < 0.05.
 
-To further explore the relationships between mitochondrial physiology and detection latency, we used a Structural Equations Modelling (SEM) approach.
-We fitted a multivariate brm from stan [@stan] in R (version 4.4.0) [@R] for each brain region/stimulus separately.
-We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem_results_OB and @fig-sem_results_OT).
-Because experience with prey did not affect lizards’ behaviour (see Tables S11, S12), and to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
+To further explore the relationships between mitochondrial physiology and detection latency, we used a Structural Equation Modelling (SEM) approach.
+We fit a multivariate brm from stan [@stan] in R (version 4.4.0) [@R] for each brain region/stimulus separately.
+We included in the model all the variables of interest and their interactions structured following a specific set of hypotheses (see @fig-sem_results_OB and @fig-sem_results_OT).
+Because experience with prey did not affect lizards’ behaviour (see Tables S11, S12), to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
 Clutch identity was included as a random factor for all variables.
@@
 
-We obtained direct, indirect, and total effects from posterior estimates in the multivariate models.
-Direct effects represent the posterior estimates of a predictor’s effect on the response variable, while indirect effects were calculated by multiplying these effects on a mediator by the mediator’s effect on the response [@kline2005principles].
+We obtained direct, indirect, and total effects from posterior distributions of parameters estimated in the multivariate models.
+Direct effects represent the posterior estimates of a predictor’s effect on the response variable, **while indirect effects were calculated by multiplying these effects on a mediator by the mediator’s effect on the response** [@kline2005principles].
 Total effects are the sum of direct and indirect effects.
@@
 ## Results
-Final sample size for mitochondrial assays was 80 lizards from a total of `r clutches` clutches (n = 20 per treatment).
+Our final sample size for mitochondrial assays was 80 lizards from a total of `r clutches` clutches (n = 20 per treatment).
 Each lizard was subjected to 4 tests for a total of n = 320 behavioural observations (n = `r sum(is.na(clean_df$t_D))` with missing data).
@@
 
-_Chemical stimulus_: lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature, nor the interaction in chemo-detection (@fig-results_behaviour A and Table S2).
+_Chemical stimulus_: lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature or the interaction between CORT and temperature (@fig-results_behaviour A and Table S2).
 
@@
 
-_Visual stimulus_: there were no significant effects of CORT, temperature, or their interaction on the detection latency of visual stimulus (@fig-results_behaviour B and Table S2).
-
-#### How are mitochondrial physiology and oxidative stress related to prey detection?
+_Visual stimulus_: there were no significant effects of CORT, temperature, or their interaction on the detection latency of visual stimuli (@fig-results_behaviour B and Table S2).
+
+#### Do metabolic function and oxidative stress compromise prey detection?
 ```{r, model_sem_OB}
@@
 
-_Detection of chemical stimulus and its relation with mitochondrial physiology in Olfactory bulbs_: we found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
+_Chemical perception (Olfactory bulbs)_: we found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
 However, there were no significant relationships between mitochondrial function or oxidative stress on prey detection, nor between other any other variables (see @fig-sem_results_OB and Table S5).
@@
 
-_Detection of visual stimulus and its relation with mitochondrial physiology in Optic tecta_: we found no significant relationships within mitochondrial variables or between mitochondrial physiology and cognitive performance (see @fig-sem_results_OT and Table S6).
+_Visual perception (Optic tecta)_: we found no significant relationships within mitochondrial variables or between mitochondrial physiology and cognitive performance (see @fig-sem_results_OT and Table S6).
 
@@
 Before disposing of the lizard, we confirmed the absence of righting response and pinching reflex in one of the front limbs.
-All the protocols complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
+All the protocols complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022/33).
 
@@
 Lizards without responses were decapitated with surgical scissors.
-This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
+This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022/33).
 After decapitation, the head was opened and the brain was dissected.

### 2025-05-29 ms.qmd
updates

@@
 Our findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
-Future research should explore alternative mechanisms linking early-life stress, metabolic function, and cognitive performance.
+Future research should explore alternative mechanisms linking early-life stress, cell metabolism, and cognitive performance.
 
@@
 The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017].
-However, the impact of early-life environments have primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other equally important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
+However, the impact of early-life environments has primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other equally important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
 In particular, perception is an often overlooked, but fundamental skill [@shettleworth].
@@
 (2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic terminal.
-These findings emphasize the role of mitochondrial action in shaping cognitive abilities.
+In conclusion, findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
+Thus, understanding how early-life conditions affect mitochondrial physiology may provide insights into the mechanisms underlying cognitive performance.
+
+Mitochondria are highly responsive to environmental conditions during early development, with potential short-term and long-lasting consequences [reviewed in @gyllenhammer2020developmental].
+For example, environmental cues such as external conditions or hormone exposure can cause immediate disruptions to metabolic function, triggering cascading effects on oxidative stress that may persist leter in life [reviewed in @gyllenhammer2020developmental].
+Among the range of environmental factors that can shape mitochondrial physiology, some exert particularly strong and consistent effects.
+Stress-related hormones and temperature are particularly relevant for mitochondrial function due to their strong effects on metabolic processes [@sapolsky_how_2000; @stier2022experimental; @crino2024eggs].
+Under stressful situations, animals instigate adaptive physiological and behavioural adjustments that involve increases in glucocorticoids (GCs), which are also important metabolic regulators [@sapolsky_how_2000].
+Similarly, temperature plays a prominent role in regulating metabolic activity in ectotherms [@stier2022experimental; @crino2024eggs].
+In fact, GC elevations or changes in the thermal environment during early development have profound and sustained effects on mitochondrial respiration efficiency or oxidative stress [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs].
 However, how the early environment influences mitochondrial physiology and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
@@
 
-**Maternal conditions can play a crucial role in shaping mitochondrial function in the offspring.
-Because mitochondria are maternally inherited, a mother’s physiological state can directly impact mitochondria through trans-generational effects [@zhu_prenatal_2004; @song2009prenatal; @picard_mitochondria_2014].
-One key factor in this process is the action of glucocorticoids (GCs), hormones that regulate energy metabolism and stress responses [@sapolsky_how_2000].
-Under stressful situations, animals instigate adaptive physiological and behavioural adjustments that involve increases in GCs.
-These GCs can be transmitted to the offspring and influence their mitochondrial functioning [@picard_mitochondria_2014].
-For instance, prenatal stress or elevated yolk GCs increase oxidative stress in Sprague-Dawley rats (_Rattus norvegicus_) and domestic chickens (_Gallus domesticus_) [@zhu_prenatal_2004; @song2009prenatal; @haussmann_embryonic_2012].**
-
-**In ectotherms, thermal stress experienced by mothers can increase maternal circulating glucocorticoids.
-Activation of the physiological stress response and higher GC concentrations can subsequently impact offspring [see @Crino_2023].
-However, temperature is also known to directly affect offspring development beyond maternal effects, particularly during early life [@noble_developmental_2018; @crino2024eggs].
-The prenatal thermal environment can play a critical role in shaping mitochondrial function, altering energy metabolism or oxidative stress [@stier2022experimental; @crino2024eggs].
-For example, Red-eared slider turtle (_Trachemys scripta elegans_) hatchlings incubated at 26.5 ºC exhibited greater oxidative damage compared to those incubated at temperatures 0.7 to 1.2 ºC warmer [@treidel2016temperature].
-Therefore, the combined effects of prenatal GCs and temperature can have profound consequences for metabolic function, with significant repercussions for brain development and cognition.
-Yet, we have little evidence of how the interactive effects of prenatal GCs and temperature can impact cognitive abilities by developmental programming of metabolic pathways.**
-
-Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial function and prey detection in a lizard (_Lampropholis delicata_).
+Here, we investigated the combined effects of prenatal temperature and corticosterone (CORT) - the main GC in reptiles - on mitochondrial physiology and prey detection in a lizard (_Lampropholis delicata_).
 We manipulated egg CORT levels and incubation in a fully factorial design.
 We then assessed hatchling prey detection ability and quantified metabolic function in brain regions related to processing chemical and visual cues.
-We hypothesized that CORT would impair prey detection through changes in metabolic function by decreasing respiration efficiency [@gong_chronic_2011;  but see @crino2024eggs] and increasing oxidative stress [reviewed in @costantini2011meta].
-**In contrast, we predicted that low incubation temperatures would only enhance behavioural performance if the benefits of mitochondrial efficiency outweigh the costs of higher oxidative stress [@treidel2016temperature; @crino2024eggs].
-We also expected complex interactions between CORT and temperature, as they have opposing effects on mitochondrial efficiency.
-Finally, we predicted region-dependent effects of early-life conditions on mitochondrial function [@coomber_independent_1997] that would lead to stimulus-specific differences in prey detection.**
+We hypothesized that prenatal CORT exposure and the incubation thermal environment would influence metabolic function and oxidative stress in the brain [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs], with significant repercussions on cognition [@zhu_prenatal_2004; @hara_presynaptic_2014].
+We predicted that metabolic deficiencies or increases in oxidative stress and damage would result in lower cognitive abilities (i.e.
+longer time to detect prey) [@picard_energetic_2018; @alexandrov_neuronal_2022].
+Furthermore, we predicted prenatal conditions to have region-dependent effects on mitochondrial function [@coomber_independent_1997] that would lead to stimulus-specific differences in prey detection.
 
@@
 C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or b) an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels **~3.7x higher than control eggs** [@crino2024eggs].
+We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
 Eggs were then incubated at either cold  (23 ± 3 ºC) or hot (28 ± 3 ºC) incubation temperatures in a 2x2 factorial design (Fig.
@@
 In panel B, we show the experimental device used to present the stimuli in the behavioural tests.
+Here, F indicates the front of the device, and Bk the back.
 In panel C, we show the experimental setup for the prey discrimination tests.
@@
 
-_Oxidative Damage_: Assays of oxidative damage from cryopreserved samples were performed **within 6 months** of the initial processing and analysis of fresh samples.
+_Oxidative Damage_: Assays of oxidative damage from cryopreserved samples were performed **5 months** after the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples, removed the cryopreservation solution (see above), and resuspended samples in 1 mL warm 1X Tris-EDTA.
@@
 We obtained direct, indirect, and total effects from posterior distributions of parameters estimated in the multivariate models.
-Direct effects represent the posterior estimates of a predictor’s effect on the response variable, **while indirect effects were calculated by multiplying these effects on a mediator by the mediator’s effect on the response** [@kline2005principles].
+Direct effects represent the posterior estimates of a predictor’s effect on the response variable, while indirect effects were computed as the product of the direct effects along the path [@kline2005principles].
 Total effects are the sum of direct and indirect effects.
@@
 ## Discussion
-In this study, we examined the impact of prenatal temperature and CORT elevations on lizards' ability to detect different stimuli from prey together with the impacts of early conditions on mitochondrial metabolic function and oxidative stress in two brain regions associated to perception of visual and chemical stimuli.
-Our findings suggest that mitochondrial metabolic function remained largely resilient to early environmental conditions, yet oxidative damage and behavioral responses exhibited region-specific and context-dependent variations.
-Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioral performance.
-
-#### Mitochondrial function and ROS production is robust to early life conditions
-We found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
-This suggests that, contrary to our predictions, mitochondrial function is robust to prenatal temperature and CORT.
-Previous studies have shown that high incubation temperatures - but not CORT - decreased mitochondrial efficiency in the liver of _L.
-delicata_ [@crino2024eggs].
-Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing metabolic efficiency [@rice2002brain; @martinez2016tca].
+We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function was the underlying reason for any perceptual differences we observed as has been shown in mammals [@hara_presynaptic_2014; @chaudhari2022early].
+Our findings suggest that metabolic processes remained largely resilient to early environmental conditions.
+However, early-life conditions have region- and context-dependent effects on oxidative damage and prey detection.
+We observed interactive effects of prenatal CORT and incubation temperature on oxidative damage in the Optic Tecta, but no effects on perception of visual stimuli.
+In contrast, CORT decrease the time to perceived chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions.
+Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioural performance.
+
+#### Metabolic processes in multiple brain regions are robust to early life conditions
+Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
+**EXAMPLES OF OTHER EXPERIMENTS WITH TEMP AND CORT IN THE BRAIN.** Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing metabolic efficiency [@rice2002brain; @martinez2016tca].
 Given the brain's high energy demands [@mcnay_decreases_2000; @mann_coupling_2021], it is plausible that mitochondria are particularly well-protected, which may help explain their resilience towards environmental inputs.
-Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.
-
-#### The interactive effects of prenatal CORT and temperature on oxidative damage are region-dependent
+
+Alternatively, the effects of early environmental conditions on metabolic function and ROS production may be immediate rather than long-lasting.
+Mitochondria are known to be highly responsive to environmental changes and energetic demands [@gyllenhammer2020developmental].
+It is possible that the metabolic effect of the early-life conditions dissipated once all animals were exposed to the same environment after hatching.
+Nonetheless, alterations in metabolic function during earlier stages would still be expected to affect oxidative damage, as these processes are cumulative and can lead to long-term consequences [@rice2002brain; @terman2006oxidative].
+
+#### Oxidative stress brought about by experiencing early life CORT are region- and temperature- dependent
 We found that DNA damage and lipid peroxidation were affected in region-specific ways.
@@
 
-CORT exposure reduced DNA damage in optic tecta from cold-incubated lizards, but had no effect at higher incubation temperatures.
-This pattern suggests that CORT might play a protective role in specific thermal conditions.
-Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — these resources can be directed towards antioxidant activity or DNA repairing mechanisms.
-In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms.
-
-In comparison, prenatal CORT elevations increased lipid peroxidation in cold incubated lizards but reduced it in those incubated at higher temperatures.
+CORT exposure reduced DNA damage in optic tecta from cold-incubated lizards but had no effect at higher incubation temperatures.
+This pattern suggests that CORT might play a protective role under specific thermal conditions.
+Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — energy mobilized by CORT may be allocated towards antioxidant activity or DNA repairing mechanisms [**REFs**].
+In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms [**REFs**].
+**A further possibility is that, under lower metabolic activity induced by cold temperatures more portion of DNA is protected within nucleosomes/histones, decreasing the effects of CORT on oxidative damage** [**REFs**].
+
+In comparison, increased prenatal CORT resulted in higher lipid peroxidation in cold incubated lizards but reduced it in those incubated at warm temperatures.
 Although the two hormonal treatments did not significantly differ within each incubation temperature, the direction of the CORT effect was the opposite between temperatures.
-One possible reason for this pattern could be temperature-dependent changes in membrane lipid composition or fluidity, which may affect how lipids respond to oxidative stress [@zeis2019temperature].
+The opposing temperature-dependent effects of CORT on DNA damage and lipid peroxidation indicates that the pathways involved in protection or repair are target-specific.
+**EXAMPLE OF ENZYMES WORKING ON LIPID PEROXIDATION AND DNA DAMAGE AT DIFFERENT TEMPERATURES**.
+An alternative mechanism might involve temperature-dependent changes in membrane lipid composition or fluidity, which may affect how lipids respond to oxidative stress [@zeis2019temperature].
 For instance, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids [@zeis2019temperature].
-Future studies should explore how temperature-dependent effects on membrane composition influence oxidative damage susceptibility.
-
-#### Detection of chemical, but not visual, stimulus improved with prenatal CORT exposure
-Detection latency decreased with CORT exposure when animals were presented with a chemical stimulus, but not with a visual stimulus.
+In that case, temperature-induced structural differences may influence the vulnerability of different molecules to oxidative stress.
+
+#### Prenatal CORT exposure enhances chemical but not visual prey detection   
+Detection latency decreased with CORT exposure when animals were presented with chemical, but not visual, stimuli.
 This effect was not related to alterations in mitochondrial physiology in the olfactory bulbs.
-Instead, we hypothesize that it may be driven by differences in motivation.
-Although lizards were fasted for two days before testing and we specifically tested for motivation, we may have overlooked factors related to the the ecological relevance of the cues, the animals' internal state, or the interaction between both factors.
+Instead, prenatal CORT may influence other mechanisms involved in chemoreception.
+For instance, in rats (_Rattus norvegicus_), GCs like dexamethasone can enchance the ability to detect chemical stimuli by upregulating gene expression in the olfactory mucosa [@meunier2020olfactory].
+Such molecular changes could alter chemosensory sensitivity in lizards, independent of mitochondrial function.
+
+Alternatively, the observed differences in chemical perception may be driven by motivation.
 Prenatal CORT is expected to increase hunger through resource mobilization [@spencer2008post; @cossin-sevrin_effect_2022; but see @crino_corticosterone_2014-learn].
-This could have lead to increase motivation in CORT-treated lizards, which could have switched to a more active foraging mode, affecting the way they respond to chemical stimuli.
+This could have led to increased motivation in CORT-treated lizards, stimulating a switch to a more active foraging mode, affecting the way they respond to chemical stimuli.
 For example, species of European lizards (family Lacertidae) that spend more time actively looking for food tend to rely more on chemoreception [@verwaijen2007relationships].
+Although we fasted lizards and controlled for motivation in our analyses, it is possible that we did not fully capture factors related to the ecological relevance of the cues, the animals' internal state, or the interaction between the two.
 Further studies need to be conducted to understand how CORT can change foraging strategies in lizards and how this can affect their perception abilities.
@@
 #### Mitochondrial function and oxidative stress do not affect detection abilities
-Despite the relevance of mitochondrial function and oxidative damage on cognitive abilities in other taxa [@zhu_prenatal_2004; @cao2014ampk; @hara_presynaptic_2014] we found no significant relationships between mitochondrial physiology and prey detection in _L.
+Mitochondria are predicted to impact cognitive performance by providing energy for neural activity or by regulating oxidative stress [@siegel1994basic; @mcnay_decreases_2000; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018; @mann_coupling_2021; @alexandrov_neuronal_2022].
+However, despite the relevance of metabolic function and oxidative damage on cognitive abilities in other taxa [@zhu_prenatal_2004; @cao2014ampk; @hara_presynaptic_2014] they did not explain variation in prey detection in _L.
 delicata_.
-This suggest that the brain regions studied are not directly related to the behaviors tested, or that other mechanisms are more relevant for preceiving prey stimuli.
-Alternatively, mitochondrial physiology may influence detection abilities only beyond a certain threshold, which may not have been reached in our study.
-
-In this regard, mitochondria are predicted to impact cognitive performance by providing energy for neural activity or by regulating oxidative stress [@siegel1994basic; @mcnay_decreases_2000; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018; @mann_coupling_2021; @alexandrov_neuronal_2022].
-However, the lack of effect of mitochondrial density and metabolic capacity on detection latency abilities in _L.
-delicata_ suggests that energy production is not a limiting factor for prey perception, or that _L.
+Metabolic function and oxidative damage did not have substantial effects on prey detection in _L.
+delicata_.
+
+Previous studies have found that **EXAMPLE ABOUT MIT DENSITY OR SIMILAR AND PREY DETECTION IF I CAN FIND ANY BECAUSE DAN DOES NOT REALIZE THAT NOONE DOES THIS**.
+In contrast, our results suggests that metabolic function is not a limiting factor for prey detection in _L.
+delicata_, or that _L.
 delicata_ can sustain this cognitive function despite fluctuations in energy availability.
-Similarly, the levels of damage in the brain regions studied may not have been high enough to impair prey perception in this species or compensatory mechanisms may have counteracted any negative effects of oxidative stress.
-
-Despite not seeing any effect of oxidative damage, since DNA damage or lipid peroxidation can accumulate over time, their effects on cognition can pronounce over time [@terman2006oxidative; @hara_presynaptic_2014].
+
+Similarly, oxidative stress did not influence _L.
+delicata_ ability to detect visual or chemical stimuli from prey.
+In this regard, oxidative damage is expected to decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+Possibly the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms - such as antioxidant activity - that counteracted the negative effects of oxidative stress, or because the studied brain regions had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].
+Alternatively, the cognitive effects of oxidative stress may not yet be detectable at this stage.
+Oxidative damage accumulates over time, and its effects on cognition can pronounce with age [@terman2006oxidative; @hara_presynaptic_2014].
 In fact, we saw a significant effect of age on DNA damage in olfactory bulbs, and on lipid peroxidation in both regions.
-Future research should investigate the long-term consequences of mitochondrial activity on brain health and the potential mechanisms that sustain cognitive resilience in _L.
-delicata_.
+Cross-sectional studies across life stages ay help reveal how oxidative damage shapes cognitive function over time.
 
@@
 delicata_.
-While mitochondrial function remained resilient to early environmental conditions, oxidative damage and behavioral responses exhibited context-dependent variations.
-Our findings suggest that the brain regions studied are not directly related to the behaviors tested, or that other factors, like motivation, are more relevant for preceiving prey stimuli.
+While mitochondrial function remained resilient to early environmental conditions, oxidative damage and behavioural responses exhibited context-dependent variations.
+Our findings suggest that the brain regions studied are not directly related to the behaviours tested, or that other factors, like motivation, are more relevant for detecting prey stimuli.
 Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.

### 2025-05-29 .vdoc.09edb856-0da8-4995-9b5a-653fd3a8cbde.r
updates

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.0f4b8747-41dc-424a-8c49-fb093d2602f5.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.10bc6297-a8ad-409f-aa88-1cb13c4cf8da.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.12984a79-ff87-4881-b97f-a4925130febf.r

@@ -1,2633 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.1b52420e-782e-4b97-98dd-416c1744e1d7.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.1d61da24-4b05-4234-9b01-edba3a336705.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.25903e27-93b7-488a-8d2f-63720eaa83a6.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.299aa805-7d34-4afc-806b-bfe4a48de7d9.r

@@ -1,2609 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.355fe8c3-40da-4c2f-aef2-1639afc7a38a.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.35d0bc13-aeb7-45a0-b5ff-b6cd899105a5.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.386b18e1-81a5-44f4-be8e-dba59345c17d.r

@@ -1,2468 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.39373d25-ae7d-4f57-9f63-0ede9f6194cc.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.3f40807f-c645-43fb-a4eb-6ce8ca04abef.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.4471f5b0-f5d3-4cb0-be30-ad3d97c90898.r

@@ -1,2494 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.49d9b9e2-1613-4034-8306-39cbf3a19917.r

@@ -1,2632 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.4ba2b539-11d8-40fa-99f2-65b85652f0d5.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.5c4ede2f-975f-4451-8748-95acc1cc2418.r

@@ -1,2695 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.63e047f8-af0d-4983-a931-cd92ba5321ef.r

@@ -1,2594 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.675860df-e6f5-4716-92d0-79dd60877e51.r

@@ -1,2609 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.69f6ef47-79c2-4547-b841-718926e1accc.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.6b146559-d45b-4ee8-bc4b-04ba62816d7a.r

@@ -1,2488 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.7271b7d2-cd48-4559-aaaa-5bcbaff2cf2a.r

@@ -1,2641 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.739e48b8-5add-410e-aac3-ccc16cfad288.r

@@ -1,2488 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.741daf43-e78d-4706-bdab-4509965804bd.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.74d21712-f04d-4bc7-b6aa-e0b5ba98e2b4.r

@@ -1,2594 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.766e1520-cdef-438b-84c5-533a82e434a3.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.786a28cd-2fbf-4b37-85c5-6fabd58c4b05.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.7c5d7eb1-c47e-44db-936e-81470500a4bf.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.87f5184e-c0f8-4127-aa48-3d9280b6a6f7.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.8b65f7b7-a8b4-4222-be00-5d56ab3a71dc.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.8c7e65a5-fe03-4c55-ae78-f220b4f237cf.r

@@ -1,2594 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.8f776bcb-1487-4c75-9275-4dfdb5382b11.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.914dc274-0f5b-4ed6-9f37-1b2f018dd3e6.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.a0ed1649-d74e-41b1-98d0-6fc1b6f0dea7.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.b473fa43-d81a-4a05-98eb-53ab58919368.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.b7ff1875-d595-4d73-84b7-32d8573e69ca.r

@@ -1,2488 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.b9f84a63-8684-4ebe-9fe8-ea6798e3a2fa.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.baf5d702-4e99-4195-bf06-3131b4601e8e.r

@@ -1,2666 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.c152e0b0-a87c-4fbe-97da-faee7da671a4.r

@@ -1,2486 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.c57a48e3-190b-47f7-880d-db5fdddad9f9.r

@@ -1,2631 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.c735b49b-3f6a-472f-b6d2-5f8467f36e0a.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.c9b0af5f-77cd-447d-b9b9-017000c88e1f.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.cae807f0-74a0-4aed-b48c-77bb90e7c1a4.r

@@ -1,2616 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.ce9d0a43-d463-4712-ab9f-128fd9b3ec29.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d019a541-0903-4560-aefc-606bd403a56c.r

@@ -1,2467 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d3b12cef-577e-454e-823d-2d2942000ef2.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d55671d3-f564-4e5a-b6fe-a86373a95df6.r

@@ -1,2609 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d62a989d-5f9b-4d2d-ab85-4c8a559af425.r

@@ -1,2581 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d74fc33d-81e1-458c-ab82-9b76d7a5b96b.r

@@ -1,2633 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.d85663b3-8e54-442f-8b29-95f2bea2c336.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.dd681af5-d296-4e7f-81f0-9a7b0a7d9eba.r

@@ -1,2493 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.df15fbc9-2367-40d7-8f7b-6f60f55fe6d6.r

@@ -1,2593 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.e0246ecf-126e-4981-83a3-b457282cd7b6.r

@@ -1,2584 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.e3197f1d-1b91-4c7c-993d-651058b256dc.r

@@ -1,2699 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.e8be67b2-15ac-40be-a60e-e86e3cfeea83.r

@@ -1,2632 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.fd258b1d-f7e7-41a3-95ec-253fd5b68aa9.r

@@ -1,2670 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-05-29 .vdoc.r

@@ -1,2175 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: cleandata
-# Obtain the main df using "./R/1_data_process.R"
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-#
-#
-#
-#| label: countclutches
-# Count the number of clutches per species
-#
-clutches <- clean_df %>% 
-  distinct(clutch) %>% 
-  nrow()
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-cat("\\newpage")
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig-cap: "Scheme of our experimental design. In panel A, we show the different stages of our experiment and the main manipulations. In panel B, we show the experimental device used to present the stimuli in the behavioural tests. In panel C, we show the experimental setup for the prey discrimination tests. In panel D, we show the relevant times from our behavioural tests."
-#
[…]

### 2025-06-01 ms.qmd
updates

@@
 
-_Oxidative Damage_: Assays of oxidative damage from cryopreserved samples were performed **5 months** after the initial processing and analysis of fresh samples.
+_Oxidative Damage_: Assays of oxidative damage from cryopreserved samples were performed 61 weeks after the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples, removed the cryopreservation solution (see above), and resuspended samples in 1 mL warm 1X Tris-EDTA.
@@
 Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
-**EXAMPLES OF OTHER EXPERIMENTS WITH TEMP AND CORT IN THE BRAIN.** Mitochondrial potential is expected to enhance ATP production per oxygen consumed, thereby increasing metabolic efficiency [@rice2002brain; @martinez2016tca].
-Given the brain's high energy demands [@mcnay_decreases_2000; @mann_coupling_2021], it is plausible that mitochondria are particularly well-protected, which may help explain their resilience towards environmental inputs.
+In the leopard gecko (_Eublepharis macularius_), incubation temperature is an important determinant of metabolic capacity in some, but not all, brain nuclei [@coomber_independent_1997], while prenatal stress decreases mitochondrial respiration efficiency in the cortex, hippocampus, and hypotalamus of rats (_Rattus norvegicus_) [@gong_chronic_2011].
+In contrast, metabolic processes in the optic tecta and olfactory bulbs of _L.
+delicata_ appear to be robust to early-life conditions, potentially to ensure that these brain regions can perform their essential functions.
+However, since the effects of early environment can be region-dependent [@coomber_independent_1997], it is possible that other brain regions not examined here may be more sensitive to prenatal CORT and temperature.
 
@@
 This pattern suggests that CORT might play a protective role under specific thermal conditions.
-Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — energy mobilized by CORT may be allocated towards antioxidant activity or DNA repairing mechanisms [**REFs**].
-In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms [**REFs**].
-**A further possibility is that, under lower metabolic activity induced by cold temperatures more portion of DNA is protected within nucleosomes/histones, decreasing the effects of CORT on oxidative damage** [**REFs**].
+Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — energy mobilised by CORT may be allocated towards antioxidant activity or DNA repairing mechanisms [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
+In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms [see @kim2019carry].
 
@@
 Although the two hormonal treatments did not significantly differ within each incubation temperature, the direction of the CORT effect was the opposite between temperatures.
-The opposing temperature-dependent effects of CORT on DNA damage and lipid peroxidation indicates that the pathways involved in protection or repair are target-specific.
-**EXAMPLE OF ENZYMES WORKING ON LIPID PEROXIDATION AND DNA DAMAGE AT DIFFERENT TEMPERATURES**.
+These contrasting CORT–temperature effects on DNA damage and lipid peroxidation suggest that the pathways protecting or repairing each molecule may vary in their thermal sensitivity.
+For instance, antioxidant enzymes differ in their response to hibernation in the soft-shelled turtle (_Pelodiscus sinensis_), indicating that thermal sensitivity varies among antioxidant system components [@tang2021antioxidant].
 An alternative mechanism might involve temperature-dependent changes in membrane lipid composition or fluidity, which may affect how lipids respond to oxidative stress [@zeis2019temperature].
-For instance, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids [@zeis2019temperature].
-In that case, temperature-induced structural differences may influence the vulnerability of different molecules to oxidative stress.
+For example, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids [@zeis2019temperature].
+Therefore, temperature-induced structural differences may influence the vulnerability of different molecules to oxidative stress.
 
@@
 Instead, prenatal CORT may influence other mechanisms involved in chemoreception.
-For instance, in rats (_Rattus norvegicus_), GCs like dexamethasone can enchance the ability to detect chemical stimuli by upregulating gene expression in the olfactory mucosa [@meunier2020olfactory].
+For instance, in rats (_R.
+norvegicus_), GCs like dexamethasone can enchance the ability to detect chemical stimuli by upregulating gene expression in the olfactory mucosa [@meunier2020olfactory].
 Such molecular changes could alter chemosensory sensitivity in lizards, independent of mitochondrial function.
@@
 Mitochondria are predicted to impact cognitive performance by providing energy for neural activity or by regulating oxidative stress [@siegel1994basic; @mcnay_decreases_2000; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018; @mann_coupling_2021; @alexandrov_neuronal_2022].
-However, despite the relevance of metabolic function and oxidative damage on cognitive abilities in other taxa [@zhu_prenatal_2004; @cao2014ampk; @hara_presynaptic_2014] they did not explain variation in prey detection in _L.
+For example, reduced cytochrome c oxidase (COX) activity in dopaminergic neurons of the olfactory bulb of mice (_Mus musculus_) is associated with impaired olfactory capacities [@pass2020impact], and links between brain metabolic function and visual acuity have also been reported in other species [reviewed in @wong2010energy].
+However, metabolic function did not explain variation in prey detection in _L.
 delicata_.
-Metabolic function and oxidative damage did not have substantial effects on prey detection in _L.
-delicata_.
-
-Previous studies have found that **EXAMPLE ABOUT MIT DENSITY OR SIMILAR AND PREY DETECTION IF I CAN FIND ANY BECAUSE DAN DOES NOT REALIZE THAT NOONE DOES THIS**.
 In contrast, our results suggests that metabolic function is not a limiting factor for prey detection in _L.
@@
 In this regard, oxidative damage is expected to decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
-Possibly the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms - such as antioxidant activity - that counteracted the negative effects of oxidative stress, or because the studied brain regions had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].
+Possibly, the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms - such as antioxidant activity - that counteracted the negative effects of oxidative stress, or because the studied brain regions had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].
 Alternatively, the cognitive effects of oxidative stress may not yet be detectable at this stage.
 Oxidative damage accumulates over time, and its effects on cognition can pronounce with age [@terman2006oxidative; @hara_presynaptic_2014].
+As such, we could find a strong relationship if older individuals were employed, but not in lizards this age.
 In fact, we saw a significant effect of age on DNA damage in olfactory bulbs, and on lipid peroxidation in both regions.
-Cross-sectional studies across life stages ay help reveal how oxidative damage shapes cognitive function over time.
+Cross-sectional studies across life stages may help reveal how oxidative damage shapes cognitive function over time.
 
@@
 
-**Staining cryopreserved samples**: Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+**Staining cryopreserved samples**: Assays of oxidative damage from cryopreserved samples were performed 61 weeks after the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in warm water.

### 2025-06-02 ms.qmd
updates

@@
 docx: 
-reference-doc: "./bib/tmpl.docx"
 csl: "./bib/proceedings-of-the-royal-society-b.csl"
 bibliography: "./bib/refs.bib"
+reference-doc: "./bib/tmpl.docx"
 execute:
@@
   warning: false
+link-citations: true
 crossref:  
-  fig-title: Figure    # (default is "Figure")
+  fig-title: Fig    # (default is "Figure")
   title-delim: —     # (default is ":")

### 2025-07-30 ms.qmd
Update ms.qmd

@@
 
-#### Does prenatal CORT and temperature affect prey detection?
+#### Does prenatal CORT and incubation temperature affect prey detection?
 Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).
@@
 We observed interactive effects of prenatal CORT and incubation temperature on oxidative damage in the Optic Tecta, but no effects on perception of visual stimuli.
-In contrast, CORT decrease the time to perceived chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions.
+In contrast, CORT decreased the time to perceive chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions.
 Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioural performance.

### 2025-09-01 R/data_process.R
updates

@@ -78,9 +78,11 @@ log_var <- c("t_D", "mean_mitodensity", "mean_dnadamage", "mean_peroxidation")
 df_merged_log <- df_merged %>%
   mutate(across(all_of(log_var), ~ log10(. + 1)))
 #
-# Convert age variables to numeric and standardize all numerical variables
+# Convert age variables to numeric, control mit potential and ROS by mitocondrial, and standardize all numerical variables
 clean_df <- df_merged_log %>%
   mutate(across(c(age_trial, age_euthanasia), as.numeric)) %>%
+  mutate(mean_conpotential = as.numeric(mean_potential / mean_mitodensity),
+         mean_conros = as.numeric(mean_ros / mean_mitodensity)) %>%
   mutate(across(where(~ is.double(.) & !is.integer(.)),
                 ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE))))
 

### 2025-09-01 ms.qmd
updates

@@
 #
-var_m <- c("mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
+var_m <- c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros", "mean_dnadamage", "mean_peroxidation")
 regions <- c("OB", "OT")
@@
 
+**____________________FROM HERE____________________**
 ```{r, organise_posteriors}

### 2025-09-06 ms.qmd
updates

@@
 
+We recorded five physiological variables from our flow cytometry assay: mitochondrial density, metabolic capacity, ROS production, DNA damage, and lipid peroxidation.
+These variables reflect average per-cell values for each sample, as fluorescence intensity was normalized by the number of gated events (cells) recorded.
+However, to assess how early environmental conditions may influence mitochondrial function at the organelle level, we created two additional response variables by dividing metabolic capacity and ROS production by mitochondrial density.
+This procedure assumes that mitochondrial density serves as a proxy for the number of mitochondria per cell, and the derived variables provide an estimate of the average activity per mitochondrion, therefore allowing us to assess mitochondrial-specific responses to early-life conditions.
+
 #### Statistical analyses
-We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and detection latency) separately.
-We first fit a set of preliminary models for all the interest variables (detection latency, mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation) where we included the main effects of hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction.
-Models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables).
+We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, metabolic capatity/mitochondrial density, ROS, ROS/mitochondrial density, DNA damage, lipid peroxidation, and detection latency) separately.
+We first fit a set of preliminary models for all the interest variables where we included the hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction.
+These models also included the sex and age of the lizards at the time where the trials started (for detection latency) or when the lizards were euthanized (for all mitochondrial-related variables).
 For detection latency, we also included previous experience with the prey (familiar versus unknown), their performance on the motivation test [if they ate the cricket (1), or not (0)], and the interaction between CORT and motivation as fixed effects.
@@
 For models of detection latency we also included a random effect of lizard identity because of our repeated measures design.
-The structure and results of these models are provided on Tables S1-S12 of the Supplementary Material.
-After the preliminary models, we simplified models by re-fitting to exclude factors that were not significant.
-However, our models always included the main interest factors: CORT, temperature and their interaction.
+The structure and results of these models are provided on Tables S1-S22 in _Supplementary Material_.
+After the preliminary models, we simplified models by excluding some of the factors that were not significant.
+However, we always included the main interest factors: CORT, temperature and their interaction.
 Random factors remain the same for the final models.
@@
 We included in the model all the variables of interest and their interactions structured following a specific set of hypotheses (see @fig-sem_results_OB and @fig-sem_results_OT).
-Because experience with prey did not affect lizards’ behaviour (see Tables S11, S12), to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
+Because experience with prey did not affect lizards’ behaviour (see Tables S21, S22), to reduce the complexity of the models, we averaged lizards detection latency across both types of prey and excluded lizard identity from the random factors.
 Clutch identity was included as a random factor for all variables.
@@
       l <- "OB"
-      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
+      if (p %in% c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros")){
         formula <- paste0(p, "~ cort*temp + (1|clutch)")
@@
       l <- "OT"
-      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
+      if (p %in% c("mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros")){
         formula <- paste0(p, "~ cort*temp + (1|clutch)")
@@
 
-**____________________FROM HERE____________________**
 ```{r, organise_posteriors}
@@
 region <- c("OB", "OT")
-names_OB <- c("m_def_mean_mitodensity_OB", "m_def_mean_potential_OB", "m_def_mean_ros_OB", "m_def_mean_dnadamage_OB", "m_def_mean_peroxidation_OB", "m_def_t_D_Chemical")
-names_OT <- c("m_def_mean_mitodensity_OT", "m_def_mean_potential_OT", "m_def_mean_ros_OT", "m_def_mean_dnadamage_OT", "m_def_mean_peroxidation_OT", "m_def_t_D_Visual")
+names_OB <- c("m_def_mean_mitodensity_OB", "m_def_mean_potential_OB", "m_def_mean_conpotential_OB", "m_def_mean_ros_OB", "m_def_mean_conros_OB", "m_def_mean_dnadamage_OB", "m_def_mean_peroxidation_OB", "m_def_t_D_Chemical")
+names_OT <- c("m_def_mean_mitodensity_OT", "m_def_mean_potential_OT", "m_def_mean_conpotential_OT", "m_def_mean_ros_OT", "m_def_mean_conros_OT", "m_def_mean_dnadamage_OT", "m_def_mean_peroxidation_OT", "m_def_t_D_Visual")
 #
@@
 MP_OB <- post_values(m_def_mean_potential_OB, "none")
+MPcon_OB <- post_values(m_def_mean_conpotential_OB, "none")
 ROS_OB <- post_values(m_def_mean_ros_OB, "none")
+ROScon_OB <- post_values(m_def_mean_conros_OB, "none")
 DNA_OB <- post_values(m_def_mean_dnadamage_OB, "none")
@@
 MP_OT <- post_values(m_def_mean_potential_OT, "none")
+MPcon_OT <- post_values(m_def_mean_conpotential_OT, "none")
 ROS_OT <- post_values(m_def_mean_ros_OT, "none")
+ROScon_OT <- post_values(m_def_mean_conros_OT, "none")
 DNA_OT <- post_values(m_def_mean_dnadamage_OT, "sex")
@@
 
-#### Does prenatal CORT and temperature affect mitochondrial function and oxidative stress in the brain?
+#### Do prenatal CORT and temperature affect mitochondrial function and oxidative stress in the brain?
 Models explained between 37.8 to 50.4 % of the variation in mitochondrial function and 39.9 to 59 % for oxidative stress (see Table S1).
@@
 _Olfactory bulbs_: age significantly increased DNA damage and lipid peroxidation (Table S3).
-However, we did not find any significant effects of CORT, temperature, or their interaction on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2).
+However, we did not find any significant effects of CORT, temperature on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2).
+The posterior distribution provided little support for an overall CORT-temperature interaction on lipid peroxidation ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OB$Control_Hot) - mean(LP_OB$CORT_Hot)) - (mean(LP_OB$Control_Cold) - mean(LP_OB$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OB$Control_Hot - LP_OB$CORT_Hot) - (LP_OB$Control_Cold - LP_OB$CORT_Cold)), 3, equal = TRUE)`).
+However, targeted contrasts showed that lipid peroxidation was lower in CORT-Hot compared to CORT-Cold lizards (β~CORT-Hot~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OB$CORT_Hot) - mean(LP_OB$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OB$CORT_Hot - LP_OB$CORT_Cold), 3, equal = TRUE)`).
 
@@
 plot_potential_OB <- plotting(MP_OB, "Mit potential")
-#
-fig_OB_energy <- plot_grid(plot_density_OB, NULL, plot_potential_OB, NULL,
-                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
+plot_conpotential_OB <- plotting(MPcon_OB, "Mit potential/Mit density")
+#
+fig_OB_energy <- plot_grid(plot_density_OB, NULL, 
+                          plot_potential_OB, plot_conpotential_OB,
+                          nrow = 2, ncol = 2, rel_widths = c(1))
 #
@@
 plot_potential_OT <- plotting(MP_OT, "Mit potential")
-#
-fig_OT_energy <- plot_grid(plot_density_OT, NULL, plot_potential_OT, NULL,
-                    nrow = 1, rel_widths = c(0.9, 0.1, 0.9, 0.3))
+plot_conpotential_OT <- plotting(MPcon_OT, "Mit potential/Mit density")
+#
+fig_OT_energy <- plot_grid(plot_density_OB, NULL, 
+                          plot_potential_OB, plot_conpotential_OB,
+                          nrow = 2, ncol = 2, rel_widths = c(1))
 #
@@
 fig_results_energy <- plot_grid(
-  fig_OB_energy, fig_OT_energy, NULL,
-  nrow = 3, rel_heights = c(0.9, 0.9, 0.15))
+  fig_OB_energy, fig_OT_energy, legend_plot,
+  nrow = 3, rel_heights = c(1, 1, 0.1))
 # Final composition: Merge everything, adding images and legend
@@
   # Insert OB image in the top right
-  draw_image("./Others/OB.png", x = 0.87, y = 0.79, width = 0.12, height = 0.2) +
+  draw_image("./Others/OB.png", x = 0.75, y = 0.75, width = 0.21, height = 0.3) +
   # Insert OT image in the middle right
-  draw_image("./Others/OT.png", x = 0.87, y = 0.34, width = 0.11, height = 0.18) +
-  # Insert legend at the bottom-right
-  draw_grob(legend_plot, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001) +
+  draw_image("./Others/OT.png", x = 0.75, y = 0.28, width = 0.19, height = 0.3) +
   # Insert title for each plot
-  annotate("text", x = 0.032, y = 0.97, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.487, y = 0.97, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.032, y = 0.515, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.49, y = 0.515, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold")
+  annotate("text", x = 0.035, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.035, y = 0.755, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.535, y = 0.755, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.035, y = 0.52, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.035, y = 0.29, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.535, y = 0.29, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold")
+
 #
 # Print final plot
-ggsave(here("./output/figures/text/results_energy.png"), plot = final_plot_energy, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/results_energy.png")
+ggsave(here("./output/figures/text/results_energy.tiff"), plot = final_plot_energy, width = 18, height = 24, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/results_energy.tiff")
 ```
@@
 _Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
-However, the interaction of CORT and temperature have significant effects on DNA damage in the optic tecta (see @fig-results_oxidative).
-For DNA damage, CORT elevations had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`), but it decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E).
-Prenatal CORT had opposite effects on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), but differences between groups were not significant (see @fig-results_oxidative C).
+However, the interaction of CORT and temperature have significant effects on DNA damage ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot)) - (mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OT$Control_Hot - DNA_OT$CORT_Hot) - (DNA_OT$Control_Cold - DNA_OT$CORT_Cold)), 3, equal = TRUE)`), and lipid peroxidation ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot)) - (mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OT$Control_Hot - LP_OT$CORT_Hot) - (LP_OT$Control_Cold - LP_OT$CORT_Cold)), 3, equal = TRUE)`) in the optic tecta.
+CORT elevations had no effect on hot-incubated lizards (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Hot - DNA_OT$CORT_Hot), 3, equal = TRUE)`), but it decreased total DNA damage in cold-incubated animals (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(DNA_OT$Control_Cold - DNA_OT$CORT_Cold), 3, equal = TRUE)`) (see @fig-results_oxidative C, E).
+The direction of the effects of CORT on lipid peroxidation were the opposite on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), with significant differences between CORT-Hot and CORT-Cold lizards (β~CORT-Hot~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$CORT_Hot) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$CORT_Hot - LP_OT$CORT_Cold), 3, equal = TRUE)`; see @fig-results_oxidative C).
 In addition, we found that females had higher DNA damage levels, and lipid peroxidation increased with age (see Table S4).
@@
 plot_ros_OB <- plotting(ROS_OB, "ROS")
+plot_conros_OB <- plotting(ROScon_OB, "ROS/Mit density")
 plot_dnadamage_OB <- plotting(DNA_OB, "DNA damage")
@@
 #
-fig_OB_oxidative_top <- plot_grid(plot_ros_OB, NULL, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
-fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, NULL, plot_peroxidation_OB,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
+fig_OB_oxidative_top <- plot_grid(plot_ros_OB, plot_conros_OB, NULL,
+                    nrow = 1, rel_widths = c(1, 1, 0.35))
+fig_OB_oxidative_bottom <- plot_grid(plot_dnadamage_OB, plot_peroxidation_OB, NULL,
+                    nrow = 1, rel_widths = c(1, 1, 0.35))
 fig_OB_oxidative <- plot_grid(fig_OB_oxidative_top, fig_OB_oxidative_bottom,
@@
 plot_ros_OT <- plotting(ROS_OT, "ROS")
+plot_conros_OT <- plotting(ROScon_OT, "ROS/Mit density")
 plot_dnadamage_OT <- plotting(DNA_OT, "DNA damage")
@@
 #
-fig_OT_oxidative_top <- plot_grid(plot_ros_OT, NULL, NULL,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
-fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, NULL, plot_peroxidation_OT,
-                    nrow = 1, rel_widths = c(0.8, 0.1, 0.8))
+fig_OT_oxidative_top <- plot_grid(plot_ros_OT, plot_conros_OT, NULL,
+                    nrow = 1, rel_widths = c(1, 1, 0.35))
+fig_OT_oxidative_bottom <- plot_grid(plot_dnadamage_OT, plot_peroxidation_OT, NULL,
+                    nrow = 1, rel_widths = c(1, 1, 0.35))
 fig_OT_oxidative <- plot_grid(fig_OT_oxidative_top, fig_OT_oxidative_bottom,
@@
 fig_results_oxidative <- plot_grid(
-  fig_OB_oxidative, NULL, fig_OT_oxidative, NULL,
-  nrow = 4, rel_heights = c(1, 0.05, 1, 0.1))
+  fig_OB_oxidative, fig_OT_oxidative, legend_plot,
+  nrow = 3, rel_heights = c(1, 1, 0.1))
 # Final composition: Merge everything, adding images and legend
@@
   # Insert OB image in the top right
-  draw_image("./Others/OB.png", x = 0.56, y = 0.755, width = 0.2, height = 0.3) +
+  draw_image("./Others/OB.png", x = 0.85, y = 0.77, width = 0.17, height = 0.24) +
   # Insert OT image in the middle right
-  draw_image("./Others/OT.png", x = 0.56, y = 0.28, width = 0.19, height = 0.28) +
-  # Insert legend at the bottom-right
-  draw_grob(legend_plot, x = 0.5, y = 0.025, width = 0.0001, height = 0.0001) +
+  draw_image("./Others/OT.png", x = 0.85, y = 0.295, width = 0.15, height = 0.245) +
   # Insert title for each plot
-  annotate("text", x = 0.0365, y = 0.986, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0365, y = 0.755, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.567, y = 0.755, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0365, y = 0.5, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0365, y = 0.268, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.567, y = 0.268, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("segment", x = 0.46, xend= 0.46, y = 0.21, yend = 0.26, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.45, xend= 0.46, y = 0.21, yend = 0.21, size = 0.5, colour = "black") +
-  annotate("segment", x = 0.45, xend= 0.46, y = 0.26, yend = 0.26, size = 0.5, colour = "black") +
-  annotate("text", x = 0.47, y = 0.228, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
+  annotate("text", x = 0.028, y = 0.986, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.455, y = 0.986, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.028, y = 0.745, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.455, y = 0.745, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.028, y = 0.51, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.455, y = 0.51, label = "F", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.028, y = 0.27, label = "G", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.455, y = 0.27, label = "H", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("segment", x = 0.39, xend= 0.39, y = 0.217, yend = 0.262, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.38, xend= 0.39, y = 0.217, yend = 0.217, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.38, xend= 0.39, y = 0.262, yend = 0.262, size = 0.5, colour = "black") +
+  annotate("text", x = 0.4, y = 0.24, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
+  annotate("segment", x = 0.83, xend= 0.83, y = 0.173, yend = 0.262, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.82, xend= 0.83, y = 0.173, yend = 0.173, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.82, xend= 0.83, y = 0.262, yend = 0.262, size = 0.5, colour = "black") +
+  annotate("text", x = 0.84, y = 0.21, label = "*", hjust = 0.5, vjust = 0.5, size = 7) +
+  annotate("segment", x = 0.83, xend= 0.83, y = 0.648, yend = 0.737, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.82, xend= 0.83, y = 0.648, yend = 0.648, size = 0.5, colour = "black") +
+  annotate("segment", x = 0.82, xend= 0.83, y = 0.737, yend = 0.737, size = 0.5, colour = "black") +
+  annotate("text", x = 0.84, y = 0.685, label = "*", hjust = 0.5, vjust = 0.5, size = 7)
 #
 # Print final plot
-ggsave(here("./output/figures/text/results_oxidative.png"), plot = final_plot_oxidative, width = 18, height = 24, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/results_oxidative.png")
-```
-
-#### Does prenatal CORT and incubation temperature affect prey detection?
+ggsave(here("./output/figures/text/results_oxidative.tiff"), plot = final_plot_oxidative, width = 24, height = 24, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/results_oxidative.tiff")
+```
+
+#### Do prenatal CORT and incubation temperature affect prey detection?
 Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).
@@
 # Print final plot
-ggsave(here("./output/figures/text/results_behaviour.png"), plot = final_plot_behaviour, width = 21, height = 7, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/results_behaviour.png")
+ggsave(here("./output/figures/text/results_behaviour.tiff"), plot = final_plot_behaviour, width = 21, height = 7, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/results_behaviour.tiff")
 ```
@@
   annotate("text", x = 0.282, y = 0.82, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
-ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/SEM_OB.png")
+ggsave(here("./output/figures/text/SEM_OB.tiff"), plot = fig_SEM_OB, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/SEM_OB.tiff")
 ```
@@
   annotate("text", x = 0.285, y = 0.82, label = OT_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
-ggsave(here("./output/figures/text/SEM_OT.png"), plot = fig_SEM_OT, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/SEM_OT.png")
+ggsave(here("./output/figures/text/SEM_OT.tiff"), plot = fig_SEM_OT, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/SEM_OT.tiff")
 ```
@@
             "mean_potential_def_OB",
+            "mean_conpotential_def_OB",
             "mean_ros_def_OB",
+            "mean_conros_def_OB",
             "mean_dnadamage_def_OB",
@@
             "mean_potential_def_OT",
+            "mean_conpotential_def_OT",
             "mean_ros_def_OT",
+            "mean_conros_def_OT",
             "mean_dnadamage_def_OT",
@@
                                   "mean_potential_def_OB",
+                                  "mean_conpotential_def_OB",
                                   "mean_ros_def_OB",
+                                  "mean_conros_def_OB",
                                   "mean_dnadamage_def_OB",
@@
                                   "mean_potential_def_OT",
+                                  "mean_conpotential_def_OT",
                                   "mean_ros_def_OT",
+                                  "mean_conros_def_OT",
                                   "mean_dnadamage_def_OT",
@@
                                   "m_def_mean_potential_OB" = "Mit potential",
+                                  "m_def_mean_conpotential_OB" = "Mit potential/Mit density",
                                   "m_def_mean_ros_OB" = "ROS",
+                                  "m_def_mean_conros_OB" = "ROS/Mit density",
                                   "m_def_mean_dnadamage_OB" = "DNA damage",
@@
                                   "m_def_mean_potential_OT" = "Mit potential",
+                                  "m_def_mean_conpotential_OT" = "Mit potential/Mit density",
                                   "m_def_mean_ros_OT" = "ROS",
+                                  "m_def_mean_conros_OT" = "ROS/Mit density",
                                   "m_def_mean_dnadamage_OT" = "DNA damage",
@@
   align(align = "center", j = c(1:5), part = "header") %>%
-  flextable::compose(i = c(2:5,8:11), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = 6, part = "body") %>% 
+  flextable::compose(i = c(2:7,10:15), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = 8, part = "body") %>% 
   autofit()
@@
 # A) Organise df 
-var <- c("MD", "MP", "ROS", "DNA", "LP", "DET")
+var <- c("MD", "MP", "MPcon", "ROS", "ROScon", "DNA", "LP", "DET")
 region <- c("OB", "OT")
@@
       Variable == "MP" ~ "Mit potential",
+      Variable == "MPcon" ~ "Mit potential/Mit density",
       Variable == "ROS" ~ "ROS",
+      Variable == "ROScon" ~ "ROS/Mit density",
       Variable == "DNA" ~ "DNA damage",
@@
     ) %>%
-  mutate(Variable = factor(Variable, levels = c("Mit density", "Mit potential", "ROS", "DNA damage", "Lipid peroxidation", "Detection latency"))) %>%
+  mutate(Variable = factor(Variable, levels = c("Mit density", "Mit potential", "Mit potential/Mit density", "ROS", "ROS/Mit density","DNA damage", "Lipid peroxidation", "Detection latency"))) %>%
   mutate(
@@
   bold(~`pMCMC contrast` <0.001, j = c("pMCMC contrast", "Contrast", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:18,20:36), j = 1, value = as_paragraph(""), part = "body") %>%
-  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24,26,27,29,30,32,33,35,36),
+  flextable::compose(i = c(2:24,26:48), j = 1, value = as_paragraph(""), part = "body") %>%
+  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24,26,27,29,30,32,33,35,36,38,39,41,42,44,45,47,48),
                     j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
-  flextable::hline(i = c(3,6,9,12,15,21,24,27,30,33,36), j = c(2:5), part = "body") %>% 
-  flextable::hline(i = 18, part = "body") %>%
+  flextable::hline(i = c(3,6,9,12,15,18,21,24,27,30,33,36,39,42,45), j = c(2:5), part = "body") %>% 
+  flextable::hline(i = 24, part = "body") %>%
   autofit()
@@
   bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:12,14:18,20:23,25:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = c(2:4,6:8,10:12,14:16,18:20,22:26,28:31,33:35), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   autofit()
@@
   bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:12,14:17,19:22,24:26), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = c(2:4,6:8,10:12,14:16,18:20,22:25,27:30,32:34), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   autofit()
@@
 
-Table S11. Summary of the preliminary model for ROS Production in Olfactory Bulbs.
+Table S11. Summary of the preliminary model for Metabolic capacity/Mitochondrial Density in Olfactory Bulbs.
+```{r, results_preliminary_conpotential_OB}
+#| tbl-cap: "Preliminary results of the models testing for Metabolic capacity/Mitochondrial Density in OB"
+#| label: results_preliminary_conpotential_OB
+#
+sum_m_conpotential_OB_prel <- m_prel_mean_conpotential_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_conpotential_OB_prel)
+```
+
+Model formula: mean_conpotential ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S12. Summary of the preliminary model for Metabolic capacity/Mitochondrial Density in Optic Tecta.
+```{r, results_preliminary_conpotential_OT}
+#| tbl-cap: "Preliminary results of the models testing for Metabolic capacity/Mitochondrial Density in OT"
+#| label: results_preliminary_conpotential_OT
+#
+sum_m_conpotential_OT_prel <- m_prel_mean_conpotential_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_conpotential_OT_prel)
+```
+
+Model formula: mean__conpotential ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S13. Summary of the preliminary model for ROS Production in Olfactory Bulbs.
 ```{r, results_preliminary_ros_OB}
@@
 
-Table S12. Summary of the preliminary model for ROS Production in Optic Tecta.
+Table S14. Summary of the preliminary model for ROS Production in Optic Tecta.
 ```{r, results_preliminary_ros_OT}
@@
 
-Table S13. Summary of the preliminary model for DNA Damage in Olfactory Bulbs.
+Table S15. Summary of the preliminary model for ROS Production/Mitochondrial density in Olfactory Bulbs.
+```{r, results_preliminary_conros_OB}
+#| tbl-cap: "Preliminary results of the models testing for ROS Production/Mitochondrial density in OB"
+#| label: results_preliminary_conros_OB
+#
+sum_m_conros_OB_prel <- m_prel_mean_conros_OB %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_conros_OB_prel)
+```
+
+Model formula: mean_ros ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S16. Summary of the preliminary model for ROS Production/Mitochondrial density in Optic Tecta.
+```{r, results_preliminary_conros_OT}
+#| tbl-cap: "Preliminary results of the models testing for ROS Production/mit density in OT"
+#| label: results_preliminary_conros_OT
+#
+sum_m_conros_OT_prel <- m_prel_mean_conros_OT %>%
+  dplyr::select(-starts_with("r_"), -sigma, -Intercept, -lprior, -lp__) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_conros_OT_prel)
+```
+
+Model formula: mean_conros ~ cort * temp + age_euthanasia + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S17. Summary of the preliminary model for DNA Damage in Olfactory Bulbs.
 ```{r, results_preliminary_dnadamage_OB}
@@
 
-Table S14. Summary of the preliminary model for DNA Damage in Optic Tecta.
+Table S18. Summary of the preliminary model for DNA Damage in Optic Tecta.
 ```{r, results_preliminary_dnadamage_OT}
@@
 
-Table S15. Summary of the preliminary model for Lipid Peroxidation in Olfactory Bulbs.
+Table S19. Summary of the preliminary model for Lipid Peroxidation in Olfactory Bulbs.
 ```{r, results_preliminary_peroxidation_OB}
@@
 
-Table S16. Summary of the preliminary model for Lipid Peroxidation in Optic Tecta.
+Table S20. Summary of the preliminary model for Lipid Peroxidation in Optic Tecta.
 ```{r, results_preliminary_peroxidation_OT}
@@
 
-Table S17. Summary of the preliminary model for Detection Latency (t_D) of Chemical stimuli. 
+Table S21. Summary of the preliminary model for Detection Latency (t_D) of Chemical stimuli. 
 ```{r, results_preliminary_t_D_Chemical}
@@
 
-Table S18. Summary of the preliminary model for Detection Latency (t_D) of Visual stimuli.
+Table S22. Summary of the preliminary model for Detection Latency (t_D) of Visual stimuli.
 ```{r, results_preliminary_t_D_Visual}
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S1.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
+knitr::include_graphics("./output/figures/suppl/Figure_S1.tiff")
 ```
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S2.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
+knitr::include_graphics("./output/figures/suppl/Figure_S2.tiff")
 ```
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S3.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
+knitr::include_graphics("./output/figures/suppl/Figure_S3.tiff")
 ```
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S4.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
+knitr::include_graphics("./output/figures/suppl/Figure_S4.tiff")
 ```
@@
 Fig S4. Posterior predictive checks for the model of Metabolic capacity in Optic Tecta.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_conpotential_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_conpotential_OB
+#| caption: "Posterior predictive checks for the model of Metabolic capacity/Mi density in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_conpotential_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S5.tiff", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S5.tiff")
+```
+
+Fig S5. Posterior predictive checks for the model of Metabolic capacity/Mitochondrial density in Olfactory Bulbs.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_conpotential_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_conpotential_OT
+#| caption: "Posterior predictive checks for the model of Metabolic capacity/Mit density in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_conpotential_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S6.tiff", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S6.tiff")
+```
+
+Fig S6. Posterior predictive checks for the model of Metabolic capacity/Mitochondrial density in Optic Tecta.
 Formula: mean_potential ~ cort * temp + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S7.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
-```
-
-Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
+knitr::include_graphics("./output/figures/suppl/Figure_S7.tiff")
+```
+
+Fig S7. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
 Formula: mean_ros ~ cort * temp + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S8.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
-```
-
-Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
+knitr::include_graphics("./output/figures/suppl/Figure_S8.tiff")
+```
+
+Fig S8. Posterior predictive checks for the model of ROS Production in Optic Tecta.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_conros_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_conros_OB
+#| caption: "Posterior predictive checks for the model of ROS Production/Mitochondrial density in Olfactory bulbs."
+#
+mod <- readRDS(here("output/models/mean_conros_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S9.tiff", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S9.tiff")
+```
+
+Fig S9. Posterior predictive checks for the model of ROS Production/Mitochondrial density in Olfactory Bulbs.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_conros_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_conros_OT
+#| caption: "Posterior predictive checks for the model of ROS Production/Mitochondrial density in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_conros_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S10.tiff", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S10.tiff")
+```
+
+Fig S10. Posterior predictive checks for the model of ROS Production/Mitochondrial density in Optic Tecta.
 Formula: mean_ros ~ cort * temp + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S11.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
-```
-
-Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
+knitr::include_graphics("./output/figures/suppl/Figure_S11.tiff")
+```
+
+Fig S11. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
 Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S12.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
-```
-
-Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
+knitr::include_graphics("./output/figures/suppl/Figure_S12.tiff")
+```
+
+Fig S12. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
 Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S13.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
-```
-
-Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
+knitr::include_graphics("./output/figures/suppl/Figure_S13.tiff")
+```
+
+Fig S13. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
 Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S14.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
-```
-
-Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
+knitr::include_graphics("./output/figures/suppl/Figure_S14.tiff")
+```
+
+Fig S14. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
 Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S15.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
-```
-
-Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
+knitr::include_graphics("./output/figures/suppl/Figure_S15.tiff")
+```
+
+Fig S15. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
 Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)
@@
 fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
+ggsave("./output/figures/suppl/Figure_S16.tiff", plot = fig_mod,
        width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
-```
-
-Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
+knitr::include_graphics("./output/figures/suppl/Figure_S16.tiff")
+```
+
+Fig S16. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
 Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-#### What did average mitochondria do?
-
-When obtaining data using FlowJo, we got the geometric mean value for each event recorded in the flow cytometer after gatting for positive values in the dyes employed (see _Methods: Flow cytometry_).
-However, this is not informative about the effects per mitochondria.
-In other words, we might be loosing some resolution about the effects at a mitochondrial level rather than a cellular level.
-To further investigate the effects per mitochondrial unit, we repeated the models above but dividing each variable by mean mitochondrial density obtained for each individual.
-
-```{r, data_mitochondrial_controlled}
-#| label: data_mitochondrial_controlled
-# 
-mit_controlled <- mit_df %>%
-    mutate(Id_flow = as.character(Id_flow)) %>%
-    mutate(mean_dnadamage = log10(mean_dnadamage + 1)) %>%
-    mutate(mean_peroxidation = log10(mean_peroxidation + 1)) %>%
-    mutate(across(where(~ is.numeric(.)), ~.x / mean_mitodensity))
-#
-df_controlled <- merge(data_discr, mit_controlled, by = c("stimulus", "Id_flow")) %>%
-  mutate(across(c(age_trial, age_euthanasia), as.numeric)) %>%
-  mutate(across(where(~ is.double(.) & !is.integer(.)),
-                ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE))))
-```
-
-```{r, models_mitochondrial_controlled}
-#| label: models_mitochondrial_controlled
-#
-# Run models mitochondrial physiology from df_controlled (each region separately)
-#
-var_m_controlled <- c("mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
-regions <- c("OB", "OT")
-for (p in var_m_controlled){
-  for (h in regions){
-    if (h == "OB"){
-      df <- df_controlled %>% filter(region == "OB")
-      l <- "OB"
-      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
-        formula <- paste0(p, "~ cort*temp + (1|clutch)")
-      } else if (p == "mean_dnadamage"){
-        formula <- paste0(p, "~ cort*temp + age_euthanasia + sex + (1|clutch)")
-      } else {
-        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
-      }
-    } else {
-      df <- df_controlled %>% filter(region == "OT")
-      l <- "OT"
-      if (p %in% c("mean_mitodensity", "mean_potential", "mean_ros")){
-        formula <- paste0(p, "~ cort*temp + (1|clutch)")
-      } else if (p == "mean_dnadamage"){
-        formula <- paste0(p, "~ cort*temp + sex + (1|clutch)")
-      } else {
-        formula <- paste0(p, "~ cort*temp + age_euthanasia + (1|clutch)")
-      }
-    }
-
-    pmodel_name <- paste0("m_controlled_", p, "_", h)
-    assign(pmodel_name, fit_m(df = df,
-                              cat = "controlled",
-                              var = p,
-                              formula = formula,
-                              fam = gaussian(),
-                              label = l,
-                              refit = FALSE),
-          envir = .GlobalEnv)  # Assign to the global environment
-  }
-} 
-```
-
-```{r, organise_posteriors_controlled}
-#| label: organise_posteriors_controlled
-#
-# Organising the posteriors of the previous models to fit the tables below
-#
-source(here("R", "func.R"))
-#
-# Databases for each region/stimulus
-post_OB_cont <- data.frame()
-post_OT_cont <- data.frame()
-# Names posteriors:
-region <- c("OB", "OT")
-names_OB_cont <- c("m_controlled_mean_potential_OB", "m_controlled_mean_ros_OB", "m_controlled_mean_dnadamage_OB", "m_controlled_mean_peroxidation_OB")
-names_OT_cont <- c("m_controlled_mean_potential_OT", "m_controlled_mean_ros_OT", "m_controlled_mean_dnadamage_OT", "m_controlled_mean_peroxidation_OT")
-#
-# Organising the results
-for (r in region) {
-  model_select <- get(paste0("names_", r, "_cont"))   
-  for (pos in model_select) {
-    model <- get(pos)      # Get the model from the global environment
-    post_result <- tidy_post(model)        # Apply tidy_post to each model
-    # Add a new column to identify the region and model
-    post_result$Region <- r
-    post_result$Model <- pos
-    # Append to the appropriate data frame
-    if (r == "OB") {
-      post_OB_cont <- bind_rows(post_OB_cont, post_result)
-    } else {
-      post_OT_cont <- bind_rows(post_OT_cont, post_result)
-    }
-  }
-}
-```
-
-```{r, values_posteriors_controlled}
-#| label: values_posteriors_controlled
-#
-# Extracting the posteriors for the models and the values of interest.
-Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
-#
-source(here("R", "func.R"))
-#
-# A) Olfactory bulbs/Chemical stimulus
-#
-MP_OB_con <- post_values(m_controlled_mean_potential_OB, "none")
-ROS_OB_con <- post_values(m_controlled_mean_ros_OB, "none")
-DNA_OB_con <- post_values(m_controlled_mean_dnadamage_OB, "none")
-LP_OB_con <- post_values(m_controlled_mean_peroxidation_OB, "none")
-#
-# B) Optic tecta/Visual stimulus
-MP_OT_con <- post_values(m_controlled_mean_potential_OT, "none")
-ROS_OT_con <- post_values(m_controlled_mean_ros_OT, "none")
-DNA_OT_con <- post_values(m_controlled_mean_dnadamage_OT, "sex")
-LP_OT_con <- post_values(m_controlled_mean_peroxidation_OT, "none")
-```
-
-Table SXX.
-Estimate of the contrasts between treatments for all the variables analysed, and pMCMC values testing the hypothesis that contrast is different from 0.
-In bold, the values that are significant at pMCMC < 0.05.
-```{r tbl-contrasts_controlled}
-#| label: tbl-contrasts_controlled
-#| tbl-cap: "Contrasts between prenatal conditions for mitochondrial physiology and learning."
-#
-source(here("R", "func.R"))
-#
-# A) Organise df 
-var <- c("MP", "ROS", "DNA", "LP")
-region <- c("OB", "OT")
-#
-data_table_controlled <- data.frame()
-for(v in var){
-  for(r in region){
-    x <- paste0(v, "_", r, "_con")
-    df <- get(x)
-    Temperature <- format_dec(mean(c(df$CORT_Hot, df$Control_Hot)) - mean(c(df$CORT_Cold, df$Control_Cold)), 3)
-    pMCMC_temp <- format_p(pmcmc(c(df$CORT_Hot, df$Control_Hot) - c(df$CORT_Cold, df$Control_Cold)), 3, equal = FALSE)
-    CORT <- format_dec(mean(c(df$Control_Hot, df$Control_Cold)) - mean(c(df$CORT_Hot, df$CORT_Cold)), 3)
-    pMCMC_cort <- format_p(pmcmc(c(df$Control_Hot, df$Control_Cold) - c(df$CORT_Hot, df$CORT_Cold)), 3, equal = FALSE)
-    Interaction <- format_dec((mean(df$Control_Hot) - mean(df$CORT_Hot)) - (mean(df$Control_Cold) - mean(df$CORT_Cold)), 3)
-    pMCMC_int <- format_p(pmcmc((df$Control_Hot - df$CORT_Hot) - (df$Control_Cold - df$CORT_Cold)), 3, equal = FALSE)
-    data_temp <- data.frame(Variable = x,
-                          Temperature = as.numeric(Temperature),
-                          pMCMC_temp = pMCMC_temp,
-                          CORT = as.numeric(CORT),
-                          pMCMC_cort = pMCMC_cort,
-                          Interaction = as.numeric(Interaction),
-                          pMCMC_int = pMCMC_int)
-    data_table_controlled <- dplyr::bind_rows(data_table_controlled, data_temp)
-  }
-}
-# Modify the df
-data_table_controlled_final <- data_table_controlled %>%
-  pivot_longer(cols = c(Temperature, CORT, Interaction), 
-               names_to = "Predictor", 
-               values_to = "Contrast") %>%
-  mutate(
-    # Extract the pMCMC values from the corresponding columns
-    `pMCMC contrast` = case_when(
-      Predictor == "Temperature" ~ pMCMC_temp,
-      Predictor == "CORT" ~ pMCMC_cort,
-      Predictor == "Interaction" ~ pMCMC_int
-    )
-  ) %>%
-  separate(col = Variable, into = c("Variable", "Region", "control"), sep = "_") %>%
-  mutate(
-    Variable = case_when(
-      Variable == "MP" ~ "Mit potential",
-      Variable == "ROS" ~ "ROS",
-      Variable == "DNA" ~ "DNA damage",
-      Variable == "LP" ~ "Lipid peroxidation",
-      Variable == "DET" ~ "Detection latency",
-      TRUE ~ Variable
-      )
-    ) %>%
-  mutate(Variable = factor(Variable, levels = c("Mit potential", "ROS", "DNA damage", "Lipid peroxidation"))) %>%
-  mutate(
-    Region = case_when(
-      Region == "OB" ~ "Olfactory bulbs",
-      Region == "OT" ~ "Optic tecta",
-      TRUE ~ Region
-    )
-  ) %>%
-  dplyr::select(Region, Variable, Predictor, Contrast, `pMCMC contrast`) %>%
-  arrange(Region, Variable)
-#
-# C) Make the contrasts table:
-#
-set_flextable_defaults(
- font.family = "Times New Roman",
- font.size = 10)
-#
-contrast_table_controlled <- flextable(data_table_controlled_final) %>%
-  align(align = "center", j = c(4,5), part = "body") %>%
-  align(align = "center", j = c(1:4), part = "header") %>%
-  bold(~`pMCMC contrast` < 0.05, j = c("pMCMC contrast", "Contrast", "Predictor"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`pMCMC contrast` <0.001, j = c("pMCMC contrast", "Contrast", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:12,14:24), j = 1, value = as_paragraph(""), part = "body") %>%
-  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18,20,21,23,24),
-                    j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
-  flextable::hline(i = c(3,6,9,12,15,18,21), j = c(2:5), part = "body") %>% 
-  flextable::hline(i = 12, part = "body") %>%
-  autofit()
-#
-contrast_table_controlled
-```
-
-Contrasts were done by:  
-_Temperature_: β~Hot~ - β~Cold~  
-_CORT_: β~CORT~ - β~Control~    
-_Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table SXX+1. Summary of the final models for Olfactory Bulbs/Chemical stimulus.
-```{r, tbl-results_OB_table_cont}
-#| label: results_OB_table_cont
-#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
-#| tbl-name: "results_OB_cont"
-#| tbl-label: "results_OB_cont"
-source(here("R", "func.R"))
-# 
-# A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB_cont)
-variable_order <- c("Mit potential",
-                     "ROS",
-                     "DNA damage",
-                     "Peroxidation")
-predictor_order <- c("b_Intercept",
-                     "b_cortCORT",
-                     "b_tempHot",
-                     "b_cortCORT:tempHot",
-                     "b_age_euthanasia",
-                     "b_sexFemale")
-post_OB_cont_refined <- post_OB_cont %>%
-    mutate(
-      Model = case_when(
-        Model == "m_controlled_mean_potential_OB" ~ "Mit potential",
-        Model == "m_controlled_mean_ros_OB" ~ "ROS",
-        Model == "m_controlled_mean_dnadamage_OB" ~ "DNA damage",
-        Model == "m_controlled_mean_peroxidation_OB" ~ "Peroxidation",
-        Model == "m_controlled_mean_potential_OT" ~ "Mit potential",
-        Model == "m_controlled_mean_ros_OT" ~ "ROS",
-        Model == "m_controlled_mean_dnadamage_OT" ~ "DNA damage",
-        Model == "m_controlled_mean_peroxidation_OT" ~ "Peroxidation",
-        TRUE ~ Model
-      )
-    ) %>%
-    mutate(CI_95 = paste0("[", CI_lower, " , ", CI_upper, "]")) %>%
-    rename(
-      Variable = Model,
-      Predictors = Predictor,
-      'Estimate Mean' = Mean, 
-      '95% CI' = CI_95, 
-      PMCMC = PMCMC
-    ) %>%
-    dplyr::select(
-      Variable,
-      Predictors,
-      'Estimate Mean',
-      '95% CI',
-      PMCMC) %>%
-    mutate(Variable = factor(Variable, levels = variable_order)) %>%
-    mutate(Predictors = factor(Predictors, levels = predictor_order)) %>%
-    arrange(Variable, Predictors)
-#
-# B) Create table
-## Table format
-set_flextable_defaults(
- font.family = "Times New Roman",
- font.size = 10)
-#
-OB_table_controlled <- flextable(post_OB_cont_refined) %>%
-  align(align = "center", j = c(3:5), part = "body") %>%
-  align(align = "center", j = c(1:5), part = "header") %>%
-  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:14,16:19), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  autofit()
-#
-OB_table_controlled
-```
-
-```{r, fig-results_OBcont}
-#| label: fig-results_OBcont
-#| fig-cap: "NEW_OB"
-#
-source(here("R", "func.R"))
-#
-# A) Plotting the results for OB/Chemical stimulus using the df from before
-plot_potential_OB_con <- plotting(MP_OB_con, "Mit potential")
-plot_ROS_OB_con <- plotting(ROS_OB_con, "ROS")
-plot_DNA_OB_con <- plotting(DNA_OB_con, "DNA damage")
-plot_LP_OB_con <- plotting(LP_OB_con, "Lipid peroxidation")
-#
-fig_OB_con <- plot_grid(plot_potential_OB_con, plot_ROS_OB_con,
-                    plot_DNA_OB_con, plot_LP_OB_con, NULL, NULL,
-                    ncol = 2, nrow = 3, rel_heights = c(1, 1, 0.2))
-#
-# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
-plot_legend_bottom_con <- plotting(MP_OB_con, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
-gtable_con <- ggplot_gtable(ggplot_build(plot_legend_bottom_con))
-legend_plot_con <- gtable_con$grobs[[which(sapply(gtable_con$grobs, function(x) x$name) == "guide-box")]]
-# 
-# Final composition: Merge everything, adding images and legend
-plot_OB_con <- ggdraw(fig_OB_con +
-  # Insert legend at the bottom-centered
-  draw_grob(legend_plot_con, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001))
-# Print final plot
-ggsave(here("./output/figures/suppl/new_models/plot_OB_con.png"), plot = plot_OB_con, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/new_models/plot_OB_con.png")
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table SXX+2. Summary of the final models for Optic Tecta/Visual stimulus.
-```{r, tbl-results_OT_table_cont}
-#| label: results_OT_table_cont
-#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
-#| tbl-name: "results_OT_cont"
-#| tbl-label: "results_OT_cont"
-source(here("R", "func.R"))
-# 
-# A) Refining the df summarizing the posteriors for OT/Visual stimulus (post_OT_cont)
-variable_order <- c("Mit potential",
-                     "ROS",
-                     "DNA damage",
-                     "Peroxidation")
-predictor_order <- c("b_Intercept",
-                     "b_cortCORT",
-                     "b_tempHot",
-                     "b_cortCORT:tempHot",
-                     "b_age_euthanasia",
-                     "b_sexFemale")
-post_OT_cont_refined <- post_OT_cont %>%
-    mutate(
-      Model = case_when(
-        Model == "m_controlled_mean_potential_OB" ~ "Mit potential",
-        Model == "m_controlled_mean_ros_OB" ~ "ROS",
-        Model == "m_controlled_mean_dnadamage_OB" ~ "DNA damage",
-        Model == "m_controlled_mean_peroxidation_OB" ~ "Peroxidation",
-        Model == "m_controlled_mean_potential_OT" ~ "Mit potential",
-        Model == "m_controlled_mean_ros_OT" ~ "ROS",
-        Model == "m_controlled_mean_dnadamage_OT" ~ "DNA damage",
-        Model == "m_controlled_mean_peroxidation_OT" ~ "Peroxidation",
-        TRUE ~ Model
-      )
-    ) %>%
-    mutate(CI_95 = paste0("[", CI_lower, " , ", CI_upper, "]")) %>%
-    rename(
-      Variable = Model,
-      Predictors = Predictor,
-      'Estimate Mean' = Mean, 
-      '95% CI' = CI_95, 
-      PMCMC = PMCMC
-    ) %>%
-    dplyr::select(
-      Variable,
-      Predictors,
-      'Estimate Mean',
-      '95% CI',
-      PMCMC) %>%
-    mutate(Variable = factor(Variable, levels = variable_order)) %>%
-    mutate(Predictors = factor(Predictors, levels = predictor_order)) %>%
-    arrange(Variable, Predictors)
-#
-# B) Create table
-## Table format
-set_flextable_defaults(
- font.family = "Times New Roman",
- font.size = 10)
-#
-OT_table_controlled <- flextable(post_OT_cont_refined) %>%
-  align(align = "center", j = c(3:5), part = "body") %>%
-  align(align = "center", j = c(1:5), part = "header") %>%
-  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:13,15:18), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  autofit()
-#
-OT_table_controlled
-```
-
-```{r, fig-results_OTcont}
-#| label: fig-results_OTcont
-#| fig-cap: "NEW_OT"
-#
-source(here("R", "func.R"))
-#
-# A) Plotting the results for OT/Visual stimulus using the df from before
-plot_potential_OT_con <- plotting(MP_OT_con, "Mit potential")
-plot_ROS_OT_con <- plotting(ROS_OT_con, "ROS")
-plot_DNA_OT_con <- plotting(DNA_OT_con, "DNA damage")
-plot_LP_OT_con <- plotting(LP_OT_con, "Lipid peroxidation")
-#
-fig_OT_con <- plot_grid(plot_potential_OT_con, plot_ROS_OT_con,
-                    plot_DNA_OT_con, plot_LP_OT_con, NULL, NULL,
-                    ncol = 2, nrow = 3, rel_heights = c(1, 1, 0.2))
-#
-# Getting the legend (get_legend() does not work if the legend is on the top or bottom)
-plot_legend_bottom_con <- plotting(MP_OT_con, "Mit potential") + theme(legend.position = "bottom", legend.title = element_blank())
-gtable_con <- ggplot_gtable(ggplot_build(plot_legend_bottom_con))
-legend_plot_con <- gtable_con$grobs[[which(sapply(gtable_con$grobs, function(x) x$name) == "guide-box")]]
-# 
-# Final composition: Merge everything, adding images and legend
-plot_OT_con <- ggdraw(fig_OT_con +
-  # Insert legend at the bottom-centered
-  draw_grob(legend_plot_con, x = 0.45, y = 0.05, width = 0.0001, height = 0.0001))
-# Print final plot
-ggsave(here("./output/figures/suppl/new_models/plot_OT_con.png"), plot = plot_OT_con, width = 21, height = 14, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/new_models/plot_OT_con.png")
-```
 
@@
   annotate("text", x = 0.55, y = 0.92, label = "B", hjust = 1, vjust = 1, size = 4, family = "Times", color = "white") 
-ggsave(here("./output/figures/suppl/Figure_SX13.png"), plot = fig_neun, width = 15, height = 5, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX13.png")
-```
-
-Figure S13.
+ggsave(here("./output/figures/suppl/Figure_SX13.tiff"), plot = fig_neun, width = 15, height = 5, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX13.tiff")
+```
+
+Figure S17.
 Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488.

### 2025-09-06 R/data_process.R

@@ -66,7 +66,9 @@ mit_df <- merge(mit_act, mit_damage,
 write.csv(mit_df, "./output/checking/mit_df.csv")
 # C) Merging all three dfs
 #
-df_merged <- merge(data_discr, mit_df, by = c("stimulus", "Id_flow"))
+df_merged <- merge(data_discr, mit_df, by = c("stimulus", "Id_flow")) %>%
+  mutate(mean_conpotential = as.numeric(mean_potential / mean_mitodensity),
+         mean_conros = as.numeric(mean_ros / mean_mitodensity))
 #
 #
 # Transform variables according to distribution (see below and exploratory analyses)
@@ -76,13 +78,11 @@ df_merged <- merge(data_discr, mit_df, by = c("stimulus", "Id_flow"))
 #
 log_var <- c("t_D", "mean_mitodensity", "mean_dnadamage", "mean_peroxidation")
 df_merged_log <- df_merged %>%
-  mutate(across(all_of(log_var), ~ log10(. + 1)))
+  mutate(across(all_of(log_var), ~ log10(. + 1))) 
 #
 # Convert age variables to numeric, control mit potential and ROS by mitocondrial, and standardize all numerical variables
 clean_df <- df_merged_log %>%
   mutate(across(c(age_trial, age_euthanasia), as.numeric)) %>%
-  mutate(mean_conpotential = as.numeric(mean_potential / mean_mitodensity),
-         mean_conros = as.numeric(mean_ros / mean_mitodensity)) %>%
   mutate(across(where(~ is.double(.) & !is.integer(.)),
                 ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE))))
 

### 2025-09-06 R/exploratory.R

@@ -11,7 +11,7 @@ data_expl <- df_merged %>%
        mutate(treatment = paste(cort, temp, sep = "-"))
 #
 # A) Make histograms for the main variables and figures factorizing main predictors
-var <- c("t_D", "mean_mitodensity", "mean_potential", "mean_ros", "mean_dnadamage", "mean_peroxidation")
+var <- c("t_D", "mean_mitodensity", "mean_potential", "mean_conpotential", "mean_ros", "mean_conros", "mean_dnadamage", "mean_peroxidation")
 
 hist_list <- list()
 for (i in var) {
@@ -29,6 +29,10 @@ for (i in var) {
      lab <- "DNA damage"
   } else if (i == "mean_peroxidation") {
      lab <- "Lipid peroxidation"
+  } else if (i == "mean_conpotential") {
+    lab <- "Mit potential/Mit density"
+  } else if (i == "mean_conros"){
+    lab <- "ROS/Mit density"
   }
   # Histograms
   hist_plot <- ggplot(data_expl, aes(x = .data[[i]])) +
@@ -74,17 +78,23 @@ for (i in var) {
             axis.text = element_text(size = 10, family = "Times"))
     fig_expl_t_D <- plot_grid(plot_age, plot_sex, plot_prey, nrow = 1, rel_widths = c(0.34, 0.33, 0.33))
   } else {
-    if (i == "mean_mitodensity") {
+  if (i == "t_D"){
+     lab <- "Detection latency (s)"
+  } else if (i == "mean_mitodensity") {
      lab <- "Mitochondrial density"
-    } else if (i == "mean_potential") {
+  } else if (i == "mean_potential") {
      lab <- "Mitochondrial potential"
-    } else if (i == "mean_ros") {
+  } else if (i == "mean_ros") {
      lab <- "ROS production"
-    } else if (i == "mean_dnadamage") {
+  } else if (i == "mean_dnadamage") {
      lab <- "DNA damage"
-    } else if (i == "mean_peroxidation") {
+  } else if (i == "mean_peroxidation") {
      lab <- "Lipid peroxidation"
-    }
+  } else if (i == "mean_conpotential") {
+    lab <- "Mit potential/Mit density"
+  } else if (i == "mean_conros"){
+    lab <- "ROS/Mit density"
+  }
     plot_age <- ggplot(data_expl, aes(x = age_trial, y = .data[[i]], color = treatment)) +
[…]

### 2025-09-06 R/func.r

@@ -172,7 +172,9 @@ tidy_post <- function(df) {
 refine_post <- function(df) {
   variable_order <- c("Mit density",
                      "Mit potential",
+                     "Mit potential/Mit density",
                      "ROS",
+                     "ROS/Mit density",
                      "DNA damage",
                      "Peroxidation",
                      "Detection lat")
@@ -189,13 +191,17 @@ refine_post <- function(df) {
       Model = case_when(
         Model == "m_def_mean_mitodensity_OB" ~ "Mit density",
         Model == "m_def_mean_potential_OB" ~ "Mit potential",
+        Model == "m_def_mean_conpotential_OB" ~ "Mit potential/Mit density",
         Model == "m_def_mean_ros_OB" ~ "ROS",
+        Model == "m_def_mean_conros_OB" ~ "ROS/Mit density",
         Model == "m_def_mean_dnadamage_OB" ~ "DNA damage",
         Model == "m_def_mean_peroxidation_OB" ~ "Peroxidation",
         Model == "m_def_t_D_Chemical" ~ "Detection lat",
         Model == "m_def_mean_mitodensity_OT" ~ "Mit density",
         Model == "m_def_mean_potential_OT" ~ "Mit potential",
+        Model == "m_def_mean_conpotential_OT" ~ "Mit potential/Mit density",
         Model == "m_def_mean_ros_OT" ~ "ROS",
+        Model == "m_def_mean_conros_OT" ~ "ROS/Mit density",
         Model == "m_def_mean_dnadamage_OT" ~ "DNA damage",
         Model == "m_def_mean_peroxidation_OT" ~ "Peroxidation",
         Model == "m_def_t_D_Visual" ~ "Detection lat",

### 2025-09-07 ms.qmd
updates

@@
 Mitochondrial physiology is thought to be a central mechanism underlying these effects.
-Here, we investigated the combined influence of prenatal corticosterone (CORT) — a key stress hormone in reptiles — and incubation temperature on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
+Here, we investigated the combined influence of prenatal corticosterone (CORT)-a key stress hormone in reptiles-and incubation temperature on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
 We manipulated egg CORT levels and incubation temperature, then assessed the ability of lizards to detect chemical and visual prey stimuli.
-Using flow cytometry, we measured metabolic function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta — two brain regions involved in sensory processing.
+Using flow cytometry, we measured metabolic function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta-two brain regions involved in sensory processing.
 We predicted that CORT exposure and high temperatures would interact to shape perception of prey via changes in mitochondrial respiration and oxidative damage.
@@
 For instance, Hara et al.
-(2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria — associated with higher oxidative stress — per presynaptic terminal.
+(2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria-associated with higher oxidative stress-per presynaptic terminal.
 In conclusion, findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
@@
 ## Discussion
-We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function was the underlying reason for any perceptual differences we observed as has been shown in mammals [@hara_presynaptic_2014; @chaudhari2022early].
-Our findings suggest that metabolic processes remained largely resilient to early environmental conditions.
-However, early-life conditions have region- and context-dependent effects on oxidative damage and prey detection.
-We observed interactive effects of prenatal CORT and incubation temperature on oxidative damage in the Optic Tecta, but no effects on perception of visual stimuli.
-In contrast, CORT decreased the time to perceive chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions.
-Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioural performance.
+We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function underpinned any perceptual differences, as previously shown in mammals [@hara_presynaptic_2014; @chaudhari2022early].
+Overall, metabolic processes appeared largely resilient to prenatal environment.
+However, early-life conditions produced region- and context-dependent effects on oxidative damage and prey detection.
+We observed interactive effects of prenatal CORT and incubation temperature on lipid proxidation in both regions, while the interaction affected DNA damage only in the optic tecta.
+Behaviourally, CORT enhanced detection of chemical-but not visual-prey stimuli.
+Yet, we found no evidence that metabolic explained behavioural variation.
+Below, we discuss these findings in the context of mitochondrial metabolic function, oxidative stress, and perception.
 
@@
 Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
-In the leopard gecko (_Eublepharis macularius_), incubation temperature is an important determinant of metabolic capacity in some, but not all, brain nuclei [@coomber_independent_1997], while prenatal stress decreases mitochondrial respiration efficiency in the cortex, hippocampus, and hypotalamus of rats (_Rattus norvegicus_) [@gong_chronic_2011].
-In contrast, metabolic processes in the optic tecta and olfactory bulbs of _L.
-delicata_ appear to be robust to early-life conditions, potentially to ensure that these brain regions can perform their essential functions.
-However, since the effects of early environment can be region-dependent [@coomber_independent_1997], it is possible that other brain regions not examined here may be more sensitive to prenatal CORT and temperature.
+These results were consistent when the effect on the average mitochondrion was analysed (i.e.
+when metabolic capacity and ROS production were divided by mitochondrial density).
+Our findings suggest that metabolic function in _L.
+delicata_'s brain is relatively resilient to early-life conditions.
+However, evidence from other species indicates that the effects of CORT and temperature can be region-specific: incubation temperature alters metabolic capacity in some, but not all, brain nuclei of leopard geckos (_Eublepharis macularius_) [@coomber_independent_1997], and prenatal stress reduces mitochondrial efficiency in select nuclei of rats (_Rattus norvegicus_) [@gong_chronic_2011].
+Thus, while our results point to robustness in the regions studied, other brain areas in _L.
+delicata_ may be more sensitive to prenatal CORT and temperature.
 
 Alternatively, the effects of early environmental conditions on metabolic function and ROS production may be immediate rather than long-lasting.
-Mitochondria are known to be highly responsive to environmental changes and energetic demands [@gyllenhammer2020developmental].
-It is possible that the metabolic effect of the early-life conditions dissipated once all animals were exposed to the same environment after hatching.
-Nonetheless, alterations in metabolic function during earlier stages would still be expected to affect oxidative damage, as these processes are cumulative and can lead to long-term consequences [@rice2002brain; @terman2006oxidative].
+Mitochondria are highly responsive to environmental changes and energetic demands [@gyllenhammer2020developmental], and any alterations present during incubation may have dissipated once all animals developed under the same post-hatching environment.
+Nonetheless, even temporary changes in metabolic function during earlier stages would still be expected to affect oxidative damage, as these processes are cumulative and can lead to long-term consequences [@rice2002brain; @terman2006oxidative].
 
@@
 We found that DNA damage and lipid peroxidation were affected in region-specific ways.
-Early conditions did not significantly influence oxidative stress in the olfactory bulbs, but had a pronounced effect on DNA damage and lipid peroxidation in the optic tecta.
-
-CORT exposure reduced DNA damage in optic tecta from cold-incubated lizards but had no effect at higher incubation temperatures.
-This pattern suggests that CORT might play a protective role under specific thermal conditions.
-Since CORT helps mobilize energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000], it is possible that under colder incubation conditions — where metabolic rates are expected to be lower — energy mobilised by CORT may be allocated towards antioxidant activity or DNA repairing mechanisms [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
-In contrast, at higher temperatures, the increased metabolic demands may limit the availability of resources for such protective mechanisms [see @kim2019carry].
-
-In comparison, increased prenatal CORT resulted in higher lipid peroxidation in cold incubated lizards but reduced it in those incubated at warm temperatures.
-Although the two hormonal treatments did not significantly differ within each incubation temperature, the direction of the CORT effect was the opposite between temperatures.
-These contrasting CORT–temperature effects on DNA damage and lipid peroxidation suggest that the pathways protecting or repairing each molecule may vary in their thermal sensitivity.
-For instance, antioxidant enzymes differ in their response to hibernation in the soft-shelled turtle (_Pelodiscus sinensis_), indicating that thermal sensitivity varies among antioxidant system components [@tang2021antioxidant].
-An alternative mechanism might involve temperature-dependent changes in membrane lipid composition or fluidity, which may affect how lipids respond to oxidative stress [@zeis2019temperature].
-For example, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids [@zeis2019temperature].
-Therefore, temperature-induced structural differences may influence the vulnerability of different molecules to oxidative stress.
+Early conditions did not significantly influence DNA damage in the olfactory bulbs, but they did alter lipid peroxidation in both brain nuclei studied.
+In the optic tecta, prenatal CORT reduced DNA damage in cold-incubated lizards but had no effect at higher incubation temperatures.
+In both regions, CORT also interacted with incubation temperature to influence lipid peroxidation, increasing it in CORT-treated cold-incubated lizards compared to the CORT-Hot group.
+
+These contrasting patterns suggest that prenatal CORT and temperature may shape oxidative damage through molecule-specific protective pathways, potentially by upregulating different antioxidant systems.
+Glucocorticoids and temperature each appear to influence antioxidant systems in enzyme-specific ways rather than uniformly across pathways.
+For example, prenatal stress in rats increases Cu/Zn superoxide dismutase (SOD) and glutathione peroxidase (GPx) activity in the brain while catalase (CAT) was unaffected [@mcintosh1998glucocorticoids].
+Similarly, in the soft-shelled turtle (_Pelodiscus sinensis_), CAT expression rises during arousal compared to hibernation, whereas GPx activity remains stable across these temperature-associated states [@tang2021antioxidant].
+Such differential regulation could explain why CORT reduces DNA damage under cold incubation but reduces lipid peroxidation under warm incubation, if distinct antioxidant pathways are preferentially activated under different environmental conditions.
+
+Alternative non-exclusive mechanisms could also contribute.
+CORT mobilises energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000].
+Under cold incubation-when metabolic rates are expected to be lower—energy may be allocated towards antioxidant activity or DNA repair [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
+In contrast, at warmer temperatures, increased metabolic demands may limit the resources available for such protective processes [see @kim2019carry].
+These context-dependent changes may explain lower DNA damage in the optic tecta of CORT-Cold lizards, but they do not fully account for the reduced lipid peroxidation in CORT-Hot animals.
+Additional factors, such as temperature-dependent differences in membrane lipid composition [@zeis2019temperature], may further modulate molecule-specific sensitivity to oxidative stress.
+For example, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids, which are more susceptible to oxidative damage [@zeis2019temperature].
+Future studies should examine the effects of prenatal CORT and temperature on metabolic rate, antioxidant regulation, and molecule structure and composition in differen brain nuclei of reptiles.
 
@@
 ### Declaration of AI use
-We declare Chat GPT was used mainly for questions related to coding and data analyses.
+We declare Chat GPT was used for checking coding errors.
 All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
@@
 
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Results of preliminary models

### 2025-09-17 ms.qmd
Update ms.qmd

@@
 However, early-life conditions like stress-related hormones or thermal environments, can shape cognitive abilities by influencing brain structure and function.
-Mitochondrial physiology is thought to be a central mechanism underlying these effects.
-Here, we investigated the combined influence of prenatal corticosterone (CORT)-a key stress hormone in reptiles-and incubation temperature on mitochondrial function and prey detection in the delicate skink (_Lampropholis delicata_).
+Mitochondrial physiology is thought to be a central mechanism underlying these effects given that it is crucial for the energy production needed for brain function.
+Here, we investigated the combined influence of prenatal corticosterone (CORT) – a key stress hormone in reptiles – and incubation temperature on brain mitochondrial function and impacts on prey detection in the delicate skink (_Lampropholis delicata_).
 We manipulated egg CORT levels and incubation temperature, then assessed the ability of lizards to detect chemical and visual prey stimuli.
@@
 However, the impact of early-life environments has primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other equally important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
-In particular, perception is an often overlooked, but fundamental skill [@shettleworth].
+Perception is an often overlooked, but fundamental skill [@shettleworth].
 Perceiving different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
-Variations in the ability to detect or recognize relevant stimuli can have significant consequences for an individual’s fitness [@pyke1984optimal].
+Variation in the ability to detect or recognize relevant stimuli can have significant consequences for fitness [@pyke1984optimal].
 For example, visual acuity in the bluegill sunfish (_Lepomis macrochirus_) allow animals to detect prey at large distances, which can increase their foraging efficiency and energy intake [@hairston1982fish].
@@
 
-Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental inputs during early stages of development [@zhu_prenatal_2004].
+Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental insults during early stages of development [@zhu_prenatal_2004].
 Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production [@amani2021perinatal], or brain structure and function [@amiel_effects_2017].
@@
 (2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria-associated with higher oxidative stress-per presynaptic terminal.
-In conclusion, findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
+Findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
 Thus, understanding how early-life conditions affect mitochondrial physiology may provide insights into the mechanisms underlying cognitive performance.
@@
 Mitochondria are highly responsive to environmental conditions during early development, with potential short-term and long-lasting consequences [reviewed in @gyllenhammer2020developmental].
-For example, environmental cues such as external conditions or hormone exposure can cause immediate disruptions to metabolic function, triggering cascading effects on oxidative stress that may persist leter in life [reviewed in @gyllenhammer2020developmental].
+For example, environmental cues such as external conditions or hormone exposure can cause immediate disruptions to metabolic function, triggering cascading effects on oxidative stress that may persist later in life [reviewed in @gyllenhammer2020developmental].
 Among the range of environmental factors that can shape mitochondrial physiology, some exert particularly strong and consistent effects.
@@
 Similarly, temperature plays a prominent role in regulating metabolic activity in ectotherms [@stier2022experimental; @crino2024eggs].
-In fact, GC elevations or changes in the thermal environment during early development have profound and sustained effects on mitochondrial respiration efficiency or oxidative stress [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs].
-However, how the early environment influences mitochondrial physiology and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
-In particular, whether and how these interactions affect perceptual abilities, such as prey detection, is still largely unexplored.
+In fact, heightened GC levels or changes in the thermal environment during early development have profound and sustained effects on mitochondrial respiration efficiency or oxidative stress [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs].
+However, how the early environment influences mitochondrial physiology, and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
+In addition, whether and how these interactions affect perceptual abilities, such as prey detection, is still largely unexplored.
 
@@
 
-_Olfactory bulbs_: age significantly increased DNA damage and lipid peroxidation (Table S3).
+_Olfactory bulbs_: Age significantly increased DNA damage and lipid peroxidation (Table S3).
 However, we did not find any significant effects of CORT, temperature on mitochondrial function nor oxidative stress in the olfactory bulbs (see @fig-results_energy, @fig-results_oxidative, and Table S2).
@@
 
-_Optic tecta_: we did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
+_Optic tecta_: We did not find significant effects of CORT, temperature, or their interaction on mitochondrial density, metabolic capacity, or ROS production in the optic tecta, neither of CORT or incubation temperature on DNA damage or lipid peroxidation (see @fig-results_energy, @fig-results_oxidative, and Table S2).
 However, the interaction of CORT and temperature have significant effects on DNA damage ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(DNA_OT$Control_Hot) - mean(DNA_OT$CORT_Hot)) - (mean(DNA_OT$Control_Cold) - mean(DNA_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((DNA_OT$Control_Hot - DNA_OT$CORT_Hot) - (DNA_OT$Control_Cold - DNA_OT$CORT_Cold)), 3, equal = TRUE)`), and lipid peroxidation ([(β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)]: mean = `r format_dec((mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot)) - (mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold)), 3)`, pMCMC `r format_p(pmcmc((LP_OT$Control_Hot - LP_OT$CORT_Hot) - (LP_OT$Control_Cold - LP_OT$CORT_Cold)), 3, equal = TRUE)`) in the optic tecta.
@@
 
-_Chemical stimulus_: lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature or the interaction between CORT and temperature (@fig-results_behaviour A and Table S2).
+_Chemical stimulus_: Lizards detected chemical stimulus faster when exposed to prenatal CORT (β~Control~ - β~CORT~: mean = `r format_dec(mean(c(DET_OB$Control_Cold, DET_OB$Control_Hot)) - mean(c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3)`, pMCMC `r format_p(pmcmc(c(DET_OB$Control_Cold, DET_OB$Control_Hot) - c(DET_OB$CORT_Cold, DET_OB$CORT_Hot)), 3, equal = TRUE)`), but there was no effect of temperature or the interaction between CORT and temperature (@fig-results_behaviour A and Table S2).
 
@@
 
-_Visual stimulus_: there were no significant effects of CORT, temperature, or their interaction on the detection latency of visual stimuli (@fig-results_behaviour B and Table S2).
+_Visual stimulus_: There were no significant effects of CORT, temperature, or their interaction on the detection latency of visual stimuli (@fig-results_behaviour B and Table S2).
 
@@
 
-_Chemical perception (Olfactory bulbs)_: we found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
+_Chemical perception (Olfactory bulbs)_: We found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
 However, there were no significant relationships between mitochondrial function or oxidative stress on prey detection, nor between other any other variables (see @fig-sem_results_OB and Table S5).
@@
 
-_Visual perception (Optic tecta)_: we found no significant relationships within mitochondrial variables or between mitochondrial physiology and cognitive performance (see @fig-sem_results_OT and Table S6).
+_Visual perception (Optic tecta)_: We found no significant relationships within mitochondrial variables or between mitochondrial physiology and cognitive performance (see @fig-sem_results_OT and Table S6).
 
@@
 We observed interactive effects of prenatal CORT and incubation temperature on lipid proxidation in both regions, while the interaction affected DNA damage only in the optic tecta.
-Behaviourally, CORT enhanced detection of chemical-but not visual-prey stimuli.
-Yet, we found no evidence that metabolic explained behavioural variation.
+Behaviourally, CORT enhanced detection of chemical but not visual prey stimuli.
+Yet, we found no evidence that metabolism explained behavioural variation.
 Below, we discuss these findings in the context of mitochondrial metabolic function, oxidative stress, and perception.
@@
 Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
-These results were consistent when the effect on the average mitochondrion was analysed (i.e.
-when metabolic capacity and ROS production were divided by mitochondrial density).
-Our findings suggest that metabolic function in _L.
-delicata_'s brain is relatively resilient to early-life conditions.
-However, evidence from other species indicates that the effects of CORT and temperature can be region-specific: incubation temperature alters metabolic capacity in some, but not all, brain nuclei of leopard geckos (_Eublepharis macularius_) [@coomber_independent_1997], and prenatal stress reduces mitochondrial efficiency in select nuclei of rats (_Rattus norvegicus_) [@gong_chronic_2011].
+These results were consistent when metabolic capacity and ROS production were standardised by mitochondrial density.
+Our findings suggest that metabolic function in the brain of _L.
+delicata_ is relatively resilient to early-life conditions.
+While we did not observe effects in the optic tecta and olfactory bulbs, evidence from other species indicates that the effects of CORT and temperature can be region-specific: incubation temperature alters metabolic capacity in some, but not all, brain nuclei of leopard geckos (_Eublepharis macularius_) [@coomber_independent_1997], and prenatal stress reduces mitochondrial efficiency in select nuclei of rats (_Rattus norvegicus_) [@gong_chronic_2011].
 Thus, while our results point to robustness in the regions studied, other brain areas in _L.
@@
 CORT mobilises energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000].
-Under cold incubation-when metabolic rates are expected to be lower—energy may be allocated towards antioxidant activity or DNA repair [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
+Under cold incubation, when metabolic rates are expected to be lower, energy may be allocated towards antioxidant activity or DNA repair [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
 In contrast, at warmer temperatures, increased metabolic demands may limit the resources available for such protective processes [see @kim2019carry].
@@
 Additional factors, such as temperature-dependent differences in membrane lipid composition [@zeis2019temperature], may further modulate molecule-specific sensitivity to oxidative stress.
-For example, lipid peroxidation levels increased in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids, which are more susceptible to oxidative damage [@zeis2019temperature].
-Future studies should examine the effects of prenatal CORT and temperature on metabolic rate, antioxidant regulation, and molecule structure and composition in differen brain nuclei of reptiles.
+For example, lipid peroxidation levels have been shown to increase in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids, which are more susceptible to oxidative damage [@zeis2019temperature].
+Future studies should examine the effects of prenatal CORT and temperature on metabolic rate, antioxidant regulation, and molecule structure and composition in differen brain nuclei of reptiles to distinguish between these hypotheses.
 
@@
 Similarly, oxidative stress did not influence _L.
-delicata_ ability to detect visual or chemical stimuli from prey.
-In this regard, oxidative damage is expected to decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+delicata_ ability to detect visual or chemical stimuli from prey despite finding that oxidative damage varied across brain regions.
+While oxidative damage is expected to decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018], our Structural Equation Models showed no relationship between lipid peroxidation or DNA damage and prey perception abilities.
 Possibly, the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms - such as antioxidant activity - that counteracted the negative effects of oxidative stress, or because the studied brain regions had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].

### 2025-10-12 ms.qmd
updates

@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
+Pablo Recio^1,2^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
 
@@
 Cognitive processes such as the ability to perceive prey, are crucial for survival and reproduction.
-However, early-life conditions like stress-related hormones or thermal environments, can shape cognitive abilities by influencing brain structure and function.
+Early-life conditions like stress-related hormones or thermal environments, can shape cognitive abilities by influencing brain structure and function.
 Mitochondrial physiology is thought to be a central mechanism underlying these effects given that it is crucial for the energy production needed for brain function.
-Here, we investigated the combined influence of prenatal corticosterone (CORT) – a key stress hormone in reptiles – and incubation temperature on brain mitochondrial function and impacts on prey detection in the delicate skink (_Lampropholis delicata_).
+Here, we investigated the combined influence of prenatal corticosterone (CORT)–a key stress hormone in reptiles–and incubation temperature on brain mitochondrial function and impacts on prey detection in the delicate skink (_Lampropholis delicata_).
 We manipulated egg CORT levels and incubation temperature, then assessed the ability of lizards to detect chemical and visual prey stimuli.
@@
 We predicted that CORT exposure and high temperatures would interact to shape perception of prey via changes in mitochondrial respiration and oxidative damage.
-While metabolic function remained robust to early life conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta but not the olfactory bulbs.
+While metabolic function remained robust to early life conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta.
 Additionally, CORT-treated lizards responded faster when presented with chemical cues but not visual stimuli.
 Our findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
-Future research should explore alternative mechanisms linking early-life stress, cell metabolism, and cognitive performance.
+Future research should explore alternative mechanisms that could link early-life stress to cell metabolism and cognitive performance.
 
@@
 The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017].
-However, the impact of early-life environments has primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other equally important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
-Perception is an often overlooked, but fundamental skill [@shettleworth].
-Perceiving different stimuli is essential for locating food, avoiding predators, or interacting with conspecifics [@burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
-Variation in the ability to detect or recognize relevant stimuli can have significant consequences for fitness [@pyke1984optimal].
-For example, visual acuity in the bluegill sunfish (_Lepomis macrochirus_) allow animals to detect prey at large distances, which can increase their foraging efficiency and energy intake [@hairston1982fish].
+However, the impact of early-life environments has primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
+For example, perception is an often overlooked, but fundamental cognitive process [@shettleworth].
+Perceiving different stimuli is essential for behaviours critical to survival and reproductive succes such as locating food, avoiding predators, or interacting with conspecifics [@hairston1982fish; @pyke1984optimal; @burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
 Therefore, understanding how prenatal conditions affect the ability to detect prey can be important to understanding the broader consequences of early-life environments on fitness.
 
-Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental insults during early stages of development [@zhu_prenatal_2004].
-Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production [@amani2021perinatal], or brain structure and function [@amiel_effects_2017].
-Long-lasting effects on energy production and oxidative stress in the brain may be a key mechanism by which cognitive function is impacted by developmental environments [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
-Given the high energetic demands of cognitive processes [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], cognitive performance is expected to depend on efficient mitochondrial respiration.
-Additionally, high cognitive abilities are typically associated with increased neuron density and functionality [@lefebvre_taxonomic_2011; @amiel_effects_2017], both of which can be impaired by excessive reactive oxygen species (ROS) production and oxidative stress [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental conditions during early stages of development [@zhu_prenatal_2004].
+Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production and neuroendocrine pathways [@amani2021perinatal], or brain structure and function [@amiel_effects_2017].
+Additionally, developmental environments can have long-lasting effects on energy production and oxidative stress in the brain, which may be a key mechanism influencing cognitive function [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+First, given the high energetic demands of cognitive processes [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], cognitive performance is intrinsically linked to efficient mitochondrial respiration.
+Likewise, high cognitive abilities are typically associated with increased neuron density and functionality [@lefebvre_taxonomic_2011; @amiel_effects_2017], both of which can be impaired by excessive reactive oxygen species (ROS) production and oxidative stress [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
 For instance, Hara et al.
-(2014) found that performance in visuospatial working-memory tasks correlate negatively with the number of round-shaped mitochondria-associated with higher oxidative stress-per presynaptic terminal.
+(2014) found that performance in visuospatial working-memory tasks is negatively correlated to the number of round-shaped mitochondria-associated with higher oxidative stress-per presynaptic terminal.
 Findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
@@
 Mitochondria are highly responsive to environmental conditions during early development, with potential short-term and long-lasting consequences [reviewed in @gyllenhammer2020developmental].
-For example, environmental cues such as external conditions or hormone exposure can cause immediate disruptions to metabolic function, triggering cascading effects on oxidative stress that may persist later in life [reviewed in @gyllenhammer2020developmental].
+For example, environmental conditions or hormone exposure can cause immediate disruptions to metabolic function, triggering cascading effects on oxidative stress that may persist later in life [reviewed in @gyllenhammer2020developmental].
 Among the range of environmental factors that can shape mitochondrial physiology, some exert particularly strong and consistent effects.
 Stress-related hormones and temperature are particularly relevant for mitochondrial function due to their strong effects on metabolic processes [@sapolsky_how_2000; @stier2022experimental; @crino2024eggs].
-Under stressful situations, animals instigate adaptive physiological and behavioural adjustments that involve increases in glucocorticoids (GCs), which are also important metabolic regulators [@sapolsky_how_2000].
-Similarly, temperature plays a prominent role in regulating metabolic activity in ectotherms [@stier2022experimental; @crino2024eggs].
+Vertebrates respond to stressful situations through the activation of physiological processes including the production of  glucocorticoids (GCs), which are also important metabolic regulators [@sapolsky_how_2000].
+In endothermic vertebrates, temperature plays a similar role in regulating metabolic activity [@stier2022experimental; @crino2024eggs].
 In fact, heightened GC levels or changes in the thermal environment during early development have profound and sustained effects on mitochondrial respiration efficiency or oxidative stress [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs].
@@
 #### Statistical analyses
-We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, metabolic capatity/mitochondrial density, ROS, ROS/mitochondrial density, DNA damage, lipid peroxidation, and detection latency) separately.
+We performed the analyses for each brain region/stimulus and each variable (mitochondrial density, metabolic capacity, metabolic capacity/mitochondrial density, ROS, ROS/mitochondrial density, DNA damage, lipid peroxidation, and detection latency) separately.
 We first fit a set of preliminary models for all the interest variables where we included the hormone (CORT versus Control), temperature (Cold versus Hot), and their interaction.
@@
 The direction of the effects of CORT on lipid peroxidation were the opposite on eggs incubated at high temperatures (β~Control-Hot~ - β~CORT-Hot~: mean = `r format_dec(mean(LP_OT$Control_Hot) - mean(LP_OT$CORT_Hot),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Hot - LP_OT$CORT_Hot), 3, equal = TRUE)`) than at low temperatures (β~Control-Cold~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$Control_Cold) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$Control_Cold - LP_OT$CORT_Cold), 3, equal = TRUE)`), with significant differences between CORT-Hot and CORT-Cold lizards (β~CORT-Hot~ - β~CORT-Cold~: mean = `r format_dec(mean(LP_OT$CORT_Hot) - mean(LP_OT$CORT_Cold),3)`, pMCMC `r format_p(pmcmc(LP_OT$CORT_Hot - LP_OT$CORT_Cold), 3, equal = TRUE)`; see @fig-results_oxidative C).
-In addition, we found that females had higher DNA damage levels, and lipid peroxidation increased with age (see Table S4).
+In addition, we found that females had higher DNA damage levels than males, and lipid peroxidation increased with age (see Table S4).
 
@@
 
-_Chemical perception (Olfactory bulbs)_: We found a significant effect of mitochondrial density on ROS production (mean β~Mit density - ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
+_Chemical perception (Olfactory bulbs)_: We found a positive effect of mitochondrial density on ROS production (mean β~Mit density~ ~-~ ~ROS~ = `r format_dec(mean(OB_coeff_mitodensity_ros), 3)`, pMCMC `r format_p(pmcmc(OB_coeff_mitodensity_ros), 3, equal = TRUE)`).
 However, there were no significant relationships between mitochondrial function or oxidative stress on prey detection, nor between other any other variables (see @fig-sem_results_OB and Table S5).
@@
 ## Discussion
-We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function underpinned any perceptual differences, as previously shown in mammals [@hara_presynaptic_2014; @chaudhari2022early].
-Overall, metabolic processes appeared largely resilient to prenatal environment.
+We examined the impact of prenatal temperature and CORT on visual and chemical perception to test whether metabolic function underpinned perception in lizards, as previously shown in mammals [@hara_presynaptic_2014; @chaudhari2022early].
+Overall, metabolic processes appeared largely resilient to prenatal conditions.
 However, early-life conditions produced region- and context-dependent effects on oxidative damage and prey detection.
-We observed interactive effects of prenatal CORT and incubation temperature on lipid proxidation in both regions, while the interaction affected DNA damage only in the optic tecta.
-Behaviourally, CORT enhanced detection of chemical but not visual prey stimuli.
-Yet, we found no evidence that metabolism explained behavioural variation.
-Below, we discuss these findings in the context of mitochondrial metabolic function, oxidative stress, and perception.
+We observed interactive effects of prenatal CORT and incubation temperature on lipid peroxidation in the two brain regions examined, while the interaction affected DNA damage only in the optic tecta.
+Behaviourally, prenatal CORT exposure enhanced detection of chemical but not visual prey stimuli.
+Yet, we found no evidence that metabolism explained behavioural variation in prey detection behaviours.
 
 #### Metabolic processes in multiple brain regions are robust to early life conditions
-Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in either of the brain regions studied.
+Contrary to our predictions, we found no significant effects of early environmental conditions on mitochondrial density, metabolic capacity, or ROS production in the optic tecta or the olfactory bulbs.
 These results were consistent when metabolic capacity and ROS production were standardised by mitochondrial density.
@@
 delicata_ is relatively resilient to early-life conditions.
-While we did not observe effects in the optic tecta and olfactory bulbs, evidence from other species indicates that the effects of CORT and temperature can be region-specific: incubation temperature alters metabolic capacity in some, but not all, brain nuclei of leopard geckos (_Eublepharis macularius_) [@coomber_independent_1997], and prenatal stress reduces mitochondrial efficiency in select nuclei of rats (_Rattus norvegicus_) [@gong_chronic_2011].
+While we did not observe effects in the optic tecta and olfactory bulbs, evidence from other species indicates that the effects of prenatal CORT and temperature can be region-specific.
+For example, incubation temperature alters metabolic capacity in some, but not all, brain nuclei of leopard geckos (_Eublepharis macularius_) [@coomber_independent_1997], and prenatal stress reduces mitochondrial efficiency in select nuclei of rats (_Rattus norvegicus_) [@gong_chronic_2011].
 Thus, while our results point to robustness in the regions studied, other brain areas in _L.
@@
 
-Alternatively, the effects of early environmental conditions on metabolic function and ROS production may be immediate rather than long-lasting.
+Alternatively, the effects of early environmental conditions on metabolic function and ROS production may be present during incubation, but diminish or dissapear over time.
 Mitochondria are highly responsive to environmental changes and energetic demands [@gyllenhammer2020developmental], and any alterations present during incubation may have dissipated once all animals developed under the same post-hatching environment.
+In our experiment, lizards were housed in standardized conditions prior to euthanasia at 238-356 days post-hatching (mean = 273.25).
+It is possible that the effects of prenatal CORT and temperature exposure had lessened by the time metabolic function and ROS production were measured.
 Nonetheless, even temporary changes in metabolic function during earlier stages would still be expected to affect oxidative damage, as these processes are cumulative and can lead to long-term consequences [@rice2002brain; @terman2006oxidative].
@@
 We found that DNA damage and lipid peroxidation were affected in region-specific ways.
-Early conditions did not significantly influence DNA damage in the olfactory bulbs, but they did alter lipid peroxidation in both brain nuclei studied.
-In the optic tecta, prenatal CORT reduced DNA damage in cold-incubated lizards but had no effect at higher incubation temperatures.
-In both regions, CORT also interacted with incubation temperature to influence lipid peroxidation, increasing it in CORT-treated cold-incubated lizards compared to the CORT-Hot group.
-
+Early conditions did not significantly influence DNA damage in the olfactory bulbs, but they did alter lipid peroxidation in the olfactory bulbs and the optic tecta.
+In the optic tecta, prenatal CORT reduced DNA damage in cold-incubated lizards but had no effect in hot-incubated animals.
+In both regions, CORT also interacted with incubation temperature to influence lipid peroxidation.
+Lizards treated with prenatal CORT and incubated at cold temperatures had higher levels of lipid peroxidation compared to the lizards from the CORT-Hot group.
 These contrasting patterns suggest that prenatal CORT and temperature may shape oxidative damage through molecule-specific protective pathways, potentially by upregulating different antioxidant systems.
-Glucocorticoids and temperature each appear to influence antioxidant systems in enzyme-specific ways rather than uniformly across pathways.
 For example, prenatal stress in rats increases Cu/Zn superoxide dismutase (SOD) and glutathione peroxidase (GPx) activity in the brain while catalase (CAT) was unaffected [@mcintosh1998glucocorticoids].
-Similarly, in the soft-shelled turtle (_Pelodiscus sinensis_), CAT expression rises during arousal compared to hibernation, whereas GPx activity remains stable across these temperature-associated states [@tang2021antioxidant].
-Such differential regulation could explain why CORT reduces DNA damage under cold incubation but reduces lipid peroxidation under warm incubation, if distinct antioxidant pathways are preferentially activated under different environmental conditions.
-
-Alternative non-exclusive mechanisms could also contribute.
-CORT mobilises energy resources to facilitate physiological adjustments to stress [@sapolsky_how_2000].
+Similarly, in the soft-shelled turtle (_Pelodiscus sinensis_), CAT expression elevates with increasing temperatures as turtles emerge from hibernation, whereas GPx activity remains stable [@tang2021antioxidant].
+Such differential regulation of antioxidant systems could explain why CORT reduces DNA damage under cold incubation but reduces lipid peroxidation under hot incubation.
+
+However, other mechanisms could be contributing to our results.
+For example, CORT mobilises energy resources to facilitate physiological adjustments to stress and environmental challenges [@sapolsky_how_2000].
 Under cold incubation, when metabolic rates are expected to be lower, energy may be allocated towards antioxidant activity or DNA repair [@kawamura1991glucocorticoid; but see @mcintosh1998glucocorticoids].
@@
 For example, lipid peroxidation levels have been shown to increase in cold-acclimated _Daphnia magna_ likely because of higher concentration of polyunsaturated fatty acids, which are more susceptible to oxidative damage [@zeis2019temperature].
-Future studies should examine the effects of prenatal CORT and temperature on metabolic rate, antioxidant regulation, and molecule structure and composition in differen brain nuclei of reptiles to distinguish between these hypotheses.
+Future studies should examine the effects of prenatal CORT and temperature on metabolic rate, antioxidant regulation, and molecule structure and composition in differen brain nuclei of reptiles to parse between these hypotheses.
 
@@
 This effect was not related to alterations in mitochondrial physiology in the olfactory bulbs.
-Instead, prenatal CORT may influence other mechanisms involved in chemoreception.
+Instead, prenatal CORT could influence other mechanisms involved in chemoreception.
 For instance, in rats (_R.
-norvegicus_), GCs like dexamethasone can enchance the ability to detect chemical stimuli by upregulating gene expression in the olfactory mucosa [@meunier2020olfactory].
-Such molecular changes could alter chemosensory sensitivity in lizards, independent of mitochondrial function.
-
-Alternatively, the observed differences in chemical perception may be driven by motivation.
-Prenatal CORT is expected to increase hunger through resource mobilization [@spencer2008post; @cossin-sevrin_effect_2022; but see @crino_corticosterone_2014-learn].
-This could have led to increased motivation in CORT-treated lizards, stimulating a switch to a more active foraging mode, affecting the way they respond to chemical stimuli.
-For example, species of European lizards (family Lacertidae) that spend more time actively looking for food tend to rely more on chemoreception [@verwaijen2007relationships].
-Although we fasted lizards and controlled for motivation in our analyses, it is possible that we did not fully capture factors related to the ecological relevance of the cues, the animals' internal state, or the interaction between the two.
+norvegicus_), dexamethsone (a synthetic GC) enhances the ability to detect chemical stimuli by upregulating gene expression in the olfactory mucosa [@meunier2020olfactory].
+If similar GC-dependent transcriptional changes occur in lizards,they could alter chemosensory sensitivity independent of mitochondrial function.
+
+Alternatively, the observed differences in chemical perception may be driven by treatment effects on motivation.
+We fasted lizards in our experiment for two days prior to behavioural assays to standardize hunger levels.
+However, the experimental treatments could have affected factors related to the animals’ internal state that we did not measure that influenced motivation.
+For example, prenatal CORT can increase hunger through resource mobilisation [@spencer2008post; @cossin-sevrin_effect_2022; but see @crino_corticosterone_2014-learn].
+This could have led to increased motivation in CORT-treated lizards decreasing their time to respond to prey chemical stimuli.
+Furthermore, the absence of CORT-related effects on visual perception could occur because actively foraging lizards often prioritize chemosensory cues [@verwaijen2007relationships], so changes in motivation may selectively affect chemical perception.
 Further studies need to be conducted to understand how CORT can change foraging strategies in lizards and how this can affect their perception abilities.
@@
 
-Similarly, oxidative stress did not influence _L.
-delicata_ ability to detect visual or chemical stimuli from prey despite finding that oxidative damage varied across brain regions.
-While oxidative damage is expected to decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018], our Structural Equation Models showed no relationship between lipid peroxidation or DNA damage and prey perception abilities.
-Possibly, the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms - such as antioxidant activity - that counteracted the negative effects of oxidative stress, or because the studied brain regions had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].
-Alternatively, the cognitive effects of oxidative stress may not yet be detectable at this stage.
+Similarly, oxidative stress did not influence the ability of lizards to detect visual or chemical stimuli from prey despite finding that oxidative damage varied across brain regions.
+Oxidative damage can decrease cognitive abilities through neuron death or senescense [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+However, our Structural Equation Models showed no relationship between lipid peroxidation or DNA damage and prey perception abilities.
+Possibly, the levels of oxidative damage observed in our study were not sufficient to impair prey detection either because of compensatory mechanisms-such as antioxidant activity-that counteracted the negative effects of oxidative stress, or because the brain regions we studied had a high neural density that provided functional resilience [@rice2002brain; @amiel_smart_2011].
+Alternatively, the cognitive effects of oxidative stress may not be detectable at the age when we tested lizards.
 Oxidative damage accumulates over time, and its effects on cognition can pronounce with age [@terman2006oxidative; @hara_presynaptic_2014].
-As such, we could find a strong relationship if older individuals were employed, but not in lizards this age.
+As such, we could find strong relationships between oxidative damage and perception if older individuals were tested.
 In fact, we saw a significant effect of age on DNA damage in olfactory bulbs, and on lipid peroxidation in both regions.
@@
 #### Conclusions
-Prenatal incubation and CORT exposure had region-specific effects in the brain of _L.
+Prenatal CORT exposure and incubation temperature produced region-specific effects in the brain of _L.
 delicata_.
-While mitochondrial function remained resilient to early environmental conditions, oxidative damage and behavioural responses exhibited context-dependent variations.
-Our findings suggest that the brain regions studied are not directly related to the behaviours tested, or that other factors, like motivation, are more relevant for detecting prey stimuli.
+While mitochondrial function remained remarkably resilient to prenatal conditions, oxidative damage and behavioural responses varied depending on treatment and brain region.
+The interaction between prenatal CORT exposure and incubation temperature influenced DNA damage and lipid peroxidation in the olfactory bulbs and the optic tecta.
+However, these effects were not related to the perception of visual or chemical prey stimuli.
+Notably, chemical perception was affected by prenatal CORT alone, highlighting a selective sensitivity of perceptual abilities independent of measured brain physiology.
+Our findings suggest that differences in prey detection may result from treatment effects on other factors, like motivation.
 Future studies should explore how the environment influences mitochondrial function accross tissues, as well as the mechanisms underlying the brain’s resilience to early-life conditions.
@@
 ### Declaration of AI use
-We declare Chat GPT was used for checking coding errors.
-All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+We declare Chat GPT was used for inspecting the code for errors when necessary. All other parts of the manuscript were written by the authors.
 

### 2025-10-12 ms.qmd
Update ms.qmd

@@
 Using flow cytometry, we measured metabolic function, reactive oxygen species (ROS) production, and oxidative stress in the olfactory bulbs and optic tecta-two brain regions involved in sensory processing.
-We predicted that CORT exposure and high temperatures would interact to shape perception of prey via changes in mitochondrial respiration and oxidative damage.
-While metabolic function remained robust to early life conditions, CORT and temperature interacted to influence oxidative stress in the optic tecta.
+While metabolic function remained robust to early life conditions, CORT and temperature interacted to influence oxidative stress.
 Additionally, CORT-treated lizards responded faster when presented with chemical cues but not visual stimuli.
-Our findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, though the role of mitochondrial function in these processes remains unclear.
-Future research should explore alternative mechanisms that could link early-life stress to cell metabolism and cognitive performance.
+Our findings suggest that prenatal conditions can have lasting effects on brain physiology and cognition, highlighting a surprising decoupling between mitochondrial function and perception in lizards.
 

### 2025-10-13 ms.qmd
updates

@@
 It is essential for survival and reproduction, allowing individuals to adapt to changing environments [@dukas_evolutionary_2004].
-However, cognitive abilities can vary considerably between individuals, with differences arising from genetic factors, environmental conditions, or a combination of both [@sakata_neural_2000; @dukas_evolutionary_2004].
+However, cognitive abilities can vary considerably between individuals, with differences arising from genetic factors, environmental conditions, or a combination [@sakata_neural_2000; @dukas_evolutionary_2004].
 The prenatal environment, in particular, plays a critical role in shaping brain development and cognition across a wide range of species [@sakata_neural_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @amiel_effects_2017].
 However, the impact of early-life environments has primarily been studied in the context of learning and memory [@zhu_prenatal_2004; @crino_corticosterone_2014-learn; @bebus_associative_2016; @abayarathna_effects_2020], neglecting how other important cognitive domains might be affected [but see @burger_antipredator_1998; @vila_pouca_quantity_2019].
-For example, perception is an often overlooked, but fundamental cognitive process [@shettleworth].
-Perceiving different stimuli is essential for behaviours critical to survival and reproductive succes such as locating food, avoiding predators, or interacting with conspecifics [@hairston1982fish; @pyke1984optimal; @burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
-Therefore, understanding how prenatal conditions affect the ability to detect prey can be important to understanding the broader consequences of early-life environments on fitness.
-
-Prenatal environments influence cognitive abilities because the brain is particularly sensitive to environmental conditions during early stages of development [@zhu_prenatal_2004].
+In this vein, perception is an often overlooked, but fundamental cognitive process, involved in behaviours critical to survival and reproductive success such as locating food, avoiding predators, or interacting with conspecifics [@shettleworth; @hairston1982fish; @pyke1984optimal; @burger_antipredator_1998; @desfilis2003stimulus; @recio2023conspecific].
+Understanding how prenatal conditions affect the ability to detect different stimuli can be fundamental to understanding the broader consequences of early-life environments on fitness.
+
+Prenatal environments influence cognitive abilities because the brain is particularly sensitive during early stages of development [@zhu_prenatal_2004].
 Early-life conditions can shape cognition by altering gene expression [@zhou2020effects], neurotransmitter production and neuroendocrine pathways [@amani2021perinatal], or brain structure and function [@amiel_effects_2017].
@@
 (2014) found that performance in visuospatial working-memory tasks is negatively correlated to the number of round-shaped mitochondria-associated with higher oxidative stress-per presynaptic terminal.
-Findings from different studies suggest that mitocondrial activity can be essential in shaping cognitive abilities.
 Thus, understanding how early-life conditions affect mitochondrial physiology may provide insights into the mechanisms underlying cognitive performance.
@@
 Vertebrates respond to stressful situations through the activation of physiological processes including the production of  glucocorticoids (GCs), which are also important metabolic regulators [@sapolsky_how_2000].
-In endothermic vertebrates, temperature plays a similar role in regulating metabolic activity [@stier2022experimental; @crino2024eggs].
+In ectothermic vertebrates, temperature plays a similar role in regulating metabolic activity [@stier2022experimental; @crino2024eggs].
 In fact, heightened GC levels or changes in the thermal environment during early development have profound and sustained effects on mitochondrial respiration efficiency or oxidative stress [@zhu_prenatal_2004; @song2009prenatal; @costantini2011meta; @haussmann_embryonic_2012; @treidel2016temperature; @stier2022experimental; @crino2024eggs].
-However, how the early environment influences mitochondrial physiology, and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
+However, how the early environment influences mitochondrial physiology and its associated cognitive consequences remains largely unknown outside of mammals [see @chaudhari2022early for a review].
 In addition, whether and how these interactions affect perceptual abilities, such as prey detection, is still largely unexplored.
@@
 Crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperatures were set to 22-24 ºC, but to ensure a temperature gradient, we provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
+Room temperatures were set to 22-24 ºC, but enclosures were also provided with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
 
@@
 These boxes were checked three days a week for eggs.
-After collection, we measured the length and width of eggs with a digital caliper to the nearest 0.1 mm and weighed them with a digital scale (± 0.001 g).
-Eggs were treated with either CORT or a vehicle control (see CORT and temperature manipulation below) and placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+After collection, eggs were treated with either CORT or a vehicle control (see CORT and temperature manipulation below) and placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
 Cups were covered with cling wrap to retain moisture and left at one of two different incubation temperatures (see CORT and temperature manipulation below) until hatching.
@@
 *Hatchlings* -- Incubators were checked three times a week for hatchlings.
-We measured and weighted lizards immediately after hatching.
-We used a ruler to measure snout-vent length (SVL) and tail length (TL) to the nearest mm, and a digital scale to measure mass (± 0.001 g).
-We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with nonstick matting and a small water dish until the beginning of the experiment.
+Hatched lizards were placed in individual enclosures in individual enclosures (18.7L x 13.2W x 6.3H cm) provided with nonstick matting and a small water dish until the beginning of the experiment.
 During this period, lizards were given water daily and received 3-6 small *A.
@@
 C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), or b) an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on our previous study where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
-Eggs were then incubated at either cold  (23 ± 3 ºC) or hot (28 ± 3 ºC) incubation temperatures in a 2x2 factorial design (Fig.
-1 A).
+We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels by approximately 2 standard deviations above the mean natural concentration [@crino2024eggs].
+Eggs were then incubated at either cold  (23 ± 3 ºC) or hot (28 ± 3 ºC) (see @fig-Methods).
 These temperatures are within the natural limits in _L.

### 2025-10-14 ms.qmd
Update ms.qmd

@@
 ### Data accessibility  
-All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT-Temp_PreyD](https://github.com/Pablo-Recio/CORT_Temp_PreyD).
+All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT_Temp_PreyD](https://github.com/Pablo-Recio/CORT_Temp_PreyD).
 
